### 2010-02-15 doc/slides.tex
*** empty log message ***

\documentclass{beamer}
%\documentclass[handout]{beamer}

%\usepackage{beamerjcf}
\usepackage[latin1]{inputenc}

\definecolor{kwblue}{rgb}{0.67,0.12,0.92}
\definecolor{ceruleanblue}{rgb}{0, 0.48, 0.65}
\definecolor{lightpink}{rgb}{1., 0.71, 0.75}
\definecolor{lightblue}{rgb}{0.8,0.8,1}
\definecolor{lightred}{rgb}{1,0.8,0.8}

\begin{document}

\title{Map/Reduce for the common man}
\author[Jean-Christophe]{Jean-Christophe Filli\^atre \& Kalyan Krishnamani}
\institute{
}
\date{ProVal, February 19, 2010}

\begin{frame}
  \titlepage
  \pgfimage[height=8mm]{cnrs-logo2}\hfill
  \pgfimage[height=6mm]{saclay}\hfill
  \pgfimage[height=10mm]{lrilogo}\hfill
  \pgfimage[height=8mm]{upsudlogo}
\end{frame}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2010-02-15 doc/whizzy.sh

@@ -1,0 +1,1 @@
+#export MKSLICE="latexpp -w -g color yes -e file lightgreen-tt  -e reponse lightred-tt -e pascal pascal-lightblue-tt"

### 2010-02-16 doc/slides.tex
*** empty log message ***

@@
 \definecolor{lightred}{rgb}{1,0.8,0.8}
+
+\let\emph\alert
 
@@
 
+\begin{frame}\frametitle{Overview}
+  \begin{enumerate}
+  \item Google's Map/Reduce
+  \item Ocaml's Map/Reduce
+  \item Implementation details
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Google's Map/Reduce}
+  what it is
+
+  what it offers
+
+  application domains
+
+\end{frame}
+
+\begin{frame}
+  \begin{center}
+    \hrulefill\\
+    \emph{Ocaml's Map/Reduce}
+  \end{center}
+\end{frame}
+
+\begin{frame}\frametitle{Two Ideas}
+  \begin{enumerate}
+  \item polymorphic map/fold instead of map/reduce
+
+    \bigskip
+  \item several library options
+    \begin{itemize}
+    \item sequential 
+    \item several cores on the same machine
+    \item network
+    \end{itemize}
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}[fragile]\frametitle{API}
+%\footnotesize
+  \begin{itemize}
+  \item traditional \texttt{List.map}
+\begin{verbatim}
+  val map : 
+    f:('a -> 'b) -> 'a list -> 'b list
+\end{verbatim}
+
+    \bigskip
+  \item traditional combination of \texttt{List.fold} and \texttt{List.map}
+
+\begin{verbatim}
+  val map_reduce :
+    map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+    'c -> 'a list -> 'c
+\end{verbatim}
+    \begin{displaymath}
+      \mathtt{map\_reduce}~f~g~c~l
+      \begin{array}[t]{l}
+        = \mathtt{fold}~g~c~(\mathtt{map}~f~l) \\
+        = \mathtt{fold}~(\mathtt{fun}~c~x\rightarrow g~c~(f~x))~c~l\\
+      \end{array}
+    \end{displaymath}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]\frametitle{}
+\begin{verbatim}
+val map_reduce_ac :
+  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+  'b -> 'a list -> 'b
+
+val map_reduce_a :
+  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+  'b -> 'a list -> 'b
+\end{verbatim}
+\end{frame}
+
+\begin{frame}\frametitle{Options}
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|c|c|}
+      \hline
+                 & \multicolumn{2}{|c|}{map} &
+                 \multicolumn{3}{|c|}{map/reduce} \\\cline{2-6}
+                 & map & local & remote & ac & a \\\hline\hline
+      sequential & & & & & \\\hline
+      cores      & & & & & \\\hline
+      network    & & & & & \\\hline
+    \end{tabular}
+  \end{center}
+\end{frame}
+
+\begin{frame}\frametitle{}
+  first demo
+\end{frame}
+
+\begin{frame}\frametitle{}
+  detailed API
+\end{frame}
+
+\begin{frame}\frametitle{}
+  another, more impressive demo ?
+\end{frame}
+
+\begin{frame}
+  \begin{center}
+    \hrulefill\\
+    \emph{Implementation details}
+  \end{center}
+\end{frame}
+
+\begin{frame}\frametitle{}
+  common part between cores and network
+
+  output/input\_value on core
+
+  architecture-independent protocol for network
+
+  same program for master and workers
+\end{frame}
+
 \end{document}

### 2010-02-16 doc/slides.tex
*** empty log message ***

@@
 
+\begin{frame}\frametitle{}
+  TODO: introduce cores/network idea
+\end{frame}
+
 \begin{frame}\frametitle{Two Ideas}
@@
 
-\begin{frame}[fragile]\frametitle{API}
+\begin{frame}\frametitle{API}
 %\footnotesize
@@
   \item traditional \texttt{List.map}
-\begin{verbatim}
+\begin{ocaml}
   val map : 
     f:('a -> 'b) -> 'a list -> 'b list
-\end{verbatim}
+\end{ocaml}
 
@@
 
-\begin{verbatim}
+\begin{ocaml}
   val map_reduce :
@@
     'c -> 'a list -> 'c
-\end{verbatim}
+\end{ocaml}
     \begin{displaymath}
@@
     \end{displaymath}
+    assuming $g$ is not sensitive to element ordering
   \end{itemize}
+
+  example :
+  \begin{displaymath}
+    \sum_{x\in l}f(x) = \mathtt{map\_reduce}~f~(+)~0~l
+  \end{displaymath}
 \end{frame}
 
-\begin{frame}[fragile]\frametitle{}
-\begin{verbatim}
+\begin{frame}\frametitle{Special Case}
+  \emph{idea}: reduce as early as possible, as in
+  \begin{displaymath}
+    \sum_{x\in l}f(x) = \mathtt{map\_reduce}~f~(+)~0~l
+  \end{displaymath}
+  in this case, + is \emph{associative / commutative}, so reductions are
+  order-insensitive and can be performed as soon as two intermediate
+  results are available
+
+  \begin{ocaml}
 val map_reduce_ac :
@@
   'b -> 'a list -> 'b
+  \end{ocaml}
+\end{frame}
 
+\begin{frame}\frametitle{Special Case 2}
+
+  \emph{order-sensitive} computation:
+  \begin{displaymath}
+    \mathtt{map\_reduce}~\mathtt{String.uppercase}~(\texttt{\^{}})~\texttt{""}~
+    \texttt{["h"; "e"; "l"; "l"; "o"]}
+  \end{displaymath}
+
+  operation is only \emph{associative}, yet we can reduce as soon as
+  adjacent intermediate results are available
+
+\begin{ocaml}
 val map_reduce_a :
@@
   'b -> 'a list -> 'b
-\end{verbatim}
+\end{ocaml}
 \end{frame}
 
-\begin{frame}\frametitle{Options}
+\begin{frame}\frametitle{Local/Remote reductions}
+  reductions can be performed either by the master (local) or by the
+  workers (remote)
+
+  \bigskip
+  remote reductions allow other computations to be performed in parallel
+
+  \texttt{map\_reduce} takes the following forms
+  \begin{ocaml}
+val map_local_reduce :
+  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+val map_remote_reduce :
+  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+  \end{ocaml}
+\end{frame}
+
+\begin{frame}\frametitle{API Summary}
+  \begin{ocaml}
+val map : 
+  f:('a -> 'b) -> 'a list -> 'b list
+val map_local_reduce :
+  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+val map_remote_reduce :
+  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+val map_reduce_ac :
+  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+  'b -> 'a list -> 'b
+val map_reduce_a :
+  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+  'b -> 'a list -> 'b
+  \end{ocaml}
+\end{frame}
+
+\begin{frame}\frametitle{Library Options}
+
+  3 modules, with 5 functions each
+
   \begin{center}
-    \begin{tabular}{|c|c|c|c|c|c|}
+    \begin{tabular}{|l|c|c|c|c|c|}
       \hline
@@
                  \multicolumn{3}{|c|}{map/reduce} \\\cline{2-6}
-                 & map & local & remote & ac & a \\\hline\hline
-      sequential & & & & & \\\hline
-      cores      & & & & & \\\hline
-      network    & & & & & \\\hline
+                 & map & local & remote & ~ac~ & ~~a~~ \\\hline\hline
+      \texttt{Sequential} & & & \multicolumn{3}{|c|}{(no difference)} \\\hline
+      \texttt{Cores}      & & & & & \\\hline
+      \texttt{Network}    & & & & & \\\hline
     \end{tabular}
   \end{center}
+  % TODO: fill the cells
 \end{frame}
 
-\begin{frame}\frametitle{}
-  first demo
+\begin{frame}\frametitle{Demo}
+  \emph{demo}: solving the $n$-queens puzzle
 \end{frame}
 
-\begin{frame}\frametitle{}
-  detailed API
-\end{frame}
+\begin{frame}\frametitle{Demo}
+  \emph{demo}: solving the $n$-queens puzzle
 
-\begin{frame}\frametitle{}
-  another, more impressive demo ?
+  example: $n=16$
+  \begin{center}
+    \begin{tabular}{|l|r|}
+      \hline
+      sequential & 24.4 s \\\hline
+      2 cores    & 12.6 s \\\hline
+      12 cores on moloch & 3.0 s \\\hline
+    \end{tabular}
+  \end{center}
+
+  example: $n=17$
+  \begin{center}
+    \begin{tabular}{|l|r|}
+       \hline
+      sequential & 172.0 s \\\hline
+      moloch (12), orcus (4), localhost(1) & 11.7 s \\\hline
+    \end{tabular}
+  \end{center}
 \end{frame}

### 2010-02-16 doc/slides.tex
*** empty log message ***

@@
 
+\begin{frame}\frametitle{Sequential}
+the sequential implementation is obvious
+
+  \begin{ocaml}
+let map ~f l = List.map f l
+
+let map_local_reduce ~map ~reduce acc l =
+  List.fold_left (fun a x -> reduce a (map x)) acc l
+
+let map_remote_reduce = map_local_reduce
+
+let map_reduce_ac = map_local_reduce
+
+let map_reduce_a = map_local_reduce
+  \end{ocaml}
+\end{frame}
+
+\begin{frame}\frametitle{Cores / Network}
+
+  \texttt{Cores} and \texttt{Network} share a main routine
+
+  abstract notions of \emph{workers} and \emph{tasks}; we can
+  \begin{enumerate}
+  \item assign a task to a worker
+  \item wait for a completed job, which may generate new tasks
+  \end{enumerate}
+
+  so the main loop is
+  \begin{lightblue-tt}
+    while pending tasks || pending jobs
+      while pending tasks && idle workers
+        assign any task to any idle worker
+      wait for any completed job
+        push new pending tasks
+  \end{lightblue-tt}
+
+\end{frame}
+
 \begin{frame}\frametitle{}

### 2010-02-16 doc/slides.tex
*** empty log message ***

@@
 
+  \begin{ocaml}
+master_loop : 
+  create_job:('worker -> 'task -> unit) -> 
+  wait:(unit -> 'worker * 'task list) -> 
+  'worker list -> 'task list -> unit
+  \end{ocaml}
+\end{frame}
+
+\begin{frame}\frametitle{Cores}
+  \begin{itemize}\item 
+    \texttt{create\_job} is achieved by \texttt{Unix.fork}, using a
+    local file to store the result (Unix pipes have size limitations)
+
+  \item 
+    \texttt{wait} is achieved using \texttt{Unix.wait}
+  \end{itemize}
+
+  Ocaml's marshalling is safely exploited to read/write values of any type
+
+  scheduling tasks to various cores is left to the OS
+\end{frame}
+
+\begin{frame}\frametitle{Network}
+  master and workers are the same program but distinguish using an
+  environment variable (\texttt{WORKER})
+
+  \begin{itemize}
+  \item a worker starts a server, waiting for tasks
+  \item the master connects to workers, sending tasks
+  \end{itemize}
+
+  simple, string-based protocol:
+  \begin{center}
+    \begin{tabular}{ccc}
+      \framebox{master} &  & \framebox{worker} \\
+      &     $\stackrel{\displaystyle connect}{\longrightarrow}$ &  \\
+      &     $\stackrel{\displaystyle\mathtt{Assign}~3~\mathtt{"f"}~\mathtt{"7"}}
+                      {\longrightarrow}$ &  \\
+      &     $\stackrel{\displaystyle \mathtt{Started}~3}
+                      {\longleftarrow}$ &  \\
+      &     $\stackrel{\displaystyle 
+                       \only<1>{\mathtt{Completed}~3~\mathtt{"42"}}
+                       \only<2>{\emph{\mathtt{Aborted}~3~}}
+                       \only<3>{\emph{\mathtt{Kill}~3}}}
+                      {\only<1-2>{\longleftarrow}
+                       \only<3>{\longrightarrow}}$ &  \\
+    \end{tabular}
+  \end{center}
+\end{frame}
+
+\begin{frame}\frametitle{Network: Polymorphic Map/Reduce}
+  if master and workers are \emph{exactly} the same executable, then
+  marshalling can still be used to pass values $\Rightarrow$
+  polymorphic map/reduce
+
+  if master and workers are compiled from the \emph{same source}, then
+  we are limited to strings $\Rightarrow$
+  monomorphic map/reduce
+
+  \footnotesize
+  \begin{ocaml}
+module Str : sig
+
+  val map : f:(string -> string) -> string list -> string list
+
+  val map_local_reduce :
+    map:(string -> string) -> 
+    reduce:(string -> string -> string) -> 
+    string -> string list -> string
+  ...
+  \end{ocaml}
+\end{frame}
+
+\begin{frame}\frametitle{Master}
+  whatever the implementation is (cores or network), we now 
+  provide the same generic master loop
+
+  \begin{ocaml}
+    master : (task -> task_result) -> 
+             (task -> task_result -> task list) ->
+             task list -> 
+             unit
+  \end{ocaml}
+
+  (the notion of workers is no more explicit)
+\end{frame}
+
+\begin{frame}\frametitle{Implementing \texttt{map}}
+  \begin{ocaml}
+val map : f:('a -> 'b) -> 'a list -> 'b list
+  \end{ocaml}
+
+  \begin{enumerate}
+  \item initially, the $i$-th element $x_i$ of the list is the task $(i,x_i)$
+  \item when task $(i,x_i)$ completes, store the result $f(x_i)$ in a table
+  \item when done, rebuild the list from the table contents
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Implementing \texttt{map\_local\_reduce}}
+  \begin{ocaml}
+ val map_local_reduce :
+  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+  \end{ocaml}
+
+  \begin{enumerate}
+  \item initially, each element $x_i$ of the list is a task
+  \item when task $x_i$ completes, the result $f(x_i)$ is
+    \emph{locally} combined with the current accumulator (a local reference)
+  \item when done, return the accumulator
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Implementing \texttt{map\_remote\_reduce}}
+  \begin{ocaml}
+val map_remote_reduce :
+  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+  \end{ocaml}
+
+  \begin{enumerate}
+  \item initially, each element $x_i$ in the list is a task $Map(x_i)$
+  \item when a task is completed with result $r$,
+    \begin{itemize}
+    \item either it is $Map(x_i)$,
+      \begin{itemize}
+      \item if accumulator is $\texttt{Some}~v$, generate the task
+        $Reduce(v,r)$ and set accumulator to \texttt{None}
+      \item if accumulator is \texttt{None}, push $r$ into $\mathit{stack}_r$
+      \end{itemize}
+    \item or it is $Reduce(r',v')$
+      \begin{itemize}
+      \item if $\mathit{stack}_r$ is empty, the accumulator is now
+        $\texttt{Some}~r$ 
+      \item otherwise, pop $v$ from $\mathit{stack}_r$, and generate
+        $Reduce(r,v)$ 
+      \end{itemize}
+    \end{itemize}
+  \item when done, return the accumulator
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Implementing \texttt{map\_reduce\_ac}}
+  \begin{ocaml}
+val map_reduce_ac :
+  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+  'b -> 'a list -> 'b
+  \end{ocaml}
+
+  \begin{enumerate}
+  \item initially, each element $x_i$ in the list is a task $Map(x_i)$
+  \item when a task is completed with result $r$,
+    \begin{itemize}
+    \item if the accumulator is \texttt{None}, store $r$ in the accumulator
+    \item else, combine $r$ with the accumulator $v$ into a new task
+      $Reduce(v, r)$ and set accumulator to \texttt{None}
+    \end{itemize}
+  \item when done, return the accumulator
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Implementing \texttt{map\_reduce\_a}}
+  \begin{ocaml}
+val map_reduce_a :
+  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+  'b -> 'a list -> 'b
+  \end{ocaml}
+
+  \begin{enumerate}
+  \item initially, each element $x_i$ in the list is a task $(i,i,Map(x_i))$
+  \item when a task $(i,j)$ is completed with result $r$,
+    \begin{itemize}
+    \item if there is some result adjacent to $(i,j)$ available,
+      merge them into a new reduce task
+    \item else, store it
+    \end{itemize}
+  \item when done, return the only remaining result
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Independent Master and Workers}
+  the library actually provides tools for independent development of
+  master \small
+
+  \begin{ocaml}
+module Master : sig
+  ...
+  val map_remote_reduce : string -> string list -> string
+  ...
+end
+  \end{ocaml}
+
+and workers\small
+
+  \begin{ocaml}
+module Worker : sig
+  val register_computation : 
+    string -> (string -> string) -> unit
+  val register_computation2 : 
+    string -> (string -> string -> string) -> unit
+  val compute : 
+    ?stop:bool -> ?port:int -> unit -> string
+end
+  \end{ocaml}
+\end{frame}
+
+\begin{frame}\frametitle{Future Work}
+
 \end{frame}
@@
 \begin{frame}\frametitle{}
-  common part between cores and network
-
-  output/input\_value on core
-
-  architecture-independent protocol for network
-
-  same program for master and workers
+  \begin{center}
+    \emph{thank you!}
+  \end{center}
 \end{frame}

### 2010-02-17 doc/slides.tex
*** empty log message ***

@@
 \begin{frame}\frametitle{Google's Map/Reduce}
+  \begin{itemize}
+  \item 
     what it is
+    \begin{itemize}
+    \item a paradigm for distributed computing at Google
+    \end{itemize}
 
+    \bigskip
+  \item 
     what it offers
+    \begin{itemize}
+    \item abstraction to hide parallelization and distribution from
+      user level code
+    \end{itemize}
 
+    \bigskip
+  \item 
     application domains
+    \begin{itemize}
+    \item large scale data processing
+    \end{itemize}
+  \end{itemize}
 
+  \vfill
+  [MapReduce: Simplified Data Processing on Large Clusters. \par
+  J. Dean and S. Ghemawat. OSDI 2004]
+\end{frame}
+
+\begin{frame}\frametitle{Main Idea}
+  the user provides two functions
+  \begin{itemize}
+  \item \emph{map}: takes a key/value pair and returns a list of key/value
+    pairs
+    \begin{ocaml}
+    map : k1 * v1 -> (k2 * v2) list
+    \end{ocaml}
+
+  \item \emph{reduce}: takes a list of values for one key and returns
+    another list for that key
+    \begin{ocaml}
+    reduce : k2 * v2 list -> v2 list
+    \end{ocaml}
+  \end{itemize}
+
+  and the library takes care of the rest
+\end{frame}
+
+\begin{frame}\frametitle{}
+  \begin{center}
+    \includegraphics[width=\textwidth]{mrwordcount}
+  \end{center}
+\end{frame}
+
+\begin{frame}\frametitle{}
+  \begin{center}
+    \includegraphics[width=\textwidth]{mrfigure}
+  \end{center}
 \end{frame}
@@
   \end{center}
+\end{frame}
+
+\begin{frame}\frametitle{Ocaml's Map/Reduce}
+  \begin{itemize}
+  \item 
+    what it is
+    \begin{itemize}
+    \item a distributed computing library for Ocaml users
+    \end{itemize}
+
+    \bigskip
+  \item 
+    what it offers
+    \begin{itemize}
+    \item abstraction to hide parallelization and distribution from
+      user level code
+    \item \emph{polymorphism}
+    \item \emph{multi-cores or clusters}
+    \end{itemize}
+
+    \bigskip
+  \item 
+    application domains
+    \begin{itemize}
+    \item large scale data processing
+    \item \emph{and much more}
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}\frametitle{Future Work}
-
+  \begin{itemize}
+  \item fault tolerance
+    \begin{itemize}
+    \item blacklist faulty workers
+    \item ping workers periodically
+    \end{itemize}
+  \item use idle workers to speedup remaining tasks
+  \end{itemize}
 \end{frame}
@@
 %%% End:
+
+% LocalWords:  parallelization Google

### 2010-02-18 doc/slides.tex
a new figure and intro to cores/network slide added [KK]

@@
 
+\begin{frame}\frametitle{}
+  \begin{center}
+    \includegraphics[width=\textwidth]{mr-new}
+  \end{center}
+\end{frame}
+
 \begin{frame}
@@
 
-\begin{frame}\frametitle{}
-  TODO: introduce cores/network idea
+\begin{frame}\frametitle{Cores \& Networks}
+  Our map/reduce library works seemlessly across
+
+  \bigskip
+  \begin{itemize}
+  \item multiple processor cores on the same (module \texttt{Cores})
+    machine
+    \begin{itemize}
+    \item user just needs to specify the number of cores to be used
+    \item uses Ocaml's \texttt{Unix} module
+    \end{itemize}
+
+    \bigskip
+  \item networked cluster of machines (module \texttt{Network})
+    \begin{itemize}
+    \item user specifies the machine name(s) to be used
+    \item relies on OCaml's \texttt{Unix} and \texttt{Marshal} modules
+    \end{itemize}
+  \end{itemize}
 \end{frame}

### 2010-02-18 doc/whizzy.sh
*** empty log message ***

@@ -1,1 +1,1 @@
-#export MKSLICE="latexpp -w -g color yes -e file lightgreen-tt  -e reponse lightred-tt -e pascal pascal-lightblue-tt"
+export MKSLICE="latexpp -w -g color yes -e file lightgreen-tt  -e reponse lightred-tt -e pascal pascal-lightblue-tt"

### 2010-02-18 doc/slides.tex
*** empty log message ***
reduce -> fold

@@
 \author[Jean-Christophe]{Jean-Christophe Filli\^atre \& Kalyan Krishnamani}
-\institute{
+\institute{\includegraphics[scale=0.5]{Common_man}
 }
@@
 
-\begin{frame}\frametitle{}
-  \begin{center}
-    \includegraphics[width=\textwidth]{mrwordcount}
-  \end{center}
-\end{frame}
+% \begin{frame}\frametitle{}
+%   \begin{center}
+%     \includegraphics[width=\textwidth]{mrwordcount}
+%   \end{center}
+% \end{frame}
 
@@
 \begin{frame}\frametitle{Cores \& Networks}
-  Our map/reduce library works seemlessly across
+  Ocaml's map/reduce library works seemlessly across
 
@@
 \begin{ocaml}
- val map_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+ val map_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
@@
     \begin{displaymath}
-      \mathtt{map\_reduce}~f~g~c~l
+      \mathtt{map\_fold}~f~g~c~l
       \begin{array}[t]{l}
@@
   \begin{displaymath}
-    \sum_{x\in l}f(x) = \mathtt{map\_reduce}~f~(+)~0~l
+    \sum_{x\in l}f(x) = \mathtt{map\_fold}~f~(+)~0~l
   \end{displaymath}
@@
 \begin{frame}\frametitle{Special Case}
-  \emph{idea}: reduce as early as possible, as in
+  \emph{idea}: fold as early as possible, as in
   \begin{displaymath}
-    \sum_{x\in l}f(x) = \mathtt{map\_reduce}~f~(+)~0~l
+    \sum_{x\in l}f(x) = \mathtt{map\_fold}~f~(+)~0~l
   \end{displaymath}
@@
   \begin{ocaml}
-val map_reduce_ac :
-  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+val map_fold_ac :
+  map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
   'b -> 'a list -> 'b
@@
   \begin{displaymath}
-    \mathtt{map\_reduce}~\mathtt{String.uppercase}~(\texttt{\^{}})~\texttt{""}~
+    \mathtt{map\_fold}~\mathtt{String.uppercase}~(\texttt{\^{}})~\texttt{""}~
     \texttt{["h"; "e"; "l"; "l"; "o"]}
@@
 
-  operation is only \emph{associative}, yet we can reduce as soon as
+  operation is only \emph{associative}, yet we can fold as soon as
   adjacent intermediate results are available
@@
 \begin{ocaml}
-val map_reduce_a :
-  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+val map_fold_a :
+  map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
   'b -> 'a list -> 'b
@@
 
-  \texttt{map\_reduce} takes the following forms
+  \texttt{map\_fold} takes the following forms
   \begin{ocaml}
-val map_local_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+val map_local_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
-val map_remote_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+val map_remote_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
@@
   f:('a -> 'b) -> 'a list -> 'b list
-val map_local_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+val map_local_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
-val map_remote_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+val map_remote_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
-val map_reduce_ac :
-  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+val map_fold_ac :
+  map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
   'b -> 'a list -> 'b
-val map_reduce_a :
-  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+val map_fold_a :
+  map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
   'b -> 'a list -> 'b
@@
                  & \multicolumn{2}{|c|}{map} &
-                 \multicolumn{3}{|c|}{map/reduce} \\\cline{2-6}
+                 \multicolumn{3}{|c|}{map/fold} \\\cline{2-6}
                  & map & local & remote & ~ac~ & ~~a~~ \\\hline\hline
@@
 
-let map_local_reduce ~map ~reduce acc l =
-  List.fold_left (fun a x -> reduce a (map x)) acc l
+let map_local_fold ~map ~fold acc l =
+  List.fold_left (fun a x -> fold a (map x)) acc l
 
-let map_remote_reduce = map_local_reduce
+let map_remote_fold = map_local_fold
 
-let map_reduce_ac = map_local_reduce
+let map_fold_ac = map_local_fold
 
-let map_reduce_a = map_local_reduce
+let map_fold_a = map_local_fold
   \end{ocaml}
@@
 
-  val map_local_reduce :
+  val map_local_fold :
     map:(string -> string) -> 
-    reduce:(string -> string -> string) -> 
+    fold:(string -> string -> string) -> 
     string -> string list -> string
@@
 
-\begin{frame}\frametitle{Implementing \texttt{map\_local\_reduce}}
+\begin{frame}\frametitle{Implementing \texttt{map\_local\_fold}}
   \begin{ocaml}
- val map_local_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+ val map_local_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
@@
 
-\begin{frame}\frametitle{Implementing \texttt{map\_remote\_reduce}}
+\begin{frame}\frametitle{Implementing \texttt{map\_remote\_fold}}
   \begin{ocaml}
-val map_remote_reduce :
-  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
+val map_remote_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
@@
       \item if accumulator is $\texttt{Some}~v$, generate the task
-        $Reduce(v,r)$ and set accumulator to \texttt{None}
+        $Fold(v,r)$ and set accumulator to \texttt{None}
       \item if accumulator is \texttt{None}, push $r$ into $\mathit{stack}_r$
       \end{itemize}
-    \item or it is $Reduce(r',v')$
+    \item or it is $Fold(r',v')$
       \begin{itemize}
@@
       \item otherwise, pop $v$ from $\mathit{stack}_r$, and generate
-        $Reduce(r,v)$ 
+        $Fold(r,v)$ 
       \end{itemize}
@@
 
-\begin{frame}\frametitle{Implementing \texttt{map\_reduce\_ac}}
+\begin{frame}\frametitle{Implementing \texttt{map\_fold\_ac}}
   \begin{ocaml}
-val map_reduce_ac :
-  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+val map_fold_ac :
+  map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
   'b -> 'a list -> 'b
@@
     \item else, combine $r$ with the accumulator $v$ into a new task
-      $Reduce(v, r)$ and set accumulator to \texttt{None}
+      $Fold(v, r)$ and set accumulator to \texttt{None}
     \end{itemize}
@@
 
-\begin{frame}\frametitle{Implementing \texttt{map\_reduce\_a}}
+\begin{frame}\frametitle{Implementing \texttt{map\_fold\_a}}
   \begin{ocaml}
-val map_reduce_a :
-  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
+val map_fold_a :
+  map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
   'b -> 'a list -> 'b
@@
     \item if there is some result adjacent to $(i,j)$ available,
-      merge them into a new reduce task
+      merge them into a new fold task
     \item else, store it
@@
   ...
-  val map_remote_reduce : string -> string list -> string
+  val map_remote_fold : string -> string list -> string
   ...

### 2010-02-19 doc/slides.tex
swapped slides regarding cores/network [KK]
slight change of API for map_reduce
*** empty log message ***

@@
 
-\begin{frame}\frametitle{}
+\begin{frame}\frametitle{Google's Map/Reduce}
   \begin{center}
@@
 
-\begin{frame}\frametitle{}
+\begin{frame}\frametitle{Example}
   \begin{center}
-    \includegraphics[width=\textwidth]{mr-new}
+    \includegraphics[width=0.8\textwidth]{mr-new}
   \end{center}
@@
 
-\begin{frame}\frametitle{Cores \& Networks}
-  Ocaml's map/reduce library works seemlessly across
-
-  \bigskip
-  \begin{itemize}
-  \item multiple processor cores on the same (module \texttt{Cores})
-    machine
-    \begin{itemize}
-    \item user just needs to specify the number of cores to be used
-    \item uses Ocaml's \texttt{Unix} module
-    \end{itemize}
-
-    \bigskip
-  \item networked cluster of machines (module \texttt{Network})
-    \begin{itemize}
-    \item user specifies the machine name(s) to be used
-    \item relies on OCaml's \texttt{Unix} and \texttt{Marshal} modules
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
 \begin{frame}\frametitle{Two Ideas}
@@
   \end{enumerate}
+\end{frame}
+
+\begin{frame}\frametitle{Library Options}
+  % Ocaml's map/reduce library works seemlessly across
+
+  \begin{itemize}
+  \item \emph{module \texttt{Sequential}}: pure sequential evaluation
+    \begin{itemize}\item 
+      used as a reference
+    \end{itemize}
+
+    \bigskip
+  \item \emph{module \texttt{Cores}}: multiple processor cores on the same 
+    machine
+    \begin{itemize}
+    \item user just needs to specify the number of cores to be used
+    \item only uses Ocaml's \texttt{Unix} module
+    \end{itemize}
+
+    \bigskip
+  \item \emph{module \texttt{Network}}: networked cluster of machines
+    \begin{itemize}
+    \item user specifies the machine name(s) and number of cores on
+      each machine
+    \item only uses Ocaml's \texttt{Unix} module
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
   'c -> 'a list -> 'c
+
 val map_remote_fold :
@@
 
+  map/fold is performed 
+  \begin{itemize}
+  \item \textcolor{orange}{locally} / \textcolor{red}{remotely}
+  \end{itemize}
+
+    \newcommand{\Lo}{\colorbox{orange}{~~}\,}
+    \renewcommand{\Re}{\colorbox{red}{~~}\,}
+    \newcommand{\Se}{\colorbox{blue}{~~}\,}
+    \newcommand{\Pa}{\colorbox{kwblue}{~~}\,}
+
   \begin{center}
@@
       \hline
-                 & \multicolumn{2}{|c|}{map} &
-                 \multicolumn{3}{|c|}{map/fold} \\\cline{2-6}
-                 & map & local & remote & ~ac~ & ~~a~~ \\\hline\hline
-      \texttt{Sequential} & & & \multicolumn{3}{|c|}{(no difference)} \\\hline
-      \texttt{Cores}      & & & & & \\\hline
-      \texttt{Network}    & & & & & \\\hline
+        map    & map & local & remote & ac & a 
+            \\\hline\hline
+      \texttt{Sequential} & \Lo & \Lo & \Lo & \Lo & \Lo \\\hline
+      \texttt{Cores}   & \Re & \Re & \Re & \Re & \Re \\\hline
+      \texttt{Network}    & \Re & \Re & \Re & \Re & \Re \\\hline
+    \end{tabular} \\[1em]
+  \end{center}
+
+  \begin{itemize}
+    \item \textcolor{blue}{sequentially} / \textcolor{kwblue}{in parallel}
+  \end{itemize}
+
+  \begin{center}
+    \begin{tabular}{|l|c|c|c|c|}
+      \hline
+        fold    & local & remote & ac / a 
+            \\\hline\hline
+      \texttt{Sequential} & \Lo\Se & \Lo\Se & \Lo\Se  
+            \\\hline
+      \texttt{Cores}   & \Lo\Se & \Re\Se & \Re\Pa  
+            \\\hline
+      \texttt{Network}    & \Lo\Se & \Re\Se & \Re\Pa  \\\hline
     \end{tabular}
@@
 \begin{frame}\frametitle{Network}
-  master and workers are the same program but distinguish using an
+  master and workers are the same program but distinguished using an
   environment variable (\texttt{WORKER})
@@
     \begin{itemize}
-    \item either it is $Map(x_i)$,
+    \item either it was $Map(x_i)$,
       \begin{itemize}
@@
       \end{itemize}
-    \item or it is $Fold(r',v')$
+    \item or it was $Fold(r',v')$
       \begin{itemize}
@@
 
+\begin{frame}\frametitle{What about Map/Reduce a la Google?}
+  we also provide map/reduce a la Google
+
+  \begin{ocaml}
+  val map_reduce :
+    map:('v1 -> ('k2 * 'v2) list) ->
+    reduce:('k2 -> 'v2 list list -> 'v2 list) ->
+    'v1 list -> ('k2 * 'v2 list) list
+  \end{ocaml}
+\end{frame}
+
 \begin{frame}\frametitle{Future Work}

### 2010-02-19 doc/slides.tex
*** empty log message ***

@@
 
+\begin{frame}\frametitle{Local/Remote folds}
+  folds can be performed either by the master (local) or by the
+  workers (remote)
+
+  \bigskip
+  remote folds allow other computations to be performed in parallel
+
+  \texttt{map\_fold} takes the following forms
+  \begin{ocaml}
+val map_local_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+
+val map_remote_fold :
+  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+  'c -> 'a list -> 'c
+  \end{ocaml}
+\end{frame}
+
 \begin{frame}\frametitle{Special Case}
@@
   \end{displaymath}
-  in this case, + is \emph{associative / commutative}, so reductions are
+  in this case, + is \emph{associative / commutative}, so folds are
   order-insensitive and can be performed as soon as two intermediate
@@
   'b -> 'a list -> 'b
-\end{ocaml}
-\end{frame}
-
-\begin{frame}\frametitle{Local/Remote reductions}
-  reductions can be performed either by the master (local) or by the
-  workers (remote)
-
-  \bigskip
-  remote reductions allow other computations to be performed in parallel
-
-  \texttt{map\_fold} takes the following forms
-  \begin{ocaml}
-val map_local_fold :
-  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-  'c -> 'a list -> 'c
-
-val map_remote_fold :
-  map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-  'c -> 'a list -> 'c
 \end{ocaml}
@@
 
-    \newcommand{\Lo}{\colorbox{orange}{~~}\,}
-    \renewcommand{\Re}{\colorbox{red}{~~}\,}
-    \newcommand{\Se}{\colorbox{blue}{~~}\,}
-    \newcommand{\Pa}{\colorbox{kwblue}{~~}\,}
+    \newcommand{\Lo}{\colorbox{orange}{~\textcolor{white}{\small L}~}\,}
+    \renewcommand{\Re}{\colorbox{red}{~\textcolor{white}{\small R}~}\,}
+    \newcommand{\Se}{\colorbox{blue}{~\textcolor{white}{\small S}~}\,}
+    \newcommand{\Pa}{\colorbox{kwblue}{~\textcolor{white}{\small P}~}\,}
 
@@
   \item the master connects to workers, sending tasks
+  \item when done, the master sends result to the workers
   \end{itemize}
@@
       &     $\stackrel{\displaystyle connect}{\longrightarrow}$ &  \\
-      &     $\stackrel{\displaystyle\mathtt{Assign}~3~\mathtt{"f"}~\mathtt{"7"}}
+      &     \pause $\stackrel{\displaystyle\mathtt{Assign}~3~\mathtt{"f"}~\mathtt{"7"}}
                       {\longrightarrow}$ &  \\
-      &     $\stackrel{\displaystyle \mathtt{Started}~3}
+      &     \pause $\stackrel{\displaystyle \mathtt{Started}~3}
                       {\longleftarrow}$ &  \\
-      &     $\stackrel{\displaystyle 
-                       \only<1>{\mathtt{Completed}~3~\mathtt{"42"}}
-                       \only<2>{\emph{\mathtt{Aborted}~3~}}
-                       \only<3>{\emph{\mathtt{Kill}~3}}}
-                      {\only<1-2>{\longleftarrow}
-                       \only<3>{\longrightarrow}}$ &  \\
+      &     \pause $\stackrel{\displaystyle 
+                       \only<4>{\mathtt{Completed}~3~\mathtt{"42"}}
+                       \only<5>{\mathtt{Result}~\mathtt{"89"}~}
+                       \only<6>{\emph{\mathtt{Aborted}~3~}}
+                       \only<7>{\emph{\mathtt{Kill}~3}}
+                      }
+                      {\only<4,6>{\longleftarrow}
+                       \only<5,7>{\longrightarrow}}$ &  \\
     \end{tabular}
@@
 \begin{frame}\frametitle{Network: Polymorphic Map/Reduce}
-  if master and workers are \emph{exactly} the same executable, then
+  if master and workers are compiled from the \emph{same source} using the
+  \emph{same Ocaml version}, then
   marshalling can still be used to pass values $\Rightarrow$

### 2010-03-15 doc/icfp-2010/main.tex
*** empty log message ***

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}

\begin{document}

\conferenceinfo{ICFP '10}{September 27-29, 2010, Baltimore, Maryland.} 
\copyrightyear{2010} 
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Experience Report: }
\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\terms
term1, term2

\keywords
keyword1, keyword2

\section{Introduction}

what it is and what it is not
\begin{itemize}
\item it is not a library which does parallelization
\item it is a library for reliable distributed execution of
  parallelizable computation 
\end{itemize}

inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
programming, ironically) ;  however, there are differences:
\begin{itemize}
\item polymorphic, higher-order API
\item more generic (cores/network)
\item does not focus on association lists
\item we don't have data locality (future work)
\end{itemize}

what is the target audience/applications: 
for instance, use if automatic provers on thousands of verification
conditions, on computing infrastructures which can be
\begin{itemize}
\item a single machine with multiple cores, possibly remote
\item several machines, small or large in number, over a network
\end{itemize}

\section{API}

main idea of the master function

four scenarios:
\begin{itemize}
\item cores
\item same binary
\item same version of ocaml
\item completely different programs
\end{itemize}

\section{Derived API}

map, map\_fold, etc.

generic programming using a functor, which is then applied in each of
the 4 cases

\section{Implementation Details}

cores?

network: protocol, fault tolerance, etc.

\section{Experiments}

n-queens, SMT solvers, etc.

\section{Future Work}

data locality: scheduling supplied by the user

% \appendix
% \section{Appendix Title}

% This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\nocite{*}
\bibliographystyle{abbrvnat}
\bibliography{./biblio}

\end{document}

### 2010-03-15 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-\usepackage{amsmath}
+\usepackage{amsmath,graphicx, color}
+
+\newcommand{\Ocaml}{OCaml}
+\newcommand{\unix}{\textsc{Unix}}
 
@@
 
+master and worker terminology introduced here
+
+a few technical details about marshaling should be explained
+
 \section{API}
 
-main idea of the master function
+The main idea is to start with a generic function, with the
+following interface: 
+% TODO: rename this function
+\begin{ocaml}
+  val compute : 
+    f:('a -> 'b) -> 
+    handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
+    ('a * 'c) list -> unit
+\end{ocaml}
+This function takes two functions, \ocaml{f} and \ocaml{handle}, and a
+list of \emph{tasks} of type \ocaml{('a * 'c) list}.
+The function \ocaml{f} is applied to the first component for each task.
+Whenever it completes, the task and the result are passed to the
+function \ocaml{handle}. This may possibly generate a new list of
+tasks. The process ends when all computations are done.
 
-four scenarios:
-\begin{itemize}
-\item cores
-\item same binary
-\item same version of ocaml
-\item completely different programs
-\end{itemize}
+Function \ocaml{f} should be pure and is executed in parallel.
+Function \ocaml{handle}, on the contrary, can be impure and is never
+executed in parallel with other \ocaml{handle}s. Typically,
+\ocaml{handle} stores and processes information locally.
+The next section will describe the usage of this generic interface to
+perform traditional map and fold operations.
+
+Actually, our library provides more than just a single \ocaml{compute}
+function as above, but instead provides five different versions
+depending on the execution context. The five possible contexts are the
+following: 
+\begin{enumerate}
+\item \textbf{purely sequential execution:}
+  this is intended for validating the results;
+
+\item \textbf{several cores on the same machine:} 
+  this implementation makes use of \unix\ processes and thus can
+  provide a polymorphic \ocaml{compute} function;
+
+\item \textbf{same executable run on master and worker machines:}
+  this implementation makes use of the ability to marshal \Ocaml\
+  closures and thus can still provide a polymorphic \ocaml{compute} function;
+
+\item \textbf{master and workers are different programs, compiled with
+    the same version of \Ocaml:} 
+  in this case, we can no longer marshal closures but we can still
+  marshal polymorphic values. As a consequence, 
+  the \ocaml{compute} function is split into two
+  polymorphic functions, to implement the master and workers respectively:
+  \begin{ocaml}
+val master : 
+  handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
+  ('a * 'c) list -> unit
+val worker : ('a -> 'b) -> unit -> unit
+  \end{ocaml}
+
+\item \textbf{master and workers are different programs, not even
+    compiled with the same version of \Ocaml:} in this case, we can no
+  longer use marshaling, so the
+  \ocaml{compute} function is split into two monomorphic functions
+  over strings: 
+  \begin{ocaml}
+val master : 
+  handle:(string * 'c -> string -> (string * 'c) list) -> 
+  (string * 'c) list -> unit
+val worker : (string -> string) -> unit -> unit
+  \end{ocaml}
+  Note that the second component of each task is still polymorphic (of
+  type \ocaml{'c} here).
+\end{enumerate}
+
+Our library is organized into three modules: \of{Sequential} for the
+pure sequential implementation, \of{Cores} for multiple cores on the
+same machine and \of{Network} for a network of machines, respectively.
+The \of{Network} module itself is organized into three sub-modules,
+corresponding to contexts 3, 4 and 5 above. The different
+API functions are summarized in the following table (their signatures
+being given above):
+\begin{center}
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \ocaml{Sequential} & \ocaml{Cores} &
+    \multicolumn{3}{|c|}{\ocaml{Network}} 
+    \\\cline{3-5}
+    &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
+    \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
+    \ocaml{master}  & \ocaml{master}  \\
+    & & & \ocaml{worker}  & \ocaml{worker} \\\hline
+  \end{tabular}
+\end{center}
 
 \section{Derived API}
+
+In most cases, the easiest way to parallelize execution it to make use
+of map and fold operations over lists, where processing of the list
+elements are done in parallel.  Using the generic API described in the
+previous section to do that is not very convenient.
+Hence we provide the most commonly used list operations in our API.
+
+The most obvious operation is the traditional map operation over
+lists, that is:
+\begin{ocaml}
+  val map : ('a -> 'b) -> 'a list -> 'b list
+\end{ocaml}
+The next natural operation is a combination of map and fold
+operations, that is a function like
+\begin{ocaml}
+ val map_fold :
+   map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+   'c -> 'a list -> 'c
+\end{ocaml}
+which, given two functions, an accumulator \of{a} and a list \of{l}, computes
+\begin{equation}\label{eq:map-fold}
+  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{map} ~ x_1)) (\of{map} ~ x_2))
+  ... (\of{map} ~ x_n)
+\end{equation}
+for some permutation $[x_1,x_2,...,x_n]$ of the list \of{l}.
+
+We assume that the \of{map} operations are always performed in parallel.
+Regarding \of{fold} operations, we distinguish two cases:
+\begin{itemize}
+\item either \of{fold} operations are computationally less expensive
+  than \of{map} and we perform them locally;
+\item or \of{fold} operations are computationally expensive and we
+  perform them in parallel.
+\end{itemize}
+We thus provide two functions \of{map_local_fold} and \of{map_remote_fold}.
+
+In the case of \of{map_remote_fold}, only one \of{fold} operation can
+be performed at a time (possibly in parallel with \of{map}
+operations), as obvious from~(\ref{eq:map-fold}). However, there are
+cases where several \of{fold} operations can be performed in parallel,
+as early as intermediate results of \of{fold} operations are available.
+This is the case when \of{fold} is an associative operation (which
+implies that type \of{'b} and \of{'c} are the same  and that the
+accumulator should be a neutral element for \of{fold}). Whenever
+\of{fold} is also commutative, we can perform even more \of{fold}
+operations in parallel. Thus our API provides two functions
+\of{map_fold_a} and \of{map_fold_ac} for these two particular cases.
+The five operations of the derived API are summarized
+in Figure~\ref{fig:derived}.
+\begin{figure}[t]
+  \begin{ocaml}
+    val map : 
+      f:('a -> 'b) -> 'a list -> 'b list 
+    val map_local_fold : 
+      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+      'c -> 'a list -> 'c 
+    val map_remote_fold : 
+      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+      'c -> 'a list -> 'c 
+    val map_fold_ac : 
+      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+      'b -> 'a list -> 'b 
+    val map_fold_a : 
+      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+      'b -> 'a list -> 'b
+  \end{ocaml}
+  \caption{Derived API.}
+\label{fig:derived}
+\end{figure}
+
+These functions are actually derived from the generic API.
 
@@
 \end{document}
+
+% LocalWords:  parallelize

### 2010-03-16 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-These functions are actually derived from the generic API.
+These five functions are actually derived from the generic API.
+The generality is achieved by implementing theses derivations as a functor.
+This functor, \of{Derive}, is parameterized by the \of{compute}
+function, as follows:
+\begin{ocaml}
+module Derive
+  (X : sig
+     val compute : 
+       f:('a -> 'b) -> 
+       handle:('a * 'c -> 'b -> ('a * 'c) list) ->
+       ('a * 'c) list -> unit
+   end) = struct ... end
+\end{ocaml}
+We now explain how each function from Figure~\ref{fig:derived} is
+implemented using \of{compute} inside the functor body.
 
-map, map\_fold, etc.
+\paragraph{\of{map}.} 
+Function \of{map} is easily implemented by associating an integer
+index with each element (of type \of{'a}) of the input list. Thus the
+list of tasks passed to \of{compute} is a list of pairs of type \of{'a * int}.
+As soon as an intermediate result is available, it is passed to
+\of{handle}, which stores it in a local table using the index as a key
+and generates no new task. Whenever \of{compute} returns, we easily
+recover the output list from the values stored in the table.
 
-generic programming using a functor, which is then applied in each of
-the 4 cases
+\paragraph{\of{map_local_fold}.} 
+Function \of{map_local_fold} is implemented using a local reference
+storing the current accumulator (of type \of{'c}).
+As soon as an intermediate result is available, \of{handle} 
+combines it with the accumulator using function \of{fold}. Whenever \of{compute}
+returns, we simply return the current accumulator.
+
+\paragraph{\of{map_remote_fold}.}
+In this case, combining the accumulator with an intermediate result is
+itself a task to be performed in parallel with \of{map} operations.
+Thus, when an intermediate \of{map} result is available, the
+accumulator may be already in use in a \of{fold} operation. Hence we
+need to store the intermediate \of{map} result in a local table, to be
+used as soon as the accumulator becomes available.
+The \of{handle} function distinguishes between \of{map} and \of{fold}
+results and, accordingly, updates the table and the accumulator. For a
+\of{map} result, either we combine it with the accumulator in a new
+task or we simply store it in the local table. For a \of{fold} result,
+either we simply store it into the accumulator or we immediately
+combine it with a pending \of{map} result from the table.
+Whenever \of{compute} returns, the accumulator must be available and
+we simply return it.
+
+\paragraph{\of{map_fold_ac}.}
+Function \of{map_fold_ac} is easier than the previous one, since we do
+not need to store more than one intermediate result. Indeed, as soon
+as two intermediate results are available, we immediately combine them
+using a \of{fold} operation. Thus the \of{handle} function 
+either stores it or combines it, depending on the availability/non
+availability of the accumulator.
+Whenever \of{compute} returns, the accumulator must be available and
+we simply return it.
+
+\paragraph{\of{map_fold_a}.}
+Function \of{map_fold_a} is definitely the trickiest. Indeed, the
+operation being only associative, we can only combine \emph{adjacent}
+intermediate results. In order to perform this, we store intermediate
+results in a local table according to indices of the input list. More
+precisely, if we have already computed
+\begin{displaymath}
+  x_i \oplus x_{i+1} \oplus \dots \oplus x_j
+\end{displaymath}
+where $\oplus$ denotes the \of{fold} operation, then we associate this
+value with indices $i,j$. As soon as a result of indices $k,i-1$ or
+$j+1,k$ is available, it can be combined with the result of $i,j$.
+Whenever \of{compute} returns, the table should only contain a result
+corresponding to the full input list and we simply return it.
 
@@
 
-% LocalWords:  parallelize
+% LocalWords:  parallelize functor parameterized indices

### 2010-03-16 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-\section{API}
+\section{API}\label{sec:API}
 
@@
 following interface: 
-% TODO: rename this function
+% TODO: rename this function?
 \begin{ocaml}
@@
 \end{ocaml}
+% TODO be even more explicit about the introduction of the word `task'
 This function takes two functions, \ocaml{f} and \ocaml{handle}, and a
@@
 
-\section{Derived API}
+\section{Derived API}\label{sec:derived}
 
@@
 
-\section{Implementation Details}
+These five functions are the most natural derivations of the
+\of{compute} functions. There are obviously other possible derivations
+(e.g. a variant of \of{map_fold} where only \of{fold} is meaningful);
+in most cases, they could be derived in a straightforward way from our
+generic API.
 
-cores?
+% TODO cannot be instantiated in the monomorphic case
 
-network: protocol, fault tolerance, etc.
+\section{Implementation Details}\label{sec:implem}
 
-\section{Experiments}
+We now describe the implementations of the various modules introduced
+in Section~\ref{sec:API}. The implementation of the \of{Sequential}
+module is straightforward and does not require any explanation.
+
+\subsection{Multiple Cores}
+
+The \of{Cores} module implements the distributed computing library for
+multiple cores on the same machine. It provides a function
+\of{set_number_of_cores: int -> unit} to indicate the number of cores
+to be used. The number passed to this function may actually be
+different from the actual number of cores in the machine; it is rather
+the number of tasks to be performed simultaneously. 
+% TODO comment more on the number of cores?
+Thus a program using the \of{Cores} module typically begins with 
+\begin{ocaml}
+open Mapreduce.Cores
+let () = set_number_of_cores 8
+\end{ocaml}
+% TODO rename Mapreduce?
+
+The \of{Cores} module is implemented with \unix\ processes, using the
+\of{fork} and \of{wait} system calls provided by the \of{Unix} library
+of \Ocaml. The idea is pretty simple. 
+The \of{compute} function maintains a global table of pending tasks
+and keeps track of the number of idle cores.
+Whenever there a pending task and an idle core, 
+a new sub-process is created using \of{Unix.fork}; 
+once completed, the sub-process marshals the result 
+into a local file. The main process maintains a table
+mapping each sub-process ID to the input task and the local file
+name. It waits for any completed sub-process using \of{Unix.wait}
+and recovers the result from the local file. Then function \of{handle}
+is applied, which may generate new tasks.
+The main loop can be depicted in the following way:
+\begin{flushleft}
+  \quad  \textbf{while} pending tasks $\lor$ pending sub-processes \\
+  \quad  \quad \textbf{while} pending tasks $\land$ idle cores \\
+  \quad  \quad \quad create new sub-process for some task \\
+  \quad  \quad \textbf{wait} for any completed sub-process \\
+  \quad  \quad \quad push new tasks generated by handle \\
+\end{flushleft}
+
+It would be more efficient to use \unix\ pipes instead of local files
+but pipes are severely restricted w.r.t. data size. To be more
+generic, we make no assumption about the size of the data. Providing
+an alternative implementation using pipes would be straightforward.
+
+The scheduling of tasks to the different cores is left to the
+operating system, through \of{Unix.fork}. Thus it may be the case that
+two tasks are executed on the same core, even if the declared number
+of cores is less or equal than the actual number of cores on the machine.
+% TODO: comment on user-provided scheduling?
+
+\subsection{Network of Machines}
+
+The \of{Network} module implements the distributed computing library for
+a network of machines. 
+It provides a function \of{declare_workers: n:int -> string -> unit} 
+to fill a table of worker machines. The argument \of{n} indicates the
+number of cores to be used in each worker machine, analogous to 
+\of{Cores} module. Thus a program using the \of{Network} module
+typically begins with
+\begin{ocaml}
+open Mapreduce.Network
+let () = declare_workers ~n:12 "machine1"
+let () = declare_workers ~n:8  "machine2"
+...
+\end{ocaml}
+
+The \of{Network} module is based on a traditional client/server
+architecture, where each worker is a server and the master is the
+client of each worker. The main execution loop is similar to the one
+in the \of{Cores} module, where distant processes on remote machines
+correspond to sub-processes and idle workers correspond to idle
+cores. In addition to this, it also involves issues of message
+transfer and fault tolerance, which are subsequently described.
+
+\subsubsection{Protocol}
+
+\subsubsection{Fault Tolerance}\label{sec:fault}
+
+\section{Experiments}\label{sec:experiments}
 
@@
 
-\section{Future Work}
+\section{Future Work}\label{sec:future}
 

### 2010-03-16 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
+Messages sent from master to workers could be any of the following kinds:
+\begin{description}
+\item[\of{Assign(id:int, f:string, x:string)}] This message assigns a
+  new task to the worker, the task being identified by the unique
+  integer \of{id}. The task to be performed is given by strings \of{f}
+  and \of{x}, which are interpreted depending on the context.
+
+\item[\of{Kill(id:int)}] This message indicates the worker that task
+  identified by \of{id} is to be killed.
+
+\item[\of{Stop}] This message informs the worker about completion of
+  the computation, so that it may choose to exit.
+
+\item[\of{Ping}] This message is used to check if the worker is still
+  alive, expecting a \of{Pong} message from the worker in return.
+\end{description}
+Messages sent by the workers could be any of the following kinds:
+\begin{description}
+\item[\of{Pong}] This message is an acknowledgment for a \of{Ping}
+  message from the master.
+
+\item[\of{Completed(id:int, s:string)}] This message indicates the
+  completion of a task identified by \of{id}, with result \of{s}.
+
+\item[\of{Aborted(id:int)}] This message informs the master that the
+  task identified by \of{id} is aborted, either as a response to a
+  \of{Kill} message or because of a worker malfunction.
+\end{description}
+Our implementation of the protocol works across different
+architectures, so that master and workers could be run on completely
+different platforms w.r.t. endianness, version of \Ocaml\ and
+operating system.
+
 \subsubsection{Fault Tolerance}\label{sec:fault}
+
+\subsubsection{Same, Poly and Mono} % TODO: change title
 
@@
 
-% LocalWords:  parallelize functor parameterized indices
+% LocalWords:  parallelize functor parameterized indices endianness

### 2010-03-16 doc/icfp-2010/main.tex
more fault tolerance on fault tolerance

@@
 
+The main issue in any distributed computing environment is ability to
+handle faults, which is also a distinguishing feature of our library.
+Faults are mainly of two kinds: either a worker is stopped, and
+possibly later restarted; or a worker is temporarily or permanently
+unreachable on the network. To provide fault tolerance, our master
+implementation is keeping track of the status of each worker. 
+This status is controlled by two timeout parameters $T_1$ and $T_2$
+and \of{Ping} and \of{Pong} messages sent by master and workers,
+respectively. There are four possible statuses for a worker: 
+\begin{description}
+\item[\of{not connected}:] the worker is still to be
+  connected to perform any useful task;
+\item[\of{alive}:] the worker has sent some message recently,
+  within $T_1$ seconds;
+\item[\of{pinged}:] the worker has not sent any message recently (for more
+  than $T_1$ seconds) and the master hash sent it a \of{Ping} message
+  within the last $T_2$ seconds;
+\item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
+  message (for more than $T_2$ seconds).
+\end{description}
+% TODO picture
+Whenever we receive a message from a worker, its status changes to
+\of{alive} and its timeout value is reset.
+
+Fault tolerance is achieved by exploiting the statues of workers as
+follows. First, tasks are only assigned to workers with either
+\of{alive} or \of{pinged} status. Second, whenever a worker executing
+a task $t$ moves to status \of{not connected} or \of{unreachable}, the
+task $t$ is rescheduled, which means
+
 \subsubsection{Same, Poly and Mono} % TODO: change title
@@
 data locality: scheduling supplied by the user
+
+speeding up the end of computation with duplicate tasks on idle workers
+(straightforward since already done for fault tolerance)
 

### 2010-03-16 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-\subsubsection{Protocol}
+\subsubsection{Protocol}\label{sec:protocol}
 
@@
 a task $t$ moves to status \of{not connected} or \of{unreachable}, the
-task $t$ is rescheduled, which means
+task $t$ is rescheduled, which means it is put back in the set of
+pending tasks. Whenever a task is completed, any rescheduled copy of
+this task is either removed from the set of pending tasks or killed if
+it was already assigned to an idle worker.
 
-\subsubsection{Same, Poly and Mono} % TODO: change title
+\subsubsection{Three Network Implementations} % TODO: change title
+
+As explained in Section~\ref{sec:API}, the \of{Network} module is
+actually providing three different implementations, according to three
+different execution scenarios (namely, \of{Same}, \of{Poly} and
+\of{Mono}). There is actually a single, generic implementation, which
+is then instantiated in three different ways.
+The generic implementation assumes that we are only transmitting
+strings over the network, as evident from \of{Assign} and
+\of{Completed} messages in the protocol section~\ref{sec:protocol}.
+Therefore, the signature for the generic implementation is as follows:
+\begin{ocaml}
+  generic_compute:
+    assign_job:('a -> string * string) ->
+    handle:('a * 'c -> string -> ('a * 'c) list)  ->
+    ('a * 'c) list -> unit
+\end{ocaml}
+The first argument \of{assign_job} is a function indicating 
+the two strings to pass inside the \of{Assign} message of the protocol.
+We can now distinguish the three cases:
+\begin{description}
+\item[\of{Same}:] in this case, \of{assign_job} is simply returning
+  the marshaled closure together with the marshaled value.
+\item[\of{Poly}:] in this case, \of{assign_job} is simply
+  returning the marshaled value (the first string being meaningless in
+  this context).
+\item[\of{Mono}:] in this case, \of{assign_job} is immediately
+  returning its argument (the first string being meaningless in
+  this context).
+\end{description}
 

### 2010-03-17 doc/icfp-2010/main.tex
master -> compute ; handle -> master ; f -> worker / shutdown workers at exit

@@
   val compute : 
-    f:('a -> 'b) -> 
-    handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
+    worker:('a -> 'b) -> 
+    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
     ('a * 'c) list -> unit
@@
 % TODO be even more explicit about the introduction of the word `task'
-This function takes two functions, \ocaml{f} and \ocaml{handle}, and a
+This function takes two functions, \ocaml{worker} and \ocaml{master}, and a
 list of \emph{tasks} of type \ocaml{('a * 'c) list}.
-The function \ocaml{f} is applied to the first component for each task.
+The function \ocaml{worker} is applied to the first component for each task.
 Whenever it completes, the task and the result are passed to the
-function \ocaml{handle}. This may possibly generate a new list of
+function \ocaml{master}. This may possibly generate a new list of
 tasks. The process ends when all computations are done.
 
-Function \ocaml{f} should be pure and is executed in parallel.
-Function \ocaml{handle}, on the contrary, can be impure and is never
-executed in parallel with other \ocaml{handle}s. Typically,
-\ocaml{handle} stores and processes information locally.
+Function \ocaml{worker} should be pure and is executed in parallel.
+Function \ocaml{master}, on the contrary, can be impure and is never
+executed in parallel with other \ocaml{master}s. Typically,
+\ocaml{master} stores and processes information locally.
 The next section will describe the usage of this generic interface to
@@
 val master : 
-  handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
+  ('a * 'c -> 'b -> ('a * 'c) list) -> 
   ('a * 'c) list -> unit
@@
 val master : 
-  handle:(string * 'c -> string -> (string * 'c) list) -> 
+  (string * 'c -> string -> (string * 'c) list) -> 
   (string * 'c) list -> unit
@@
      val compute : 
-       f:('a -> 'b) -> 
-       handle:('a * 'c -> 'b -> ('a * 'c) list) ->
+       worker:('a -> 'b) -> 
+       master:('a * 'c -> 'b -> ('a * 'c) list) ->
        ('a * 'c) list -> unit
@@
 As soon as an intermediate result is available, it is passed to
-\of{handle}, which stores it in a local table using the index as a key
+\of{master}, which stores it in a local table using the index as a key
 and generates no new task. Whenever \of{compute} returns, we easily
@@
 storing the current accumulator (of type \of{'c}).
-As soon as an intermediate result is available, \of{handle} 
+As soon as an intermediate result is available, \of{master} 
 combines it with the accumulator using function \of{fold}. Whenever \of{compute}
@@
 used as soon as the accumulator becomes available.
-The \of{handle} function distinguishes between \of{map} and \of{fold}
+The \of{master} function distinguishes between \of{map} and \of{fold}
 results and, accordingly, updates the table and the accumulator. For a
@@
 as two intermediate results are available, we immediately combine them
-using a \of{fold} operation. Thus the \of{handle} function 
+using a \of{fold} operation. Thus the \of{master} function 
 either stores it or combines it, depending on the availability/non
@@
 name. It waits for any completed sub-process using \of{Unix.wait}
-and recovers the result from the local file. Then function \of{handle}
+and recovers the result from the local file. Then function \of{master}
 is applied, which may generate new tasks.
@@
   \quad  \quad \textbf{wait} for any completed sub-process \\
-  \quad  \quad \quad push new tasks generated by handle \\
+  \quad  \quad \quad push new tasks generated by \of{master} \\
 \end{flushleft}
@@
 \begin{ocaml}
-  generic_compute:
+  generic_master:
     assign_job:('a -> string * string) ->
-    handle:('a * 'c -> string -> ('a * 'c) list)  ->
+    master:('a * 'c -> string -> ('a * 'c) list)  ->
     ('a * 'c) list -> unit
@@
 
-% LocalWords:  parallelize functor parameterized indices endianness
+% LocalWords:  parallelize functor parameterized indices endianness monomorphic

### 2010-03-17 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-\titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{short description of paper}   % 'preprint' option specified.
+\titlebanner{submitted to ICFP 2010}        % These are ignored unless
+\preprintfooter{Experience Report: Distributed Computing for the Common Man}   % 'preprint' option specified.
 
-\title{Experience Report: }
-\subtitle{Subtitle Text, if any}
+\title{Experience Report: Distributed Computing for the Common Man}
+%\subtitle{Subtitle Text, if any}
 
@@
 \section{Introduction}
+
+This paper introduces a generic distributed computing library.
+It was initially motivated by computing needs that exist in our own
+research team. Our applications include large-scale deductive program
+verification, which amounts to check the validity of a huge number of
+logical formulas using a variety of automated theorem
+provers~\cite{filliatre07cav}. Our computing infrastructure consists
+of a few powerful multi-core machines (typically 8 to 16 cores) and
+several desktop PCs (typically dual-core). However, there is no
+library that helps in exploiting such a computing infrastructure in
+our favorite functional programming language. Hence we designed and
+implemented such a library\footnote{This library is implemented in
+  \Ocaml, but the implementation should be straightforward in any
+  functional programming language.}, which is the subject of this paper.
 

### 2010-03-18 doc/icfp-2010/main.tex
*** empty log message ***
added more contents in introduction section; fixed some grammatical errors [KK]

@@
 
-This paper introduces a generic distributed computing library.
-It was initially motivated by computing needs that exist in our own
+This paper introduces a generic distributed computing library. This
+work was initially motivated by computing needs that exist in our own
 research team. Our applications include large-scale deductive program
-verification, which amounts to check the validity of a huge number of
-logical formulas using a variety of automated theorem
+verification, which amounts to checking the validity of a large number
+of logical formulas using a variety of automated theorem
 provers~\cite{filliatre07cav}. Our computing infrastructure consists
 of a few powerful multi-core machines (typically 8 to 16 cores) and
-several desktop PCs (typically dual-core). However, there is no
-library that helps in exploiting such a computing infrastructure in
-our favorite functional programming language. Hence we designed and
-implemented such a library\footnote{This library is implemented in
-  \Ocaml, but the implementation should be straightforward in any
-  functional programming language.}, which is the subject of this paper.
+several desktop PCs (typically dual-core). However, for our
+application needs, there is no library that helps in exploiting such a
+computing infrastructure in our favorite functional programming
+language. Hence we designed and implemented such a
+library\footnote{This library is implemented in \Ocaml, but the
+  implementation should be straightforward in any functional
+  programming language.}, which is the subject of this paper.
 
-what it is and what it is not
+The distributing computing library presented in this paper, is not a
+library that helps in parallelizing computations. Rather, it provides
+facilities for reliable, distributed execution of parallelizable
+computations. In particular, it provides a set of APIs that allows
+execution of large-scale parallelizable computations, very relevant to
+our application needs, (and presumably relevant to a variety of
+real-world applications) over multiple cores in the same machine or
+over a network of machines. It relieves the user from implementing
+details of fault-tolerance, task distribution and parallelization. The
+most striking feature of the library is its generality. The user needs
+to change no more than a couple of lines in his code when he moves his
+large-scale computation from a multi-core machine to a network of
+machines. Further, the library provides exciting flexibilities as it
+supports polymorphic functions, and rich features such as
+fault-tolerance that become a cause for concern in any distributed
+programming environment. It is worth noting that the library is not
+targeted at applications running on server farms, crunching enormous
+amounts of data. Neither it is limited to research labs possessing a
+few number of powerful computing infrastructure. It serves a variety
+of users and a wide spectrum of needs, from desktop PCs to network of
+machines, and hence we find title \emph{Distributed Computing for the
+  Common Man}, apt and appropriate.
+
+% what it is and what it is not
+% \begin{itemize}
+% \item it is not a library which does parallelization
+% \item it is a library for reliable distributed execution of
+%   parallelizable computation 
+% \end{itemize}
+
+% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
+% programming, ironically) ;  however, there are differences:
+% \begin{itemize}
+% \item polymorphic, higher-order API
+% \item more generic (cores/network)
+% \item does not focus on association lists
+% \item we don't have data locality (future work)
+% \end{itemize}
+
+% what is the target audience/applications: 
+% for instance, use if automatic provers on thousands of verification
+% conditions, on computing infrastructures which can be
+% \begin{itemize}
+% \item a single machine with multiple cores, possibly remote
+% \item several machines, small or large in number, over a network
+% \end{itemize}
+
+\emph{Distributed Computing:} 
+A typical distributed computing library, in its minimal implementation
+should provide the following:
 \begin{itemize}
-\item it is not a library which does parallelization
-\item it is a library for reliable distributed execution of
-  parallelizable computation 
+\item A notion of a \of{task} which denotes the computation to be
+  performed in a distributed manner
+\item A set of machines or programs called \of{workers} that perform
+  the tasks, producing results
+\item A machine or a program called a \of{master} that is incharge
+  of distributing the tasks among the workers and managing results
+  produced by the workers
 \end{itemize}
+In addition to the above, distributed computing environments also
+implement mechanisms for fault-tolerance and efficient storage and
+distribution of tasks. This is required to handle network failures
+that may occur, as well as to optimize the usage of machines in the
+network. Another concern of importance is the transmission of messages
+over the network. This requires efficient serialization (or
+marshalling) of data that can be reliably reconstructed when
+transmitted over different computing environments.  It is desirable to
+maintain architecture independence while transmitting marshalled data,
+as machines in a distributed computing environment more often run on
+different hardware architectures and make use of varying software
+platforms. For example, a computer running on a hardware architecture
+should be able to reliably reconstruct the data irrespective of
+endianness. We carefully exploit the marshalling provided by \Ocaml\,
+in particular the restrictions involved in marshalling functions, in
+order to achieve maximally reliable transmission of messages over a
+network.
 
-inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
-programming, ironically) ;  however, there are differences:
-\begin{itemize}
-\item polymorphic, higher-order API
-\item more generic (cores/network)
-\item does not focus on association lists
-\item we don't have data locality (future work)
-\end{itemize}
+The application domain of such a distributed computing library is
+manyfold.  The most striking application, as also relevant to our
+research endeavours, is in validating thousands of verification
+conditions using automated theorem provers. Having noted this, it
+should be apparent that the library finds use in efficient execution
+of any large-scale expensive computation, that is parallelizable.
 
-what is the target audience/applications: 
-for instance, use if automatic provers on thousands of verification
-conditions, on computing infrastructures which can be
-\begin{itemize}
-\item a single machine with multiple cores, possibly remote
-\item several machines, small or large in number, over a network
-\end{itemize}
+A couple of caveats are in order. (1) We do not focus on issues like
+data-locality or distributed file system like Google's, rather we
+assume a \unix\ Network File System (NFS). (2) Our focus is not
+exactly on Google's Mapreduce paradigm~\cite{mapreduce} that works on
+association lists, but we also provide an API similar to Google's that
+works on association lists, just for satisfying popular interest.
 
-master and worker terminology introduced here
+Google's approach was also one of our inspirations to design and
+implement a generic distributed computing library for functional
+programming. Ironically, Google's approach itself was inspired by
+functional programming primitives.
 
-a few technical details about marshaling should be explained
+The paper is organized as follows. In Section~\ref{sec:API}, we
+describe the main idea behind the generic API, and introduce
+\of{master} and \of{worker} programs. In Section~\ref{sec:derived}, we
+derive more interesting functions from the generic
+API. Section~\ref{sec:implem} delves into the implementation details
+of our library, while Section~\ref{sec:experiments} illustrates the
+potential of the presented library through experimental evaluation. We
+draw some conclusions in Section~\ref{sec:future} and outline our
+future work.
+
+% master and worker terminology introduced here
+% a few technical details about marshaling should be explained
 

### 2010-03-18 doc/icfp-2010/main.tex
*** empty log message ***

@@
 computing infrastructure in our favorite functional programming
-language. Hence we designed and implemented such a
+language. 
+Hence we designed and implemented such a
 library\footnote{This library is implemented in \Ocaml, but the
@@
 
-The distributing computing library presented in this paper, is not a
+The distributing computing library presented in this paper is not a
 library that helps in parallelizing computations. Rather, it provides
@@
 execution of large-scale parallelizable computations, very relevant to
-our application needs, (and presumably relevant to a variety of
-real-world applications) over multiple cores in the same machine or
-over a network of machines. It relieves the user from implementing
-details of fault-tolerance, task distribution and parallelization. The
-most striking feature of the library is its generality. The user needs
-to change no more than a couple of lines in his code when he moves his
-large-scale computation from a multi-core machine to a network of
-machines. Further, the library provides exciting flexibilities as it
-supports polymorphic functions, and rich features such as
-fault-tolerance that become a cause for concern in any distributed
-programming environment. It is worth noting that the library is not
+our application needs (and presumably relevant to a variety of
+real-world applications), over multiple cores in the same machine or
+over a network of machines. The most important features of our library
+are the following:
+\begin{itemize}
+\item \emph{genericity}: 
+  it allows various patterns of polymorphic computations;
+\item \emph{simplicity}: 
+  switching between multiple cores and a network of
+  machine is as simple as changing a couple of lines of code;
+\item \emph{task distribution and fault-tolerance}: 
+  it relieves the user from implementing task distribution routines
+  and reliable fault-tolerance mechanism.
+\end{itemize}
+It is worth noting that the library is not
 targeted at applications running on server farms, crunching enormous
@@
 of users and a wide spectrum of needs, from desktop PCs to network of
-machines, and hence we find title \emph{Distributed Computing for the
-  Common Man}, apt and appropriate.
+machines, and hence the title.
+
+The application domain of such a distributed computing library is
+manyfold.  The most striking application, as also relevant to our
+research endeavours, is in validating thousands of verification
+conditions using automated theorem provers. Having noted this, it
+should be apparent that the library finds use in efficient execution
+of any large-scale expensive computation, that is parallelizable.
 
@@
 
-\emph{Distributed Computing:} 
-A typical distributed computing library, in its minimal implementation
-should provide the following:
+\paragraph{Distributed Computing.}
+A typical distributed computing library
+provides the following (borrowing some terminology from Google's
+Mapreduce~\cite{mapreduce}):
 \begin{itemize}
-\item A notion of a \of{task} which denotes the computation to be
-  performed in a distributed manner
-\item A set of machines or programs called \of{workers} that perform
-  the tasks, producing results
-\item A machine or a program called a \of{master} that is incharge
+\item A notion of a \emph{task} which denotes the atomic computation to be
+  performed in a distributed manner;
+\item A set of machines or programs called \emph{workers} that perform
+  the tasks, producing results;
+\item A single program called a \emph{master} that is in charge
   of distributing the tasks among the workers and managing results
-  produced by the workers
+  produced by the workers.
 \end{itemize}
 In addition to the above, distributed computing environments also
-implement mechanisms for fault-tolerance and efficient storage and
+implement mechanisms for fault-tolerance, efficient storage and
 distribution of tasks. This is required to handle network failures
@@
 maintain architecture independence while transmitting marshalled data,
-as machines in a distributed computing environment more often run on
+as machines in a distributed computing environment often run on
 different hardware architectures and make use of varying software
-platforms. For example, a computer running on a hardware architecture
-should be able to reliably reconstruct the data irrespective of
-endianness. We carefully exploit the marshalling provided by \Ocaml\,
-in particular the restrictions involved in marshalling functions, in
-order to achieve maximally reliable transmission of messages over a
-network.
+platforms. For example, endianness may be different across machines on
+the network.
 
-The application domain of such a distributed computing library is
-manyfold.  The most striking application, as also relevant to our
-research endeavours, is in validating thousands of verification
-conditions using automated theorem provers. Having noted this, it
-should be apparent that the library finds use in efficient execution
-of any large-scale expensive computation, that is parallelizable.
+\paragraph{A Functional Programming Approach.}
+Our work was initially inspired by Google's
+Mapreduce\footnote{Ironically, Google's approach itself was inspired
+  by functional programming primitives.}. However, our functional
+programming environment allows us to be more generic. 
+The main idea behind our approach is that
+workers may implement any polymorphic function
+\begin{ocaml}
+  worker: 'a -> 'b
+\end{ocaml}
+where \of{'a} denotes the type of tasks and \of{'b} the type of results.
+Then the master is given a list of initial tasks, together with a
+function to handle the results:
+\begin{ocaml}
+  master: ('a -> 'b -> 'a list) -> 'a list -> unit
+\end{ocaml}
+The function passed to the master is applied whenever a result is
+available and may in turn generate new tasks (hence the return type
+\of{'a list}).  The master is executed as long as there are incomplete
+tasks.
 
-A couple of caveats are in order. (1) We do not focus on issues like
-data-locality or distributed file system like Google's, rather we
-assume a \unix\ Network File System (NFS). (2) Our focus is not
-exactly on Google's Mapreduce paradigm~\cite{mapreduce} that works on
-association lists, but we also provide an API similar to Google's that
-works on association lists, just for satisfying popular interest.
-
-Google's approach was also one of our inspirations to design and
-implement a generic distributed computing library for functional
-programming. Ironically, Google's approach itself was inspired by
-functional programming primitives.
+Our library makes use of \Ocaml's marshaling capabilities, as much as
+possible. Whenever master and worker executables are exactly the same,
+we can marshal polymorphic values and closures. However, it is not
+always possible to have master and workers running the same
+executable. In that case, we cannot marshal closures anymore but we
+can still marshal polymorphic values as long as the same version of
+\Ocaml\ is used to compile master and workers. When different versions
+of \Ocaml\ are used, we can no longer marshal values but we can still
+transmit strings between master and workers. Our library adapts to all
+these situations, by providing several APIs.
 
@@
 potential of the presented library through experimental evaluation. We
-draw some conclusions in Section~\ref{sec:future} and outline our
-future work.
-
-% master and worker terminology introduced here
-% a few technical details about marshaling should be explained
+compare our approach with relevant related work in
+Section~\ref{sec:future} and outline our future work.
 
@@
 
-The main idea is to start with a generic function, with the
-following interface: 
-% TODO: rename this function?
+The main function in our API follows the idea sketched in the
+introduction. It has the following signature:
 \begin{ocaml}
@@
 \end{ocaml}
-% TODO be even more explicit about the introduction of the word `task'
-This function takes two functions, \ocaml{worker} and \ocaml{master}, and a
-list of \emph{tasks} of type \ocaml{('a * 'c) list}.
-The function \ocaml{worker} is applied to the first component for each task.
-Whenever it completes, the task and the result are passed to the
-function \ocaml{master}. This may possibly generate a new list of
-tasks. The process ends when all computations are done.
-
+Tasks are pairs, of type \of{'a * 'c}, where the first component is
+passed to the worker and the second component is local to the master.
 Function \ocaml{worker} should be pure and is executed in parallel.
 Function \ocaml{master}, on the contrary, can be impure and is never
-executed in parallel with other \ocaml{master}s. Typically,
-\ocaml{master} stores and processes information locally.
+executed in parallel with other \ocaml{master}s. Function
+\ocaml{master} always processes information locally, typically storing
+it to internal data structures.
 The next section will describe the usage of this generic interface to
@@
 
-Actually, our library provides more than just a single \ocaml{compute}
-function as above, but instead provides five different versions
-depending on the execution context. The five possible contexts are the
-following: 
+Actually, our library provides not just a single \ocaml{compute}
+function as above, but instead five different versions
+depending on the execution context. 
+The five possible contexts are the following: 
 \begin{enumerate}
@@
     the same version of \Ocaml:} 
-  in this case, we can no longer marshal closures but we can still
+  we can no longer marshal closures but we can still
   marshal polymorphic values. As a consequence, 
   the \ocaml{compute} function is split into two
-  polymorphic functions, to implement the master and workers respectively:
+  polymorphic functions, to implement the master and workers 
+  respectively:%
 \begin{ocaml}
@@
 \end{ocaml}
-
 \item \textbf{master and workers are different programs, not even
-    compiled with the same version of \Ocaml:} in this case, we can no
+    compiled with the same version of \Ocaml:} we can no
   longer use marshaling, so the
   \ocaml{compute} function is split into two monomorphic functions
-  over strings: 
+  over strings:%
 \begin{ocaml}
@@
   Note that the second component of each task is still polymorphic (of
-  type \ocaml{'c} here).
+  type \ocaml{'c} here), since it is local to the master.
 \end{enumerate}
@@
 \end{center}
+The following section introduces other interesting APIs that can be
+derived from this generic API.
 
@@
 of map and fold operations over lists, where processing of the list
-elements are done in parallel.  Using the generic API described in the
-previous section to do that is not very convenient.
-Hence we provide the most commonly used list operations in our API.
+elements are done in parallel.  To facilitate such a processing, we
+now derive the most commonly used list operations from our generic
+API.
 
@@
 for some permutation $[x_1,x_2,...,x_n]$ of the list \of{l}.
-
 We assume that the \of{map} operations are always performed in parallel.
@@
 
+A couple of caveats are in order. (1) We do not focus on issues like
+data-locality or distributed file system like Google's, rather we
+assume a \unix\ Network File System (NFS). (2) Our focus is not
+exactly on Google's Mapreduce paradigm~\cite{mapreduce} that works on
+association lists, but we also provide an API similar to Google's that
+works on association lists, just for satisfying popular interest.
+
 data locality: scheduling supplied by the user
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
+% LocalWords:  genericity executables

### 2010-03-21 doc/icfp-2010/main.tex
*** empty log message ***
Some grammatical errors rectified [KK]

@@
 
-\authorinfo{Name1}
-           {Affiliation1}
-           {Email1}
-\authorinfo{Name2\and Name3}
-           {Affiliation2/3}
-           {Email2/3}
+\authorinfo{Jean-Christophe Filliâtre}
+           {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
+            INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
+           {filliatr@lri.fr}
+\authorinfo{Kalyan Krishnamani}
+           {INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
+           {kalyan@lri.fr}
 
@@
 This paper introduces a generic distributed computing library. This
-work was initially motivated by computing needs that exist in our own
-research team. Our applications include large-scale deductive program
-verification, which amounts to checking the validity of a large number
-of logical formulas using a variety of automated theorem
+work was initially motivated by the computing needs that exist in our
+own research team. Our applications include large-scale deductive
+program verification, which amounts to checking the validity of a
+large number of logical formulas using a variety of automated theorem
 provers~\cite{filliatre07cav}. Our computing infrastructure consists
@@
 computing infrastructure in our favorite functional programming
-language. 
-Hence we designed and implemented such a
+language.  Hence we designed and implemented such a
 library\footnote{This library is implemented in \Ocaml, but the
@@
 facilities for reliable, distributed execution of parallelizable
-computations. In particular, it provides a set of APIs that allows
+computations. In particular, it provides a set of APIs that allows the
 execution of large-scale parallelizable computations, very relevant to
@@
   it allows various patterns of polymorphic computations;
-\item \emph{simplicity}: 
-  switching between multiple cores and a network of
-  machine is as simple as changing a couple of lines of code;
+\item \emph{simplicity}: switching between multiple cores on the same
+  machine and a network of machines is as simple as changing a couple
+  of lines of code;
 \item \emph{task distribution and fault-tolerance}: 
   it relieves the user from implementing task distribution routines
-  and reliable fault-tolerance mechanism.
+  and reliable fault-tolerance mechanisms.
 \end{itemize}
-It is worth noting that the library is not
-targeted at applications running on server farms, crunching enormous
-amounts of data. Neither it is limited to research labs possessing a
-few number of powerful computing infrastructure. It serves a variety
-of users and a wide spectrum of needs, from desktop PCs to network of
-machines, and hence the title.
+It is worth noting that the library is not targeted at applications
+running on server farms, crunching enormous amounts of data. Neither
+is it limited to research labs possessing a few number of powerful
+computing infrastructure. It serves a variety of users and a wide
+spectrum of needs, from desktop PCs to networks of machines, and hence
+the title.
 
@@
 should be apparent that the library finds use in efficient execution
-of any large-scale expensive computation, that is parallelizable.
+of any large-scale expensive computation that is parallelizable.
 
@@
 \paragraph{Distributed Computing.}
-A typical distributed computing library
-provides the following (borrowing some terminology from Google's
-Mapreduce~\cite{mapreduce}):
+A typical distributed computing library provides the following (we
+borrow some terminology from Google's Mapreduce~\cite{mapreduce}):
 \begin{itemize}
@@
 always possible to have master and workers running the same
-executable. In that case, we cannot marshal closures anymore but we
+executable. In this case, we cannot marshal closures anymore but we
 can still marshal polymorphic values as long as the same version of
@@
 passed to the worker and the second component is local to the master.
-Function \ocaml{worker} should be pure and is executed in parallel.
-Function \ocaml{master}, on the contrary, can be impure and is never
-executed in parallel with other \ocaml{master}s. Function
-\ocaml{master} always processes information locally, typically storing
-it to internal data structures.
+The function \ocaml{worker} should be pure and is executed in parallel
+with other \ocaml{worker}s. The function \ocaml{master}, on the
+contrary, can be impure and is never executed in parallel with other
+\ocaml{master}s. The function \ocaml{master} always processes
+information locally, typically storing it to internal data structures.
 The next section will describe the usage of this generic interface to
@@
 
-In most cases, the easiest way to parallelize execution it to make use
-of map and fold operations over lists, where processing of the list
-elements are done in parallel.  To facilitate such a processing, we
-now derive the most commonly used list operations from our generic
+In most cases, the easiest way to parallelize an execution it to make
+use of map and fold operations over lists, where processing of the
+list elements are done in parallel.  To facilitate such a processing,
+we now derive the most commonly used list operations from our generic
 API.
@@
 
-These five functions are actually derived from the generic API.
-The generality is achieved by implementing theses derivations as a functor.
+These five functions are actually derived from the generic API.  The
+generality is achieved by implementing these derivations as a functor.
 This functor, \of{Derive}, is parameterized by the \of{compute}
@@
 Function \of{map_local_fold} is implemented using a local reference
-storing the current accumulator (of type \of{'c}).
-As soon as an intermediate result is available, \of{master} 
-combines it with the accumulator using function \of{fold}. Whenever \of{compute}
+storing the current accumulator (of type \of{'c}).  As soon as an
+intermediate result is available, the \of{master} combines it with the
+accumulator using the function \of{fold}. Whenever \of{compute}
 returns, we simply return the current accumulator.
@@
 \paragraph{\of{map_fold_ac}.}
-Function \of{map_fold_ac} is easier than the previous one, since we do
-not need to store more than one intermediate result. Indeed, as soon
-as two intermediate results are available, we immediately combine them
-using a \of{fold} operation. Thus the \of{master} function 
-either stores it or combines it, depending on the availability/non
-availability of the accumulator.
-Whenever \of{compute} returns, the accumulator must be available and
-we simply return it.
+The function \of{map_fold_ac} is easier than the previous one, since
+we do not need to store more than one intermediate result. Indeed, as
+soon as two intermediate results are available, we immediately combine
+them using a \of{fold} operation. Thus the \of{master} function either
+stores it or combines it, depending on the availability/non
+availability of the accumulator.  Whenever \of{compute} returns, the
+accumulator must be available and we simply return it.
 
@@
 \of{set_number_of_cores: int -> unit} to indicate the number of cores
-to be used. The number passed to this function may actually be
-different from the actual number of cores in the machine; it is rather
-the number of tasks to be performed simultaneously. 
+to be used. The number passed to this function may be different from
+the actual number of cores in the machine; it is rather the number of
+tasks to be performed simultaneously.
 % TODO comment more on the number of cores?
@@
 \of{fork} and \of{wait} system calls provided by the \of{Unix} library
-of \Ocaml. The idea is pretty simple. 
-The \of{compute} function maintains a global table of pending tasks
-and keeps track of the number of idle cores.
-Whenever there a pending task and an idle core, 
-a new sub-process is created using \of{Unix.fork}; 
-once completed, the sub-process marshals the result 
-into a local file. The main process maintains a table
-mapping each sub-process ID to the input task and the local file
-name. It waits for any completed sub-process using \of{Unix.wait}
-and recovers the result from the local file. Then function \of{master}
-is applied, which may generate new tasks.
-The main loop can be depicted in the following way:
+of \Ocaml. The idea is pretty simple.  The \of{compute} function
+maintains a global table of pending tasks and keeps track of the
+number of idle cores.  Whenever there is a pending task and an idle
+core, a new sub-process is created using \of{Unix.fork}; once
+completed, the sub-process marshals the result into a local file. The
+main process maintains a table mapping each sub-process ID to the
+input task and the local file name. It waits for any completed
+sub-process using \of{Unix.wait} and recovers the result from the
+local file. Then function \of{master} is applied, which may generate
+new tasks.  The main loop can be depicted in the following way:
 \begin{flushleft}
@@
 but pipes are severely restricted w.r.t. data size. To be more
-generic, we make no assumption about the size of the data. Providing
+generic, we make no assumptions about the size of the data. Providing
 an alternative implementation using pipes would be straightforward.
@@
 
-The \of{Network} module implements the distributed computing library for
-a network of machines. 
-It provides a function \of{declare_workers: n:int -> string -> unit} 
-to fill a table of worker machines. The argument \of{n} indicates the
-number of cores to be used in each worker machine, analogous to 
-\of{Cores} module. Thus a program using the \of{Network} module
-typically begins with
+The \of{Network} module implements the distributed computing library
+for a network of machines.  It provides a function
+\of{declare_workers: n:int -> string -> unit} to fill a table of
+worker machines. The argument \of{n} indicates the number of cores to
+be used in each worker machine, analogous to the \of{Cores}
+module. Thus a program using the \of{Network} module typically begins
+with
 \begin{ocaml}
@@
 
-The main issue in any distributed computing environment is ability to
-handle faults, which is also a distinguishing feature of our library.
-Faults are mainly of two kinds: either a worker is stopped, and
-possibly later restarted; or a worker is temporarily or permanently
-unreachable on the network. To provide fault tolerance, our master
-implementation is keeping track of the status of each worker. 
-This status is controlled by two timeout parameters $T_1$ and $T_2$
-and \of{Ping} and \of{Pong} messages sent by master and workers,
+The main issue in any distributed computing environment is the ability
+to handle faults, which is also a distinguishing feature of our
+library.  Faults are mainly of two kinds: either a worker is stopped,
+and possibly later restarted; or a worker is temporarily or
+permanently unreachable on the network. To provide fault tolerance,
+our master implementation is keeping track of the status of each
+worker.  This status is controlled by two timeout parameters $T_1$ and
+$T_2$ and \of{Ping} and \of{Pong} messages sent by master and workers,
 respectively. There are four possible statuses for a worker:
@@
   within $T_1$ seconds;
-\item[\of{pinged}:] the worker has not sent any message recently (for more
-  than $T_1$ seconds) and the master hash sent it a \of{Ping} message
-  within the last $T_2$ seconds;
+\item[\of{pinged}:] the worker has not sent any message recently (for
+  more than $T_1$ seconds) and the master has sent the worker a
+  \of{Ping} message within the last $T_2$ seconds;
 \item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
@@
 \of{alive} and its timeout value is reset.
+\begin{center}
+  \includegraphics{state.mps}
+\end{center}
 
-Fault tolerance is achieved by exploiting the statues of workers as
+Fault tolerance is achieved by exploiting the status of workers as
 follows. First, tasks are only assigned to workers with either
@@
 A couple of caveats are in order. (1) We do not focus on issues like
-data-locality or distributed file system like Google's, rather we
+data-locality or a distributed file system like Google's, rather we
 assume a \unix\ Network File System (NFS). (2) Our focus is not
@@
 association lists, but we also provide an API similar to Google's that
-works on association lists, just for satisfying popular interest.
+works on association lists, just to satisfy popular interest.
 
+%TODO
 data locality: scheduling supplied by the user

### 2010-03-22 doc/icfp-2010/main.tex
just started correcting/modifying sec.4.2.3 [KK]

@@
 
-As explained in Section~\ref{sec:API}, the \of{Network} module is
+As mentioned in Section~\ref{sec:API}, the \of{Network} module is
 actually providing three different implementations, according to three
 different execution scenarios (namely, \of{Same}, \of{Poly} and
-\of{Mono}). There is actually a single, generic implementation, which
-is then instantiated in three different ways.
-The generic implementation assumes that we are only transmitting
-strings over the network, as evident from \of{Assign} and
-\of{Completed} messages in the protocol section~\ref{sec:protocol}.
-Therefore, the signature for the generic implementation is as follows:
+\of{Mono}). The exciting thing about our library is that there is
+actually a single, generic implementation, which is then instantiated
+in three different ways, corresponding to the three scenarios.  The
+generic implementation assumes that only string are being transmitting
+over the network, as evident from \of{Assign} and \of{Completed}
+messages in Section~\ref{sec:protocol}.  The signature for the generic
+implementation is therefore as follows:
 \begin{ocaml}

### 2010-03-22 doc/icfp-2010/main.tex
just started correcting/modifying sec.4.2.3 [KK]

@@
 
-As mentioned in Section~\ref{sec:API}, the \of{Network} module is
-actually providing three different implementations, according to three
+As mentioned in Section~\ref{sec:API}, the \of{Network} module
+actually provides three different implementations, according to three
 different execution scenarios (namely, \of{Same}, \of{Poly} and
@@
 \end{ocaml}
-The first argument \of{assign_job} is a function indicating 
-the two strings to pass inside the \of{Assign} message of the protocol.
-We can now distinguish the three cases:
+The first argument \of{assign_job} is a function indicating the two
+strings that are passed in the \of{Assign} message of the protocol.
+We can now distinguish the three cases, based on \of{assign_job}:
 \begin{description}

### 2010-03-22 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-n-queens, SMT solvers, etc.
+In this section, we show the potential of our library with two 
+computationally-extensive examples.
+
+\subsection{N-queens}
+
+The first example is the classical $N$-queens problem, where we
+compute the total number of solutions for a given $N$.
+The following table shows execution timings for various values of $N$
+and our three different implementations modules \of{Sequential}, \of{Cores} and
+\of{Network}. The purpose of this experiment is to measure the speedup
+ratio w.r.t. the sequential implementation. Therefore, all
+computations are performed on the same machine, a 8 cores Intel Xeon
+3.2 GHz running Debian Linux. The sequential implementation uses a
+single core. The cores implementation uses the 8 cores of
+the machine. The network implementation uses 8 workers running locally and
+a master running on a remote machine (incurs communication cost).
+
+The first column shows the value of $N$.
+Each computation is divided into a certain number of tasks,
+which is shown in second column.
+Then the last three column show execution timings in seconds for the
+three implementations. The figures within brackets show the speedup
+ratio w.r.t. sequential implementation.
+\begin{center}
+  \begin{tabular}{|r|r|r|r|r|}
+    \hline
+    N & \#tasks         & sequential& cores         & network 
+    \\\hline\hline
+    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
+    \\\hline
+       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
+    \\\hline
+    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
+    \\\hline
+       &  240    & 107.0     &  14.00                & 24.90    
+    \\\hline
+    18 &   18    &           & 123.00               & 125.00     
+    \\\hline
+       &  272    &           & 103.00               & 124.00
+    \\\hline
+    19 &   19    &           &                      & 940.00
+    \\\hline
+       &  306    &           &                      & 819.00      \\\hline
+  \end{tabular}
+\end{center}
+From the table above, it is clear that the \of{Cores} implementation
+provides a significant speedup. When the number of tasks is large
+enough, the speedup ratio is almost 8, which is also the number of
+cores we use. 
+% (When the number of tasks is small and not a multiple of
+% the number of cores,  for example as in third row with 17 tasks
+% executing on 8 cores, the
+
+% TODO comment on network ratios
+
+Where computation cost for each tasks is too small, then communication
+cost dominates and could even slow down the execution
+w.r.t. sequential implementation, as evident from second row.
+
+\subsection{SMT Solvers}
+
+Here we demonstrate the potential of our library for our application
+needs as mentioned in the introduction. Our computing
+infrastructure consists of 3 machines with 4, 8 and 8 cores
+respectively. The benchmark we use for
+experimentation consists of 80 challenging verification conditions
+(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
+VC is stored in a file, which is accessible over
+NFS. The purpose of the experiment is to check the validity of each VC
+using several automated provers (namely Alt-Ergo, Simplify and Z3),
+which are all installed on each of the machines in use.
+
+The master program proceeds by reading the 80 file names, turning them
+into tasks by multiplying them by the number of provers (240 tasks in total).
+Each worker in turn invokes the given prover on the given file.
+We set a timeout limit of 600 seconds.
+Each task completes with one of the four possible outcomes: \emph{valid},
+\emph{invalid}, \emph{cannot-decide} (depending on
+whether the VC is valid, invalid or undecided by the prover)
+and \emph{timeout}.
+
+\begin{center}
+  \begin{tabular}{|r|r|r|r|r|}
+    \hline
+    prover   & valid & invalid & cannot & timeout 
+    \\\hline\hline
+    Alt-ergo & 
+    \\\hline
+    Simplify & 
+    \\\hline
+    Z3       & 
+    \\\hline
+  \end{tabular}
+\end{center}
+
+% TODO compare total amount of time wrt sequential time
 

### 2010-03-24 doc/icfp-2010/main.tex
*** empty log message ***

@@
 which is shown in second column.
-Then the last three column show execution timings in seconds for the
+Then the last three columns show execution timings in seconds for the
 three implementations. The figures within brackets show the speedup
@@
     \\\hline
-       &  240    & 107.0     &  14.00                & 24.90    
+       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
     \\\hline
-    18 &   18    &           & 123.00               & 125.00     
+    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
     \\\hline
-       &  272    &           & 103.00               & 124.00
+       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
     \\\hline
-    19 &   19    &           &                      & 940.00
+    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
     \\\hline
-       &  306    &           &                      & 819.00      \\\hline
+       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
+    \\\hline
   \end{tabular}
 \end{center}
-From the table above, it is clear that the \of{Cores} implementation
-provides a significant speedup. When the number of tasks is large
-enough, the speedup ratio is almost 8, which is also the number of
+From the table above, it is clear that the \of{Cores} and \of{Network}
+implementations
+provides a significant speedup. As evident from the last row,
+the speedup ratio is almost 8, which is also the number of
 cores we use. 
-% (When the number of tasks is small and not a multiple of
-% the number of cores,  for example as in third row with 17 tasks
-% executing on 8 cores, the
-
-% TODO comment on network ratios
-
-Where computation cost for each tasks is too small, then communication
-cost dominates and could even slow down the execution
-w.r.t. sequential implementation, as evident from second row.
+It is also evident from the last column that \of{Network}
+implementation performs significantly better when the computation time
+dominates the total execution time.
+The two extreme cases correspond to second and last rows: in the
+second row, the communication time dominates and is in fact more than
+91\%\ of the total execution time; on the other hand, for the last row
+communication time amounts to just 4.6\%\ of the total execution time.
+As expected, the network implementation is only beneficial when the
+computation time for each individual task is significant, which is the
+case in realistic examples.
 
@@
 Each worker in turn invokes the given prover on the given file.
-We set a timeout limit of 600 seconds.
+We set a timeout limit of 10 minutes.
 Each task completes with one of the four possible outcomes: \emph{valid},
-\emph{invalid}, \emph{cannot-decide} (depending on
-whether the VC is valid, invalid or undecided by the prover)
-and \emph{timeout}.
-
+\emph{cannot-decide} (depending on
+whether the VC is valid or undecided by the prover), 
+\emph{timeout} and \emph{failure}.
+The figure below shows the total time in minutes spent by each prover
+for each possible outcomes.
 \begin{center}
@@
     \hline
-    prover   & valid & invalid & cannot & timeout 
+    prover   & valid & cannot & timeout & failure
     \\\hline\hline
-    Alt-ergo & 
+    Z3       &   4.20 &   0.00 &    160  &  20.0    
     \\\hline
-    Simplify & 
+    Alt-ergo &   3.60 &   0.57 &    190  &   8.2     
     \\\hline
-    Z3       & 
+    Simplify &   0.02 &   0.00 &     40  &   1.0   
     \\\hline
@@
 \end{center}
+These figures sum up to more than 7 hours if provers were executed
+sequentially. However, using our library and our 3-machines
+infrastructure, it completes in 1 hour and 5 minutes, giving us a
+speedup of 6.6$\times$.
 
-% TODO compare total amount of time wrt sequential time
+% timeout = 600
+%            |  valid       | invalid      | cannot       | timeout      | failure
+% ---------------------------------------------------------------------------------------
+%         z3 |  42 (251.25s) |   0 ( 0.00s) |   0 ( 0.00s) |  16 (9548.79s) |  22 (1223.81s) | 
+%   alt-ergo |  40 (217.55s) |   0 ( 0.00s) |  19 (34.29s) |  19 (11376.70s) |   2 (489.05s) | 
+%   simplify |  32 ( 0.95s) |   0 ( 0.00s) |  34 ( 0.25s) |   4 (2394.60s) |  10 (62.76s) | 
+% 0.43user 0.33system 1:04:57elapsed 0%CPU (0avgtext+0avgdata 0maxresident)k
+
+% total amount of time = 25630 seconds = 7 hours 7 minutes 10 seconds
+% ratio = 6.57
+
+% timeout = 120
+%            |  valid       | invalid      | cannot       | timeout      | failure
+% ---------------------------------------------------------------------------------------
+%         z3 |  41 (211.08s) |   0 ( 0.00s) |   0 ( 0.00s) |  20 (2383.89s) |  19 (262.30s) | 
+%   alt-ergo |  40 (237.00s) |   0 ( 0.00s) |  19 (34.45s) |  21 (2514.42s) |   0 ( 0.00s) | 
+%   simplify |  32 ( 0.98s) |   0 ( 0.00s) |  34 ( 0.31s) |   5 (598.92s) |   9 (18.98s) |
 

### 2010-03-24 doc/icfp-2010/main.tex
*** empty log message ***

@@
 \newcommand{\Ocaml}{OCaml}
+\newcommand{\JoCaml}{Jo{\&\!}Caml}
 \newcommand{\unix}{\textsc{Unix}}
@@
 \begin{abstract}
-This is the text of the abstract.
+
 \end{abstract}
 
-\category{CR-number}{subcategory}{third-level}
+\category{D1.1}{Programming techniques}{Applicative (Functional) Programming}
 
 \terms
-term1, term2
+Distributed Computing, Functional Programming
 
-\keywords
-keyword1, keyword2
+%\keywords
 
@@
 
-\section{Future Work}\label{sec:future}
+\section{Conclusion and Future Work}\label{sec:future}
 
-A couple of caveats are in order. (1) We do not focus on issues like
-data-locality or a distributed file system like Google's, rather we
-assume a \unix\ Network File System (NFS). (2) Our focus is not
-exactly on Google's Mapreduce paradigm~\cite{mapreduce} that works on
-association lists, but we also provide an API similar to Google's that
-works on association lists, just to satisfy popular interest.
+We presented a distributed programming environment for functional
+programming. The main features are the genericity of the interface,
+which makes use of polymorphic higher-order functions, and the ability
+to easily switch between sequential, multi-cores and network
+implementations. A distinguishing feature of our library is also a
+robust fault tolerance mechanism implemented within, which relieves
+the user from cumbersome implementation details.
 
-%TODO
-data locality: scheduling supplied by the user
+Closest to the approach in this paper is Yohann Padioleau's mapreduce
+implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
+top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
+protocol for message passing.  On one hand, we have less flexibility
+w.r.t. deployment of the user program; on the other hand, we provide a
+more generic API together with fault tolerance.
+There are other distributed computing libraries on top of which one
+could implement the library discussed in this paper. \JoCaml\
+is one of them~\cite{jocaml}. However, \JoCaml\ does not allow
+arbitrary closures to be passed between machines, which would limit
+the genericity of the API. Additionally, \JoCaml\ does not provide
+fault tolerance, which is essential in a distributed setting.
+% TODO check with Louis
 
-speeding up the end of computation with duplicate tasks on idle workers
-(straightforward since already done for fault tolerance)
+There are other implementations of distributed computing in the
+context of functional programming. One is the Disco
+project~\cite{disco}, which implements exactly Google's mapreduce in
+Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
+implementation of Google's mapreduce; rather it is more generic and
+not restricted to just key/value pairs.
+There are other ways to exploit multi-core architectures. One of which
+is data parallelism, which is also relevant in the functional
+programming setting~\cite{parallel-haskell}. Our work
+does not target data parallelism at all.
+
+There are still some interesting features that could be added to our
+library. 
+\begin{itemize}
+\item 
+  One is the ability to efficiently assign tasks to workers depending
+  on resource parameters, such as data locality, CPU power, memory,
+  etc. This could be achieved by providing the user ways to control
+  task scheduling.  
+\item 
+  Another interesting feature could be the ability
+  to add or remove workers dynamically. Currently, our library assumes
+  the list of workers to be used is given \emph{a priori}.
+\item
+  Our library provides limited support for displaying real-time
+  information about the computations. Processing and storing
+  information about tasks locally in the master is straightforward; 
+  monitoring it in real-time is less obvious.
+\item 
+  One very nice feature of Google's mapreduce is the possibility to
+  use idle workers on duplicated tasks to speedup the end of computation.
+  Since we already have the fault tolerance implemented, this
+  optimization should be straightforward to add to our library.
+\end{itemize}
+We intend to enrich our library with all above features.
 
@@
 
-\acks
+% \acks
 
-Acknowledgments, if needed.
+% Acknowledgments, if needed.
 

### 2010-03-24 doc/icfp-2010/main.tex
*** empty log message ***

@@
 \begin{abstract}
-
+  In this paper, we present a distributed computing library for
+  functional programming languages. The main features of this library
+  include (1) a highly-generic API, (2) several implementations to
+  adapt to different computing environments such as sequential,
+  multi-core or network, and (3) a reliable fault tolerance mechanism.
+  This paper describes the motivation behind this work, as well as
+  the design and implementation of the library. It also demonstrates
+  the potential of the library with realistic experiments.
 \end{abstract}
@@
 % \end{itemize}
-
 % inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
@@
 % \end{itemize}
-
 % what is the target audience/applications: 
@@
 
+\medskip
 The paper is organized as follows. In Section~\ref{sec:API}, we
@@
 \of{master} and \of{worker} programs. In Section~\ref{sec:derived}, we
-derive more interesting functions from the generic
+derive some interesting functions from the generic
 API. Section~\ref{sec:implem} delves into the implementation details
@@
 
+\medskip
 These five functions are the most natural derivations of the

### 2010-03-25 doc/icfp-2010/main.tex
petit corrections, but it introduced an orphan :( [KK]

@@
 
-\authorinfo{Jean-Christophe Filliâtre}
+\authorinfo{Jean-Christophe Filli\^{a}tre}
            {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
@@
 of any large-scale expensive computation that is parallelizable.
-
 % what it is and what it is not
@@
 % \end{itemize}
-
 \paragraph{Distributed Computing.}
@@
 \section{API}\label{sec:API}
-
-The main function in our API follows the idea sketched in the
-introduction. It has the following signature:
+In this section, we describe the uniqueness of approach, namely the
+generic API which we shall extend to a plethora of more specialized
+APIs in later sections. The main function in our API follows the idea
+sketched in the introduction. It has the following signature:
 \begin{ocaml}
@@
 
-In this section, we show the potential of our library with two 
+In this section, we show the potential of our library using two 
 computationally-extensive examples.
@@
 From the table above, it is clear that the \of{Cores} and \of{Network}
-implementations
-provides a significant speedup. As evident from the last row,
-the speedup ratio is almost 8, which is also the number of
-cores we use. 
-It is also evident from the last column that \of{Network}
-implementation performs significantly better when the computation time
-dominates the total execution time.
-The two extreme cases correspond to second and last rows: in the
-second row, the communication time dominates and is in fact more than
-91\%\ of the total execution time; on the other hand, for the last row
+implementations provide a significant speedup. As evident from the
+last row, the speedup ratio is almost 8, which is also the number of
+cores we use.  It is also evident from the last column that
+\of{Network} implementation performs significantly better when the
+computation time dominates the total execution time.  The two extreme
+cases correspond to second and last rows: in the second row, the
+communication time dominates and is in fact more than 91\%\ of the
+total execution time; on the other hand, for the last row
 communication time amounts to just 4.6\%\ of the total execution time.
@@
 These figures sum up to more than 7 hours if provers were executed
-sequentially. However, using our library and our 3-machines
+sequentially. However, using our library and our 3-machine
 infrastructure, it completes in 1 hour and 5 minutes, giving us a
@@
 w.r.t. deployment of the user program; on the other hand, we provide a
-more generic API together with fault tolerance.
-There are other distributed computing libraries on top of which one
-could implement the library discussed in this paper. \JoCaml\
-is one of them~\cite{jocaml}. However, \JoCaml\ does not allow
-arbitrary closures to be passed between machines, which would limit
-the genericity of the API. Additionally, \JoCaml\ does not provide
-fault tolerance, which is essential in a distributed setting.
+more generic API together with fault tolerance.  There are other
+distributed computing libraries on top of which one could implement
+the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
+them. However, \JoCaml\ does not allow arbitrary closures to be passed
+between machines, which would limit the genericity of the
+API. Additionally, \JoCaml\ does not provide fault tolerance, which is
+indispensible in a distributed setting.
 % TODO check with Louis

### 2010-03-25 doc/icfp-2010/main.tex
mapreduce -> factory
*** empty log message ***

@@
            {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
-            INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
+            INRIA Saclay -- \^Ile-de-France, ProVal, Orsay, F-91893}
            {filliatr@lri.fr}
 \authorinfo{Kalyan Krishnamani}
-           {INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
+           {INRIA Saclay -- \^Ile-de-France, ProVal, Orsay, F-91893}
            {kalyan@lri.fr}

### 2010-03-28 doc/icfp-2010/main.tex
some corrections as suggested by Claude [KK]

--- 2010-03-25 doc/icfp-2010/main.tex
@@
            {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
-            INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
+            INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
            {filliatr@lri.fr}
 \authorinfo{Kalyan Krishnamani}
-           {INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
-           {kalyan@lri.fr}
+           {INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
+           {kalyan@inria.fr}
 
@@
 \begin{abstract}
-  In this paper, we present a distributed computing library for
+  We present a distributed computing library for
   functional programming languages. The main features of this library
@@
 
-This paper introduces a generic distributed computing library. This
+This paper introduces a generic library for distributed computing. This
 work was initially motivated by the computing needs that exist in our
@@
 
-The distributing computing library presented in this paper is not a
+The distributed computing library presented in this paper is not a
 library that helps in parallelizing computations. Rather, it provides

### 2010-03-29 doc/icfp-2010/main.tex
factory -> functory

@@
 \newcommand{\Ocaml}{OCaml}
+\newcommand{\functory}{\textsf{Functory}}
 \newcommand{\JoCaml}{Jo{\&\!}Caml}
@@
 real-world applications), over multiple cores in the same machine or
-over a network of machines. The most important features of our library
-are the following:
+over a network of machines. 
+The most important features of our library are the following:
 \begin{itemize}
@@
 network. Another concern of importance is the transmission of messages
-over the network. This requires efficient serialization (or
-marshalling) of data that can be reliably reconstructed when
-transmitted over different computing environments.  It is desirable to
+over the network. This requires efficient 
+\emph{marshalling} of data, that is encoding and decoding of data 
+for transmission over different computing environments.  It is desirable to
 maintain architecture independence while transmitting marshalled data,
@@
 
+Our library, \functory, 
 In this section, we show the potential of our library using two 
@@
 respectively. The benchmark we use for
-experimentation consists of 80 challenging verification conditions
+experimentation consists of around 80 challenging verification conditions
 (VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
@@
 NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify and Z3),
+using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
 which are all installed on each of the machines in use.
 
-The master program proceeds by reading the 80 file names, turning them
-into tasks by multiplying them by the number of provers (240 tasks in total).
+The master program proceeds by reading the file names, turning them
+into tasks by multiplying them by the number of provers (more than
+300 tasks in total).
 Each worker in turn invokes the given prover on the given file.
-We set a timeout limit of 10 minutes.
+We set a timeout limit of 1 minute.
 Each task completes with one of the four possible outcomes: \emph{valid},
@@
     \\\hline\hline
-    Z3       &   4.20 &   0.00 &    160  &  20.0    
+    Alt-ergo &  24.7  &  2.4  & 1260.0 &   0.0       
     \\\hline
-    Alt-ergo &   3.60 &   0.57 &    190  &   8.2     
+    Simplify &   0.6  &  0.3  &  300.0 &  43.0   
     \\\hline
-    Simplify &   0.02 &   0.00 &     40  &   1.0   
+    Z3       &  98.6  &  0.0  & 1200.0 &  66.3    
+    \\\hline
+    CVC3     &  46.7  & 28.0  &  720.0 &   0.0   
     \\\hline
@@
 \end{center}
-These figures sum up to more than 7 hours if provers were executed
+These figures sum up to more than 1 hour if provers were executed
 sequentially. However, using our library and our 3-machine
-infrastructure, it completes in 1 hour and 5 minutes, giving us a
-speedup of 6.6$\times$.
-
-% timeout = 600
-%            |  valid       | invalid      | cannot       | timeout      | failure
-% ---------------------------------------------------------------------------------------
-%         z3 |  42 (251.25s) |   0 ( 0.00s) |   0 ( 0.00s) |  16 (9548.79s) |  22 (1223.81s) | 
-%   alt-ergo |  40 (217.55s) |   0 ( 0.00s) |  19 (34.29s) |  19 (11376.70s) |   2 (489.05s) | 
-%   simplify |  32 ( 0.95s) |   0 ( 0.00s) |  34 ( 0.25s) |   4 (2394.60s) |  10 (62.76s) | 
-% 0.43user 0.33system 1:04:57elapsed 0%CPU (0avgtext+0avgdata 0maxresident)k
-
-% total amount of time = 25630 seconds = 7 hours 7 minutes 10 seconds
-% ratio = 6.57
-
-% timeout = 120
-%            |  valid       | invalid      | cannot       | timeout      | failure
-% ---------------------------------------------------------------------------------------
-%         z3 |  41 (211.08s) |   0 ( 0.00s) |   0 ( 0.00s) |  20 (2383.89s) |  19 (262.30s) | 
-%   alt-ergo |  40 (237.00s) |   0 ( 0.00s) |  19 (34.45s) |  21 (2514.42s) |   0 ( 0.00s) | 
-%   simplify |  32 ( 0.98s) |   0 ( 0.00s) |  34 ( 0.31s) |   5 (598.92s) |   9 (18.98s) |
+infrastructure, it completes in 4 minutes and 8 seconds, giving us a
+speedup of more than 15$\times$.
 
@@
 the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
-them. However, \JoCaml\ does not allow arbitrary closures to be passed
-between machines, which would limit the genericity of the
-API. Additionally, \JoCaml\ does not provide fault tolerance, which is
-indispensible in a distributed setting.
-% TODO check with Louis
+them. However, \JoCaml\ does not provide fault tolerance, which is
+indispensible in a distributed setting. The user has to include code
+for fault tolerance, as already demonstrated in some \JoCaml\
+experiments~\cite{mandel2008}.
 
@@
 
-\nocite{*}
 \bibliographystyle{abbrvnat}

### 2010-03-29 doc/icfp-2010/main.tex
*** empty log message ***

@@
 
-\usepackage{amsmath,graphicx, color}
+\usepackage{amsmath,graphicx,color,url}
 
@@
 
-Our library, \functory, 
-In this section, we show the potential of our library using two 
+Our library, \functory, is implemented in \Ocaml\ and is available
+from \url{http://www.lri.fr/~filliatr/functory/}.
+In this section, we show the potential of this library using two 
 computationally-extensive examples.
@@
 
-The first column shows the value of $N$.
-Each computation is divided into a certain number of tasks,
-which is shown in second column.
-Then the last three columns show execution timings in seconds for the
-three implementations. The figures within brackets show the speedup
-ratio w.r.t. sequential implementation.
+The first column shows the value of $N$.  Each computation is divided
+into a certain number of tasks, by placing some of the queens before
+starting the distributed computation.  The number of tasks is shown in
+second column.  Then the last three columns show execution timings in
+seconds for the three implementations. The figures within brackets
+show the speedup ratio w.r.t. sequential implementation.
 \begin{center}
@@
 infrastructure, it completes in 4 minutes and 8 seconds, giving us a
-speedup of more than 15$\times$.
+speedup of more than 15$\times$. We are still far away from the ideal
+ratio of 20$\times$ (we are using 20 cores), since some provers are
+allocating a lot of memory and time spent in system calls is not
+accounted for in the total observed time. However, a ratio of
+15$\times$ is already a significant improvement for our day-to-day
+experiments.
 
@@
 
-% \acks
+\acks
+The authors are grateful to the ProVal team for support and comments
+on early versions of the library and of this paper.
 
-% Acknowledgments, if needed.
-
+\vfill\pagebreak
 % We recommend abbrvnat bibliography style.

### 2010-03-31 doc/icfp-2010/main.tex
*** empty log message ***
minor corrections [KK]

@@
            {filliatr@lri.fr}
-\authorinfo{Kalyan Krishnamani}
-           {INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
-           {kalyan@inria.fr}
+\authorinfo{K. Kalyanasundaram}
+           {ProVal, INRIA Saclay - \^{I}le-de-France, Orsay, F-91893}
+           {kalyan.krishnamani@inria.fr}
 
@@
   the design and implementation of the library. It also demonstrates
-  the potential of the library with realistic experiments.
+  the potential of the library using realistic experiments.
 \end{abstract}
@@
 \begin{itemize}
-\item \emph{genericity}: 
+\item \emph{Genericity}: 
   it allows various patterns of polymorphic computations;
-\item \emph{simplicity}: switching between multiple cores on the same
+\item \emph{Simplicity}: switching between multiple cores on the same
   machine and a network of machines is as simple as changing a couple
   of lines of code;
-\item \emph{task distribution and fault-tolerance}: 
+\item \emph{Task distribution and fault-tolerance}: 
   it relieves the user from implementing task distribution routines
@@
 with other \ocaml{worker}s. The function \ocaml{master}, on the
-contrary, can be impure and is never executed in parallel with other
-\ocaml{master}s. The function \ocaml{master} always processes
+contrary, can be impure and is never executed in parallel.
+The function \ocaml{master} always processes
 information locally, typically storing it to internal data structures.
@@
 \begin{enumerate}
-\item \textbf{purely sequential execution:}
+\item \textbf{Purely sequential execution:}
   this is intended for validating the results;
 
-\item \textbf{several cores on the same machine:} 
+\item \textbf{Several cores on the same machine:} 
   this implementation makes use of \unix\ processes and thus can
@@
 
-\item \textbf{same executable run on master and worker machines:}
+\item \textbf{Same executable run on master and worker machines:}
   this implementation makes use of the ability to marshal \Ocaml\
@@
 
-\item \textbf{master and workers are different programs, compiled with
+\item \textbf{Master and workers are different programs, compiled with
     the same version of \Ocaml:} 
@@
 \end{ocaml}
-\item \textbf{master and workers are different programs, not even
+\item \textbf{Master and workers are different programs, not even
     compiled with the same version of \Ocaml:} we can no
@@
 \end{center}
-The following section introduces other interesting APIs that can be
-derived from this generic API.
+The following section introduces other interesting APIs that we
+derive from this generic API.
 
@@
   to add or remove workers dynamically. Currently, our library assumes
-  the list of workers to be used is given \emph{a priori}.
+  that the list of workers to be used is given \emph{a priori}.
 \item

### 2010-04-01 doc/icfp-2010/main.tex
*** empty log message ***

@@
   functional programming languages. The main features of this library
-  include (1) a highly-generic API, (2) several implementations to
+  include (1) a generic API, (2) several implementations to
   adapt to different computing environments such as sequential,
@@
 A typical distributed computing library provides the following (we
-borrow some terminology from Google's Mapreduce~\cite{mapreduce}):
+borrow some terminology from Google's mapreduce~\cite{mapreduce}):
 \begin{itemize}
@@
 Our work was initially inspired by Google's
-Mapreduce\footnote{Ironically, Google's approach itself was inspired
+mapreduce\footnote{Ironically, Google's approach itself was inspired
   by functional programming primitives.}. However, our functional
@@
 available and may in turn generate new tasks (hence the return type
-\of{'a list}).  The master is executed as long as there are incomplete
+\of{'a list}).  The master is executed as long as there are pending
 tasks.
@@
   this implementation makes use of the ability to marshal \Ocaml\
-  closures and thus can still provide a polymorphic \ocaml{compute} function;
+  closures and thus can still provide a polymorphic \ocaml{compute}
+  function.
+  Depending on whether the program is run as a master or as a worker,
+  the relevant arguments of \ocaml{compute} are used.
 
@@
   respectively:%
+\vspace{-1em}
 \begin{ocaml}
+val worker : ('a -> 'b) -> unit -> unit
 val master : 
@@
   ('a * 'c) list -> unit
-val worker : ('a -> 'b) -> unit -> unit
 \end{ocaml}
@@
   over strings:%
+\vspace{-1em}
 \begin{ocaml}
+val worker : (string -> string) -> unit -> unit
 val master : 
@@
   (string * 'c) list -> unit
-val worker : (string -> string) -> unit -> unit
 \end{ocaml}
@@
 \end{itemize}
-We thus provide two functions \of{map_local_fold} and \of{map_remote_fold}.
+We thus provide two functions \of{map_local_fold} and \of{map_remote_}
+\of{fold}.
 
@@
 soon as two intermediate results are available, we immediately combine
-them using a \of{fold} operation. Thus the \of{master} function either
-stores it or combines it, depending on the availability/non
-availability of the accumulator.  Whenever \of{compute} returns, the
-accumulator must be available and we simply return it.
+them using a \of{fold} operation. Thus, whenever the \of{master}
+receives some result, it either stores it or combines it with the
+accumulator, depending on the availability/non availability of the
+accumulator.  Whenever \of{compute} returns, the accumulator must be
+available and we simply return it.
 
@@
 \begin{ocaml}
-open Mapreduce.Cores
+open Cores
 let () = set_number_of_cores 8
 \end{ocaml}
-% TODO rename Mapreduce?
 
@@
 \begin{ocaml}
-open Mapreduce.Network
+open Network
 let () = declare_workers ~n:12 "machine1"
@@
 in three different ways, corresponding to the three scenarios.  The
-generic implementation assumes that only string are being transmitting
+generic implementation assumes that only strings are being transmitted
 over the network, as evident from \of{Assign} and \of{Completed}
@@
 The first example is the classical $N$-queens problem, where we
-compute the total number of solutions for a given $N$.
+compute the total number of ways to place $N$ queens on a $N\times N$
+chessboard in such a way no two queens attack each other.
 The following table shows execution timings for various values of $N$
@@
 the machine. The network implementation uses 8 workers running locally and
-a master running on a remote machine (incurs communication cost).
+a master running on a remote machine (which incurs communication cost).
 
@@
     \\\hline\hline
-    Alt-ergo &  24.7  &  2.4  & 1260.0 &   0.0       
+    Alt-ergo & 406.0 & 3.0   &  11400.0 & 0.0       
     \\\hline
-    Simplify &   0.6  &  0.3  &  300.0 &  43.0   
+    Simplify &  0.5   & 0.4   &  1200.0 & 222.0   
     \\\hline
-    Z3       &  98.6  &  0.0  & 1200.0 &  66.3    
+    Z3       & 80.7   & 0.0   &  1800.0 & 1695.0   
     \\\hline
-    CVC3     &  46.7  & 28.0  &  720.0 &   0.0   
+    CVC3     & 303.0  & 82.7  &  4200.0 & 659.0   
     \\\hline
@@
 \end{center}
-These figures sum up to more than 1 hour if provers were executed
+These figures sum up to more than 6 hours if provers were executed
 sequentially. However, using our library and our 3-machine
-infrastructure, it completes in 4 minutes and 8 seconds, giving us a
-speedup of more than 15$\times$. We are still far away from the ideal
+infrastructure, it completes in 22 minutes and 37 seconds, giving us a
+speedup of more than 16$\times$. We are still far away from the ideal
 ratio of 20$\times$ (we are using 20 cores), since some provers are
@@
 accounted for in the total observed time. However, a ratio of
-15$\times$ is already a significant improvement for our day-to-day
+16$\times$ is already a significant improvement for our day-to-day
 experiments.

### 2010-04-01 doc/icfp-2010/main.tex
*** empty log message ***

@@
 \authorinfo{Jean-Christophe Filli\^{a}tre}
-           {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
+           {CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
             INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
@@
   adapt to different computing environments such as sequential,
-  multi-core or network, and (3) a reliable fault tolerance mechanism.
+  multi-core or network, and (3) a reliable fault-tolerance mechanism.
   This paper describes the motivation behind this work, as well as
@@
 running on server farms, crunching enormous amounts of data. Neither
-is it limited to research labs possessing a few number of powerful
-computing infrastructure. It serves a variety of users and a wide
+is it limited to research labs possessing few powerful
+machines. It serves a variety of users and a wide
 spectrum of needs, from desktop PCs to networks of machines, and hence
@@
 A typical distributed computing library provides the following (we
-borrow some terminology from Google's mapreduce~\cite{mapreduce}):
+borrow some terminology from Google's MapReduce~\cite{mapreduce}):
 \begin{itemize}
-\item A notion of a \emph{task} which denotes the atomic computation to be
-  performed in a distributed manner;
-\item A set of machines or programs called \emph{workers} that perform
+\item A notion of \emph{task} which denotes an atomic computation
+  to be performed in a distributed manner; 
+\item A set of processes (possibly executing on remote machines)
+  called \emph{workers} that perform
   the tasks, producing results;
-\item A single program called a \emph{master} that is in charge
+\item A single process called a \emph{master} which is in charge
   of distributing the tasks among the workers and managing results
@@
 In addition to the above, distributed computing environments also
-implement mechanisms for fault-tolerance, efficient storage and
+implement mechanisms for fault tolerance, efficient storage, and
 distribution of tasks. This is required to handle network failures
@@
 over the network. This requires efficient 
-\emph{marshalling} of data, that is encoding and decoding of data 
+\emph{marshaling} of data, that is encoding and decoding of data 
 for transmission over different computing environments.  It is desirable to
@@
 different hardware architectures and make use of varying software
-platforms. For example, endianness may be different across machines on
-the network.
+platforms. For example, machine word size or endianness may be different
+across machines on the network.
 
@@
 Our work was initially inspired by Google's
-mapreduce\footnote{Ironically, Google's approach itself was inspired
+MapReduce\footnote{Ironically, Google's approach itself was inspired
   by functional programming primitives.}. However, our functional
@@
 The main idea behind our approach is that
-workers may implement any polymorphic function
+workers may implement any polymorphic function:
 \begin{ocaml}
@@
 
-Our library makes use of \Ocaml's marshaling capabilities, as much as
+Our library makes use of \Ocaml's marshaling capabilities as much as
 possible. Whenever master and worker executables are exactly the same,
@@
 \section{API}\label{sec:API}
-In this section, we describe the uniqueness of approach, namely the
-generic API which we shall extend to a plethora of more specialized
-APIs in later sections. The main function in our API follows the idea
+In this section, we describe the
+generic API which we shall extend with several specialized
+functions in later sections. The main function in our API follows the idea
 sketched in the introduction. It has the following signature:
@@
 passed to the worker and the second component is local to the master.
-The function \ocaml{worker} should be pure and is executed in parallel
+The \ocaml{worker} function should be pure and is executed in parallel
 with other \ocaml{worker}s. The function \ocaml{master}, on the
 contrary, can be impure and is never executed in parallel.
-The function \ocaml{master} always processes
+The \ocaml{master} function always processes
 information locally, typically storing it to internal data structures.
 The next section will describe the usage of this generic interface to
-perform traditional map and fold operations.
+perform the traditional map and fold operations.
 
@@
 \item \textbf{Purely sequential execution:}
-  this is intended for validating the results;
+  this is mostly intended for debugging;
 
 \item \textbf{Several cores on the same machine:} 
-  this implementation makes use of \unix\ processes and thus can
-  provide a polymorphic \ocaml{compute} function;
+  this implementation makes use of \unix\ processes and 
+  provides a polymorphic \ocaml{compute} function;
 
@@
   this implementation makes use of the ability to marshal \Ocaml\
-  closures and thus can still provide a polymorphic \ocaml{compute}
+  closures and  provides a polymorphic \ocaml{compute}
   function.
@@
 In most cases, the easiest way to parallelize an execution it to make
-use of map and fold operations over lists, where processing of the
+use of operations over lists, where processing of the
 list elements are done in parallel.  To facilitate such a processing,
@@
 accumulator may be already in use in a \of{fold} operation. Hence we
-need to store the intermediate \of{map} result in a local table, to be
+need to store the intermediate \of{map} result in a table local to the
+master, to be
 used as soon as the accumulator becomes available.
@@
 open Network
-let () = declare_workers ~n:12 "machine1"
-let () = declare_workers ~n:8  "machine2"
+let () = declare_workers ~n:12 "hostname1"
+let () = declare_workers ~n:8  "192.168.24.2"
 ...
@@
 
-The \of{Network} module is based on a traditional client/server
+The \of{Network} module is based on a traditional TCP-based client/server
 architecture, where each worker is a server and the master is the
@@
 in the \of{Cores} module, where distant processes on remote machines
-correspond to sub-processes and idle workers correspond to idle
-cores. In addition to this, it also involves issues of message
+correspond to sub-processes and idle cores are the idle cores 
+of remote workers. In addition to this, it also involves issues of message
 transfer and fault tolerance, which are subsequently described.
@@
 
-\item[\of{Kill(id:int)}] This message indicates the worker that task
-  identified by \of{id} is to be killed.
+\item[\of{Kill(id:int)}] This message tells the worker to kill the task
+  identified by \of{id}.
 
@@
 \end{description}
-Messages sent by the workers could be any of the following kinds:
+Messages sent by workers could be any of the following kinds:
 \begin{description}
@@
 \begin{description}
-\item[\of{not connected}:] the worker is still to be
-  connected to perform any useful task;
-\item[\of{alive}:] the worker has sent some message recently,
+\item[\of{not connected}:] there is no ongoing TCP connection between
+  the master and the worker;
+\item[\of{alive}:] the worker has sent some message
   within $T_1$ seconds;
-\item[\of{pinged}:] the worker has not sent any message recently (for
-  more than $T_1$ seconds) and the master has sent the worker a
-  \of{Ping} message within the last $T_2$ seconds;
+\item[\of{pinged}:] the worker has not sent any message within $T_1$
+  seconds and the master has sent the worker a
+  \of{Ping} message within $T_2$ seconds;
 \item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
@@
 this task is either removed from the set of pending tasks or killed if
-it was already assigned to an idle worker.
+it was already assigned to another worker.
 
@@
 different execution scenarios (namely, \of{Same}, \of{Poly} and
-\of{Mono}). The exciting thing about our library is that there is
+\of{Mono}). There is
 actually a single, generic implementation, which is then instantiated
@@
 generic implementation assumes that only strings are being transmitted
-over the network, as evident from \of{Assign} and \of{Completed}
+over the network, as evident from the \of{Assign} and \of{Completed}
 messages in Section~\ref{sec:protocol}.  The signature for the generic
@@
 \begin{description}
-\item[\of{Same}:] in this case, \of{assign_job} is simply returning
+\item[\of{Same}:] in this case, \of{assign_job} simply returns
   the marshaled closure together with the marshaled value.
@@
 In this section, we show the potential of this library using two 
-computationally-extensive examples.
+computationally extensive examples.
 
@@
 chessboard in such a way no two queens attack each other.
-The following table shows execution timings for various values of $N$
-and our three different implementations modules \of{Sequential}, \of{Cores} and
+The following table shows execution times for various values of $N$
+and our three different implementations: \of{Sequential}, \of{Cores}, and
 \of{Network}. The purpose of this experiment is to measure the speedup
-ratio w.r.t. the sequential implementation. Therefore, all
+w.r.t. the sequential implementation. Therefore, all
 computations are performed on the same machine, a 8 cores Intel Xeon
 3.2 GHz running Debian Linux. The sequential implementation uses a
-single core. The cores implementation uses the 8 cores of
+single core. The multi-core implementation uses the 8 cores of
 the machine. The network implementation uses 8 workers running locally and
@@
 starting the distributed computation.  The number of tasks is shown in
-second column.  Then the last three columns show execution timings in
+second column.  Then the last three columns show execution times in
 seconds for the three implementations. The figures within brackets
-show the speedup ratio w.r.t. sequential implementation.
+show the speedup w.r.t. sequential implementation.
 \begin{center}
@@
 implementations provide a significant speedup. As evident from the
-last row, the speedup ratio is almost 8, which is also the number of
-cores we use.  It is also evident from the last column that
+last row, the speedup is almost 8, which is also the number of
+cores we use.  It is also evident from the last column that the
 \of{Network} implementation performs significantly better when the
-computation time dominates the total execution time.  The two extreme
-cases correspond to second and last rows: in the second row, the
+computation time dominates in the total execution time.  The two extreme
+cases correspond to the second and the last row: in the second row, the
 communication time dominates and is in fact more than 91\%\ of the
@@
 Each task completes with one of the four possible outcomes: \emph{valid},
-\emph{cannot-decide} (depending on
+\emph{unknown} (depending on
 whether the VC is valid or undecided by the prover), 
@@
 \begin{center}
-  \begin{tabular}{|r|r|r|r|r|}
+  \begin{tabular}{|r||r|r|r|r|}
     \hline
-    prover   & valid & cannot & timeout & failure
+    prover   & valid & unknown & timeout & failure
     \\\hline\hline
@@
 which makes use of polymorphic higher-order functions, and the ability
-to easily switch between sequential, multi-cores and network
-implementations. A distinguishing feature of our library is also a
-robust fault tolerance mechanism implemented within, which relieves
-the user from cumbersome implementation details.
+to easily switch between sequential, multi-core, and network
+implementations. A distinguishing feature of our library is a
+robust fault-tolerance mechanism which relieves
+the user of cumbersome implementation details.
 
-Closest to the approach in this paper is Yohann Padioleau's mapreduce
+Closest to the approach in this paper is Yohann Padioleau's MapReduce
 implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
-top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
-protocol for message passing.  On one hand, we have less flexibility
-w.r.t. deployment of the user program; on the other hand, we provide a
+top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
+protocol for message passing.  Currently, we have less flexibility
+w.r.t. deployment of the user program than OCamlMPI; on the other
+hand, we provide a 
 more generic API together with fault tolerance.  There are other
@@
 them. However, \JoCaml\ does not provide fault tolerance, which is
-indispensible in a distributed setting. The user has to include code
+indispensable in a distributed setting. The user has to include code
 for fault tolerance, as already demonstrated in some \JoCaml\
@@
 context of functional programming. One is the Disco
-project~\cite{disco}, which implements exactly Google's mapreduce in
+project~\cite{disco}, which implements exactly Google's MapReduce in
 Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
-implementation of Google's mapreduce; rather it is more generic and
+implementation of Google's MapReduce; rather it is more generic and
 not restricted to just key/value pairs.
-There are other ways to exploit multi-core architectures. One of which
+There are other ways to exploit multi-core architectures. One of these
 is data parallelism, which is also relevant in the functional
@@
   on resource parameters, such as data locality, CPU power, memory,
-  etc. This could be achieved by providing the user ways to control
+  etc. This could be achieved by providing the user with the means to control
   task scheduling.  
@@
 \item 
-  One very nice feature of Google's mapreduce is the possibility to
-  use idle workers on duplicated tasks to speedup the end of computation.
+  One very nice feature of Google's MapReduce is the possibility to
+  use redundantly several idle workers on the same tasks
+  for speedup when reaching the end of computation.
   Since we already have the fault tolerance implemented, this
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables
+% LocalWords:  genericity executables OCamlMPI MapReduce

### 2010-04-02 doc/icfp-2010/main.tex
illustrative example

@@
 describe the main idea behind the generic API, and introduce
-\of{master} and \of{worker} programs. In Section~\ref{sec:derived}, we
+\of{master} and \of{worker} programs. Section~\ref{sec:example}
+illustrates the API usage with an example.
+In Section~\ref{sec:derived}, we
 derive some interesting functions from the generic
@@
 \begin{ocaml}
-val worker : ('a -> 'b) -> unit -> unit
-val master : 
+val worker_compute : ('a -> 'b) -> unit -> unit
+val master_compute : 
   ('a * 'c -> 'b -> ('a * 'c) list) -> 
@@
 \begin{ocaml}
-val worker : (string -> string) -> unit -> unit
-val master : 
+val worker_compute : (string -> string) -> unit -> unit
+val master_compute : 
   (string * 'c -> string -> (string * 'c) list) -> 
@@
 The \of{Network} module itself is organized into three sub-modules,
-corresponding to contexts 3, 4 and 5 above. The different
-API functions are summarized in the following table (their signatures
-being given above):
-\begin{center}
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    \ocaml{Sequential} & \ocaml{Cores} &
-    \multicolumn{3}{|c|}{\ocaml{Network}} 
-    \\\cline{3-5}
-    &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
-    \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
-    \ocaml{master}  & \ocaml{master}  \\
-    & & & \ocaml{worker}  & \ocaml{worker} \\\hline
-  \end{tabular}
-\end{center}
-The following section introduces other interesting APIs that we
-derive from this generic API.
+called \of{Same}, \of{Poly} and \of{Mono},
+corresponding to contexts 3, 4 and 5 above. 
+The next section demonstrates the use of these modules.
+
+% The different
+% API functions are summarized in the following table (their signatures
+% being given above):
+% \begin{center}
+%   \begin{tabular}{|c|c|c|c|c|}
+%     \hline
+%     \ocaml{Sequential} & \ocaml{Cores} &
+%     \multicolumn{3}{|c|}{\ocaml{Network}} 
+%     \\\cline{3-5}
+%     &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
+%     \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
+%     \ocaml{master}  & \ocaml{master}  \\
+%     & & & \ocaml{worker}  & \ocaml{worker} \\\hline
+%   \end{tabular}
+% \end{center}
+% The following section introduces other interesting APIs that we
+% derive from this generic API.
+
+\section{Illustrative Example}\label{sec:example}
+
+Let us assume we need to compute a sum such as
+\begin{displaymath}
+  s(n) = \sum_{i=0}^{n}f(i)
+\end{displaymath}
+where $f$ is a pure and computationally expensive function.
+We start with the sequential implementation provided by the library:
+\begin{ocaml}
+  open Sequential
+\end{ocaml}
+The idea is to test our code on small values of $n$ before
+parallelizing the computation for larger values.
+Let us assume that $f$ is defined as a toplevel function
+\begin{ocaml}
+  let f i = ...
+\end{ocaml}
+and that $n$ is obtained from the command line
+\begin{ocaml}
+  let n = int_of_string Sys.argv.(1)
+\end{ocaml}
+The list of tasks is simply the list of integers 0,1,\dots,$n$,
+computed as follows:
+\begin{ocaml}
+  let tasks = 
+    let l = ref [] in 
+    for i = 0 to n do l := (i, ()) :: !l done; 
+    !l
+\end{ocaml}
+Each task is actually a pair, where the second component contains
+some information which is kept local to the master process.
+It is irrelevant here and thus the value \of{()} of type \of{unit} is used.
+The current sum is kept into a global reference \of{s}:
+\begin{ocaml}
+  let s = ref 0
+\end{ocaml}
+The master function receives a task \of{(i,())} together with its
+result \of{fi} and simply adds \of{fi} to \of{s}. There is no new task
+generated, hence the returned empty list.
+\begin{ocaml}
+  let master (i,()) fi = s := !s + fi; []
+\end{ocaml}
+Finally, the whole computation takes place when we call the
+\of{compute} function, passing \of{f} as the worker function:
+\begin{ocaml}
+  let () = compute ~worker:f ~master tasks
+\end{ocaml}
+
+Now, let us assume we want to make use of a 4-core machine to compute
+$s(n)$. It is as easy as just replacing the very first line with
+\begin{ocaml}
+  open Cores
+  let () = set_number_of_cores 4
+\end{ocaml}
+while the rest of the code remains unchanged.
+
+Later, we may want to use a network of machines instead, for example
+two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
+cores respectively. Similar to the above, we just replace the first
+lines with
+\begin{ocaml}
+  open Network
+  let () = declare_workers ~n:4 "orcus"
+  let () = declare_workers ~n:8 "belzebuth"
+  open Same
+\end{ocaml}
+Here, \of{Same} is a module which is to be used when master and worker
+are running the same executable. It provides a \of{compute} function
+which has the same signature as in modules \of{Sequential} and
+\of{Cores}, so that the rest of the code remains unchanged. Master and
+worker processes are distinguished at run-time using an environment
+variable \texttt{WORKER} which is set/unset.
+
+If we need to write two different programs for master and worker, for
+reason of binary incompatibility or any other reason, the library API
+is providing functions to do that. If master and worker are still
+compiled with the same version of \Ocaml, we use the \of{Poly} module
+which provides a polymorphic API. Let us start with the worker program.
+It is as simple as
+\begin{ocaml}
+  open Poly
+  let f i = ...
+  let () = Worker.compute f ()
+\end{ocaml}
+The \of{Worker.compute} function enters a loop which waits for
+tasks sent by the master and returns results computed using \of{f}.
+The master program is almost the same as before. First, we replace
+module \of{Same} with module \of{Poly}:
+\begin{ocaml}
+  open Network
+  let () = declare_workers ~n:4 "orcus"
+  let () = declare_workers ~n:8 "belzebuth"
+  open Poly
+\end{ocaml}
+Tasks and \of{master} function are unchanged:
+\begin{ocaml}
+  let tasks = ...
+  let s = ref 0
+  let master (i,()) fi = ...
+\end{ocaml}
+Finally, we start the computation with \of{Master.compute}, which does
+not have a \of{worker} parameter anymore:
+\begin{ocaml}
+  let () = Master.compute ~master tasks
+\end{ocaml}
+
+When master and worker programs are compiled with different
+versions of the \Ocaml\ compiler, our library still provides
+a monomorphic API over strings. As a consequence, we need to convert
+integers to and from strings in both master and worker.
+The modified worker program is as follows:
+\begin{ocaml}
+  open Mono
+  let f = ...
+  let f_string i = string_of_int (f (int_of_string i))
+  let () = Worker.compute f_string ()
+\end{ocaml}
+The master program is almost unchanged. We simply need to replace
+\of{Poly} with \of{Mono} and to encode tasks as strings:
+\begin{ocaml}
+  let tasks = 
+    let l = ref [] in 
+    for i = 0 to n do l := (string_of_int i, ()) :: !l done; 
+    !l
+\end{ocaml}
+Consequently, \of{master} is now receiving a string as argument:
+\begin{ocaml}
+  let master (i,()) fi = s := !s + int_of_string fi; []
+\end{ocaml}
 
@@
 use of operations over lists, where processing of the
-list elements are done in parallel.  To facilitate such a processing,
+list elements are done in parallel. 
+The previous section illustrated one such case.
+To facilitate such a processing,
 we now derive the most commonly used list operations from our generic
@@
 The \of{Cores} module implements the distributed computing library for
-multiple cores on the same machine. It provides a function
+multiple cores on the same machine. As illustrated in
+Section~\ref{sec:example}, 
+it provides a function
 \of{set_number_of_cores: int -> unit} to indicate the number of cores
@@
 % TODO comment more on the number of cores?
-Thus a program using the \of{Cores} module typically begins with 
-\begin{ocaml}
-open Cores
-let () = set_number_of_cores 8
-\end{ocaml}
+% Thus a program using the \of{Cores} module typically begins with 
+% \begin{ocaml}
+% open Cores
+% let () = set_number_of_cores 8
+% \end{ocaml}
 
@@
 The \of{Network} module implements the distributed computing library
-for a network of machines.  It provides a function
+for a network of machines. 
+As illustrated in Section~\ref{sec:example}, it
+provides a function
 \of{declare_workers: n:int -> string -> unit} to fill a table of
-worker machines. The argument \of{n} indicates the number of cores to
-be used in each worker machine, analogous to the \of{Cores}
-module. Thus a program using the \of{Network} module typically begins
-with
-\begin{ocaml}
-open Network
-let () = declare_workers ~n:12 "hostname1"
-let () = declare_workers ~n:8  "192.168.24.2"
-...
-\end{ocaml}
+worker machines. 
+% The argument \of{n} indicates the number of cores to
+% be used in each worker machine, analogous to the \of{Cores}
+% module. Thus a program using the \of{Network} module typically begins
+% with
+% \begin{ocaml}
+% open Network
+% let () = declare_workers ~n:12 "hostname1"
+% let () = declare_workers ~n:8  "192.168.24.2"
+% ...
+% \end{ocaml}
 
@@
 over the network, as evident from the \of{Assign} and \of{Completed}
-messages in Section~\ref{sec:protocol}.  The signature for the generic
-implementation is therefore as follows:
-\begin{ocaml}
-  generic_master:
-    assign_job:('a -> string * string) ->
-    master:('a * 'c -> string -> ('a * 'c) list)  ->
-    ('a * 'c) list -> unit
-\end{ocaml}
-The first argument \of{assign_job} is a function indicating the two
-strings that are passed in the \of{Assign} message of the protocol.
-We can now distinguish the three cases, based on \of{assign_job}:
-\begin{description}
-\item[\of{Same}:] in this case, \of{assign_job} simply returns
-  the marshaled closure together with the marshaled value.
-\item[\of{Poly}:] in this case, \of{assign_job} is simply
-  returning the marshaled value (the first string being meaningless in
-  this context).
-\item[\of{Mono}:] in this case, \of{assign_job} is immediately
-  returning its argument (the first string being meaningless in
-  this context).
-\end{description}
+messages in Section~\ref{sec:protocol}. Depending on the
+implementation, these strings could be marshaled closures, marshaled
+values or simply user strings.
+
+% The signature for the generic
+% implementation is therefore as follows:
+% \begin{ocaml}
+%   generic_master:
+%     assign_job:('a -> string * string) ->
+%     master:('a * 'c -> string -> ('a * 'c) list)  ->
+%     ('a * 'c) list -> unit
+% \end{ocaml}
+% The first argument \of{assign_job} is a function indicating the two
+% strings that are passed in the \of{Assign} message of the protocol.
+% We can now distinguish the three cases, based on \of{assign_job}:
+% \begin{description}
+% \item[\of{Same}:] in this case, \of{assign_job} simply returns
+%   the marshaled closure together with the marshaled value.
+% \item[\of{Poly}:] in this case, \of{assign_job} is simply
+%   returning the marshaled value (the first string being meaningless in
+%   this context).
+% \item[\of{Mono}:] in this case, \of{assign_job} is immediately
+%   returning its argument (the first string being meaningless in
+%   this context).
+% \end{description}
 
@@
 
-% \appendix
-% \section{Appendix Title}
-
-% This is the text of the appendix, if you need one.
+%\appendix
 
@@
 
-\vfill\pagebreak
+%\vfill\pagebreak
 % We recommend abbrvnat bibliography style.
-
 \bibliographystyle{abbrvnat}

### 2010-04-02 doc/icfp-2010/main.tex
*** empty log message ***

@@
 \begin{itemize}
-\item A notion of \emph{task} which denotes an atomic computation
+\item A notion of \emph{tasks} which denote atomic computations
   to be performed in a distributed manner; 
@@
 The \ocaml{worker} function should be pure and is executed in parallel
-with other \ocaml{worker}s. The function \ocaml{master}, on the
-contrary, can be impure and is never executed in parallel.
-The \ocaml{master} function always processes
-information locally, typically storing it to internal data structures.
+in all worker processes. The function \ocaml{master}, on the
+contrary, can be impure and is only executed in the master process.
+The \ocaml{master} function typically stores results 
+in some internal data structure.
 The next section will describe the usage of this generic interface to
@@
 \begin{ocaml}
-val worker_compute : ('a -> 'b) -> unit -> unit
-val master_compute : 
+val Worker.compute : ('a -> 'b) -> unit -> unit
+val Master.compute : 
   ('a * 'c -> 'b -> ('a * 'c) list) -> 
@@
 \begin{ocaml}
-val worker_compute : (string -> string) -> unit -> unit
-val master_compute : 
+val Worker.compute : (string -> string) -> unit -> unit
+val Master.compute : 
   (string * 'c -> string -> (string * 'c) list) -> 
@@
 parallelizing the computation for larger values.
-Let us assume that $f$ is defined as a toplevel function
+Let us assume that $f$ is implemented as a top-level \of{worker} function:
 \begin{ocaml}
-  let f i = ...
+  let worker i = ...
 \end{ocaml}
-and that $n$ is obtained from the command line
+and that $n$ is obtained from the command line:
 \begin{ocaml}
@@
 \end{ocaml}
-The list of tasks is simply the list of integers 0,1,\dots,$n$,
+The list of tasks is simply the list of integers $0,1,\dots,n$,
 computed as follows:
@@
 Finally, the whole computation takes place when we call the
-\of{compute} function, passing \of{f} as the worker function:
+\of{compute} function:
 \begin{ocaml}
-  let () = compute ~worker:f ~master tasks
+  let () = compute ~worker ~master tasks
 \end{ocaml}
@@
   open Poly
-  let f i = ...
-  let () = Worker.compute f ()
+  let worker i = ...
+  let () = Worker.compute worker ()
 \end{ocaml}
 The \of{Worker.compute} function enters a loop which waits for
-tasks sent by the master and returns results computed using \of{f}.
+tasks sent by the master and returns results computed using \of{worker}.
 The master program is almost the same as before. First, we replace
@@
   open Mono
-  let f = ...
-  let f_string i = string_of_int (f (int_of_string i))
-  let () = Worker.compute f_string ()
+  let worker i = ...
+  let worker_string i = string_of_int (worker (int_of_string i))
+  let () = Worker.compute worker_string ()
 \end{ocaml}
-The master program is almost unchanged. We simply need to replace
-\of{Poly} with \of{Mono} and to encode tasks as strings:
+The master program is modified in a similar way.
+We simply replace \of{Poly} with \of{Mono} and encode/decode integers
+as strings, as follows:
 \begin{ocaml}
@@
     !l
-\end{ocaml}
-Consequently, \of{master} is now receiving a string as argument:
-\begin{ocaml}
   let master (i,()) fi = s := !s + int_of_string fi; []
@@
 This is the case when \of{fold} is an associative operation (which
-implies that type \of{'b} and \of{'c} are the same  and that the
-accumulator should be a neutral element for \of{fold}). Whenever
+implies that type \of{'b} and \of{'c} are the same).
+Whenever
 \of{fold} is also commutative, we can perform even more \of{fold}
@@
 itself a task to be performed in parallel with \of{map} operations.
-Thus, when an intermediate \of{map} result is available, the
+Thus tasks are of two different kinds and we use a sum type to make
+the distinction.
+When an intermediate \of{map} result is available, the
 accumulator may be already in use in a \of{fold} operation. Hence we
@@
 core, a new sub-process is created using \of{Unix.fork}; once
-completed, the sub-process marshals the result into a local file. The
+completed, the sub-process marshals the result into a 
+local file. The
 main process maintains a table mapping each sub-process ID to the
@@
 \end{flushleft}
-
-It would be more efficient to use \unix\ pipes instead of local files
-but pipes are severely restricted w.r.t. data size. To be more
-generic, we make no assumptions about the size of the data. Providing
-an alternative implementation using pipes would be straightforward.
+We also have an alternative implementation using
+\unix\ pipes instead of local files.
 
@@
 Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
-implementation of Google's MapReduce; rather it is more generic and
-not restricted to just key/value pairs.
+implementation of Google's MapReduce.
 There are other ways to exploit multi-core architectures. One of these

### 2010-05-25 doc/icfp-2010/response.txt
*** empty log message ***

Response to review #5A:
-----------------------

> The major weakness however is that the task creation is
> completely static, i.e., all tasks must be created ahead of time.

It is true that in the derived API (Figure 1 on page 4) the tasks are
completely static. But in the low-level API, new tasks can be
generated dynamically, as mentioned on page 2, first paragraph (right
below the type of the master function).

> The author claims that strategy utilizes efficient storage of data

Actually, we did not, and only mentioned that as an interesting
extension to our library (last paragraph of Section 7, first bullet).

Response to review #5B:
-----------------------

> I wondered why oc4mc wasn't used for the single machine version

As far as we understand, oc4mc is a modification of OCaml (available
for a single architecture). On the contrary, our library does not
require any modification of the compiler.

Response to review #5C:
-----------------------

> (also, the example code is heavily imperative with lots of unnecessary refs).

It is true that the example code in Section 3 is making use of
imperative features. The reason is that it uses the low-level API of
Section 2. However, the derived API of Section 4 provides fold-like
functions and thus allows pure functional code, even if we did not show
in the paper.

> the connections to functional programming are tenuous

Our library API is mostly composed of higher-order functions: in the
low-level API (Section 2), master is a higher-order function; in the
derived API (Figure 1), all functions are higher-order map/fold.

Additionally, our library is heavily relying on OCaml module
system: first, to provide three different implementations with the
same interface; second, to achieve maximal code reuse in the library
itself using functors (as explained at the end of page 3 for instance).

Responses for explicit questions for authors:
---------------------------------------------

> What happens if a worker raises an exception?

It is considered as a worker failure. We leave it to the user to catch
it and handle it (using a sum type for instance).

> It appears that your local implementation does not handle faults in the
> workers (e.g., if you run out of a system resource). Is that the case?

Yes, it is. One possibility to overcome this would be to change the
resource limits in the user worker function (right after the fork).

> In the network implementation, how to you initiate the workers?

As of now, we do not provide any facility in order to do this.
Currently, we assume the user deploys the workers manually.

### 2010-07-01 doc/ml-workshop-2010/main.tex
*** empty log message ***

--- 2010-04-02 doc/icfp-2010/main.tex
@@
 
-\usepackage{amsmath,graphicx,color,url}
+\usepackage{amsmath,color,url}
 
@@
 
-\conferenceinfo{ICFP '10}{September 27-29, 2010, Baltimore, Maryland.} 
+\conferenceinfo{Workshop on ML 2010}{September 26, 2010, Baltimore, Maryland.} 
 \copyrightyear{2010} 
@@
 
-\titlebanner{submitted to ICFP 2010}        % These are ignored unless
-\preprintfooter{Experience Report: Distributed Computing for the Common Man}   % 'preprint' option specified.
+\titlebanner{submitted to Workshop on ML 2010}        % These are ignored unless
+\preprintfooter{Distributed Computing for the Common Man}   % 'preprint' option specified.
 
-\title{Experience Report: Distributed Computing for the Common Man}
+\title{Distributed Computing for the Common Man}
 %\subtitle{Subtitle Text, if any}
@@
 
-\category{D1.1}{Programming techniques}{Applicative (Functional) Programming}
+% \category{D1.1}{Programming techniques}{Applicative (Functional) Programming}
 
-\terms
-Distributed Computing, Functional Programming
+% \terms
+% Distributed Computing, Functional Programming
 
@@
 
-The application domain of such a distributed computing library is
-manyfold.  The most striking application, as also relevant to our
-research endeavours, is in validating thousands of verification
-conditions using automated theorem provers. Having noted this, it
-should be apparent that the library finds use in efficient execution
-of any large-scale expensive computation that is parallelizable.
-% what it is and what it is not
-% \begin{itemize}
-% \item it is not a library which does parallelization
-% \item it is a library for reliable distributed execution of
-%   parallelizable computation 
-% \end{itemize}
-% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
-% programming, ironically) ;  however, there are differences:
-% \begin{itemize}
-% \item polymorphic, higher-order API
-% \item more generic (cores/network)
-% \item does not focus on association lists
-% \item we don't have data locality (future work)
-% \end{itemize}
-% what is the target audience/applications: 
-% for instance, use if automatic provers on thousands of verification
-% conditions, on computing infrastructures which can be
-% \begin{itemize}
-% \item a single machine with multiple cores, possibly remote
-% \item several machines, small or large in number, over a network
-% \end{itemize}
-\paragraph{Distributed Computing.}
-A typical distributed computing library provides the following (we
-borrow some terminology from Google's MapReduce~\cite{mapreduce}):
-\begin{itemize}
-\item A notion of \emph{tasks} which denote atomic computations
-  to be performed in a distributed manner; 
-\item A set of processes (possibly executing on remote machines)
-  called \emph{workers} that perform
-  the tasks, producing results;
-\item A single process called a \emph{master} which is in charge
-  of distributing the tasks among the workers and managing results
-  produced by the workers.
-\end{itemize}
 In addition to the above, distributed computing environments also
@@
 
+Our library, \functory, is implemented in \Ocaml\ and is available
+from \url{http://www.lri.fr/~filliatr/functory/}.
+
+TODO : refer to the technical report available on-line
+
 \section{API}\label{sec:API}
-In this section, we describe the
-generic API which we shall extend with several specialized
-functions in later sections. The main function in our API follows the idea
-sketched in the introduction. It has the following signature:
+
+Our work was initially inspired by Google's
+MapReduce\footnote{Ironically, Google's approach itself was inspired
+  by functional programming primitives.}, from where we borrow some 
+terminology:
+\begin{itemize}
+\item A notion of \emph{tasks} which denote atomic computations
+  to be performed in a distributed manner; 
+\item A set of processes (possibly executing on remote machines)
+  called \emph{workers} that perform
+  the tasks, producing results;
+\item A single process called a \emph{master} which is in charge
+  of distributing the tasks among the workers and managing results
+  produced by the workers.
+\end{itemize}
+The main function of our API combines these three notions of tasks,
+master and workers into a single higher-order polymorphic function
+\of{compute}:
 \begin{ocaml}
@@
 in some internal data structure.
-The next section will describe the usage of this generic interface to
-perform the traditional map and fold operations.
 
-Actually, our library provides not just a single \ocaml{compute}
+Our library provides not just a single \ocaml{compute}
 function as above, but instead five different versions
-depending on the execution context. 
-The five possible contexts are the following: 
+depending on the execution context:
 \begin{enumerate}
@@
 
-Our library is organized into three modules: \of{Sequential} for the
-pure sequential implementation, \of{Cores} for multiple cores on the
-same machine and \of{Network} for a network of machines, respectively.
-The \of{Network} module itself is organized into three sub-modules,
-called \of{Same}, \of{Poly} and \of{Mono},
-corresponding to contexts 3, 4 and 5 above. 
-The next section demonstrates the use of these modules.
+% Our library is organized into three modules: \of{Sequential} for the
+% pure sequential implementation, \of{Cores} for multiple cores on the
+% same machine and \of{Network} for a network of machines, respectively.
+% The \of{Network} module itself is organized into three sub-modules,
+% called \of{Same}, \of{Poly} and \of{Mono},
+% corresponding to contexts 3, 4 and 5 above. 
+% The next section demonstrates the use of these modules.
 
-% The different
-% API functions are summarized in the following table (their signatures
-% being given above):
-% \begin{center}
-%   \begin{tabular}{|c|c|c|c|c|}
-%     \hline
-%     \ocaml{Sequential} & \ocaml{Cores} &
-%     \multicolumn{3}{|c|}{\ocaml{Network}} 
-%     \\\cline{3-5}
-%     &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
-%     \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
-%     \ocaml{master}  & \ocaml{master}  \\
-%     & & & \ocaml{worker}  & \ocaml{worker} \\\hline
-%   \end{tabular}
-% \end{center}
-% The following section introduces other interesting APIs that we
-% derive from this generic API.
+  \begin{ocaml}
+    val map : 
+      f:('a -> 'b) -> 'a list -> 'b list 
+    val map_local_fold : 
+      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+      'c -> 'a list -> 'c 
+    val map_remote_fold : 
+      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+      'c -> 'a list -> 'c 
+    val map_fold_ac : 
+      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+      'b -> 'a list -> 'b 
+    val map_fold_a : 
+      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+      'b -> 'a list -> 'b
+  \end{ocaml}
 
@@
 
-\section{Derived API}\label{sec:derived}
-
-In most cases, the easiest way to parallelize an execution it to make
-use of operations over lists, where processing of the
-list elements are done in parallel. 
-The previous section illustrated one such case.
-To facilitate such a processing,
-we now derive the most commonly used list operations from our generic
-API.
-
-The most obvious operation is the traditional map operation over
-lists, that is:
-\begin{ocaml}
-  val map : ('a -> 'b) -> 'a list -> 'b list
-\end{ocaml}
-The next natural operation is a combination of map and fold
-operations, that is a function like
-\begin{ocaml}
- val map_fold :
-   map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-   'c -> 'a list -> 'c
-\end{ocaml}
-which, given two functions, an accumulator \of{a} and a list \of{l}, computes
-\begin{equation}\label{eq:map-fold}
-  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{map} ~ x_1)) (\of{map} ~ x_2))
-  ... (\of{map} ~ x_n)
-\end{equation}
-for some permutation $[x_1,x_2,...,x_n]$ of the list \of{l}.
-We assume that the \of{map} operations are always performed in parallel.
-Regarding \of{fold} operations, we distinguish two cases:
-\begin{itemize}
-\item either \of{fold} operations are computationally less expensive
-  than \of{map} and we perform them locally;
-\item or \of{fold} operations are computationally expensive and we
-  perform them in parallel.
-\end{itemize}
-We thus provide two functions \of{map_local_fold} and \of{map_remote_}
-\of{fold}.
-
-In the case of \of{map_remote_fold}, only one \of{fold} operation can
-be performed at a time (possibly in parallel with \of{map}
-operations), as obvious from~(\ref{eq:map-fold}). However, there are
-cases where several \of{fold} operations can be performed in parallel,
-as early as intermediate results of \of{fold} operations are available.
-This is the case when \of{fold} is an associative operation (which
-implies that type \of{'b} and \of{'c} are the same).
-Whenever
-\of{fold} is also commutative, we can perform even more \of{fold}
-operations in parallel. Thus our API provides two functions
-\of{map_fold_a} and \of{map_fold_ac} for these two particular cases.
-The five operations of the derived API are summarized
-in Figure~\ref{fig:derived}.
-\begin{figure}[t]
-  \begin{ocaml}
-    val map : 
-      f:('a -> 'b) -> 'a list -> 'b list 
-    val map_local_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_remote_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_fold_ac : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b 
-    val map_fold_a : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b
-  \end{ocaml}
-  \caption{Derived API.}
-\label{fig:derived}
-\end{figure}
-
-These five functions are actually derived from the generic API.  The
-generality is achieved by implementing these derivations as a functor.
-This functor, \of{Derive}, is parameterized by the \of{compute}
-function, as follows:
-\begin{ocaml}
-module Derive
-  (X : sig
-     val compute : 
-       worker:('a -> 'b) -> 
-       master:('a * 'c -> 'b -> ('a * 'c) list) ->
-       ('a * 'c) list -> unit
-   end) = struct ... end
-\end{ocaml}
-We now explain how each function from Figure~\ref{fig:derived} is
-implemented using \of{compute} inside the functor body.
-
-\paragraph{\of{map}.} 
-Function \of{map} is easily implemented by associating an integer
-index with each element (of type \of{'a}) of the input list. Thus the
-list of tasks passed to \of{compute} is a list of pairs of type \of{'a * int}.
-As soon as an intermediate result is available, it is passed to
-\of{master}, which stores it in a local table using the index as a key
-and generates no new task. Whenever \of{compute} returns, we easily
-recover the output list from the values stored in the table.
-
-\paragraph{\of{map_local_fold}.} 
-Function \of{map_local_fold} is implemented using a local reference
-storing the current accumulator (of type \of{'c}).  As soon as an
-intermediate result is available, the \of{master} combines it with the
-accumulator using the function \of{fold}. Whenever \of{compute}
-returns, we simply return the current accumulator.
-
-\paragraph{\of{map_remote_fold}.}
-In this case, combining the accumulator with an intermediate result is
-itself a task to be performed in parallel with \of{map} operations.
-Thus tasks are of two different kinds and we use a sum type to make
-the distinction.
-When an intermediate \of{map} result is available, the
-accumulator may be already in use in a \of{fold} operation. Hence we
-need to store the intermediate \of{map} result in a table local to the
-master, to be
-used as soon as the accumulator becomes available.
-The \of{master} function distinguishes between \of{map} and \of{fold}
-results and, accordingly, updates the table and the accumulator. For a
-\of{map} result, either we combine it with the accumulator in a new
-task or we simply store it in the local table. For a \of{fold} result,
-either we simply store it into the accumulator or we immediately
-combine it with a pending \of{map} result from the table.
-Whenever \of{compute} returns, the accumulator must be available and
-we simply return it.
-
-\paragraph{\of{map_fold_ac}.}
-The function \of{map_fold_ac} is easier than the previous one, since
-we do not need to store more than one intermediate result. Indeed, as
-soon as two intermediate results are available, we immediately combine
-them using a \of{fold} operation. Thus, whenever the \of{master}
-receives some result, it either stores it or combines it with the
-accumulator, depending on the availability/non availability of the
-accumulator.  Whenever \of{compute} returns, the accumulator must be
-available and we simply return it.
-
-\paragraph{\of{map_fold_a}.}
-Function \of{map_fold_a} is definitely the trickiest. Indeed, the
-operation being only associative, we can only combine \emph{adjacent}
-intermediate results. In order to perform this, we store intermediate
-results in a local table according to indices of the input list. More
-precisely, if we have already computed
-\begin{displaymath}
-  x_i \oplus x_{i+1} \oplus \dots \oplus x_j
-\end{displaymath}
-where $\oplus$ denotes the \of{fold} operation, then we associate this
-value with indices $i,j$. As soon as a result of indices $k,i-1$ or
-$j+1,k$ is available, it can be combined with the result of $i,j$.
-Whenever \of{compute} returns, the table should only contain a result
-corresponding to the full input list and we simply return it.
-
-\medskip
-These five functions are the most natural derivations of the
-\of{compute} functions. There are obviously other possible derivations
-(e.g. a variant of \of{map_fold} where only \of{fold} is meaningful);
-in most cases, they could be derived in a straightforward way from our
-generic API.
-
-% TODO cannot be instantiated in the monomorphic case
-
-\section{Implementation Details}\label{sec:implem}
-
-We now describe the implementations of the various modules introduced
-in Section~\ref{sec:API}. The implementation of the \of{Sequential}
-module is straightforward and does not require any explanation.
-
-\subsection{Multiple Cores}
-
-The \of{Cores} module implements the distributed computing library for
-multiple cores on the same machine. As illustrated in
-Section~\ref{sec:example}, 
-it provides a function
-\of{set_number_of_cores: int -> unit} to indicate the number of cores
-to be used. The number passed to this function may be different from
-the actual number of cores in the machine; it is rather the number of
-tasks to be performed simultaneously.
-% TODO comment more on the number of cores?
-% Thus a program using the \of{Cores} module typically begins with 
-% \begin{ocaml}
-% open Cores
-% let () = set_number_of_cores 8
-% \end{ocaml}
-
-The \of{Cores} module is implemented with \unix\ processes, using the
-\of{fork} and \of{wait} system calls provided by the \of{Unix} library
-of \Ocaml. The idea is pretty simple.  The \of{compute} function
-maintains a global table of pending tasks and keeps track of the
-number of idle cores.  Whenever there is a pending task and an idle
-core, a new sub-process is created using \of{Unix.fork}; once
-completed, the sub-process marshals the result into a 
-local file. The
-main process maintains a table mapping each sub-process ID to the
-input task and the local file name. It waits for any completed
-sub-process using \of{Unix.wait} and recovers the result from the
-local file. Then function \of{master} is applied, which may generate
-new tasks.  The main loop can be depicted in the following way:
-\begin{flushleft}
-  \quad  \textbf{while} pending tasks $\lor$ pending sub-processes \\
-  \quad  \quad \textbf{while} pending tasks $\land$ idle cores \\
-  \quad  \quad \quad create new sub-process for some task \\
-  \quad  \quad \textbf{wait} for any completed sub-process \\
-  \quad  \quad \quad push new tasks generated by \of{master} \\
-\end{flushleft}
-We also have an alternative implementation using
-\unix\ pipes instead of local files.
-
-The scheduling of tasks to the different cores is left to the
-operating system, through \of{Unix.fork}. Thus it may be the case that
-two tasks are executed on the same core, even if the declared number
-of cores is less or equal than the actual number of cores on the machine.
-% TODO: comment on user-provided scheduling?
-
-\subsection{Network of Machines}
-
-The \of{Network} module implements the distributed computing library
-for a network of machines. 
-As illustrated in Section~\ref{sec:example}, it
-provides a function
-\of{declare_workers: n:int -> string -> unit} to fill a table of
-worker machines. 
-% The argument \of{n} indicates the number of cores to
-% be used in each worker machine, analogous to the \of{Cores}
-% module. Thus a program using the \of{Network} module typically begins
-% with
-% \begin{ocaml}
-% open Network
-% let () = declare_workers ~n:12 "hostname1"
-% let () = declare_workers ~n:8  "192.168.24.2"
-% ...
-% \end{ocaml}
-
-The \of{Network} module is based on a traditional TCP-based client/server
-architecture, where each worker is a server and the master is the
-client of each worker. The main execution loop is similar to the one
-in the \of{Cores} module, where distant processes on remote machines
-correspond to sub-processes and idle cores are the idle cores 
-of remote workers. In addition to this, it also involves issues of message
-transfer and fault tolerance, which are subsequently described.
-
-\subsubsection{Protocol}\label{sec:protocol}
-
-Messages sent from master to workers could be any of the following kinds:
-\begin{description}
-\item[\of{Assign(id:int, f:string, x:string)}] This message assigns a
-  new task to the worker, the task being identified by the unique
-  integer \of{id}. The task to be performed is given by strings \of{f}
-  and \of{x}, which are interpreted depending on the context.
-
-\item[\of{Kill(id:int)}] This message tells the worker to kill the task
-  identified by \of{id}.
-
-\item[\of{Stop}] This message informs the worker about completion of
-  the computation, so that it may choose to exit.
-
-\item[\of{Ping}] This message is used to check if the worker is still
-  alive, expecting a \of{Pong} message from the worker in return.
-\end{description}
-Messages sent by workers could be any of the following kinds:
-\begin{description}
-\item[\of{Pong}] This message is an acknowledgment for a \of{Ping}
-  message from the master.
-
-\item[\of{Completed(id:int, s:string)}] This message indicates the
-  completion of a task identified by \of{id}, with result \of{s}.
-
-\item[\of{Aborted(id:int)}] This message informs the master that the
-  task identified by \of{id} is aborted, either as a response to a
-  \of{Kill} message or because of a worker malfunction.
-\end{description}
-Our implementation of the protocol works across different
-architectures, so that master and workers could be run on completely
-different platforms w.r.t. endianness, version of \Ocaml\ and
-operating system.
-
-\subsubsection{Fault Tolerance}\label{sec:fault}
-
-The main issue in any distributed computing environment is the ability
-to handle faults, which is also a distinguishing feature of our
-library.  Faults are mainly of two kinds: either a worker is stopped,
-and possibly later restarted; or a worker is temporarily or
-permanently unreachable on the network. To provide fault tolerance,
-our master implementation is keeping track of the status of each
-worker.  This status is controlled by two timeout parameters $T_1$ and
-$T_2$ and \of{Ping} and \of{Pong} messages sent by master and workers,
-respectively. There are four possible statuses for a worker:
-\begin{description}
-\item[\of{not connected}:] there is no ongoing TCP connection between
-  the master and the worker;
-\item[\of{alive}:] the worker has sent some message
-  within $T_1$ seconds;
-\item[\of{pinged}:] the worker has not sent any message within $T_1$
-  seconds and the master has sent the worker a
-  \of{Ping} message within $T_2$ seconds;
-\item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
-  message (for more than $T_2$ seconds).
-\end{description}
-% TODO picture
-Whenever we receive a message from a worker, its status changes to
-\of{alive} and its timeout value is reset.
-\begin{center}
-  \includegraphics{state.mps}
-\end{center}
-
-Fault tolerance is achieved by exploiting the status of workers as
-follows. First, tasks are only assigned to workers with either
-\of{alive} or \of{pinged} status. Second, whenever a worker executing
-a task $t$ moves to status \of{not connected} or \of{unreachable}, the
-task $t$ is rescheduled, which means it is put back in the set of
-pending tasks. Whenever a task is completed, any rescheduled copy of
-this task is either removed from the set of pending tasks or killed if
-it was already assigned to another worker.
-
-\subsubsection{Three Network Implementations} % TODO: change title
-
-As mentioned in Section~\ref{sec:API}, the \of{Network} module
-actually provides three different implementations, according to three
-different execution scenarios (namely, \of{Same}, \of{Poly} and
-\of{Mono}). There is
-actually a single, generic implementation, which is then instantiated
-in three different ways, corresponding to the three scenarios.  The
-generic implementation assumes that only strings are being transmitted
-over the network, as evident from the \of{Assign} and \of{Completed}
-messages in Section~\ref{sec:protocol}. Depending on the
-implementation, these strings could be marshaled closures, marshaled
-values or simply user strings.
-
-% The signature for the generic
-% implementation is therefore as follows:
-% \begin{ocaml}
-%   generic_master:
-%     assign_job:('a -> string * string) ->
-%     master:('a * 'c -> string -> ('a * 'c) list)  ->
-%     ('a * 'c) list -> unit
-% \end{ocaml}
-% The first argument \of{assign_job} is a function indicating the two
-% strings that are passed in the \of{Assign} message of the protocol.
-% We can now distinguish the three cases, based on \of{assign_job}:
-% \begin{description}
-% \item[\of{Same}:] in this case, \of{assign_job} simply returns
-%   the marshaled closure together with the marshaled value.
-% \item[\of{Poly}:] in this case, \of{assign_job} is simply
-%   returning the marshaled value (the first string being meaningless in
-%   this context).
-% \item[\of{Mono}:] in this case, \of{assign_job} is immediately
-%   returning its argument (the first string being meaningless in
-%   this context).
-% \end{description}
-
 \section{Experiments}\label{sec:experiments}
 
-Our library, \functory, is implemented in \Ocaml\ and is available
-from \url{http://www.lri.fr/~filliatr/functory/}.
-In this section, we show the potential of this library using two 
-computationally extensive examples.
+Here we demonstrate the potential of our library for using the
+classical $N$-queens problem and large-scale verification using SMT-solvers.
 
-\subsection{N-queens}
-
-The first example is the classical $N$-queens problem, where we
-compute the total number of ways to place $N$ queens on a $N\times N$
-chessboard in such a way no two queens attack each other.
-The following table shows execution times for various values of $N$
-and our three different implementations: \of{Sequential}, \of{Cores}, and
-\of{Network}. The purpose of this experiment is to measure the speedup
-w.r.t. the sequential implementation. Therefore, all
-computations are performed on the same machine, a 8 cores Intel Xeon
-3.2 GHz running Debian Linux. The sequential implementation uses a
-single core. The multi-core implementation uses the 8 cores of
-the machine. The network implementation uses 8 workers running locally and
-a master running on a remote machine (which incurs communication cost).
-
-The first column shows the value of $N$.  Each computation is divided
-into a certain number of tasks, by placing some of the queens before
-starting the distributed computation.  The number of tasks is shown in
-second column.  Then the last three columns show execution times in
-seconds for the three implementations. The figures within brackets
-show the speedup w.r.t. sequential implementation.
+\paragraph{N-queens.} 
+The following table shows execution times\footnote{machine
+  characteristics: 8 cores Intel Xeon 3.2 GHz running Debian Linux.} 
+in seconds for various values of $N$
+and our three different implementations (sequential, 8 cores on a
+single machine and a network of 8 workers).
 \begin{center}
@@
 \end{center}
-From the table above, it is clear that the \of{Cores} and \of{Network}
-implementations provide a significant speedup. As evident from the
-last row, the speedup is almost 8, which is also the number of
-cores we use.  It is also evident from the last column that the
-\of{Network} implementation performs significantly better when the
-computation time dominates in the total execution time.  The two extreme
-cases correspond to the second and the last row: in the second row, the
-communication time dominates and is in fact more than 91\%\ of the
-total execution time; on the other hand, for the last row
-communication time amounts to just 4.6\%\ of the total execution time.
-As expected, the network implementation is only beneficial when the
-computation time for each individual task is significant, which is the
-case in realistic examples.
+% From the table above, it is clear that the \of{Cores} and \of{Network}
+% implementations provide a significant speedup. As evident from the
+% last row, the speedup is almost 8, which is also the number of
+% cores we use.  It is also evident from the last column that the
+% \of{Network} implementation performs significantly better when the
+% computation time dominates in the total execution time.  The two extreme
+% cases correspond to the second and the last row: in the second row, the
+% communication time dominates and is in fact more than 91\%\ of the
+% total execution time; on the other hand, for the last row
+% communication time amounts to just 4.6\%\ of the total execution time.
+% As expected, the network implementation is only beneficial when the
+% computation time for each individual task is significant, which is the
+% case in realistic examples.
 
-\subsection{SMT Solvers}
-
-Here we demonstrate the potential of our library for our application
-needs as mentioned in the introduction. Our computing
-infrastructure consists of 3 machines with 4, 8 and 8 cores
-respectively. The benchmark we use for
-experimentation consists of around 80 challenging verification conditions
-(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
-VC is stored in a file, which is accessible over
-NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
-which are all installed on each of the machines in use.
-
-The master program proceeds by reading the file names, turning them
-into tasks by multiplying them by the number of provers (more than
-300 tasks in total).
-Each worker in turn invokes the given prover on the given file.
-We set a timeout limit of 1 minute.
-Each task completes with one of the four possible outcomes: \emph{valid},
+\paragraph{SMT Solvers.}
+Our setup consists of 3 machines with 4, 8 and 8 cores
+respectively and the benchmark consists of
+80 challenging verification conditions (VC).  
+The purpose of the experiment is to check the validity of each VC
+using the four SMT-solvers Alt-Ergo, Simplify, Z3 and CVC3.
+With a timeout of 1 minute, each task completes 
+with one of the four possible outcomes: \emph{valid},
 \emph{unknown} (depending on
@@
 \emph{timeout} and \emph{failure}.
-The figure below shows the total time in minutes spent by each prover
-for each possible outcomes.
+Timings in minutes are the following:
 \begin{center}
@@
 infrastructure, it completes in 22 minutes and 37 seconds, giving us a
-speedup of more than 16$\times$. We are still far away from the ideal
-ratio of 20$\times$ (we are using 20 cores), since some provers are
-allocating a lot of memory and time spent in system calls is not
-accounted for in the total observed time. However, a ratio of
-16$\times$ is already a significant improvement for our day-to-day
-experiments.
-
-\section{Conclusion and Future Work}\label{sec:future}
-
-We presented a distributed programming environment for functional
-programming. The main features are the genericity of the interface,
-which makes use of polymorphic higher-order functions, and the ability
-to easily switch between sequential, multi-core, and network
-implementations. A distinguishing feature of our library is a
-robust fault-tolerance mechanism which relieves
-the user of cumbersome implementation details.
-
-Closest to the approach in this paper is Yohann Padioleau's MapReduce
-implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
-top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
-protocol for message passing.  Currently, we have less flexibility
-w.r.t. deployment of the user program than OCamlMPI; on the other
-hand, we provide a 
-more generic API together with fault tolerance.  There are other
-distributed computing libraries on top of which one could implement
-the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
-them. However, \JoCaml\ does not provide fault tolerance, which is
-indispensable in a distributed setting. The user has to include code
-for fault tolerance, as already demonstrated in some \JoCaml\
-experiments~\cite{mandel2008}.
-
-There are other implementations of distributed computing in the
-context of functional programming. One is the Disco
-project~\cite{disco}, which implements exactly Google's MapReduce in
-Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
-implementation of Google's MapReduce.
-There are other ways to exploit multi-core architectures. One of these
-is data parallelism, which is also relevant in the functional
-programming setting~\cite{parallel-haskell}. Our work
-does not target data parallelism at all.
-
-There are still some interesting features that could be added to our
-library. 
-\begin{itemize}
-\item 
-  One is the ability to efficiently assign tasks to workers depending
-  on resource parameters, such as data locality, CPU power, memory,
-  etc. This could be achieved by providing the user with the means to control
-  task scheduling.  
-\item 
-  Another interesting feature could be the ability
-  to add or remove workers dynamically. Currently, our library assumes
-  that the list of workers to be used is given \emph{a priori}.
-\item
-  Our library provides limited support for displaying real-time
-  information about the computations. Processing and storing
-  information about tasks locally in the master is straightforward; 
-  monitoring it in real-time is less obvious.
-\item 
-  One very nice feature of Google's MapReduce is the possibility to
-  use redundantly several idle workers on the same tasks
-  for speedup when reaching the end of computation.
-  Since we already have the fault tolerance implemented, this
-  optimization should be straightforward to add to our library.
-\end{itemize}
-We intend to enrich our library with all above features.
-
-%\appendix
-
-\acks
-The authors are grateful to the ProVal team for support and comments
-on early versions of the library and of this paper.
-
-%\vfill\pagebreak
-% We recommend abbrvnat bibliography style.
-\bibliographystyle{abbrvnat}
-\bibliography{./biblio}
+speedup of more than 16$\times$.
 

### 2010-07-01 doc/ml-workshop-2010/main.tex
*** empty log message ***

@@
 
-In addition to the above, distributed computing environments also
-implement mechanisms for fault tolerance, efficient storage, and
-distribution of tasks. This is required to handle network failures
-that may occur, as well as to optimize the usage of machines in the
-network. Another concern of importance is the transmission of messages
-over the network. This requires efficient 
-\emph{marshaling} of data, that is encoding and decoding of data 
-for transmission over different computing environments.  It is desirable to
-maintain architecture independence while transmitting marshalled data,
-as machines in a distributed computing environment often run on
-different hardware architectures and make use of varying software
-platforms. For example, machine word size or endianness may be different
-across machines on the network.
+% In addition to the above, distributed computing environments also
+% implement mechanisms for fault tolerance, efficient storage, and
+% distribution of tasks. This is required to handle network failures
+% that may occur, as well as to optimize the usage of machines in the
+% network. Another concern of importance is the transmission of messages
+% over the network. This requires efficient 
+% \emph{marshaling} of data, that is encoding and decoding of data 
+% for transmission over different computing environments.  It is desirable to
+% maintain architecture independence while transmitting marshalled data,
+% as machines in a distributed computing environment often run on
+% different hardware architectures and make use of varying software
+% platforms. For example, machine word size or endianness may be different
+% across machines on the network.
 
-\paragraph{A Functional Programming Approach.}
-Our work was initially inspired by Google's
-MapReduce\footnote{Ironically, Google's approach itself was inspired
-  by functional programming primitives.}. However, our functional
-programming environment allows us to be more generic. 
-The main idea behind our approach is that
-workers may implement any polymorphic function:
-\begin{ocaml}
-  worker: 'a -> 'b
-\end{ocaml}
-where \of{'a} denotes the type of tasks and \of{'b} the type of results.
-Then the master is given a list of initial tasks, together with a
-function to handle the results:
-\begin{ocaml}
-  master: ('a -> 'b -> 'a list) -> 'a list -> unit
-\end{ocaml}
-The function passed to the master is applied whenever a result is
-available and may in turn generate new tasks (hence the return type
-\of{'a list}).  The master is executed as long as there are pending
-tasks.
+% \paragraph{A Functional Programming Approach.}
+% Our work was initially inspired by Google's
+% MapReduce\footnote{Ironically, Google's approach itself was inspired
+%   by functional programming primitives.}. However, our functional
+% programming environment allows us to be more generic. 
+% The main idea behind our approach is that
+% workers may implement any polymorphic function:
+% \begin{ocaml}
+%   worker: 'a -> 'b
+% \end{ocaml}
+% where \of{'a} denotes the type of tasks and \of{'b} the type of results.
+% Then the master is given a list of initial tasks, together with a
+% function to handle the results:
+% \begin{ocaml}
+%   master: ('a -> 'b -> 'a list) -> 'a list -> unit
+% \end{ocaml}
+% The function passed to the master is applied whenever a result is
+% available and may in turn generate new tasks (hence the return type
+% \of{'a list}).  The master is executed as long as there are pending
+% tasks.
 
-Our library makes use of \Ocaml's marshaling capabilities as much as
-possible. Whenever master and worker executables are exactly the same,
-we can marshal polymorphic values and closures. However, it is not
-always possible to have master and workers running the same
-executable. In this case, we cannot marshal closures anymore but we
-can still marshal polymorphic values as long as the same version of
-\Ocaml\ is used to compile master and workers. When different versions
-of \Ocaml\ are used, we can no longer marshal values but we can still
-transmit strings between master and workers. Our library adapts to all
-these situations, by providing several APIs.
+% Our library makes use of \Ocaml's marshaling capabilities as much as
+% possible. Whenever master and worker executables are exactly the same,
+% we can marshal polymorphic values and closures. However, it is not
+% always possible to have master and workers running the same
+% executable. In this case, we cannot marshal closures anymore but we
+% can still marshal polymorphic values as long as the same version of
+% \Ocaml\ is used to compile master and workers. When different versions
+% of \Ocaml\ are used, we can no longer marshal values but we can still
+% transmit strings between master and workers. Our library adapts to all
+% these situations, by providing several APIs.
 
-\medskip
-The paper is organized as follows. In Section~\ref{sec:API}, we
-describe the main idea behind the generic API, and introduce
-\of{master} and \of{worker} programs. Section~\ref{sec:example}
-illustrates the API usage with an example.
-In Section~\ref{sec:derived}, we
-derive some interesting functions from the generic
-API. Section~\ref{sec:implem} delves into the implementation details
-of our library, while Section~\ref{sec:experiments} illustrates the
-potential of the presented library through experimental evaluation. We
-compare our approach with relevant related work in
-Section~\ref{sec:future} and outline our future work.
-
-Our library, \functory, is implemented in \Ocaml\ and is available
-from \url{http://www.lri.fr/~filliatr/functory/}.
-
-TODO : refer to the technical report available on-line
-
-\section{API}\label{sec:API}
+% \medskip
+% The paper is organized as follows. In Section~\ref{sec:API}, we
+% describe the main idea behind the generic API, and introduce
+% \of{master} and \of{worker} programs. Section~\ref{sec:example}
+% illustrates the API usage with an example.
+% In Section~\ref{sec:derived}, we
+% derive some interesting functions from the generic
+% API. Section~\ref{sec:implem} delves into the implementation details
+% of our library, while Section~\ref{sec:experiments} illustrates the
+% potential of the presented library through experimental evaluation. We
+% compare our approach with relevant related work in
+% Section~\ref{sec:future} and outline our future work.
 
@@
 \end{itemize}
+
+Our library, \functory, is implemented in \Ocaml.
+It makes use of \Ocaml's marshaling capabilities as much as
+possible. Whenever master and worker executables are exactly the same,
+we can marshal polymorphic values and closures. However, it is not
+always possible to have master and workers running the same
+executable. In this case, we cannot marshal closures anymore but we
+can still marshal polymorphic values as long as the same version of
+\Ocaml\ is used to compile master and workers. When different versions
+of \Ocaml\ are used, we can no longer marshal values but we can still
+transmit strings between master and workers. Our library adapts to all
+these situations, by providing several APIs.
+
+Our library is available
+at \url{http://www.lri.fr/~filliatr/functory/}.
+A report is also available there, which provides more
+technical details.
+
+\section{API}\label{sec:API}
+
 The main function of our API combines these three notions of tasks,
@@
 
+In addition to this low-level API, our library also provides more
+traditional \of{map} and \of{fold} higher-order functions with
+distributed computation. For instance, 
 \begin{ocaml}
-    val map : 
-      f:('a -> 'b) -> 'a list -> 'b list 
   val map_local_fold : 
@@
     'c -> 'a list -> 'c 
-    val map_remote_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_fold_ac : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b 
-    val map_fold_a : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b
 \end{ocaml}
+applies \of{map} to each element of the list in a distributed way and
+folds the results locally to the master.
 
@@
 where $f$ is a pure and computationally expensive function.
-We start with the sequential implementation provided by the library:
+Using the sequential implementation and the low-level \of{compute}
+function, the code looks like:
 \begin{ocaml}
-  open Sequential
+1.  open Sequential
+2.  let f i = ...
+3.  let n = int_of_string Sys.argv.(1)
+4.  let tasks = ... (* the lists [0; 1; ...; n] *)
+5.  let s = ref 0
+6.  let master (i,()) fi = s := !s + fi; []
+7.  let () = compute ~worker:f ~master tasks
 \end{ocaml}
-The idea is to test our code on small values of $n$ before
-parallelizing the computation for larger values.
-Let us assume that $f$ is implemented as a top-level \of{worker} function:
+To switch to the multi-core implementation, say with 4 cores, we
+simply replace line 1 with
 \begin{ocaml}
-  let worker i = ...
+1a.  open Cores
+1b.  let () = set_number_of_cores 4
 \end{ocaml}
-and that $n$ is obtained from the command line:
+Similarly, in order to use a network of two machines \texttt{m1} and
+\texttt{m2}, with 4 and 8 cores respectively, line 1 is replaced with
 \begin{ocaml}
-  let n = int_of_string Sys.argv.(1)
-\end{ocaml}
-The list of tasks is simply the list of integers $0,1,\dots,n$,
-computed as follows:
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in 
-    for i = 0 to n do l := (i, ()) :: !l done; 
-    !l
-\end{ocaml}
-Each task is actually a pair, where the second component contains
-some information which is kept local to the master process.
-It is irrelevant here and thus the value \of{()} of type \of{unit} is used.
-The current sum is kept into a global reference \of{s}:
-\begin{ocaml}
-  let s = ref 0
-\end{ocaml}
-The master function receives a task \of{(i,())} together with its
-result \of{fi} and simply adds \of{fi} to \of{s}. There is no new task
-generated, hence the returned empty list.
-\begin{ocaml}
-  let master (i,()) fi = s := !s + fi; []
-\end{ocaml}
-Finally, the whole computation takes place when we call the
-\of{compute} function:
-\begin{ocaml}
-  let () = compute ~worker ~master tasks
+1a.  open Network
+1b.  let () = declare_workers ~n:4 "m1"
+1c.  let () = declare_workers ~n:8 "m2"
 \end{ocaml}
 
-Now, let us assume we want to make use of a 4-core machine to compute
-$s(n)$. It is as easy as just replacing the very first line with
+We could simplify the code even further using the derived API, for
+example by replacing lines 5-7 with
 \begin{ocaml}
-  open Cores
-  let () = set_number_of_cores 4
-\end{ocaml}
-while the rest of the code remains unchanged.
-
-Later, we may want to use a network of machines instead, for example
-two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
-cores respectively. Similar to the above, we just replace the first
-lines with
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Same
-\end{ocaml}
-Here, \of{Same} is a module which is to be used when master and worker
-are running the same executable. It provides a \of{compute} function
-which has the same signature as in modules \of{Sequential} and
-\of{Cores}, so that the rest of the code remains unchanged. Master and
-worker processes are distinguished at run-time using an environment
-variable \texttt{WORKER} which is set/unset.
-
-If we need to write two different programs for master and worker, for
-reason of binary incompatibility or any other reason, the library API
-is providing functions to do that. If master and worker are still
-compiled with the same version of \Ocaml, we use the \of{Poly} module
-which provides a polymorphic API. Let us start with the worker program.
-It is as simple as
-\begin{ocaml}
-  open Poly
-  let worker i = ...
-  let () = Worker.compute worker ()
-\end{ocaml}
-The \of{Worker.compute} function enters a loop which waits for
-tasks sent by the master and returns results computed using \of{worker}.
-The master program is almost the same as before. First, we replace
-module \of{Same} with module \of{Poly}:
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Poly
-\end{ocaml}
-Tasks and \of{master} function are unchanged:
-\begin{ocaml}
-  let tasks = ...
-  let s = ref 0
-  let master (i,()) fi = ...
-\end{ocaml}
-Finally, we start the computation with \of{Master.compute}, which does
-not have a \of{worker} parameter anymore:
-\begin{ocaml}
-  let () = Master.compute ~master tasks
-\end{ocaml}
-
-When master and worker programs are compiled with different
-versions of the \Ocaml\ compiler, our library still provides
-a monomorphic API over strings. As a consequence, we need to convert
-integers to and from strings in both master and worker.
-The modified worker program is as follows:
-\begin{ocaml}
-  open Mono
-  let worker i = ...
-  let worker_string i = string_of_int (worker (int_of_string i))
-  let () = Worker.compute worker_string ()
-\end{ocaml}
-The master program is modified in a similar way.
-We simply replace \of{Poly} with \of{Mono} and encode/decode integers
-as strings, as follows:
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in 
-    for i = 0 to n do l := (string_of_int i, ()) :: !l done; 
-    !l
-  let master (i,()) fi = s := !s + int_of_string fi; []
+5.  let s = map_local_fold ~map:f ~fold:(+) 0 tasks
 \end{ocaml}

### 2010-07-02 doc/icfp-2010/main.tex
submission to ML workshop

@@
 
-\titlebanner{submitted to ICFP 2010}        % These are ignored unless
-\preprintfooter{Experience Report: Distributed Computing for the Common Man}   % 'preprint' option specified.
+\titlebanner{Technical Report}        % These are ignored unless
+\preprintfooter{Functory: Distributed Computing for the Common Man}   % 'preprint' option specified.
 
-\title{Experience Report: Distributed Computing for the Common Man}
+\title{Functory: Distributed Computing for the Common Man}
 %\subtitle{Subtitle Text, if any}

### 2010-07-02 doc/ml-workshop-2010/main.tex
submission to ML workshop

@@
 \titlebanner{submitted to Workshop on ML 2010}        % These are ignored unless
-\preprintfooter{Distributed Computing for the Common Man}   % 'preprint' option specified.
+\preprintfooter{Functory: Distributed Computing for the Common Man}   % 'preprint' option specified.
 
-\title{Distributed Computing for the Common Man}
+\title{Functory: Distributed Computing for the Common Man}
 %\subtitle{Subtitle Text, if any}
@@
 large number of logical formulas using a variety of automated theorem
-provers~\cite{filliatre07cav}. Our computing infrastructure consists
+provers. Our computing infrastructure consists
 of a few powerful multi-core machines (typically 8 to 16 cores) and
@@
   it relieves the user from implementing task distribution routines
-  and reliable fault-tolerance mechanisms.
+  and fault-tolerance mechanisms.
 \end{itemize}
@@
 \end{center}
+As evident from the last row, the speedup almost reaches 8, which is
+also the number of cores we use.
+
 % From the table above, it is clear that the \of{Cores} and \of{Network}
@@
 
+\section{Conclusion}
+
+We presented \functory, a distributed programming environment for \Ocaml.
+The main features are the genericity of the interface,
+which makes use of polymorphic higher-order functions, and the ability
+to easily switch between sequential, multi-core, and network
+implementations. A distinguishing feature of our library is a
+robust fault-tolerance mechanism which relieves
+the user of cumbersome implementation details.
+
+The most relevant application domains for this library are the ones
+where computation time is much greater than communication time, which
+is the case in many real-world applications. This results in a speedup
+ratio which is close to the total number of cores.
+
 \end{document}

### 2010-10-04 doc/report/main.tex
*** empty log message ***

--- 2010-07-02 doc/icfp-2010/main.tex
@@
 
-\documentclass[preprint]{sigplanconf}
+\documentclass[a4paper,12pt]{article}
 
@@
 
-\usepackage{amsmath,graphicx,color,url}
+\usepackage{fullpage,comment,amsmath,graphicx,color,url}
 
@@
 
-\conferenceinfo{ICFP '10}{September 27-29, 2010, Baltimore, Maryland.} 
-\copyrightyear{2010} 
-\copyrightdata{[to be supplied]}
-
-\titlebanner{Technical Report}        % These are ignored unless
-\preprintfooter{Functory: Distributed Computing for the Common Man}   % 'preprint' option specified.
-
-\title{Functory: Distributed Computing for the Common Man}
+\title{Functory: Distributed Computing for the Common Man\thanks{This
+    research was partly supported by the French national project U3CAT
+    (\emph{Unification of Critical C Code Analysis Techniques},
+    ANR-08-SEGI-021).}} 
 %\subtitle{Subtitle Text, if any}
 
-\authorinfo{Jean-Christophe Filli\^{a}tre}
-           {CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
-            INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
-           {filliatr@lri.fr}
-\authorinfo{K. Kalyanasundaram}
-           {ProVal, INRIA Saclay - \^{I}le-de-France, Orsay, F-91893}
-           {kalyan.krishnamani@inria.fr}
+\author{Jean-Christophe Filli\^{a}tre \quad K. Kalyanasundaram \\
+  CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
+  INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893 \\
+   \url{filliatr@lri.fr, kalyan.krishnamani@inria.fr}}
 
@@
 \end{abstract}
-
-\category{D1.1}{Programming techniques}{Applicative (Functional) Programming}
-
-\terms
-Distributed Computing, Functional Programming
 
@@
 describe the main idea behind the generic API, and introduce
-\of{master} and \of{worker} programs. Section~\ref{sec:example}
-illustrates the API usage with an example.
+\of{master} and \of{worker} programs. 
+% Section~\ref{sec:example} illustrates the API usage with an example.
 In Section~\ref{sec:derived}, we
@@
 
+\begin{comment}
 \section{Illustrative Example}\label{sec:example}
@@
 \end{ocaml}
+\end{comment}
 
@@
 
+\section{Illustrative Case Studies}
+
+\subsection{Matrix Multiplication}
+
+As a first example, let us consider matrix multiplication.
+We assume two matrices \of{a} and \of{b}, respectively of size
+$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
+a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
+Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
+the standard matrix multiplication would be as follows:
+\begin{ocaml}
+  for i = 0 to n-1 do
+    for j = 0 to p-1 do
+      for k = 0 to m-1 do
+	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
+      done
+    done
+  done  
+\end{ocaml}
+and runs in $O(\of{n}\times\of{p}\times\of{m})$, assuming constant
+time addition and multiplication (denoted $+$ and $\times$).
+An easy way to distribute this computation is obviously to turn each
+computation of the inner loop over \of{k} into a task.
+To do so, we first prepare the list of tasks:
+\begin{ocaml}
+  let tasks = 
+    let l = ref [] in
+    for i = 0 to n-1 do for j = 0 to p-1 do
+      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
+    done done;
+    !l
+\end{ocaml}
+Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
+tuple, together with position \of{(i,j)} as a second tuple.
+The worker is receiving the first tuple as argument and computes
+the dot product:
+\begin{ocaml}
+  let worker (ai, bj) =
+    let c = ref 0 in
+    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
+    !c
+\end{ocaml}
+The master is a one line function which receives the result \of{r}
+from the worker and simply stores it into \of{c}, according to the
+position contained in the input task. It produces no new task.
+\begin{ocaml}
+  let master (_, (i,j)) r = c.(i).(j) <- r; []
+\end{ocaml}
+Finally, the overall computation is started by invoking \of{compute}
+as follows:
+\begin{ocaml}
+  let () = compute ~worker ~master !tasks
+\end{ocaml}
+The total number of tasks is $\of{n}\times\of{m}$.
+
+Using the sequential implementation provided by \of{Functory} is as
+simple as including the following line in the code:
+\begin{ocaml}
+  open Sequential
+\end{ocaml}
+Then the computation is roughly similar to the standard multiplication
+described above.
+
+Now, let us assume we want to make use of a 4-core machine to perform
+the same computation. 
+This is achieved by just replacing the line above by
+\begin{ocaml}
+  open Cores
+  let () = set_number_of_cores 4
+\end{ocaml}
+while the rest of the code remains unchanged.
+
+Later, we may want to use a network of machines instead, for example
+two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
+cores respectively. Similar to the above, we turn the above lines into
+\begin{ocaml}
+  open Network
+  let () = declare_workers ~n:4 "orcus"
+  let () = declare_workers ~n:8 "belzebuth"
+  open Same
+\end{ocaml}
+Here, \of{Same} is a module which is to be used when master and worker
+are running the same executable. It provides a \of{compute} function
+which has the same signature as in modules \of{Sequential} and
+\of{Cores}, so that the rest of the code remains unchanged. Master and
+worker processes are distinguished at run-time using an environment
+variable \texttt{WORKER} which is set/unset.
+
+If we need to write two different programs for master and worker, for
+reason of binary incompatibility or any other reason, the library API
+is providing functions to do that. If master and worker are still
+compiled with the same version of \Ocaml, we use the \of{Poly} module
+which provides a polymorphic API. Let us start with the worker program.
+It now looks like:
+\begin{ocaml}
+  open Poly
+  let worker (ai, bj) = ...
+  let () = Worker.compute worker ()
+\end{ocaml}
+The \of{Worker.compute} function enters a loop which waits for
+tasks sent by the master and returns results computed using \of{worker}.
+The master program is almost the same as before. First, we replace
+module \of{Same} with module \of{Poly}:
+\begin{ocaml}
+  open Network
+  let () = declare_workers ~n:4 "orcus"
+  let () = declare_workers ~n:8 "belzebuth"
+  open Poly
+\end{ocaml}
+Tasks and \of{master} function are unchanged:
+\begin{ocaml}
+  let tasks = ...
+  let master (_, (i,j)) r = ...
+\end{ocaml}
+Finally, we start the computation with \of{Master.compute}, which does
+not have a \of{worker} parameter anymore:
+\begin{ocaml}
+  let () = Master.compute ~master tasks
+\end{ocaml}
+
+When master and worker programs are compiled with different
+versions of the \Ocaml\ compiler, our library still provides
+a monomorphic API over strings. As a consequence, we need to convert
+tasks and results to and from strings in both master and worker.
+The modified worker program then looks as follows:
+\begin{ocaml}
+  open Mono
+  let worker (ai, bj) = ...
+  let worker_string t = string_of_int (worker (task_of_string i))
+  let () = Worker.compute worker_string ()
+\end{ocaml}
+The master program is modified in a similar way.
+We simply replace \of{Poly} with \of{Mono} and encode/decode integers
+as strings, as follows:
+\begin{ocaml}
+  let tasks = ... string_of_task ...
+  let master (_, (i,j)) r = c.(i).(j) <- int_of_string r; []
+\end{ocaml}
+where \of{string_of_task} and \of{task_of_string} are user-defined
+functions to convert tasks to and from strings.
+
+\subsection{N-queens}\label{sec:n-queens}
+
+The first example is the classical $N$-queens problem, where we
+compute the total number of ways to place $N$ queens on a $N\times N$
+chessboard in such a way no two queens attack each other.
+We use a standard backtracking algorithm for this problem, which
+places the queens one by one starting from the first row.
+Distributing the computation is thus quite easy: we consider all
+possible ways to place queens on the first $D$ rows and then perform
+the subsequent search in parallel. Choosing $D=1$ will result in
+exactly $N$ tasks. More subtly, $D=2$ will result in $N^2-3N+2$ tasks.
+% TODO: make a picture to illustrate positioning of 2 queens on the
+% first 2 rows
+
+Each tasks only consists of three integers and results in one integer,
+which is the total number of solutions for this task.
+We make use of function \of{map_local_fold} from the derived API,
+where \of{map} is performing the search and \of{fold} simply adds the
+intermediate results.
+
+\subsection{Mandelbrot Set}
+
+\subsection{Polynomial Multiplication}
+
 \section{Implementation Details}\label{sec:implem}
@@
 The \of{Cores} module implements the distributed computing library for
-multiple cores on the same machine. As illustrated in
-Section~\ref{sec:example}, 
-it provides a function
+multiple cores on the same machine. 
+% As illustrated in Section~\ref{sec:example}, 
+It provides a function
 \of{set_number_of_cores: int -> unit} to indicate the number of cores
@@
 for a network of machines. 
-As illustrated in Section~\ref{sec:example}, it
-provides a function
+% As illustrated in Section~\ref{sec:example}, 
+It provides a function
 \of{declare_workers: n:int -> string -> unit} to fill a table of
@@
 
-The first example is the classical $N$-queens problem, where we
-compute the total number of ways to place $N$ queens on a $N\times N$
-chessboard in such a way no two queens attack each other.
 The following table shows execution times for various values of $N$
@@
 
-The first column shows the value of $N$.  Each computation is divided
-into a certain number of tasks, by placing some of the queens before
-starting the distributed computation.  The number of tasks is shown in
+The first column shows the value of $N$.  The number of tasks is shown in
 second column.  Then the last three columns show execution times in
@@
 
-\acks
+\paragraph{Acknowledgements.}
 The authors are grateful to the ProVal team for support and comments

### 2010-10-06 doc/report/main.tex
*** empty log message ***

@@
 the subsequent search in parallel. Choosing $D=1$ will result in
-exactly $N$ tasks. More subtly, $D=2$ will result in $N^2-3N+2$ tasks.
+exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
+so on.
 % TODO: make a picture to illustrate positioning of 2 queens on the
@@
 
-Each tasks only consists of three integers and results in one integer,
+Each task only consists of three integers and its result is one integer,
 which is the total number of solutions for this task.
@@
 \subsection{Mandelbrot Set}
+
+Drawing the Mandelbrot set is another classical example that could be
+distributed easily, since the color of each point can be computed
+independently of the others. 
+Let us assume we are given the coordinates of the region to be drawn,
+along with the width \of{w} and height \of{h} of the resulting image
+(in pixels). 
+Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
+parameter. It is immediate to split the image into \of{t} sub-images,
+each of which is computed in parallel with and independently of the
+others. For instance, the image could be split into horizontal slices
+or, more generally, into rectangular blocks.
+
+Each task is thus four floating-point numbers denoting the region
+coordinates, together with two integers denoting the dimensions of the
+sub-image to be drawn. The result of the task is a matrix of pixels,
+of size $(\of{w}\times\of{h}) / \of{t}$. 
+For instance, drawing a $800\times 600$ image using 20 tasks will
+result in 20 sub-images of size $176,000$ bytes each,
+assuming each pixel is encoded in four bytes.
 
@@
 computationally extensive examples.
+
+\subsubsection{Matrix Multiplication}
+
+$\of{n} = \of{p} = \of{m} = 100$
+
+coefficients have several thousands of digits; we use GMP
+
+%            parall  line
+
+% # tasks    10,000   100
+
+% sequential  20.3   20.2
+% cores 2     22.7   11.3
+% cores 4     12.3    6.1
+% cores 6      8.6    4.3
+% cores 8      8.0    3.5
+
+% network is localhost(moloch) + belzebuth
+
+% network 3+3        12.7
+% network 6+  112     7.2
+% network  +6   ?    36.4
+% network 6+6  76.1  11.9
+
+\begin{center}
+  \begin{tabular}{|l|l|l|}
+    \hline
+    sequential & cores & network \\\hline
+    20.3 & 
+  \end{tabular}
+\end{center}
 
@@
 
+\subsection{Mandelbrot Set}
+
 \subsection{SMT Solvers}

### 2010-10-07 doc/report/main.tex
*** empty log message ***

@@
 
+% TODO: explain the issue with closures referring to global variables here
+
 \item \textbf{Master and workers are different programs, compiled with
@@
 \end{ocaml}
+
 \item \textbf{Master and workers are different programs, not even
@@
 
+In this section, we show how to use \functory\ on several case studies.
+We focus here on the use of the API. Experimental results of these
+case studies are given in Section~\ref{sec:experiments}.
+
 \subsection{Matrix Multiplication}
@@
 intermediate results.
+In the network configuration, we make use of the \of{Network.Same}
+module, workers and master being the same executable.
 
@@
 
-\subsection{Polynomial Multiplication}
+In the network configuration, we deliberately choose to have two
+different programs for master and workers, using the \of{Network.Poly}
+module (actually, a single source code with a command line option).
+This has to be done since the worker code makes use of global
+variables, which prevents us from using
+\of{Network.Same}, as described in Section~\ref{sec:API}. 
+Alternately, we could have refactored the worker
+code to avoid global variables.
+
+% \subsection{Polynomial Multiplication}
+% TODO?
+
+\subsection{SMT Solvers}
 
@@
     \hline
-    N & \#tasks  & sequential& cores                 & network 
+    N & \#tasks  & \of{Sequential}& \of{Cores}                 & \of{Network} 
     \\\hline\hline
@@
 \subsection{Mandelbrot Set}
+
+This benchmark consists in drawing the fragment of the Mandelbrot set
+with lower left corner $(-1.1, 0.2)$ and upper right corner $(-0.8,
+0.4)$, as a $9,000\times6,000$ image. The sequential computation of
+this image consumes 29.4 seconds. For \of{Cores} and \of{Network}
+implementations, the computation times in seconds are tabulated below.
+
+% TODO compute ratios for Network
+\begin{center}
+  \begin{tabular}{|r|r|r|r|r|}
+    \hline
+    \#cores  &\#tasks & \of{Cores} & \of{Network} \\
+    \hline
+    2       & 10 & 15.8       (1.86) &  20.3       \\
+            & 30 & 15.7       (\underline{1.87}) &  18.7       \\
+            & 100 & 16.1      (1.83) &  19.8       \\
+            & 1000 & 19.6     (1.50) &  38.6      \\
+    \hline
+    4       & 10 & 9.50       (3.09)  &  14.4       \\
+            & 30 & 8.26       (\underline{3.56})  &  11.4   \\
+            & 100 & 8.37      (3.51)  &  11.4   \\
+            & 1000 & 10.6     (2.77)  &  20.5    \\
+    \hline
+    8       & 10 & 9.40       (3.13)  &  12.6      \\
+            & 30 & 4.24       (\underline{6.93})  &   7.6      \\
+            & 100 & 4.38      (6.71)  &   7.5      \\
+            & 1000 & 6.86     (4.29)  &  11.3      \\
+    \hline
+  \end{tabular}
+\end{center}
+The best timings are achieved for the \of{Cores} configuration, where
+communications happen within the same machine and are thus cheaper.
+There are two significant differences with respect to the n-queens
+benchmark.  On one hand, the number of tasks can be controlled more
+easily than in the case of n-queens. We experimentally figured out the
+optimal number of tasks to be 30. One the other hand, each computation
+result is an image, rather than just an integer as in the case of
+n-queens. Consequently, communication costs are much greater. 
+In this particular experiment, the total size of the results
+transmitted is more than 200 Mb.
 

### 2010-10-07 doc/report/main.tex
*** empty log message ***

@@
 
-\section{Illustrative Case Studies}
+\section{Illustrative Case Studies}\label{sec:studies}
 
@@
 case studies are given in Section~\ref{sec:experiments}.
-
-\subsection{Matrix Multiplication}
+The source code for all these case studies is contained in the
+distribution, in sub-directory \texttt{tests}.
+
+\subsection{Matrix Multiplication}\label{sec:matrix}
 
@@
 \end{ocaml}
-and runs in $O(\of{n}\times\of{p}\times\of{m})$, assuming constant
-time addition and multiplication (denoted $+$ and $\times$).
+and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
+assuming addition and multiplication over coefficients to be 
+denoted $+$ and $\times$ respectively.
 An easy way to distribute this computation is obviously to turn each
@@
 The master program is modified in a similar way.
-We simply replace \of{Poly} with \of{Mono} and encode/decode integers
+We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
 as strings, as follows:
@@
   let tasks = ... string_of_task ...
-  let master (_, (i,j)) r = c.(i).(j) <- int_of_string r; []
+  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
 \end{ocaml}
@@
 
-\subsection{SMT Solvers}
+\subsection{SMT Solvers}\label{sec:SMT}
+
+Here we demonstrate the potential of our library for our application
+needs as mentioned in the introduction. In this case study,
+we consider 80 challenging verification conditions
+(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
+VC is stored in a file, which is accessible over
+NFS. The purpose of the experiment is to check the validity of each VC
+using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
+which are all installed on each of the machines in use.
+
+The master program proceeds by reading the file names, turning them
+into tasks by multiplying them by the number of provers (more than
+300 tasks in total).
+Each worker in turn invokes the given prover on the given file, within
+a timeout limit.
+Each task completes with one of the four possible outcomes: \emph{valid},
+\emph{unknown} (depending on
+whether the VC is valid or undecided by the prover), 
+\emph{timeout} and \emph{failure}.
+
+The result of each computation is a pair denoting the status and the
+time spent in the prover call. The master collects these results and
+sums up the timings for each prover and each possible status.
 
@@
 from \url{http://www.lri.fr/~filliatr/functory/}.
-In this section, we show the potential of this library using two 
-computationally extensive examples.
-
-\subsubsection{Matrix Multiplication}
-
-$\of{n} = \of{p} = \of{m} = 100$
-
-coefficients have several thousands of digits; we use GMP
-
-%            parall  line
-
-% # tasks    10,000   100
-
-% sequential  20.3   20.2
-% cores 2     22.7   11.3
-% cores 4     12.3    6.1
-% cores 6      8.6    4.3
-% cores 8      8.0    3.5
-
-% network is localhost(moloch) + belzebuth
-
-% network 3+3        12.7
-% network 6+  112     7.2
-% network  +6   ?    36.4
-% network 6+6  76.1  11.9
-
-\begin{center}
-  \begin{tabular}{|l|l|l|}
-    \hline
-    sequential & cores & network \\\hline
-    20.3 & 
-  \end{tabular}
-\end{center}
+In this section, we show the experimental results pertaining to the
+case studies of Section~\ref{sec:studies}.
 
@@
     \#cores  &\#tasks & \of{Cores} & \of{Network} \\
-    \hline
+    \hline\hline
     2       & 10 & 15.8       (1.86) &  20.3       \\
@@
 
+\subsection{Matrix Multiplication}
+
+This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
+It consists of multiplication of
+two square matrices of dimension 100, that is $\of{n} = \of{p} =
+\of{m} = 100$, with integer coefficients.
+Coefficients have several thousands of digits, hence we use
+GMP~\cite{GMP} to handle operations over coefficients.
+
+We compare the performances of two different implementations. In the
+first one, called \of{mm1}, each task consists of the computation of a single
+coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
+In the second one, called \of{mm2}, each task consists of the
+computation of a whole row of the resultant matrix.
+As a consequence, the total number of tasks in
+$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
+The experimental results (in seconds) are tabulated below.
+\begin{center}
+  \begin{tabular}{|r|r|r|}
+    \hline
+    & \of{mm1}       & \of{mm2}  \\
+    & (10,000 tasks) & (100 tasks) \\
+    \hline\hline
+\of{Sequential} & 20.3 &  20.2 \\
+\hline
+ \of{Cores} 
+ (2 cores)     &   22.7  &  11.3\\
+ (4 cores)     &   12.3  &   6.1\\
+ (6 cores)     &    8.6  &   4.3\\
+ (8 cores)     &    8.0  &   3.5\\
+ \hline
+  \end{tabular}
+\end{center}
+We do not include results for the network configuration, as they do
+not achieve any benefit with respect to the sequential
+implementation. The reason is that the communication cost dominates the
+computation cost in such a way that the total execution time is
+always greater than 30 seconds. Indeed, irrespective of the
+implementation (\of{mm1} or \of{mm2}), the total size of the
+transmitted data is 
+$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
+billions of bytes.
+
+A less naive implementation would have the worker read the input matrices
+only once, \emph{e.g.} from a file, and then have the master send only
+row and column indices. This would reduce the communication cost to
+$O(\of{n}\times\of{m})$ only.
+
+%             mm1     mm2
+
+% # tasks    10,000   100
+
+% sequential  20.3   20.2
+
+% cores 2     22.7   11.3
+% cores 4     12.3    6.1
+% cores 6      8.6    4.3
+% cores 8      8.0    3.5
+
+% network = workers on moloch + master on belzebuth
+
+% cores 2     93     32.3
+% cores 4     87     33.7
+% cores 6     92     33.5
+% cores 8     88     32.5
+
 \subsection{SMT Solvers}
 
-Here we demonstrate the potential of our library for our application
-needs as mentioned in the introduction. Our computing
-infrastructure consists of 3 machines with 4, 8 and 8 cores
-respectively. The benchmark we use for
-experimentation consists of around 80 challenging verification conditions
-(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
-VC is stored in a file, which is accessible over
-NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
-which are all installed on each of the machines in use.
-
-The master program proceeds by reading the file names, turning them
-into tasks by multiplying them by the number of provers (more than
-300 tasks in total).
-Each worker in turn invokes the given prover on the given file.
-We set a timeout limit of 1 minute.
-Each task completes with one of the four possible outcomes: \emph{valid},
-\emph{unknown} (depending on
-whether the VC is valid or undecided by the prover), 
-\emph{timeout} and \emph{failure}.
+As explained in Section~\ref{sec:SMT}, this benchmark consists of 80
+verification conditions, each being checked by 4 different SMT
+solvers. Each task is executed with a timeout limit of 1 minute.
+Our computing
+infrastructure for this experiment consists of 3 machines with 4, 8 and 8 cores
+respectively. 
 The figure below shows the total time in minutes spent by each prover
-for each possible outcomes.
+for each possible outcome.
 \begin{center}
@@
 %\vfill\pagebreak
-% We recommend abbrvnat bibliography style.
-\bibliographystyle{abbrvnat}
+\bibliographystyle{plain}
 \bibliography{./biblio}

### 2010-10-11 doc/report/main.tex
*** empty log message ***

@@
 variables, which prevents us from using
-\of{Network.Same}, as described in Section~\ref{sec:API}. 
+\of{Network.Same} (as explained below in Section~\ref{sec:implem}). 
 Alternately, we could have refactored the worker
@@
 operating system.
+
+\subsubsection{Network Sub-modules} % TODO: change title
+
+As mentioned in Section~\ref{sec:API}, the \of{Network} module
+actually provides three different implementations, according to three
+different execution scenarios. There are provided in three
+sub-modules, as described below.
+
+\paragraph{\of{Same}.} This module is used when master and workers are
+running the same executable. The master and workers have to be
+differentiated in some manner. We use an environment variable
+\texttt{WORKER} for this purpose. When set, it indicates that the
+executable acts as a worker. At run-time, a worker immediately 
+enters a loop waiting for tasks from the master, without even getting
+into the user code. 
+As explained in Section~\ref{sec:API}, 
+the master function has the following signature.
+\begin{ocaml}
+ val compute : 
+      worker:('a -> 'b) -> 
+      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit  
+\end{ocaml}
+The master uses marshaling to send both a closure of type 
+\of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting strings
+are passed as argument \of{f} and \of{x} in message
+\of{Assign}. Similarly, the worker uses marshaling to send back the
+result of the computation of type \of{'b}, which is the argument
+\of{s} in message \of{Completed}.
+
+Though the ability to run the same executable helps a lot in deploying
+the program in different machines, it comes at a small price. Since
+the worker is not getting into the user code, closures which are
+transmitted from the master can not refer to global variables in the
+user code. Indeed, the initialization code for these global variables
+is never reached on the worker side. For instance, the Mandelbrot set
+example could be written as follows:
+\begin{ocaml}
+  let max_iterations = 200
+  let worker (xmi, xma, ymi, yma, w, h) =
+    ... draw sub-image using max_iterations ...
+\end{ocaml}
+That is, the global function \of{worker} makes use of the global
+variable \of{max_iterations}. The worker gets the function to compute
+from the master, namely the closure corresponding to function
+\of{worker} in that case, but on the worker side the initialization of
+\of{max_iterations} is never executed.
+
+One obvious solution is  not to use global variables in the
+worker code. This is not always possible, though. 
+To overcome this, the \of{Same} sub-module also provides a
+\of{Worker.compute} function to start the worker loop manually from
+the user code. This way, it can be started at any point, in particular
+after the initialization of the required global variables.
+Master and worker are still running the same executable, but are 
+distinguished using a user-defined way (command-line argument,
+environment variable, etc.).
+
+\bigskip
+There are situations where it is not possible
+to run the same executable for master and workers.
+For instance, architectures or operating systems could be different
+across the network.
+For that reason, the \of{Network} module provides two other
+implementations.
+
+\paragraph{\of{Poly}.} When master and workers are compiled with the
+same version of \Ocaml, we can no longer marshal closures but we can
+still marshal polymorphic values. Indeed, an interesting property of
+marshaling in \Ocaml\ is to be fully architecture-independent, as long
+as a single version of \Ocaml\ is used. It is worth pointing out that
+absence of marshaled closures
+now enables the use of two different programs for master and workers. This
+is not mandatory, though, since master and workers could still be
+distinguished at run-time as in the previous case.
+
+On the worker side, the main loop is started manually using
+\of{Worker.compute}. The computation to be performed on each task is
+given as an argument to this function. It thus looks as follows:
+\begin{ocaml}
+  Worker.compute : ('a -> 'b) -> unit -> unit
+\end{ocaml}
+On the master side, the \of{compute} function is simpler than in the
+previous case, as it has one argument less, and thus has the following
+signature. 
+\begin{ocaml}
+  Master.compute : master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit 
+\end{ocaml}
+For realistic applications, where master and workers are completely
+different programs, possibly written by different teams, this is the
+module of choice in our library, since it can still pass polymorphic
+values over the network. The issues of marshaling are automatically
+taken care of by \Ocaml\ run-time.
+
+% TODO : explain the new derived API
+
+\paragraph{\of{Mono}.}
+
+% TODO
+
+% \bigskip
+
+% There is
+% actually a single, generic implementation, which is then instantiated
+% in three different ways, corresponding to the three scenarios.  The
+% generic implementation assumes that only strings are being transmitted
+% over the network, as evident from the \of{Assign} and \of{Completed}
+% messages in Section~\ref{sec:protocol}. Depending on the
+% implementation, these strings could be marshaled closures, marshaled
+% values or simply user strings.
 
@@
 it was already assigned to another worker.
-
-\subsubsection{Three Network Implementations} % TODO: change title
-
-As mentioned in Section~\ref{sec:API}, the \of{Network} module
-actually provides three different implementations, according to three
-different execution scenarios (namely, \of{Same}, \of{Poly} and
-\of{Mono}). There is
-actually a single, generic implementation, which is then instantiated
-in three different ways, corresponding to the three scenarios.  The
-generic implementation assumes that only strings are being transmitted
-over the network, as evident from the \of{Assign} and \of{Completed}
-messages in Section~\ref{sec:protocol}. Depending on the
-implementation, these strings could be marshaled closures, marshaled
-values or simply user strings.
 

### 2010-10-11 doc/report/main.tex
*** empty log message ***

@@
 \end{ocaml}
-  Note that the second component of each task is still polymorphic (of
-  type \ocaml{'c} here), since it is local to the master.
 \end{enumerate}
@@
 the worker is not getting into the user code, closures which are
-transmitted from the master can not refer to global variables in the
+transmitted from the master cannot refer to global variables in the
 user code. Indeed, the initialization code for these global variables
@@
 
-% TODO : explain the new derived API
-
-\paragraph{\of{Mono}.}
-
-% TODO
-
-% \bigskip
+The derived API presented in Section~\ref{sec:derived} is adapted to
+deal with the absence of closures. Exactly as the \of{compute}
+function, each API now takes two forms, one for the master and another
+for the workers. For example, \of{map_fold_ac} takes the following
+forms.
+\begin{ocaml}
+  Worker.map_fold_ac : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit -> unit
+  Master.map_fold_ac : 'b -> 'a list -> 'b
+\end{ocaml}
+It is the responsibility of the user to ensure type consistency between
+master and workers.
+
+\paragraph{\of{Mono}.} When master and workers are compiled using
+different versions of \Ocaml, we can no longer use marshaling. 
+As in the previous case, we split \of{compute} into two functions, one
+for master and one for workers. In addition, values transmitted over
+the network can only be strings. The signature thus takes the
+following form.
+\begin{ocaml}
+  Worker.compute : (string -> string) -> unit -> unit
+  Master.compute : master:(string * 'c -> string -> (string * 'c) list) -> 
+                  (string * 'c) list -> unit
+\end{ocaml}
+Any other datatype for tasks should be encoded to/from strings. This
+conversion is left to the user.
+Note that the second component of each task is still polymorphic (of
+type \of{'c} here), since it is local to the master.
 

### 2010-10-11 doc/report/main.tex
*** empty log message ***

@@
 
-\title{Functory: Distributed Computing for the Common Man\thanks{This
+\title{Functory: Distributed Computing \\ for the Common Man\thanks{This
     research was partly supported by the French national project U3CAT
@@
 \begin{abstract}
-  We present a distributed computing library for
+  We present Functory, a distributed computing library for
   functional programming languages. The main features of this library
-  include (1) a generic API, (2) several implementations to
+  include (1) a polymorphic API, (2) several implementations to
   adapt to different computing environments such as sequential,
@@
 
-%\keywords
-
 \section{Introduction}
 
-This paper introduces a generic library for distributed computing. This
-work was initially motivated by the computing needs that exist in our
+This paper introduces a generic library for distributed computing,
+targeted at functional programming. 
+This work was initially motivated by the computing needs that exist in our
 own research team. Our applications include large-scale deductive
@@
 language.  Hence we designed and implemented such a
-library\footnote{This library is implemented in \Ocaml, but the
+library, which is the subject of this paper.
+This library is implemented in \Ocaml, but the
 implementation should be straightforward in any functional
-  programming language.}, which is the subject of this paper.
+programming language.
 
@@
 \end{itemize}
+The application domain of such a distributed computing library is manyfold. 
 It is worth noting that the library is not targeted at applications
-running on server farms, crunching enormous amounts of data. Neither
-is it limited to research labs possessing few powerful
-machines. It serves a variety of users and a wide
+running on server farms, crunching enormous amounts of data. 
+It is rather more relevant for research labs, to make use of available
+computing resources efficiently. It serves a variety of users and a wide
 spectrum of needs, from desktop PCs to networks of machines, and hence
 the title.
-
-The application domain of such a distributed computing library is
-manyfold.  The most striking application, as also relevant to our
-research endeavours, is in validating thousands of verification
-conditions using automated theorem provers. Having noted this, it
-should be apparent that the library finds use in efficient execution
+The most striking application would involve efficient execution
 of any large-scale expensive computation that is parallelizable.
+This is the case in our
+research endeavours, that is validating thousands of verification
+conditions using automated theorem provers.
+
 % what it is and what it is not
@@
 % \end{itemize}
+
+\medskip
+
+The remainder of this section introduces our approach to distributed
+computing in a functional programming setting.
+
 \paragraph{Distributed Computing.}
-A typical distributed computing library provides the following (we
+A typical distributed computing library, as \functory,
+provides the following (we
 borrow some terminology from Google's MapReduce~\cite{mapreduce}):
@@
 \medskip
-The paper is organized as follows. In Section~\ref{sec:API}, we
+The paper is organized as follows. 
+In Section~\ref{sec:API}, we
 describe the main idea behind the generic API, and introduce
 \of{master} and \of{worker} programs. 
-% Section~\ref{sec:example} illustrates the API usage with an example.
 In Section~\ref{sec:derived}, we
-derive some interesting functions from the generic
-API. Section~\ref{sec:implem} delves into the implementation details
+present some additional functions provided by \functory,
+which are derived from the generic
+API. 
+Section~\ref{sec:studies} illustrates the API usage with several
+case studies.
+Section~\ref{sec:implem} delves into the implementation details
 of our library, while Section~\ref{sec:experiments} illustrates the
@@
 generic API which we shall extend with several specialized
-functions in later sections. The main function in our API follows the idea
+functions in the next section. The main function in our API follows the idea
 sketched in the introduction. It has the following signature:
@@
 in some internal data structure.
+Additionally, it may produce new tasks, as a list of type 
+\of{('a * 'c) list}, which are then appended to the current set of
+pending tasks.
 The next section will describe the usage of this generic interface to
@@
 depending on the execution context. 
-The five possible contexts are the following: 
+The first two contexts are the following: 
 \begin{enumerate}
 \item \textbf{Purely sequential execution:}
-  this is mostly intended for debugging;
+  this is mostly intended to be a reference implementation
+  for performance comparisons, as well as for debugging;
 
 \item \textbf{Several cores on the same machine:} 
-  this implementation makes use of \unix\ processes and 
-  provides a polymorphic \ocaml{compute} function;
-
+  this implementation is intended to distribute the computation over a
+  single machine and it makes use of \unix\ processes;
+\end{enumerate}
+The next three implementations are intended for distributing the
+computation over a network of machines.
+\begin{enumerate}
+\setcounter{enumi}{2}
 \item \textbf{Same executable run on master and worker machines:}
   this implementation makes use of the ability to marshal \Ocaml\
-  closures and  provides a polymorphic \ocaml{compute}
-  function.
+  closures and polymorphic values.
   Depending on whether the program is run as a master or as a worker,
   the relevant arguments of \ocaml{compute} are used.
-
-% TODO: explain the issue with closures referring to global variables here
 
@@
 corresponding to contexts 3, 4 and 5 above. 
-The next section demonstrates the use of these modules.
+The next section presents more functions provided by \functory, which
+are all derived from the generic API just presented.
 
@@
 \begin{ocaml}
- val map_fold :
-   map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-   'c -> 'a list -> 'c
+ val map_fold : map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
 \end{ocaml}
@@
 \end{itemize}
-We thus provide two functions \of{map_local_fold} and \of{map_remote_}
-\of{fold}.
+We thus provide two functions \of{map_local_fold} and \of{map_remote_fold}.
 
@@
 This is the case when \of{fold} is an associative operation (which
-implies that type \of{'b} and \of{'c} are the same).
+implies that types \of{'b} and \of{'c} are the same).
 Whenever
@@
      val compute : 
-       worker:('a -> 'b) -> 
-       master:('a * 'c -> 'b -> ('a * 'c) list) ->
+       worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a * 'c) list) ->
        ('a * 'c) list -> unit
@@
 
-The first example is the classical $N$-queens problem, where we
+The next example is the classical $N$-queens problem, where we
 compute the total number of ways to place $N$ queens on a $N\times N$
@@
 module (actually, a single source code with a command line option).
-This has to be done since the worker code makes use of global
-variables, which prevents us from using
-\of{Network.Same} (as explained below in Section~\ref{sec:implem}). 
-Alternately, we could have refactored the worker
-code to avoid global variables.
 

### 2010-10-11 doc/report/main.tex
*** empty log message ***

@@
 to easily switch between sequential, multi-core, and network
-implementations. A distinguishing feature of our library is a
+implementations. In particular, \functory\ allows to use the same
+executable for master and workers, which makes the deployment of small programs
+immediate --- master and workers being only distinguished by an
+environment variable. \functory\ also allows master and workers to be
+completely different programs, which is ideal for large scale deployment.
+Another distinguishing feature of our library is a
 robust fault-tolerance mechanism which relieves
 the user of cumbersome implementation details.
-
+Finally, \functory\ also allows to cascade several distributed
+computations inside the same program.
+
+\paragraph{Related Work.}
 Closest to the approach in this paper is Yohann Padioleau's MapReduce
@@
 
+\paragraph{Future Work.}
 There are still some interesting features that could be added to our
@@
   Another interesting feature could be the ability
-  to add or remove workers dynamically. Currently, our library assumes
-  that the list of workers to be used is given \emph{a priori}.
+  to add or remove machines dynamically. Currently, our library assumes
+  that the list of machines to be used is given \emph{a priori}, as
+  part of the code. An alternative would be to read machine names from
+  a file, watched periodically by the master.
 \item
   Our library provides limited support for displaying real-time
-  information about the computations. Processing and storing
-  information about tasks locally in the master is straightforward; 
-  monitoring it in real-time is less obvious.
+  information about computations and communications. Processing and storing
+  information about workers and tasks locally in the master is straightforward; 
+  monitoring it in real-time could be done using
+  \textsf{Ocamlviz}~\cite{ocamlviz}. 
 \item 

### 2010-10-12 doc/jfla-2001/main.tex
*** empty log message ***

--- 2010-10-11 doc/report/main.tex
@@
 
-\documentclass[a4paper,12pt]{article}
-
-% The following \documentclass options may be useful:
-%
-% 10pt          To set in 10-point type instead of 9-point.
-% 11pt          To set in 11-point type instead of 9-point.
-% authoryear    To obtain author/year citation style instead of numeric.
-
-\usepackage{fullpage,comment,amsmath,graphicx,color,url}
+\documentclass[twoside]{article}
+
+\usepackage{actes,comment,amsmath,graphicx,color,url}
+\usepackage[french]{babel}
+\usepackage[latin1]{inputenc}
+%\usepackage[T1]{fontenc}
 
@@
 
-\begin{document}
-
 \title{Functory: Distributed Computing \\ for the Common Man\thanks{This
@@
     ANR-08-SEGI-021).}}
-%\subtitle{Subtitle Text, if any}
-
-\author{Jean-Christophe Filli\^{a}tre \quad K. Kalyanasundaram \\
-  CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
-  INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893 \\
-   \url{filliatr@lri.fr, kalyan.krishnamani@inria.fr}}
-
+
+\author{Jean-Christophe Filli\^{a}tre$^{1,2,3}$ \& K. Kalyanasundaram$^3$}
+
+\titlehead{Functory}%  a droite (page impaire)
+
+\authorhead{Filliâtre \& Kalyanasundaram}% a gauche (page paire)
+
+\affiliation{\begin{tabular}{rr} 
+\\ 1:  CNRS / LRI UMR 8623 F-91405 Orsay
+\\ 2:  Université Paris Sud F-91405 Orsay
+\\ 3:  INRIA Saclay -- Île-de-France F-91893 Orsay
+\\ {\tt filliatr@lri.fr},  
+   {\tt kalyan.krishnamani@inria.fr}
+\end{tabular}}
+
+\begin{document}
+\setcounter{page}{1}
 \maketitle

### 2010-10-12 doc/jfla-2011/main.tex
*** empty log message ***

--- 2010-10-12 doc/jfla-2001/main.tex
@@
 \usepackage[latin1]{inputenc}
-%\usepackage[T1]{fontenc}
 
@@
 
-\title{Functory: Distributed Computing \\ for the Common Man\thanks{This
-    research was partly supported by the French national project U3CAT
+\title{Functory: Distributed Computing \\ for the Common Man\footnotetext{Ce
+    travail a été réalisé en partie dans le cadre du projet U3CAT
     (\emph{Unification of Critical C Code Analysis Techniques},

### 2010-10-13 doc/jfla-2011/main.tex
*** empty log message ***

@@
-%-----------------------------------------------------------------------------
-%
-%               Template for sigplanconf LaTeX Class
-%
-% Name:         sigplanconf-template.tex
-%
-% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
-%               file for SIGPLAN conference proceedings.
-%
-% Author:       Paul C. Anagnostopoulos
-%               Windfall Software
-%               978 371-2316
-%               paul@windfall.com
-%
-% Created:      15 February 2005
-%
-%-----------------------------------------------------------------------------
-
 \documentclass[twoside]{article}
@@
 \newcommand{\unix}{\textsc{Unix}}
-
-\title{Functory: Distributed Computing \\ for the Common Man\footnotetext{Ce
+\newcommand{\coeur}{c\oe ur}
+\newcommand{\coeurs}{c\oe urs}
+
+\title{Functory: Une bibliothèque de calcul distribué \\ pour
+  Objective Caml\footnotetext{Ce
     travail a été réalisé en partie dans le cadre du projet U3CAT
@@
 \begin{abstract}
-  We present Functory, a distributed computing library for
-  functional programming languages. The main features of this library
-  include (1) a polymorphic API, (2) several implementations to
-  adapt to different computing environments such as sequential,
-  multi-core or network, and (3) a reliable fault-tolerance mechanism.
-  This paper describes the motivation behind this work, as well as
-  the design and implementation of the library. It also demonstrates
-  the potential of the library using realistic experiments.
+  Cet article présente \functory, une bibliothèque de calcul distribué
+  pour Objective Caml. Les principales caractéristiques de cette
+  bibliothèque sont (1) une interface polymorphe, (2) plusieurs
+  réalisations correspondant à des contextes d'utilisation différents
+  et (3) un mécanisme de tolérance aux pannes.
+  Cet article détaille la conception et la réalisation de Functory et
+  montre son potentiel sur de nombreux exemples.
 \end{abstract}
@@
 
-This paper introduces a generic library for distributed computing,
-targeted at functional programming. 
-This work was initially motivated by the computing needs that exist in our
-own research team. Our applications include large-scale deductive
-program verification, which amounts to checking the validity of a
-large number of logical formulas using a variety of automated theorem
-provers~\cite{filliatre07cav}. Our computing infrastructure consists
-of a few powerful multi-core machines (typically 8 to 16 cores) and
-several desktop PCs (typically dual-core). However, for our
-application needs, there is no library that helps in exploiting such a
-computing infrastructure in our favorite functional programming
-language.  Hence we designed and implemented such a
-library, which is the subject of this paper.
-This library is implemented in \Ocaml, but the
-implementation should be straightforward in any functional
-programming language.
-
-The distributed computing library presented in this paper is not a
-library that helps in parallelizing computations. Rather, it provides
-facilities for reliable, distributed execution of parallelizable
-computations. In particular, it provides a set of APIs that allows the
-execution of large-scale parallelizable computations, very relevant to
-our application needs (and presumably relevant to a variety of
-real-world applications), over multiple cores in the same machine or
-over a network of machines. 
-The most important features of our library are the following:
+Cet article présente \functory, une bibliothèque de calcul distribué
+pour Objective Caml. Initialement, ce travail a été motivé par des
+besoins de calcul au sein de notre équipe de recherche, ProVal.
+Nos applications en vérification déductive de programmes incluent
+notamment la validation de très nombreuses formules logiques par des
+démonstrateurs automatiques variés~\cite{filliatre07cav}.
+Nos moyens de calcul consistent en quelques machines très puissantes
+(typiquement 8 ou 16 \coeurs) et plusieurs ordinateurs de bureau
+(typiquement 2 \coeurs). Aucune bibliothèque ne nous permettait de
+tirer facilement partie de cette infrastructure de calcul dans notre
+langage préféré. C'est pourquoi nous avons conçu et réalisé la
+bibliothèque \functory\ qui est le sujet de cet article.
+Cette bibliothèque est réalisée pour \Ocaml\ mais pourrait facilement
+être adaptée pour tout autre langage fonctionnel.
+
+\functory\ n'est pas une bibliothèque qui aide l'utilisateur à
+paralléliser ses calculs. Son rôle consiste plutôt à offrir des
+facilités pour distribuer, de manière sûre, des calculs déjà
+identifiés comme indépendants. En particulier, \functory\ offre
+plusieurs interfaces génériques pour distribuer des calculs sur différents
+\coeurs\ d'une même machine ou sur un réseau de machines. Ceci
+correspond exactement au contexte qui a motivé la construction de
+cette bibliothèque mais aussi, très probablement, à celui de
+nombreuses autres applications.
+Les principales caractéristiques de \functory\ sont les suivantes :
 \begin{itemize}
-\item \emph{Genericity}: 
-  it allows various patterns of polymorphic computations;
-\item \emph{Simplicity}: switching between multiple cores on the same
-  machine and a network of machines is as simple as changing a couple
-  of lines of code;
-\item \emph{Task distribution and fault-tolerance}: 
-  it relieves the user from implementing task distribution routines
-  and reliable fault-tolerance mechanisms.
+\item \emph{généricité} : les
+  traits fonctionnels que sont l'ordre
+  supérieur et le polymorphisme sont exploités pour fournir un maximum
+  de généricité ;
+\item \emph{simplicité} : on passe d'un calcul séquentiel à un
+  calcul multi-\coeurs puis à un calcul en réseau en ne modifiant
+  que quelques lignes dans le code ;
+\item \emph{distribution et tolérance aux pannes} : l'intégralité de
+  la gestion de la distribution et de la tolérance aux pannes est
+  prise en charge par la bibliothèque.
 \end{itemize}
+
 The application domain of such a distributed computing library is manyfold. 
@@
 conditions using automated theorem provers.
-
-% what it is and what it is not
-% \begin{itemize}
-% \item it is not a library which does parallelization
-% \item it is a library for reliable distributed execution of
-%   parallelizable computation 
-% \end{itemize}
-% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
-% programming, ironically) ;  however, there are differences:
-% \begin{itemize}
-% \item polymorphic, higher-order API
-% \item more generic (cores/network)
-% \item does not focus on association lists
-% \item we don't have data locality (future work)
-% \end{itemize}
-% what is the target audience/applications: 
-% for instance, use if automatic provers on thousands of verification
-% conditions, on computing infrastructures which can be
-% \begin{itemize}
-% \item a single machine with multiple cores, possibly remote
-% \item several machines, small or large in number, over a network
-% \end{itemize}
 
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables OCamlMPI MapReduce
+% LocalWords:  genericity executables OCamlMPI MapReduce généricité
+
+%%% Local Variables: 
+%%% mode: latex
+%%% ispell-local-dictionary: "francais"
+%%% TeX-master: t
+%%% End:

### 2010-10-13 doc/jfla-2011/main.tex
introduction done

@@
 \end{itemize}
-
-The application domain of such a distributed computing library is manyfold. 
-It is worth noting that the library is not targeted at applications
-running on server farms, crunching enormous amounts of data. 
-It is rather more relevant for research labs, to make use of available
-computing resources efficiently. It serves a variety of users and a wide
-spectrum of needs, from desktop PCs to networks of machines, and hence
-the title.
-The most striking application would involve efficient execution
-of any large-scale expensive computation that is parallelizable.
-This is the case in our
-research endeavours, that is validating thousands of verification
-conditions using automated theorem provers.
+Bien que \functory\ a été écrite avec un soucis de généralité, elle ne
+vise pas les grandes grappes de serveurs ou les fermes de calcul. 
+Elle s'adresse plutôt aux équipes de recherche qui souhaitent
+exploiter rapidement des capacités de calcul existantes allant d'un
+simple ordinateur de bureau à un réseau de machines.
+Le reste de cette introduction décrit notre approche du calcul
+distribué dans le contexte d'un langage fonctionnel.
+
+\paragraph{Calcul distribué.}
+Initialement inspirée par la bibliothèque MapReduce de
+Google~\cite{mapreduce}, notre approche lui emprunte beaucoup de
+vocabulaire.
+La bibliothèque \functory\ est centrée autour de la notion de
+\emph{tâche}. Les tâches représentent les calculs atomiques pouvant
+être réalisés de manière indépendante.
+Elles sont traitées par un \emph{patron} et des \emph{ouvriers}
+(respectivement \emph{master} and \emph{workers} en anglais).
+Les ouvriers représentent les capacités de calcul qui effectuent les
+tâches. Ils sont
+matérialisés par un ou plusieurs programmes, s'exécutant en parallèle
+sur une ou plusieurs machines.
+Le rôle du patron consiste à distribuer les tâches auprès des ouvriers
+et à récolter les résultats. Il est matérialisé par un unique
+programme s'exécutant de manière séquentielle.
+
+Une partie importante du travail de \functory\ consiste en la
+transmission des tâches et de leurs résultats. 
+Ceci implique leur sérialisation (en anglais
+\emph{marshalling}) à travers le
+réseau, sur un parc potentiellement hétérogène en termes
+d'architectures et de systèmes d'exploitation.
+La taille du mot et l'\emph{endianness} peuvent notamment
+varier\footnote{Les auteurs se refusent à utiliser le mot \og boutisme
+  \fg\ pour désigner l'\emph{endianness}.}.
+Un autre aspect essentiel du calcul distribué, et de \functory\ en
+particulier, est la \emph{tolérance aux pannes}. Les ouvriers peuvent
+ainsi être arrêtés, relancés, temporairement stoppés ou inatteignables
+à cause de problèmes liés au réseau, sans jamais compromettre le
+résultat final du calcul.
+
+\paragraph{Une approche fonctionnelle.}
+Nous avons essayé de tirer partie des spécificités de la programmation
+fonctionnelle pour proposer une interface la plus générique possible.
+L'une des idées principales de \functory\ est que chaque ouvrier est
+une fonction potentiellement polymorphe
+\begin{ocaml}
+  worker: 'a -> 'b
+\end{ocaml}
+où \of{'a} dénote le type des tâches et \of{'b} le type des résultats.
+Le patron est pour sa part une fonction à laquelle on passe la liste
+des tâches initiales, ainsi qu'une fonction pour traiter les résultats :
+\begin{ocaml}
+  master: ('a -> 'b -> 'a list) -> 'a list -> unit
+\end{ocaml}
+Cette fonction passée en argument à \of{master} est appliquée dès
+qu'un nouveau résultat est disponible. Elle peut produire de nouvelles
+tâches (d'où son type de retour \of{'a list}), qui s'ajoutent alors à
+la liste des tâches à effectuer.
+Le processus complet s'achève lorsque toutes les tâches ont été
+effectuées.
+
+Notre bibliothèque tire partie des capacités de sérialisation
+d'\Ocaml\ autant que possible. Ainsi lorsque le patron et les ouvriers
+sont matérialisés par le même exécutable, fonctions et valeurs
+polymorphes peuvent être sérialisées, ce qui permet de réaliser
+facilement le schéma ci-dessus. Il n'est cependant pas toujours
+possible d'utiliser le même programme pour le patron et les
+ouvriers. Dans ce cas, on peut tout de même continuer à sérialiser des
+valeurs polymorphes lorsque la version d'\Ocaml\ utilisée est la même
+pour tous. Sinon, on ne peut plus que transmettre des chaînes de
+caractères entre les différents acteurs. La bibliothèque \functory\
+s'adapte à toutes ces situations en proposant plusieurs interfaces.
 
@@
 
-The remainder of this section introduces our approach to distributed
-computing in a functional programming setting.
-
-\paragraph{Distributed Computing.}
-A typical distributed computing library, as \functory,
-provides the following (we
-borrow some terminology from Google's MapReduce~\cite{mapreduce}):
-\begin{itemize}
-\item A notion of \emph{tasks} which denote atomic computations
-  to be performed in a distributed manner; 
-\item A set of processes (possibly executing on remote machines)
-  called \emph{workers} that perform
-  the tasks, producing results;
-\item A single process called a \emph{master} which is in charge
-  of distributing the tasks among the workers and managing results
-  produced by the workers.
-\end{itemize}
-In addition to the above, distributed computing environments also
-implement mechanisms for fault tolerance, efficient storage, and
-distribution of tasks. This is required to handle network failures
-that may occur, as well as to optimize the usage of machines in the
-network. Another concern of importance is the transmission of messages
-over the network. This requires efficient 
-\emph{marshaling} of data, that is encoding and decoding of data 
-for transmission over different computing environments.  It is desirable to
-maintain architecture independence while transmitting marshalled data,
-as machines in a distributed computing environment often run on
-different hardware architectures and make use of varying software
-platforms. For example, machine word size or endianness may be different
-across machines on the network.
-
-\paragraph{A Functional Programming Approach.}
-Our work was initially inspired by Google's
-MapReduce\footnote{Ironically, Google's approach itself was inspired
-  by functional programming primitives.}. However, our functional
-programming environment allows us to be more generic. 
-The main idea behind our approach is that
-workers may implement any polymorphic function:
-\begin{ocaml}
-  worker: 'a -> 'b
-\end{ocaml}
-where \of{'a} denotes the type of tasks and \of{'b} the type of results.
-Then the master is given a list of initial tasks, together with a
-function to handle the results:
-\begin{ocaml}
-  master: ('a -> 'b -> 'a list) -> 'a list -> unit
-\end{ocaml}
-The function passed to the master is applied whenever a result is
-available and may in turn generate new tasks (hence the return type
-\of{'a list}).  The master is executed as long as there are pending
-tasks.
-
-Our library makes use of \Ocaml's marshaling capabilities as much as
-possible. Whenever master and worker executables are exactly the same,
-we can marshal polymorphic values and closures. However, it is not
-always possible to have master and workers running the same
-executable. In this case, we cannot marshal closures anymore but we
-can still marshal polymorphic values as long as the same version of
-\Ocaml\ is used to compile master and workers. When different versions
-of \Ocaml\ are used, we can no longer marshal values but we can still
-transmit strings between master and workers. Our library adapts to all
-these situations, by providing several APIs.
-
-\medskip
-The paper is organized as follows. 
-In Section~\ref{sec:API}, we
-describe the main idea behind the generic API, and introduce
-\of{master} and \of{worker} programs. 
-In Section~\ref{sec:derived}, we
-present some additional functions provided by \functory,
-which are derived from the generic
-API. 
-Section~\ref{sec:studies} illustrates the API usage with several
-case studies.
-Section~\ref{sec:implem} delves into the implementation details
-of our library, while Section~\ref{sec:experiments} illustrates the
-potential of the presented library through experimental evaluation. We
-compare our approach with relevant related work in
-Section~\ref{sec:future} and outline our future work.
+Cet article s'organise ainsi.
+La section~\ref{sec:API} présente l'interface de la bibliothèque.
+Son utilisation est alors illustrée sur des exemples dans la
+section~\ref{sec:studies}.
+La section~\ref{sec:implem} donne des détails techniques concernant la
+réalisation de \functory. Enfin la section~\ref{sec:experiments}
+montre le potentiel de cette bibliothèque sur des tests expérimentaux.
+%
+\functory\ est librement distribuée à l'adresse
+\url{http://functory.lri.fr/}. Un rapport plus détaillé que le présent
+article, en anglais, est également disponible sur ce site.
 
@@
 
-% The different
-% API functions are summarized in the following table (their signatures
-% being given above):
-% \begin{center}
-%   \begin{tabular}{|c|c|c|c|c|}
-%     \hline
-%     \ocaml{Sequential} & \ocaml{Cores} &
-%     \multicolumn{3}{|c|}{\ocaml{Network}} 
-%     \\\cline{3-5}
-%     &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
-%     \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
-%     \ocaml{master}  & \ocaml{master}  \\
-%     & & & \ocaml{worker}  & \ocaml{worker} \\\hline
-%   \end{tabular}
-% \end{center}
-% The following section introduces other interesting APIs that we
-% derive from this generic API.
-
-\begin{comment}
-\section{Illustrative Example}\label{sec:example}
-
-Let us assume we need to compute a sum such as
-\begin{displaymath}
-  s(n) = \sum_{i=0}^{n}f(i)
-\end{displaymath}
-where $f$ is a pure and computationally expensive function.
-We start with the sequential implementation provided by the library:
-\begin{ocaml}
-  open Sequential
-\end{ocaml}
-The idea is to test our code on small values of $n$ before
-parallelizing the computation for larger values.
-Let us assume that $f$ is implemented as a top-level \of{worker} function:
-\begin{ocaml}
-  let worker i = ...
-\end{ocaml}
-and that $n$ is obtained from the command line:
-\begin{ocaml}
-  let n = int_of_string Sys.argv.(1)
-\end{ocaml}
-The list of tasks is simply the list of integers $0,1,\dots,n$,
-computed as follows:
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in 
-    for i = 0 to n do l := (i, ()) :: !l done; 
-    !l
-\end{ocaml}
-Each task is actually a pair, where the second component contains
-some information which is kept local to the master process.
-It is irrelevant here and thus the value \of{()} of type \of{unit} is used.
-The current sum is kept into a global reference \of{s}:
-\begin{ocaml}
-  let s = ref 0
-\end{ocaml}
-The master function receives a task \of{(i,())} together with its
-result \of{fi} and simply adds \of{fi} to \of{s}. There is no new task
-generated, hence the returned empty list.
-\begin{ocaml}
-  let master (i,()) fi = s := !s + fi; []
-\end{ocaml}
-Finally, the whole computation takes place when we call the
-\of{compute} function:
-\begin{ocaml}
-  let () = compute ~worker ~master tasks
-\end{ocaml}
-
-Now, let us assume we want to make use of a 4-core machine to compute
-$s(n)$. It is as easy as just replacing the very first line with
-\begin{ocaml}
-  open Cores
-  let () = set_number_of_cores 4
-\end{ocaml}
-while the rest of the code remains unchanged.
-
-Later, we may want to use a network of machines instead, for example
-two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
-cores respectively. Similar to the above, we just replace the first
-lines with
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Same
-\end{ocaml}
-Here, \of{Same} is a module which is to be used when master and worker
-are running the same executable. It provides a \of{compute} function
-which has the same signature as in modules \of{Sequential} and
-\of{Cores}, so that the rest of the code remains unchanged. Master and
-worker processes are distinguished at run-time using an environment
-variable \texttt{WORKER} which is set/unset.
-
-If we need to write two different programs for master and worker, for
-reason of binary incompatibility or any other reason, the library API
-is providing functions to do that. If master and worker are still
-compiled with the same version of \Ocaml, we use the \of{Poly} module
-which provides a polymorphic API. Let us start with the worker program.
-It is as simple as
-\begin{ocaml}
-  open Poly
-  let worker i = ...
-  let () = Worker.compute worker ()
-\end{ocaml}
-The \of{Worker.compute} function enters a loop which waits for
-tasks sent by the master and returns results computed using \of{worker}.
-The master program is almost the same as before. First, we replace
-module \of{Same} with module \of{Poly}:
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Poly
-\end{ocaml}
-Tasks and \of{master} function are unchanged:
-\begin{ocaml}
-  let tasks = ...
-  let s = ref 0
-  let master (i,()) fi = ...
-\end{ocaml}
-Finally, we start the computation with \of{Master.compute}, which does
-not have a \of{worker} parameter anymore:
-\begin{ocaml}
-  let () = Master.compute ~master tasks
-\end{ocaml}
-
-When master and worker programs are compiled with different
-versions of the \Ocaml\ compiler, our library still provides
-a monomorphic API over strings. As a consequence, we need to convert
-integers to and from strings in both master and worker.
-The modified worker program is as follows:
-\begin{ocaml}
-  open Mono
-  let worker i = ...
-  let worker_string i = string_of_int (worker (int_of_string i))
-  let () = Worker.compute worker_string ()
-\end{ocaml}
-The master program is modified in a similar way.
-We simply replace \of{Poly} with \of{Mono} and encode/decode integers
-as strings, as follows:
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in 
-    for i = 0 to n do l := (string_of_int i, ()) :: !l done; 
-    !l
-  let master (i,()) fi = s := !s + int_of_string fi; []
-\end{ocaml}
-\end{comment}
-
-\section{Derived API}\label{sec:derived}
+% \section{Derived API}\label{sec:derived}
 
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables OCamlMPI MapReduce généricité
+% LocalWords:  genericity executables OCamlMPI MapReduce généricité marshalling
 
@@
 %%% End:
+
+% LocalWords:  sérialisation boutisme inatteignables

### 2010-10-13 doc/jfla-2011/main.tex
*** empty log message ***

@@
 \newcommand{\unix}{\textsc{Unix}}
+\newcommand{\oeuvre}{\oe uvre}
 \newcommand{\coeur}{c\oe ur}
@@
 
-\section{API}\label{sec:API}
-In this section, we describe the
-generic API which we shall extend with several specialized
-functions in the next section. The main function in our API follows the idea
-sketched in the introduction. It has the following signature:
+\section{Interface}\label{sec:API}
+Cette section décrit l'interface de la bibliothèque \functory.
+La fonctionnalité primitive est une fonction \of{compute} réalisant
+l'idée principale évoquée dans l'introduction.
 \begin{ocaml}
@@
 \end{ocaml}
-Tasks are pairs, of type \of{'a * 'c}, where the first component is
-passed to the worker and the second component is local to the master.
-The \ocaml{worker} function should be pure and is executed in parallel
-in all worker processes. The function \ocaml{master}, on the
-contrary, can be impure and is only executed in the master process.
-The \ocaml{master} function typically stores results 
-in some internal data structure.
-Additionally, it may produce new tasks, as a list of type 
-\of{('a * 'c) list}, which are then appended to the current set of
-pending tasks.
-The next section will describe the usage of this generic interface to
-perform the traditional map and fold operations.
-
-Actually, our library provides not just a single \ocaml{compute}
-function as above, but instead five different versions
-depending on the execution context. 
-The first two contexts are the following: 
+Les tâches sont des paires, de type \of{'a * 'c}, où la première
+composante sera transmise à l'ouvrier et la seconde conservée
+par le patron. La fonction \of{worker} doit être observationnellement
+pure et sera exécutée en parallèle par tous les ouvriers. La fonction
+\of{master}, au contraire, peut être impure et sera exécutée
+uniquement au sein d'un unique processus séquentiel.
+Cette fonction \of{master} accumule typiquement les résultats renvoyés
+par les ouvriers dans une structure locale. Elle peut en outre
+produire de nouvelles tâches, 
+sous la forme d'une liste de type \of{('a * 'c) list}, 
+qui sont alors ajoutées aux tâches restant à effectuer.
+Le troisième argument de \of{compute} est la liste des tâches
+initiales, qui déclenchent le calcul.
+La fonction \of{compute} rend la main lorsque toutes les tâches ont
+été effectuées. Il n'y a pas de résultat renvoyé, mais uniquement des
+effets de bord de la fonction \of{master}.
+Nous montrerons à la fin de cette section comment en déduire des
+fonctions d'ordre supérieur usuelles telles que \of{map} ou \of{fold}.
+
+En réalité, la bibliothèque \functory\ fournit \emph{cinq}
+mises en {\oe}uvre différentes de la fonction \of{compute}, correspondant à
+cinq contextes d'utilisation différents.
+Les deux premiers sont les plus simples.
 \begin{enumerate}
-\item \textbf{Purely sequential execution:}
-  this is mostly intended to be a reference implementation
-  for performance comparisons, as well as for debugging;
-
-\item \textbf{Several cores on the same machine:} 
-  this implementation is intended to distribute the computation over a
-  single machine and it makes use of \unix\ processes;
+\item \textbf{Exécution purement séquentielle :}
+  Elle permet d'obtenir un code de référence, pour mesurer des
+  performances ou mettre au point son programme.
+
+\item \textbf{Plusieurs \coeurs\ sur une même machine :} 
+  Il s'agit là de distribuer le calcul sur une unique machine,
+  uniquement en créant des processus fils.
 \end{enumerate}
-The next three implementations are intended for distributing the
-computation over a network of machines.
+Les trois contextes suivants correspondent à une distribution du
+calcul sur un réseau de machines.
 \begin{enumerate}
 \setcounter{enumi}{2}
-\item \textbf{Same executable run on master and worker machines:}
-  this implementation makes use of the ability to marshal \Ocaml\
-  closures and polymorphic values.
-  Depending on whether the program is run as a master or as a worker,
-  the relevant arguments of \ocaml{compute} are used.
-
-\item \textbf{Master and workers are different programs, compiled with
-    the same version of \Ocaml:} 
-  we can no longer marshal closures but we can still
-  marshal polymorphic values. As a consequence, 
-  the \ocaml{compute} function is split into two
-  polymorphic functions, to implement the master and workers 
-  respectively:%
-\vspace{-1em}
-\begin{ocaml}
-val Worker.compute : ('a -> 'b) -> unit -> unit
-val Master.compute : 
-  ('a * 'c -> 'b -> ('a * 'c) list) -> 
-  ('a * 'c) list -> unit
-\end{ocaml}
-
-\item \textbf{Master and workers are different programs, not even
-    compiled with the same version of \Ocaml:} we can no
-  longer use marshaling, so the
-  \ocaml{compute} function is split into two monomorphic functions
-  over strings:%
-\vspace{-1em}
-\begin{ocaml}
-val Worker.compute : (string -> string) -> unit -> unit
-val Master.compute : 
-  (string * 'c -> string -> (string * 'c) list) -> 
+\item \textbf{Patron et ouvriers sont matérialisés par un même exécutable :}
+  Cette mise en {\oe}uvre exploite la capacité d'\Ocaml\ à sérialiser
+  fonctions et valeurs polymorphes de manière portable.
+  Selon que le programme est exécuté comme le patron ou comme un
+  ouvrier, les arguments pertinents de la fonction \of{compute} sont utilisés.
+
+\item \textbf{Patron et ouvriers sont matérialisés par différents
+    programmes, compilés avec la même version d'\Ocaml\ :} 
+  Il n'est plus possible de sérialiser des fonctions mais on peut
+  encore sérialiser des valeurs polymorphes.
+  En conséquence, la fonction \of{compute} est présentée sous la forme
+  de deux fonctions, servant respectivement à réaliser le patron et
+  les ouvriers :
+\vspace{-0.5em}
+\begin{ocaml}
+val Worker.compute : ('a -> 'b) -> unit
+val Master.compute : ('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
+\end{ocaml}
+
+\item \textbf{Patron et ouvriers sont matérialisés par différents
+    programmes, qui ne sont même pas compilés avec la même version
+    d'\Ocaml\ :} Il n'est plus possible d'utiliser la sérialisation
+  d'\Ocaml\ et la fonction \of{compute} est présentée sous la forme
+  de deux fonctions ne manipulant plus que des chaînes de caractères :%
+\vspace{-0.5em}
+\begin{ocaml}
+val Worker.compute : (string -> string) -> unit
+val Master.compute : (string * 'c -> string -> (string * 'c) list) ->
                     (string * 'c) list -> unit
@@
 \end{enumerate}
-
-Our library is organized into three modules: \of{Sequential} for the
-pure sequential implementation, \of{Cores} for multiple cores on the
-same machine and \of{Network} for a network of machines, respectively.
-The \of{Network} module itself is organized into three sub-modules,
-called \of{Same}, \of{Poly} and \of{Mono},
-corresponding to contexts 3, 4 and 5 above. 
-The next section presents more functions provided by \functory, which
-are all derived from the generic API just presented.
-
-% \section{Derived API}\label{sec:derived}
-
-In most cases, the easiest way to parallelize an execution it to make
-use of operations over lists, where processing of the
-list elements are done in parallel. 
-The previous section illustrated one such case.
-To facilitate such a processing,
-we now derive the most commonly used list operations from our generic
-API.
-
-The most obvious operation is the traditional map operation over
-lists, that is:
+La bibliothèque \functory\ est donc organisée en trois modules :
+\of{Sequential} pour le calcul purement séquentiel ; \of{Cores} pour
+le calcul distribué sur plusieurs \coeurs\ d'une même machine ; et
+enfin \of{Network} pour le calcul en réseau. Ce dernier module
+comporte trois sous-modules, appelés respectivement \of{Same},
+\of{Poly} and \of{Mono}, correspondant aux situations 3, 4 et 5 ci-dessus.
+
+%%% derived API %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph{Fonctions dérivées.}\label{sec:derived}
+Dans de nombreuses situations, la mise en \oeuvre\ la plus simple du
+parallélisme consiste à appliquer une opération sur une liste, le
+traitement de chaque élément pouvant être réaliser en parallèle.
+\functory\ fournit plusieurs fonctions d'ordre supérieur
+offrant du calcul parallèle sur des listes, toutes dérivées de la fonction
+\of{compute}. En particulier, elles sont disponibles dans les cinq
+modules décrits ci-dessus.
+
+L'opération la plus naturelle est celle consistant à appliquer une
+fonction à tous les éléments d'une liste, c'est-à-dire
 \begin{ocaml}
@@
 \end{ocaml}
-The next natural operation is a combination of map and fold
-operations, that is a function like
-\begin{ocaml}
- val map_fold : map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
-\end{ocaml}
-which, given two functions, an accumulator \of{a} and a list \of{l}, computes
+Plus subtilement, on peut combiner une fonction \of{map:'a -> 'b}
+avec une fonction \of{fold:'c -> 'b -> 'c} pour calculer, à partir d'une liste
+$l$  et d'un accumulateur initial $a$, la valeur finale
 \begin{equation}\label{eq:map-fold}
-  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{map} ~ x_1)) (\of{map} ~ x_2))
-  ... (\of{map} ~ x_n)
+  \of{fold} ~ ... ~ (\of{fold} ~ (\of{fold} ~ a ~ (\of{map} ~ x_1)) ~
+  (\of{map} ~ x_2)) ~ ... ~ (\of{map} ~ x_n)
 \end{equation}
-for some permutation $[x_1,x_2,...,x_n]$ of the list \of{l}.
-We assume that the \of{map} operations are always performed in parallel.
-Regarding \of{fold} operations, we distinguish two cases:
-\begin{itemize}
-\item either \of{fold} operations are computationally less expensive
-  than \of{map} and we perform them locally;
-\item or \of{fold} operations are computationally expensive and we
-  perform them in parallel.
-\end{itemize}
-We thus provide two functions \of{map_local_fold} and \of{map_remote_fold}.
-
-In the case of \of{map_remote_fold}, only one \of{fold} operation can
-be performed at a time (possibly in parallel with \of{map}
-operations), as obvious from~(\ref{eq:map-fold}). However, there are
-cases where several \of{fold} operations can be performed in parallel,
-as early as intermediate results of \of{fold} operations are available.
-This is the case when \of{fold} is an associative operation (which
-implies that types \of{'b} and \of{'c} are the same).
-Whenever
-\of{fold} is also commutative, we can perform even more \of{fold}
-operations in parallel. Thus our API provides two functions
-\of{map_fold_a} and \of{map_fold_ac} for these two particular cases.
-The five operations of the derived API are summarized
-in Figure~\ref{fig:derived}.
-\begin{figure}[t]
-  \begin{ocaml}
-    val map : 
-      f:('a -> 'b) -> 'a list -> 'b list 
-    val map_local_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_remote_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_fold_ac : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b 
-    val map_fold_a : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b
-  \end{ocaml}
-  \caption{Derived API.}
-\label{fig:derived}
-\end{figure}
-
-These five functions are actually derived from the generic API.  The
-generality is achieved by implementing these derivations as a functor.
-This functor, \of{Derive}, is parameterized by the \of{compute}
-function, as follows:
-\begin{ocaml}
-module Derive
-  (X : sig
-     val compute : 
-       worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a * 'c) list) ->
-       ('a * 'c) list -> unit
-   end) = struct ... end
-\end{ocaml}
-We now explain how each function from Figure~\ref{fig:derived} is
-implemented using \of{compute} inside the functor body.
-
-\paragraph{\of{map}.} 
-Function \of{map} is easily implemented by associating an integer
-index with each element (of type \of{'a}) of the input list. Thus the
-list of tasks passed to \of{compute} is a list of pairs of type \of{'a * int}.
-As soon as an intermediate result is available, it is passed to
-\of{master}, which stores it in a local table using the index as a key
-and generates no new task. Whenever \of{compute} returns, we easily
-recover the output list from the values stored in the table.
-
-\paragraph{\of{map_local_fold}.} 
-Function \of{map_local_fold} is implemented using a local reference
-storing the current accumulator (of type \of{'c}).  As soon as an
-intermediate result is available, the \of{master} combines it with the
-accumulator using the function \of{fold}. Whenever \of{compute}
-returns, we simply return the current accumulator.
-
-\paragraph{\of{map_remote_fold}.}
-In this case, combining the accumulator with an intermediate result is
-itself a task to be performed in parallel with \of{map} operations.
-Thus tasks are of two different kinds and we use a sum type to make
-the distinction.
-When an intermediate \of{map} result is available, the
-accumulator may be already in use in a \of{fold} operation. Hence we
-need to store the intermediate \of{map} result in a table local to the
-master, to be
-used as soon as the accumulator becomes available.
-The \of{master} function distinguishes between \of{map} and \of{fold}
-results and, accordingly, updates the table and the accumulator. For a
-\of{map} result, either we combine it with the accumulator in a new
-task or we simply store it in the local table. For a \of{fold} result,
-either we simply store it into the accumulator or we immediately
-combine it with a pending \of{map} result from the table.
-Whenever \of{compute} returns, the accumulator must be available and
-we simply return it.
-
-\paragraph{\of{map_fold_ac}.}
-The function \of{map_fold_ac} is easier than the previous one, since
-we do not need to store more than one intermediate result. Indeed, as
-soon as two intermediate results are available, we immediately combine
-them using a \of{fold} operation. Thus, whenever the \of{master}
-receives some result, it either stores it or combines it with the
-accumulator, depending on the availability/non availability of the
-accumulator.  Whenever \of{compute} returns, the accumulator must be
-available and we simply return it.
-
-\paragraph{\of{map_fold_a}.}
-Function \of{map_fold_a} is definitely the trickiest. Indeed, the
-operation being only associative, we can only combine \emph{adjacent}
-intermediate results. In order to perform this, we store intermediate
-results in a local table according to indices of the input list. More
-precisely, if we have already computed
-\begin{displaymath}
-  x_i \oplus x_{i+1} \oplus \dots \oplus x_j
-\end{displaymath}
-where $\oplus$ denotes the \of{fold} operation, then we associate this
-value with indices $i,j$. As soon as a result of indices $k,i-1$ or
-$j+1,k$ is available, it can be combined with the result of $i,j$.
-Whenever \of{compute} returns, the table should only contain a result
-corresponding to the full input list and we simply return it.
+pour une certaine permutation non spécifiée $[x_1,x_2,...,x_n]$ 
+de la liste \of{l}.
+On peut alors distinguer deux cas, selon que l'opération \of{fold} est
+beaucoup moins coûteuse que l'opération \of{map}, et peut être
+effectuée localement par le patron, ou qu'au contraire elle peut être
+coûteuse et a donc intérêt à être effectuée en parallèle des
+opérations \of{map}. La bibliothèque fournit donc deux fonctions,
+correspondant à ces deux cas de figure :
+\begin{ocaml}
+  val map_{local,remote}_fold : map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c   
+\end{ocaml}
+
+Dans le cas de \of{map_remote_fold}, une seule opération \of{fold}
+peut être effectuée à la fois (possiblement en parallèle d'opérations
+\of{map}), comme le montre l'équation~(\ref{eq:map-fold}).
+Il existe cependant des situations dans lesquelles plusieurs
+opérations \of{fold} peuvent être effectuées en parallèle, dès que des
+résultats intermédiaires de \of{map} sont disponibles. C'est le cas
+notamment lorsque l'opération \of{fold} est associative (ce qui
+implique que les types \of{'b} et \of{'c} sont égaux).
+Lorsque l'opération \of{fold} est de plus commutative, on peut
+effectuer encore plus d'opérations \of{fold} en parallèle.
+Notre interface fournit donc deux autres fonctions pour ces cas
+particuliers :
+\begin{ocaml}
+  val map_fold_ac : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
+  val map_fold_a  : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b
+\end{ocaml}
+
+\hrulefill ICI\hrulefill
 
@@
 generic API.
+
+% TODO : insister sur la simplicité de Same, pour les petits programmes
+% notamment
 
@@
 
-% LocalWords:  sérialisation boutisme inatteignables
+% LocalWords:  sérialisation boutisme inatteignables observationnellement
+% LocalWords:  sous-modules

### 2010-10-15 doc/report/main.tex
*** empty log message ***

@@
 over the network. This requires efficient 
-\emph{marshaling} of data, that is encoding and decoding of data 
+\emph{marshalling} of data, that is encoding and decoding of data 
 for transmission over different computing environments.  It is desirable to
@@
 
-Our library makes use of \Ocaml's marshaling capabilities as much as
+Our library makes use of \Ocaml's marshalling capabilities as much as
 possible. Whenever master and worker executables are exactly the same,
@@
     compiled with the same version of \Ocaml:} we can no
-  longer use marshaling, so the
+  longer use marshalling, so the
   \ocaml{compute} function is split into two monomorphic functions
@@
 list elements are done in parallel. 
-The previous section illustrated one such case.
+% The previous section illustrated one such case.
 To facilitate such a processing,
@@
 \end{ocaml}
-The master uses marshaling to send both a closure of type 
+The master uses marshalling to send both a closure of type 
 \of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting strings
 are passed as argument \of{f} and \of{x} in message
-\of{Assign}. Similarly, the worker uses marshaling to send back the
+\of{Assign}. Similarly, the worker uses marshalling to send back the
 result of the computation of type \of{'b}, which is the argument
@@
 still marshal polymorphic values. Indeed, an interesting property of
-marshaling in \Ocaml\ is to be fully architecture-independent, as long
+marshalling in \Ocaml\ is to be fully architecture-independent, as long
 as a single version of \Ocaml\ is used. It is worth pointing out that
@@
 module of choice in our library, since it can still pass polymorphic
-values over the network. The issues of marshaling are automatically
+values over the network. The issues of marshalling are automatically
 taken care of by \Ocaml\ run-time.
@@
 \paragraph{\of{Mono}.} When master and workers are compiled using
-different versions of \Ocaml, we can no longer use marshaling. 
+different versions of \Ocaml, we can no longer use marshalling. 
 As in the previous case, we split \of{compute} into two functions, one
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables OCamlMPI MapReduce
+% LocalWords:  genericity executables OCamlMPI MapReduce marshalling

### 2010-10-15 doc/report/main.tex
*** empty log message ***

@@
 execution of large-scale parallelizable computations, very relevant to
-our application needs (and presumably relevant to a variety of
+our application needs (and actually relevant to a variety of
 real-world applications), over multiple cores in the same machine or
@@
 \item \emph{Task distribution and fault-tolerance}: 
-  it relieves the user from implementing task distribution routines
-  and reliable fault-tolerance mechanisms.
+  it provides automatic task distribution and
+  a robust fault-tolerance mechanism, thereby
+  relieving the user from implementing such routines.
 \end{itemize}
@@
 the title.
-The most striking application would involve efficient execution
-of any large-scale expensive computation that is parallelizable.
+Typical applications would involve a large number
+of computation expensive tasks.
 This is the case in our
@@
 as machines in a distributed computing environment often run on
-different hardware architectures and make use of varying software
+different hardware architectures and make use of different software
 platforms. For example, machine word size or endianness may be different
@@
 \end{ocaml}
-which, given two functions, an accumulator \of{a} and a list \of{l}, computes
+which, given two functions, an accumulator $a$ and a list $l$, computes
 \begin{equation}\label{eq:map-fold}
@@
 \end{equation}
-for some permutation $[x_1,x_2,...,x_n]$ of the list \of{l}.
+for some permutation $[x_1,x_2,...,x_n]$ of the list $l$.
 We assume that the \of{map} operations are always performed in parallel.

### 2010-10-15 doc/jfla-2011/main.tex
*** empty log message ***

@@
 \paragraph{Calcul distribué.}
-Initialement inspirée par la bibliothèque MapReduce de
+Initialement inspirée par la bibliothèque
+MapReduce\footnote{Ironiquement, l'approche de Google a été elle-même
+  inspirée par la programmation fonctionnelle.} de
 Google~\cite{mapreduce}, notre approche lui emprunte beaucoup de
@@
 pour une certaine permutation non spécifiée $[x_1,x_2,...,x_n]$ 
-de la liste \of{l}.
+de la liste $l$.
 On peut alors distinguer deux cas, selon que l'opération \of{fold} est
@@
 \begin{ocaml}
-  val map_fold_ac : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
-  val map_fold_a  : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b
+  val map_fold_{a,ac} : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
 \end{ocaml}

### 2010-10-21 doc/jfla-2011/main.tex
*** empty log message ***

@@
 Elles sont traitées par un \emph{patron} et des \emph{ouvriers}
-(respectivement \emph{master} and \emph{workers} en anglais).
+(respectivement \emph{master} et \emph{workers} en anglais).
 Les ouvriers représentent les capacités de calcul qui effectuent les
@@
 
-\hrulefill ICI\hrulefill
-
-\medskip
-These five functions are the most natural derivations of the
-\of{compute} functions. There are obviously other possible derivations
-(e.g. a variant of \of{map_fold} where only \of{fold} is meaningful);
-in most cases, they could be derived in a straightforward way from our
-generic API.
+% \hrulefill ICI\hrulefill
+
+%\medskip
+Ces cinq fonctions se dérivent facilement de la fonction \of{compute}.
+On peut évidemment en imaginer d'autres, comme par exemple une variante
+où seule la fonction \of{fold} est significative.
+L'utilisateur de \functory\ ne devrait pas avoir de mal à les dériver
+lui-même de la fonction \of{compute} ou des fonctions du type
+\of{map_fold} ci-dessus.
 
@@
 
-% TODO cannot be instantiated in the monomorphic case
-
-\section{Illustrative Case Studies}\label{sec:studies}
-
-In this section, we show how to use \functory\ on several case studies.
-We focus here on the use of the API. Experimental results of these
-case studies are given in Section~\ref{sec:experiments}.
-The source code for all these case studies is contained in the
-distribution, in sub-directory \texttt{tests}.
-
-\subsection{Matrix Multiplication}\label{sec:matrix}
-
-As a first example, let us consider matrix multiplication.
-We assume two matrices \of{a} and \of{b}, respectively of size
-$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
-a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
-Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
-the standard matrix multiplication would be as follows:
+\section{Études de cas}\label{sec:studies}
+
+Cette section présente plusieurs études de cas que nous avons
+réalisées avec \functory. On se concentre ici sur l'utilisation de la
+bibliothèque ; les résultats expérimentaux sont présentés plus loin
+(section~\ref{sec:experiments}). Le code source de tous les exemples
+ci-dessus est contenu dans la distribution de \functory, dans le
+sous-répertoire \texttt{tests}.
+
+\subsection{Multiplication de matrices}\label{sec:matrix}
+
+Comme premier exemple, on considère la multiplication de deux matrices
+\of{a} et \of{b}, de tailles respectives
+$\of{n}\times\of{p}$ et $\of{p}\times\of{m}$. 
+Le résultat sera stocké dans une matrice
+\of{c} de taille $\of{n}\times\of{m}$.
+On suppose que \of{a}, \of{b} et \of{c} sont contenues dans des
+variables globales.
+En supposant en outre \of{a} organisée en lignes et \of{b} en colonnes, une
+multiplication séquentielle s'écrirait ainsi 
 \begin{ocaml}
   for i = 0 to n-1 do
-    for j = 0 to p-1 do
-      for k = 0 to m-1 do
+    for j = 0 to m-1 do
+      for k = 0 to p-1 do
 	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
@@
 \end{ocaml}
-and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
-assuming addition and multiplication over coefficients to be 
-denoted $+$ and $\times$ respectively.
-An easy way to distribute this computation is obviously to turn each
-computation of the inner loop over \of{k} into a task.
-To do so, we first prepare the list of tasks:
+l'addition et la multiplication des coefficients étant notées
+respectivement $+$ et $\times$.
+La complexité de ce code est clairement $O(\of{n}\times\of{m}\times\of{p})$,
+
+Une manière évidente de distribuer ce calcul consiste à faire de la
+boucle interne sur \of{k} une tâche élémentaire.
+On construit alors une liste de $\of{n} \times \of{m}$ tâches, de la
+manière suivante :
 \begin{ocaml}
@@
     let l = ref [] in
-    for i = 0 to n-1 do for j = 0 to p-1 do
+    for i = 0 to n-1 do for j = 0 to m-1 do
       tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
@@
 \end{ocaml}
-Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
-tuple, together with position \of{(i,j)} as a second tuple.
-The worker is receiving the first tuple as argument and computes
-the dot product:
+Chaque tâche est une paire formée d'une part d'une ligne \of{a.(i)} et
+d'une colonne \of{b.(j)}, et d'autre part d'une position \of{(i,j)}.
+L'ouvrier reçoit la première composante et calcule le produit scalaire.
 \begin{ocaml}
@@
     let c = ref 0 in
-    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
+    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
     !c
 \end{ocaml}
-The master is a one line function which receives the result \of{r}
-from the worker and simply stores it into \of{c}, according to the
-position contained in the input task. It produces no new task.
+La patron est une fonction d'une ligne, qui reçoit le résultat \of{r}
+renvoyé par l'ouvrier et remplit la matrice \of{c}, en fonction de la
+position contenue dans la seconde composante de la tâche. Aucune 
+nouvelle tâche n'est produite.
 \begin{ocaml}
@@
 \end{ocaml}
-Finally, the overall computation is started by invoking \of{compute}
-as follows:
+Au final, l'ensemble du calcul est lancé par un appel à la fonction
+\of{compute}.
 \begin{ocaml}
@@
 \end{ocaml}
-The total number of tasks is $\of{n}\times\of{m}$.
-
-Using the sequential implementation provided by \of{Functory} is as
-simple as including the following line in the code:
-\begin{ocaml}
-  open Sequential
-\end{ocaml}
-Then the computation is roughly similar to the standard multiplication
-described above.
-
-Now, let us assume we want to make use of a 4-core machine to perform
-the same computation. 
-This is achieved by just replacing the line above by
+
+Utiliser la version séquentielle de \functory\ est aussi simple que
+d'ajouter \of{open Sequential} au début du code. Le calcul est alors
+similaire à la multiplication usuelle donnée plus haut. Cela peut être
+néanmoins utile pour vérifier la correction du code avant de chercher
+à distribuer le calcul.
+
+Supposons maintenant que l'on veuille utiliser une machine 4 \coeurs\
+pour effectuer le calcul. Il suffit alors de remplacer la ligne
+\of{open Sequential} par les deux lignes suivantes.
 \begin{ocaml}
@@
 \end{ocaml}
-while the rest of the code remains unchanged.
-
-Later, we may want to use a network of machines instead, for example
-two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
-cores respectively. Similar to the above, we turn the above lines into
+Le reste du code est inchangé.
+
+Supposons enfin que l'on souhaite utiliser plutôt des machines
+présentes sur le réseau, par exemple deux machines appelées
+\texttt{orcus} et \texttt{belzebuth} offrant respectivement 4 et 8 \coeurs.
+Il suffit de remplacer les deux lignes de code ci-dessus par les suivantes.
 \begin{ocaml}
@@
 \end{ocaml}
-Here, \of{Same} is a module which is to be used when master and worker
-are running the same executable. It provides a \of{compute} function
-which has the same signature as in modules \of{Sequential} and
-\of{Cores}, so that the rest of the code remains unchanged. Master and
-worker processes are distinguished at run-time using an environment
-variable \texttt{WORKER} which is set/unset.
-
-If we need to write two different programs for master and worker, for
-reason of binary incompatibility or any other reason, the library API
-is providing functions to do that. If master and worker are still
-compiled with the same version of \Ocaml, we use the \of{Poly} module
-which provides a polymorphic API. Let us start with the worker program.
-It now looks like:
-\begin{ocaml}
-  open Poly
+On utilise ici le sous-module \of{Same} de \of{Network}, qui permet
+d'utiliser le même exécutable pour le patron et les ouvriers. Ce
+module fournit toujours une fonction \of{compute} de même signature
+que dans les modules \of{Sequential} et \of{Cores}, et le reste du
+code est toujours inchangé. Le patron et les ouvriers sont distingués
+à l'exécution par la présence de la variable d'environnement
+\texttt{WORKER}.
+
+Si on a besoin d'écrire deux programmes différents pour le patron et
+les ouvriers, pour des raisons d'incompatibilité de binaires ou tout
+autre raison, l'interface de \functory\ permet de le faire.
+Si les deux programmes sont compilés avec la même version d'\Ocaml, on
+utilise le module \of{Poly}.
+Commençons par l'ouvrier. Son code prend la forme suivante.
+\begin{ocaml}
+  open Network.Poly
   let worker (ai, bj) = ...
@@
 \end{ocaml}
-The \of{Worker.compute} function enters a loop which waits for
-tasks sent by the master and returns results computed using \of{worker}.
-The master program is almost the same as before. First, we replace
-module \of{Same} with module \of{Poly}:
+La fonction \of{Worker.compute} entre dans une boucle qui attend les
+tâches envoyées par le patron et renvoie les résultats calculés par
+la fonction \of{worker}. Le code du patron, quant à lui, est quasiment
+le même qu'auparavant. On commence par remplacer \of{Same} par
+\of{Poly}.
 \begin{ocaml}
@@
 \end{ocaml}
-Tasks and \of{master} function are unchanged:
+La construction des tâches et la fonction \of{master} sont inchangées.
 \begin{ocaml}
@@
 \end{ocaml}
-Finally, we start the computation with \of{Master.compute}, which does
-not have a \of{worker} parameter anymore:
+Enfin on lance le calcul avec la fonction \of{Master.compute}, qui ne
+prend plus de fonction \of{worker} en argument.
 \begin{ocaml}
@@
 
-When master and worker programs are compiled with different
-versions of the \Ocaml\ compiler, our library still provides
-a monomorphic API over strings. As a consequence, we need to convert
-tasks and results to and from strings in both master and worker.
-The modified worker program then looks as follows:
+Quand le patron et les ouvriers sont compilés avec des versions
+différentes d'\Ocaml, la bibliothèque fournit une interface
+ne permettant plus que l'échange de chaînes de caractères.
+En conséquence, il faut convertir les tâches et les résultats depuis
+et vers des chaînes, dans les deux programmes.
+L'ouvrier modifié prend alors la forme suivante.
 \begin{ocaml}
@@
   let worker (ai, bj) = ...
-  let worker_string t = string_of_int (worker (task_of_string i))
+  let worker_string s = string_of_coeff (worker (task_of_string s))
   let () = Worker.compute worker_string ()
 \end{ocaml}
+Le patron est modifié de la même façon.
 The master program is modified in a similar way.
-We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
-as strings, as follows:
+On remplace \of{Poly} par \of{Mono} et on encode/décode les tâches et
+résultats.
 \begin{ocaml}
@@
 \end{ocaml}
-where \of{string_of_task} and \of{task_of_string} are user-defined
-functions to convert tasks to and from strings.
-
-\subsection{N-queens}\label{sec:n-queens}
+Les quatre fonctions de conversion \of{string_of_}\{\of{task,coeff}\} et
+\{\of{task,coeff}\}\of{_of_string} sont à la charge de l'utilisateur.
+
+\subsection{Autres études de cas}
+
+Nous présentons ici trois autres études de cas, plus rapidement.
+
+\paragraph{N-reines.}\label{sec:n-queens}
 
@@
     \hline\hline
-\of{Sequential} & 20.3 &  20.2 \\
+\of{Sequential} & 20.3 \phantom{(0.00)} &  20.2 \phantom{(0.00)} \\
 \hline
  \of{Cores} 
- (2 cores)     &   22.7  &  11.3\\
- (4 cores)     &   12.3  &   6.1\\
- (6 cores)     &    8.6  &   4.3\\
- (8 cores)     &    8.0  &   3.5\\
+ (2 cores)     &   22.7 (0.89) &  11.3 (1.79) \\
+ (4 cores)     &   12.3 (1.65) &   6.1 (3.31) \\
+ (6 cores)     &    8.6 (2.36) &   4.3 (4.70) \\
+ (8 cores)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
  \hline

### 2010-10-21 doc/report/main.tex
*** empty log message ***

@@
   for i = 0 to n-1 do
-    for j = 0 to p-1 do
-      for k = 0 to m-1 do
+    for j = 0 to m-1 do
+      for k = 0 to p-1 do
 	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
@@
     let c = ref 0 in
-    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
+    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
     !c
@@
   let worker (ai, bj) = ...
-  let worker_string t = string_of_int (worker (task_of_string i))
+  let worker_string s = string_of_coeff (worker (task_of_string s))
   let () = Worker.compute worker_string ()
@@
  \of{Cores} 
- (2 cores)     &   22.7  &  11.3\\
- (4 cores)     &   12.3  &   6.1\\
- (6 cores)     &    8.6  &   4.3\\
- (8 cores)     &    8.0  &   3.5\\
+ (2 cores)     &   22.7 (0.89) &  11.3 (1.79) \\
+ (4 cores)     &   12.3 (1.65) &   6.1 (3.31) \\
+ (6 cores)     &    8.6 (2.36) &   4.3 (4.70) \\
+ (8 cores)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
  \hline

### 2010-10-21 doc/jfla-2011/main.tex
*** empty log message ***

@@
 
-% \hrulefill ICI\hrulefill
-
 %\medskip
@@
 \paragraph{N-reines.}\label{sec:n-queens}
-
-The next example is the classical $N$-queens problem, where we
-compute the total number of ways to place $N$ queens on a $N\times N$
-chessboard in such a way no two queens attack each other.
-We use a standard backtracking algorithm for this problem, which
-places the queens one by one starting from the first row.
-Distributing the computation is thus quite easy: we consider all
-possible ways to place queens on the first $D$ rows and then perform
-the subsequent search in parallel. Choosing $D=1$ will result in
-exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
-so on.
-% TODO: make a picture to illustrate positioning of 2 queens on the
-% first 2 rows
-
-Each task only consists of three integers and its result is one integer,
-which is the total number of solutions for this task.
-We make use of function \of{map_local_fold} from the derived API,
-where \of{map} is performing the search and \of{fold} simply adds the
-intermediate results.
-In the network configuration, we make use of the \of{Network.Same}
-module, workers and master being the same executable.
-
-\subsection{Mandelbrot Set}
-
-Drawing the Mandelbrot set is another classical example that could be
-distributed easily, since the color of each point can be computed
-independently of the others. 
-Let us assume we are given the coordinates of the region to be drawn,
-along with the width \of{w} and height \of{h} of the resulting image
-(in pixels). 
-Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
-parameter. It is immediate to split the image into \of{t} sub-images,
-each of which is computed in parallel with and independently of the
-others. For instance, the image could be split into horizontal slices
-or, more generally, into rectangular blocks.
-
-Each task is thus four floating-point numbers denoting the region
-coordinates, together with two integers denoting the dimensions of the
-sub-image to be drawn. The result of the task is a matrix of pixels,
-of size $(\of{w}\times\of{h}) / \of{t}$. 
-For instance, drawing a $800\times 600$ image using 20 tasks will
-result in 20 sub-images of size $176,000$ bytes each,
-assuming each pixel is encoded in four bytes.
-
-In the network configuration, we deliberately choose to have two
-different programs for master and workers, using the \of{Network.Poly}
-module (actually, a single source code with a command line option).
-
-% \subsection{Polynomial Multiplication}
-% TODO?
-
-\subsection{SMT Solvers}\label{sec:SMT}
-
-Here we demonstrate the potential of our library for our application
-needs as mentioned in the introduction. In this case study,
-we consider 80 challenging verification conditions
-(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
-VC is stored in a file, which is accessible over
-NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
-which are all installed on each of the machines in use.
-
-The master program proceeds by reading the file names, turning them
-into tasks by multiplying them by the number of provers (more than
-300 tasks in total).
-Each worker in turn invokes the given prover on the given file, within
-a timeout limit.
-Each task completes with one of the four possible outcomes: \emph{valid},
-\emph{unknown} (depending on
-whether the VC is valid or undecided by the prover), 
-\emph{timeout} and \emph{failure}.
-
-The result of each computation is a pair denoting the status and the
-time spent in the prover call. The master collects these results and
-sums up the timings for each prover and each possible status.
+Il s'agit du problème classique consistant à calculer le nombre de
+façons de disposer $N$ reines sur un échiquier $N\times N$ sans que
+deux d'entre elles soient en prise.
+On utilise un algorithme standard, consistant à positionner une reine
+sur chaque ligne de l'échiquier, en partant de la première ligne. Il
+est facile de distribuer le calcul: on considère toutes les façons de
+placer les reines des $D$ premières lignes et on effectue le reste du
+calcul en parallèle. Pour $D=1$ on obtient ainsi $N$ tâches ; pour
+$D=2$ on obtient $N^2-3N+2$ tâches ; et ainsi de suite.
+Chaque tâche est composée de quelques entiers et son résultat est un
+entier donnant le nombre de solutions pour cette tâche.
+On utilise la fonction \of{map_local_fold} où \of{map} effectue la
+recherche et \of{fold} somme les résultats intermédiaires.
+
+\paragraph{L'ensemble de Mandelbrot.}
+Dessiner l'ensemble de Mandelbrot est un autre exemple de calcul
+aisément distribuable, puisque la couleur de chaque point peut être
+calculée indépendamment. Supposons donnée une région à dessiner, ainsi
+que la taille $\of{w} \times \of{h}$ en pixels de l'image finale.
+On se donne un nombre de tâches $\of{t} \ge 1$ comme paramètre. Il est
+immédiat de découper l'image en \of{t} sous-images, par exemple en
+bandes horizontales (si $\of{h} \ge \of{t}$) ou plus généralement en
+blocs rectangulaires.
+Chaque tâche est constituée de quatre nombre flottants définissant la
+région à dessiner, ainsi que deux entiers donnant la taille de l'image
+correspondante. Le résultat est une matrice de pixels, de taille
+$(\of{w}\times\of{h}) / \of{t}$.
+Ainsi, dessiner une image $800\times 600$ en utilisant 20 tâches
+donnera 20 sous-images de $176\,000$ octets chacune, en supposant
+chaque pixel représenté par quatre octets.
+
+\paragraph{Démonstrateurs automatiques.}\label{sec:SMT}
+Le dernier exemple correspond à celui mentionné dans l'introduction.
+Il s'agit ici de vérifier la validité de 80 obligations de preuve
+(OP), issues de la plate-forme Why~\cite{filliatre07cav}, à l'aide de
+quatre démonstrateurs automatique de la famille SMT (à savoir
+Alt-Ergo, Simplify, Z3 et CVC3).  Chaque OP est vérifiée avec chaque
+démonstrateur, ce qui fait un total de 320 tâches.
+Les OP sont contenues dans des fichiers accessibles par NFS et une
+tâche est donc un nom de fichier et un nom de démonstrateur. 
+Le démonstrateur est appelé comme un programme externe, avec un temps
+d'exécution maximal. Le
+résultat d'une tâche est la réponse du démonstrateur (valide, abandon,
+temps limite atteint, erreur pendant l'exécution) et son temps de calcul.
+Le patron collecte les résultats et les présente au final sous forme
+d'une table synthétique.
+
+% \hrulefill ICI\hrulefill
 

### 2010-10-21 doc/jfla-2011/main.tex
*** empty log message ***

@@
 
-% \hrulefill ICI\hrulefill
-
-\section{Implementation Details}\label{sec:implem}
-
-We now describe the implementations of the various modules introduced
-in Section~\ref{sec:API}. The implementation of the \of{Sequential}
-module is straightforward and does not require any explanation.
-
-\subsection{Multiple Cores}
+\section{Détails techniques}\label{sec:implem}
+
+Cette section décrit la réalisation des différents modules de la
+bibliothèque \functory\ introduits section~\ref{sec:API}, exception
+faire du module \of{Sequential}, dont la réalisation est immédiate.
+
+\hrulefill ICI\hrulefill
+
+Les deux modules \of{Cores} et \of{Network}
+
+\begin{flushleft}
+  \quad  \textbf{tant que} tâches à faire $\lor$ tâches en cours \\
+  \quad  \quad \textbf{tant que} tâches à faire $\land$ ouvriers disponibles \\
+  \quad  \quad \quad affecter une tâche à un ouvrier \\
+  \quad  \quad \textbf{attendre} la fin d'une tâche \\
+  \quad  \quad \quad ajouter les nouvelles tâches renvoyées par \of{master} \\
+\end{flushleft}
+
+\subsection{\of{Cores}}
 
@@
 new tasks.  The main loop can be depicted in the following way:
-\begin{flushleft}
-  \quad  \textbf{while} pending tasks $\lor$ pending sub-processes \\
-  \quad  \quad \textbf{while} pending tasks $\land$ idle cores \\
-  \quad  \quad \quad create new sub-process for some task \\
-  \quad  \quad \textbf{wait} for any completed sub-process \\
-  \quad  \quad \quad push new tasks generated by \of{master} \\
-\end{flushleft}
+
 We also have an alternative implementation using
@@
 
-\subsection{Network of Machines}
+\subsection{\of{Network}}
 

### 2010-10-22 doc/jfla-2011/main.tex
*** empty log message ***

@@
 
-\hrulefill ICI\hrulefill
-
-Les deux modules \of{Cores} et \of{Network}
-
+Les deux modules \of{Cores} et \of{Network} contiennent une boucle
+principale analogue, de la forme suivante :
 \begin{flushleft}
@@
 \end{flushleft}
+La différence se situe dans la technologie utilisée pour affecter une
+tâche à un ouvrier, pour attendre la fin d'une tâche et enfin pour
+ajouter de la tolérance aux pannes.
 
@@
 
-The \of{Cores} module implements the distributed computing library for
-multiple cores on the same machine. 
-% As illustrated in Section~\ref{sec:example}, 
-It provides a function
-\of{set_number_of_cores: int -> unit} to indicate the number of cores
-to be used. The number passed to this function may be different from
-the actual number of cores in the machine; it is rather the number of
-tasks to be performed simultaneously.
-% TODO comment more on the number of cores?
-% Thus a program using the \of{Cores} module typically begins with 
-% \begin{ocaml}
-% open Cores
-% let () = set_number_of_cores 8
-% \end{ocaml}
-
-The \of{Cores} module is implemented with \unix\ processes, using the
-\of{fork} and \of{wait} system calls provided by the \of{Unix} library
-of \Ocaml. The idea is pretty simple.  The \of{compute} function
-maintains a global table of pending tasks and keeps track of the
-number of idle cores.  Whenever there is a pending task and an idle
-core, a new sub-process is created using \of{Unix.fork}; once
-completed, the sub-process marshals the result into a 
-local file. The
-main process maintains a table mapping each sub-process ID to the
-input task and the local file name. It waits for any completed
-sub-process using \of{Unix.wait} and recovers the result from the
-local file. Then function \of{master} is applied, which may generate
-new tasks.  The main loop can be depicted in the following way:
-
-We also have an alternative implementation using
-\unix\ pipes instead of local files.
-
-The scheduling of tasks to the different cores is left to the
-operating system, through \of{Unix.fork}. Thus it may be the case that
-two tasks are executed on the same core, even if the declared number
-of cores is less or equal than the actual number of cores on the machine.
-% TODO: comment on user-provided scheduling?
+Le module \of{Cores} permet de distribuer le calcul sur une unique
+machine, typiquement en exploitant plusieurs \coeurs. 
+Comme illustré section~\ref{sec:studies}, la
+\of{set_number_of_cores} permet de spécifier le nombre de \coeurs\ à
+utiliser. Ce nombre peut être différent du nombre effectif de \coeurs\ 
+de la machine. Il peut même être strictement plus grand.
+En fait, ce nombre indique tout simplement combien de tâches peuvent être
+effectuées simultanément.
+
+Le module \of{Cores} est réalisé à l'aide de processus \unix, en
+utilisant la fonction \of{Unix.fork} de la bibliothèque standard d'\Ocaml.
+Plus précisément, un compteur indique le nombre d'ouvriers disponibles 
+et l'affectation d'une tâche à un ouvrier se fait en 
+créant un nouveau sous-processus avec \of{fork}.
+À la fin du calcul, le résultat de la tâche est transmis au processus
+père par sérialisation dans un tube.
+La répartition des tâches sur les différents \coeurs\ (physiques) de
+la machine, le cas échéant, est laissée au système. Il se peut donc
+que deux tâches se retrouvent être exécutées sur le même \coeur, y
+compris dans le cas où le nombre de \coeurs\ déclarés est inférieur ou
+égal au nombre de \coeurs\ effectifs.
 
@@
 
-The \of{Network} module implements the distributed computing library
-for a network of machines. 
-% As illustrated in Section~\ref{sec:example}, 
-It provides a function
-\of{declare_workers: n:int -> string -> unit} to fill a table of
-worker machines. 
-% The argument \of{n} indicates the number of cores to
-% be used in each worker machine, analogous to the \of{Cores}
-% module. Thus a program using the \of{Network} module typically begins
-% with
-% \begin{ocaml}
-% open Network
-% let () = declare_workers ~n:12 "hostname1"
-% let () = declare_workers ~n:8  "192.168.24.2"
-% ...
-% \end{ocaml}
-
-The \of{Network} module is based on a traditional TCP-based client/server
-architecture, where each worker is a server and the master is the
-client of each worker. The main execution loop is similar to the one
-in the \of{Cores} module, where distant processes on remote machines
-correspond to sub-processes and idle cores are the idle cores 
-of remote workers. In addition to this, it also involves issues of message
-transfer and fault tolerance, which are subsequently described.
-
-\subsubsection{Protocol}\label{sec:protocol}
-
-Messages sent from master to workers could be any of the following kinds:
-\begin{description}
-\item[\of{Assign(id:int, f:string, x:string)}] This message assigns a
-  new task to the worker, the task being identified by the unique
-  integer \of{id}. The task to be performed is given by strings \of{f}
-  and \of{x}, which are interpreted depending on the context.
-
-\item[\of{Kill(id:int)}] This message tells the worker to kill the task
-  identified by \of{id}.
-
-\item[\of{Stop}] This message informs the worker about completion of
-  the computation, so that it may choose to exit.
-
-\item[\of{Ping}] This message is used to check if the worker is still
-  alive, expecting a \of{Pong} message from the worker in return.
-\end{description}
-Messages sent by workers could be any of the following kinds:
-\begin{description}
-\item[\of{Pong}] This message is an acknowledgment for a \of{Ping}
-  message from the master.
-
-\item[\of{Completed(id:int, s:string)}] This message indicates the
-  completion of a task identified by \of{id}, with result \of{s}.
-
-\item[\of{Aborted(id:int)}] This message informs the master that the
-  task identified by \of{id} is aborted, either as a response to a
-  \of{Kill} message or because of a worker malfunction.
-\end{description}
-Our implementation of the protocol works across different
-architectures, so that master and workers could be run on completely
-different platforms w.r.t. endianness, version of \Ocaml\ and
-operating system.
-
-\subsubsection{Network Sub-modules} % TODO: change title
-
-As mentioned in Section~\ref{sec:API}, the \of{Network} module
-actually provides three different implementations, according to three
-different execution scenarios. There are provided in three
-sub-modules, as described below.
-
-\paragraph{\of{Same}.} This module is used when master and workers are
-running the same executable. The master and workers have to be
-differentiated in some manner. We use an environment variable
-\texttt{WORKER} for this purpose. When set, it indicates that the
-executable acts as a worker. At run-time, a worker immediately 
-enters a loop waiting for tasks from the master, without even getting
-into the user code. 
-As explained in Section~\ref{sec:API}, 
-the master function has the following signature.
-\begin{ocaml}
- val compute : 
-      worker:('a -> 'b) -> 
-      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit  
-\end{ocaml}
-The master uses marshaling to send both a closure of type 
-\of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting strings
-are passed as argument \of{f} and \of{x} in message
-\of{Assign}. Similarly, the worker uses marshaling to send back the
-result of the computation of type \of{'b}, which is the argument
-\of{s} in message \of{Completed}.
-
-Though the ability to run the same executable helps a lot in deploying
-the program in different machines, it comes at a small price. Since
-the worker is not getting into the user code, closures which are
-transmitted from the master cannot refer to global variables in the
-user code. Indeed, the initialization code for these global variables
-is never reached on the worker side. For instance, the Mandelbrot set
-example could be written as follows:
-\begin{ocaml}
-  let max_iterations = 200
-  let worker (xmi, xma, ymi, yma, w, h) =
-    ... draw sub-image using max_iterations ...
-\end{ocaml}
-That is, the global function \of{worker} makes use of the global
-variable \of{max_iterations}. The worker gets the function to compute
-from the master, namely the closure corresponding to function
-\of{worker} in that case, but on the worker side the initialization of
-\of{max_iterations} is never executed.
-
-One obvious solution is  not to use global variables in the
-worker code. This is not always possible, though. 
-To overcome this, the \of{Same} sub-module also provides a
-\of{Worker.compute} function to start the worker loop manually from
-the user code. This way, it can be started at any point, in particular
-after the initialization of the required global variables.
-Master and worker are still running the same executable, but are 
-distinguished using a user-defined way (command-line argument,
-environment variable, etc.).
-
-\bigskip
-There are situations where it is not possible
-to run the same executable for master and workers.
-For instance, architectures or operating systems could be different
-across the network.
-For that reason, the \of{Network} module provides two other
-implementations.
-
-\paragraph{\of{Poly}.} When master and workers are compiled with the
-same version of \Ocaml, we can no longer marshal closures but we can
-still marshal polymorphic values. Indeed, an interesting property of
-marshaling in \Ocaml\ is to be fully architecture-independent, as long
-as a single version of \Ocaml\ is used. It is worth pointing out that
-absence of marshaled closures
-now enables the use of two different programs for master and workers. This
-is not mandatory, though, since master and workers could still be
-distinguished at run-time as in the previous case.
-
-On the worker side, the main loop is started manually using
-\of{Worker.compute}. The computation to be performed on each task is
-given as an argument to this function. It thus looks as follows:
-\begin{ocaml}
-  Worker.compute : ('a -> 'b) -> unit -> unit
-\end{ocaml}
-On the master side, the \of{compute} function is simpler than in the
-previous case, as it has one argument less, and thus has the following
-signature. 
-\begin{ocaml}
-  Master.compute : master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit 
-\end{ocaml}
-For realistic applications, where master and workers are completely
-different programs, possibly written by different teams, this is the
-module of choice in our library, since it can still pass polymorphic
-values over the network. The issues of marshaling are automatically
-taken care of by \Ocaml\ run-time.
-
-The derived API presented in Section~\ref{sec:derived} is adapted to
-deal with the absence of closures. Exactly as the \of{compute}
-function, each API now takes two forms, one for the master and another
-for the workers. For example, \of{map_fold_ac} takes the following
-forms.
-\begin{ocaml}
-  Worker.map_fold_ac : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit -> unit
-  Master.map_fold_ac : 'b -> 'a list -> 'b
-\end{ocaml}
-It is the responsibility of the user to ensure type consistency between
-master and workers.
-
-\paragraph{\of{Mono}.} When master and workers are compiled using
-different versions of \Ocaml, we can no longer use marshaling. 
-As in the previous case, we split \of{compute} into two functions, one
-for master and one for workers. In addition, values transmitted over
-the network can only be strings. The signature thus takes the
-following form.
-\begin{ocaml}
-  Worker.compute : (string -> string) -> unit -> unit
-  Master.compute : master:(string * 'c -> string -> (string * 'c) list) -> 
-                  (string * 'c) list -> unit
-\end{ocaml}
-Any other datatype for tasks should be encoded to/from strings. This
-conversion is left to the user.
-Note that the second component of each task is still polymorphic (of
-type \of{'c} here), since it is local to the master.
-
-% There is
-% actually a single, generic implementation, which is then instantiated
-% in three different ways, corresponding to the three scenarios.  The
-% generic implementation assumes that only strings are being transmitted
-% over the network, as evident from the \of{Assign} and \of{Completed}
-% messages in Section~\ref{sec:protocol}. Depending on the
-% implementation, these strings could be marshaled closures, marshaled
-% values or simply user strings.
-
-\subsubsection{Fault Tolerance}\label{sec:fault}
+Le module \of{Network} permet de distribuer le calcul sur un réseau de
+machines. 
+Comme illustré section~\ref{sec:studies}, la fonction 
+\of{declare_workers: n:int -> string -> unit} permet de déclarer
+l'ensemble des machines du réseau sur lesquels se trouvent des
+ouvriers, en spécifiant un nombre d'ouvriers par machine (qui 
+ne coïncide pas nécessairement avec un nombre effectif de \coeurs).
+Le module \of{Network} est basé sur une architecture client/serveur
+utilisant TCP/IP, où chaque ouvrier est un serveur et où le patron
+est (ironiquement) le client de chaque ouvrier.
+
+\paragraph{Protocole.}\label{sec:protocol}
+Le protocole utilisé comporte sept messages différents.
+Les quatre messages possibles envoyés par le patron à un ouvrier sont : 
+\of{Assign(id:int, f:string, x:string)} pour affecter une nouvelle
+tâche \of{id} à l'ouvrier, sous la forme de deux chaînes \of{f} et \of{x} dont
+le sens dépend du contexte ; 
+\of{Kill(id:int)} pour demander l'interruption de la tâche \of{id} ;
+\of{Stop} pour demander l'arrêt définitif de l'ouvrier ;
+et enfin \of{Ping} pour vérifier que l'ouvrier est toujours réactif.
+Inversement, les trois messages possibles envoyés par un ouvrier au
+patron sont :
+\of{Pong} en réponse à un message \of{Ping} ;
+\of{Completed(id:int, s:string)} pour renvoyer le résultat \of{s} de
+la tâche \of{id} ;
+et enfin \of{Aborted(id:int)} pour signifier l'échec de la tâche
+\of{id}, soit en réponse à \of{Kill}, soit à cause d'une erreur lors
+de l'exécution de la fonction \of{worker}.
+
+Ce protocole est de telle
+sorte que patron et ouvriers peuvent être exécutés sur des
+plate-formes complètement différentes, vis-à-vis de
+l'\emph{endianness}, de la version d'\Ocaml\ et du système d'exploitation.
+Dans le sous-module \of{Same}, les arguments \of{f} et \of{x} du
+message \of{Assign} désignent respectivement
+la sérialisation d'une fonction et de
+son argument, et l'argument \of{s} du message \of{Completed} la
+sérialisation du résultat. Dans les sous-module \of{Poly} et
+\of{Mono}, en revanche, la
+partie \of{f} du message \of{Assign} n'est plus pertinente
+car la fonction \of{worker} est maintenant locale à l'ouvrier.
+Les chaînes \of{x} et \of{s} restent pertinentes ; pour \of{Poly},
+ce sont des sérialisations de valeurs \Ocaml\ et pour \of{Mono}  de
+simples chaînes de caractères.
+
+\subsubsection{Tolérance aux pannes}\label{sec:fault}
 
@@
 % LocalWords:  sérialisation boutisme inatteignables observationnellement
-% LocalWords:  sous-modules
+% LocalWords:  sous-modules sérialisations

### 2010-10-22 doc/jfla-2011/main.tex
*** empty log message ***

@@
 
-\section{Experiments}\label{sec:experiments}
+\section{Résultats expérimentaux}\label{sec:experiments}
 
@@
 
-\section{Conclusion and Future Work}\label{sec:future}
+\section{Conclusion et perspectives}\label{sec:future}
 
@@
 
-\paragraph{Acknowledgements.}
-The authors are grateful to the ProVal team for support and comments
-on early versions of the library and of this paper.
+\paragraph{Remerciements.}
+Les auteurs remercient Alain Mebsout et Johannes Kanig pour avoir testé
+la bibliothèque \functory\ et pour leurs commentaires.
 

### 2010-10-22 doc/jfla-2011/main.tex
*** empty log message ***

@@
 faire du module \of{Sequential}, dont la réalisation est immédiate.
-
+%
 Les deux modules \of{Cores} et \of{Network} contiennent une boucle
@@
 
-The main issue in any distributed computing environment is the ability
-to handle faults, which is also a distinguishing feature of our
-library.  Faults are mainly of two kinds: either a worker is stopped,
-and possibly later restarted; or a worker is temporarily or
-permanently unreachable on the network. To provide fault tolerance,
-our master implementation is keeping track of the status of each
-worker.  This status is controlled by two timeout parameters $T_1$ and
-$T_2$ and \of{Ping} and \of{Pong} messages sent by master and workers,
-respectively. There are four possible statuses for a worker:
-\begin{description}
-\item[\of{not connected}:] there is no ongoing TCP connection between
-  the master and the worker;
-\item[\of{alive}:] the worker has sent some message
-  within $T_1$ seconds;
-\item[\of{pinged}:] the worker has not sent any message within $T_1$
-  seconds and the master has sent the worker a
-  \of{Ping} message within $T_2$ seconds;
-\item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
-  message (for more than $T_2$ seconds).
-\end{description}
-% TODO picture
-Whenever we receive a message from a worker, its status changes to
-\of{alive} and its timeout value is reset.
+L'une des principales difficultés dans un environnement de calcul
+distribué est la tolérance aux pannes. C'est l'un des principaux
+atouts de la bibliothèque \functory.
+Les pannes concernent ici uniquement le module \of{Network} et sont de
+deux sortes : un ouvrier peut être stoppé et éventuellement plus tard
+redémarré ; ou un ouvrier peut être temporairement ou définitivement
+inaccessible sur le réseau. Dans tous les cas, on souhaite que le
+calcul parvienne à son terme, dès lors que cela reste possible, et
+idéalement en utilisant au mieux les ressources disponibles.
+
+Pour assurer cette tolérance aux pannes, le patron maintient en
+permanence l'état de chaque ouvrier. Cet état est contrôlé par deux
+délais $T_1$ et $T_2$, paramétrés par l'utilisateur,
+et par l'ensemble des messages échangés.
+Il y a quatre états possibles pour un ouvrier :
+\of{déconnecté} signifie qu'il n'y a pas de connection TCP en cours
+entre le patron et l'ouvrier ; 
+\of{vivant} signifie qu'un message de l'ouvrier a été reçu par le patron
+il y a moins de $T_1$ secondes ;
+\of{contacté} signifie que l'ouvrier n'a pas envoyé de message depuis
+plus de $T_1$ secondes et que le patron lui a envoyé un message
+\of{Ping} depuis moins de $T_2$ secondes ;
+enfin \of{inatteignable} signifie que l'ouvrier n'a toujours pas répondu au
+message \of{Ping} (depuis plus de $T_2$ secondes).
+Dès qu'un message est reçu en provenance d'un ouvrier, son état est
+mis à jour. On a donc l'automate suivant pour les états d'un ouvrier donné.
 \begin{center}
@@
 
-Fault tolerance is achieved by exploiting the status of workers as
-follows. First, tasks are only assigned to workers with either
-\of{alive} or \of{pinged} status. Second, whenever a worker executing
-a task $t$ moves to status \of{not connected} or \of{unreachable}, the
-task $t$ is rescheduled, which means it is put back in the set of
-pending tasks. Whenever a task is completed, any rescheduled copy of
-this task is either removed from the set of pending tasks or killed if
-it was already assigned to another worker.
-
-% The signature for the generic
-% implementation is therefore as follows:
-% \begin{ocaml}
-%   generic_master:
-%     assign_job:('a -> string * string) ->
-%     master:('a * 'c -> string -> ('a * 'c) list)  ->
-%     ('a * 'c) list -> unit
-% \end{ocaml}
-% The first argument \of{assign_job} is a function indicating the two
-% strings that are passed in the \of{Assign} message of the protocol.
-% We can now distinguish the three cases, based on \of{assign_job}:
-% \begin{description}
-% \item[\of{Same}:] in this case, \of{assign_job} simply returns
-%   the marshaled closure together with the marshaled value.
-% \item[\of{Poly}:] in this case, \of{assign_job} is simply
-%   returning the marshaled value (the first string being meaningless in
-%   this context).
-% \item[\of{Mono}:] in this case, \of{assign_job} is immediately
-%   returning its argument (the first string being meaningless in
-%   this context).
-% \end{description}
+La tolérance aux pannes est réalisée en exploitant l'état des ouvriers
+de la manière suivante. Les tâches ne sont envoyées qu'à des ouvriers
+se trouvant dans l'état \of{vivant} ou \of{contacté}.
+D'autre part, dès qu'un ouvrier en train d'effectuer une tâche $t$
+passe dans l'état \of{déconnecté} ou \of{inatteignable}, la tâche $t$
+et reprogrammée, ce qui signifie qu'elle est ajoutée de nouveau à
+l'ensemble des tâches restant à effectuer. Lorsque le patron reçoit un
+résultat pour une tâche $t$, il déprogramme la tâche $t$ si elle avait
+été reprogrammée et indique à tout ouvrier ayant déjà entrepris de
+recalculer $t$ de cesser son travail (avec le message \of{Kill}).
 
@@
 
-Our library, \functory, is implemented in \Ocaml\ and is available
-from \url{http://www.lri.fr/~filliatr/functory/}.
-In this section, we show the experimental results pertaining to the
-case studies of Section~\ref{sec:studies}.
+Cette section présente quelques résultats expérimentaux obtenus avec
+les quatre programmes décrits section~\ref{sec:studies}.
 
-\subsection{N-queens}
-
-The following table shows execution times for various values of $N$
-and our three different implementations: \of{Sequential}, \of{Cores}, and
-\of{Network}. The purpose of this experiment is to measure the speedup
-w.r.t. the sequential implementation. Therefore, all
-computations are performed on the same machine, a 8 cores Intel Xeon
-3.2 GHz running Debian Linux. The sequential implementation uses a
-single core. The multi-core implementation uses the 8 cores of
-the machine. The network implementation uses 8 workers running locally and
-a master running on a remote machine (which incurs communication cost).
-
-The first column shows the value of $N$.  The number of tasks is shown in
-second column.  Then the last three columns show execution times in
-seconds for the three implementations. The figures within brackets
-show the speedup w.r.t. sequential implementation.
+\paragraph{N-reines.}
+La table suivant montre les temps d'exécution pour différentes valeurs
+de $N$ et pour chacun des trois modules \of{Sequential}, \of{Cores} et
+\of{Network}. L'objectif est ici de mesurer le facteur par rapport à
+l'exécution séquentielle. En conséquence, tous les calculs sont
+effectués sur la même machine, un Intel Xeon 8 \coeurs\ 3.2 GHz sous
+Linux Debian. L'exécution séquentielle utilise un unique \coeur. La
+version multi-\coeurs\ utilisent les 8 \coeurs\ de la machine. La
+version réseau utilise également 8 ouvriers locaux à la machine, le
+patron s'exécutant sur une machine distante pour induire des
+communications réseaux réalistes.
+La première colonne donne la valeur de $N$ et la seconde le nombre de
+tâches. Les trois colonnes suivantes donnent les temps d'exécution, en
+secondes, et le facteur d'accélération entre parenthèses.
 \begin{center}
@@
     \hline
-    N & \#tasks  & \of{Sequential}& \of{Cores}                 & \of{Network} 
+    N & \#tâches  & \of{Sequential}& \of{Cores}                 & \of{Network} 
     \\\hline\hline

### 2010-10-22 doc/jfla-2011/main.tex
*** empty log message ***

@@
 \end{center}
-From the table above, it is clear that the \of{Cores} and \of{Network}
-implementations provide a significant speedup. As evident from the
-last row, the speedup is almost 8, which is also the number of
-cores we use.  It is also evident from the last column that the
-\of{Network} implementation performs significantly better when the
-computation time dominates in the total execution time.  The two extreme
-cases correspond to the second and the last row: in the second row, the
-communication time dominates and is in fact more than 91\%\ of the
-total execution time; on the other hand, for the last row
-communication time amounts to just 4.6\%\ of the total execution time.
-As expected, the network implementation is only beneficial when the
-computation time for each individual task is significant, which is the
-case in realistic examples.
+Il apparaît clairement que les modules \of{Cores} et \of{Network}
+permettent d'obtenir un facteur d'accélération significatif, qui
+atteint presque 8, c'est-à-dire le nombre de \coeurs\ utilisés.
+Il apparaît également que le module \of{Network} donne de meilleurs
+résultats lorsque le temps de calcul domine largement le temps de
+communication. Les deux cas extrêmes correspondent aux 
+deuxième et dernière lignes : dans la deuxième ligne, le temps de
+communication domine et représente en fait plus de 91\%\ du temps
+total ; dans la dernière, au contraire, il ne représente plus que
+4.6\%\ du temps total.
+D'une manière générale, le module \of{Network} n'est intéressant que
+si le temps de calcul de chaque tâche est significatif.
 
-\subsection{Mandelbrot Set}
-
-This benchmark consists in drawing the fragment of the Mandelbrot set
-with lower left corner $(-1.1, 0.2)$ and upper right corner $(-0.8,
-0.4)$, as a $9,000\times6,000$ image. The sequential computation of
-this image consumes 29.4 seconds. For \of{Cores} and \of{Network}
-implementations, the computation times in seconds are tabulated below.
-
-% TODO compute ratios for Network
+\paragraph{L'ensemble de Mandelbrot.}
+Dans cette expérience, on dessine le fragment rectangulaire de l'ensemble de
+Mandelbrot compris entre les points $(-1.1, 0.2)$ et $(-0.8, 
+0.4)$, sous la forme d'une image de résolution $9\,000\times6\,000$.
+On utilise la même machine que dans l'expérience précédente.
+Un calcul purement séquentiel prend 29,4 secondes.
+Les résultats pour les modules \of{Cores} et \of{Network} sont
+présentés dans la table ci-dessous.
 \begin{center}
@@
     \hline
-    \#cores  &\#tasks & \of{Cores} & \of{Network} \\
+    \#\coeurs  &\#tâches & \of{Cores} & \of{Network} \\
     \hline\hline
@@
 \end{center}
-The best timings are achieved for the \of{Cores} configuration, where
-communications happen within the same machine and are thus cheaper.
-There are two significant differences with respect to the n-queens
-benchmark.  On one hand, the number of tasks can be controlled more
-easily than in the case of n-queens. We experimentally figured out the
-optimal number of tasks to be 30. One the other hand, each computation
-result is an image, rather than just an integer as in the case of
-n-queens. Consequently, communication costs are much greater. 
-In this particular experiment, the total size of the results
-transmitted is more than 200 Mb.
+Les meilleurs résultats sont obtenus avec le module \of{Cores}, où la
+communication n'intervient que localement, entre processus.
+Il y a deux différences significatives avec l'expérience précédente
+des $N$-reines. D'un côté, le nombre de tâches peut être contrôlé
+beaucoup plus facilement. Expérimentalement, on constate que le nombre
+optimal de tâches est de l'ordre de 30. D'un autre côté, en revanche,
+le résultat de chaque calcul est une image, et non plus un simple
+entier. En conséquence, les coûts de communication sont bien plus
+grands. Dans cette expérience particulière, la taille totale des
+résultats dépasse les 200 Mo.
+
+\hrulefill LA TRADUCTION S'ARRÊTE ICI\hrulefill
 

### 2010-10-22 doc/jfla-2011/main.tex
*** empty log message ***

@@
 
-\hrulefill LA TRADUCTION S'ARRÊTE ICI\hrulefill
+\paragraph{Multiplication de matrices}
+Cette expérience a été inspirée par le concours de programmation
+associé à la conférence PASCO 2010~\cite{PASCO}.
+L'un des problèmes consistait à multiplier deux matrices de taille
+$100\times 100$, c'est-à-dire $\of{n} = \of{p} =
+\of{m} = 100$ avec les notations de la section~\ref{sec:matrix}.
+Les coefficients sont entiers mais peuvent avoir des milliers de
+chiffres et GMP~\cite{GMP} est utilisé ici pour les calculs relatifs
+aux coefficients.
 
-\subsection{Matrix Multiplication}
-
-This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
-It consists of multiplication of
-two square matrices of dimension 100, that is $\of{n} = \of{p} =
-\of{m} = 100$, with integer coefficients.
-Coefficients have several thousands of digits, hence we use
-GMP~\cite{GMP} to handle operations over coefficients.
-
-We compare the performances of two different implementations. In the
-first one, called \of{mm1}, each task consists of the computation of a single
-coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
-In the second one, called \of{mm2}, each task consists of the
-computation of a whole row of the resultant matrix.
-As a consequence, the total number of tasks in
-$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
-The experimental results (in seconds) are tabulated below.
+On compare les performances de deux programmes distribués utilisant
+\functory. Dans le premier, appelé \of{mm1}, chaque tâche consiste en
+le calcul d'un unique coefficient de la matrice résultat, exactement 
+comme décrit section~\ref{sec:matrix}. 
+Dans le second, appelé \of{mm2}, chaque tâche consiste en le calcul de
+toute une ligne de la matrice résultat.
+En conséquence, le nombre total de tâches est $\of{n}\times\of{m}=10,000$ 
+pour \of{mm1} et seulement $\of{n}=100$ pour \of{mm2}.
+Les résultats expérimentaux, toujours en secondes et toujours sur la
+même machine, sont les suivants.
 \begin{center}
@@
     & \of{mm1}       & \of{mm2}  \\
-    & (10,000 tasks) & (100 tasks) \\
+    & (10,000 tâches) & (100 tâches) \\
     \hline\hline
@@
  \of{Cores} 
- (2 cores)     &   22.7 (0.89) &  11.3 (1.79) \\
- (4 cores)     &   12.3 (1.65) &   6.1 (3.31) \\
- (6 cores)     &    8.6 (2.36) &   4.3 (4.70) \\
- (8 cores)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
+ (2 \coeurs)     &   22.7 (0.89) &  11.3 (1.79) \\
+ (4 \coeurs)     &   12.3 (1.65) &   6.1 (3.31) \\
+ (6 \coeurs)     &    8.6 (2.36) &   4.3 (4.70) \\
+ (8 \coeurs)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
  \hline
@@
 \end{center}
-We do not include results for the network configuration, as they do
-not achieve any benefit with respect to the sequential
-implementation. The reason is that the communication cost dominates the
-computation cost in such a way that the total execution time is
-always greater than 30 seconds. Indeed, irrespective of the
-implementation (\of{mm1} or \of{mm2}), the total size of the
-transmitted data is 
-$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
-billions of bytes.
+On exclut les résultats pour la configuration en réseau, car ils
+n'apportent aucune amélioration par rapport à l'exécution séquentielle.
+La raison en est que les coûts de communication dominent largement les
+coûts de calcul, donnant un temps d'exécution total supérieur à 30
+secondes. Que ce soit pour \of{mm1} ou \of{mm2}, le nombre total de
+coefficients transmis est en effet en
+$O(\of{n}\times\of{m}\times\of{p})$, ce qui représente ici des
+milliards d'octets.
+Une réalisation moins naïve pourrait voir les ouvriers lire les deux
+matrices une seule fois, par exemple depuis un fichier, et le patron
+envoyer uniquement des indices de lignes et de colonnes. Cela
+réduirait le volume des communications à seulement $O(\of{n}\times\of{m})$.
 
-A less naive implementation would have the worker read the input matrices
-only once, \emph{e.g.} from a file, and then have the master send only
-row and column indices. This would reduce the communication cost to
-$O(\of{n}\times\of{m})$ only.
-
-%             mm1     mm2
-
-% # tasks    10,000   100
-
-% sequential  20.3   20.2
-
-% cores 2     22.7   11.3
-% cores 4     12.3    6.1
-% cores 6      8.6    4.3
-% cores 8      8.0    3.5
-
-% network = workers on moloch + master on belzebuth
-
-% cores 2     93     32.3
-% cores 4     87     33.7
-% cores 6     92     33.5
-% cores 8     88     32.5
-
-\subsection{SMT Solvers}
-
-As explained in Section~\ref{sec:SMT}, this benchmark consists of 80
-verification conditions, each being checked by 4 different SMT
-solvers. Each task is executed with a timeout limit of 1 minute.
-Our computing
-infrastructure for this experiment consists of 3 machines with 4, 8 and 8 cores
-respectively. 
-The figure below shows the total time in minutes spent by each prover
-for each possible outcome.
+\paragraph{Démonstrateur automatiques.}
+Comme expliqué plus haut, cette expérience consiste à vérifier la
+validité de 80 formules logiques à l'aide de quatre démonstrateurs
+automatiques de la famille SMT. Chacune de ces 320 tâches est exécutée
+dans une limite de temps fixée à une minute.
+La plate-forme de test est ici constituée de trois machines en réseau,
+de 4, 8 et 8 \coeurs\ respectivement.
+Le tableau ci-dessus indique le temps total passé dans chaque
+démonstrateur automatique, cumulé pour chacune des réponses possibles.
 \begin{center}
@@
     \hline
-    prover   & valid & unknown & timeout & failure
+    démonstrateur   & valide & abandon & temps limite & erreur
     \\\hline\hline
-    Alt-ergo & 406.0 & 3.0   &  11400.0 & 0.0       
+    Alt-ergo & 406.0 & 3.0   &  11\,400.0 & 0.0       
     \\\hline
-    Simplify &  0.5   & 0.4   &  1200.0 & 222.0   
+    Simplify &  0.5   & 0.4   &  1\,200.0 & 222.0   
     \\\hline
-    Z3       & 80.7   & 0.0   &  1800.0 & 1695.0   
+    Z3       & 80.7   & 0.0   &  1\,800.0 & 1\,695.0   
     \\\hline
-    CVC3     & 303.0  & 82.7  &  4200.0 & 659.0   
+    CVC3     & 303.0  & 82.7  &  4\,200.0 & 659.0   
     \\\hline
@@
 \end{center}
-These figures sum up to more than 6 hours if provers were executed
-sequentially. However, using our library and our 3-machine
-infrastructure, it completes in 22 minutes and 37 seconds, giving us a
-speedup of more than 16$\times$. We are still far away from the ideal
-ratio of 20$\times$ (we are using 20 cores), since some provers are
-allocating a lot of memory and time spent in system calls is not
-accounted for in the total observed time. However, a ratio of
-16$\times$ is already a significant improvement for our day-to-day
-experiments.
+Ce qui nous intéresse ici est en particulier le temps total
+d'exécution séquentielle cumulée, qui dépasse les 6 heures de calcul.
+Cependant, en utilisant \functory\ et notre petit réseau de trois
+machines, le temps total de calcul n'a pas dépassé 22 minutes et 30
+secondes, ce qui représente une accélération d'un facteur 16.
+On est encore loin du facteur optimal 20 (on utilise ici 20 \coeurs\
+de puissance équivalente), en particulier parce que certains
+démonstrateurs allouent beaucoup de mémoire et que le temps passé dans
+les appels systèmes n'est pas pris en compte ici.
+Un facteur 16 est cependant une accélération significative pour une
+utilisation au jour le jour.
 
@@
 
-We presented a distributed programming environment for functional
-programming. The main features are the genericity of the interface,
-which makes use of polymorphic higher-order functions, and the ability
-to easily switch between sequential, multi-core, and network
-implementations. In particular, \functory\ allows to use the same
-executable for master and workers, which makes the deployment of small programs
-immediate --- master and workers being only distinguished by an
-environment variable. \functory\ also allows master and workers to be
-completely different programs, which is ideal for large scale deployment.
-Another distinguishing feature of our library is a
-robust fault-tolerance mechanism which relieves
-the user of cumbersome implementation details.
-Finally, \functory\ also allows to cascade several distributed
-computations inside the same program.
+Nous avons présenté \functory, une bibliothèque de calcul distribué
+pour \Ocaml. Les caractéristiques principales de cette
+bibliothèque sont son interface générique, sous forme de fonctions
+polymorphes d'ordre supérieur, et la possibilité de basculer aisément
+entre exécutions séquentielle, multi-\coeurs\ et réseau.
+En particulier, \functory\ permet d'utiliser le même exécutable pour
+le patron et les ouvriers, ce qui rend le déploiement de petits
+programmes immédiat --- les deux ne sont distingués que par une
+variable d'environnement. Bien entendu, \functory\ permet également
+l'utilisation de deux programmes complètement différents, ce qui est
+toujours le cas dans une application à grande échelle.
+Une autre caractéristique essentielle de \functory\ est son mécanisme
+robuste de tolérance aux pannes, qui simplifie grandement le code de
+l'utilisateur. 
+% Enfin, une dernière caractéristique non évoquée dans
+% cet article, est la possibilité d'enchaîner plusieurs calculs
+% distribués dans un même programme.
 
-\paragraph{Related Work.}
-Closest to the approach in this paper is Yohann Padioleau's MapReduce
-implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
-top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
-protocol for message passing.  Currently, we have less flexibility
-w.r.t. deployment of the user program than OCamlMPI; on the other
-hand, we provide a 
-more generic API together with fault tolerance.  There are other
-distributed computing libraries on top of which one could implement
-the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
-them. However, \JoCaml\ does not provide fault tolerance, which is
-indispensable in a distributed setting. The user has to include code
-for fault tolerance, as already demonstrated in some \JoCaml\
-experiments~\cite{mandel2008}.
+\paragraph{Travaux connexes.}
+Une bibliothèque relativement proche de la nôtre est l'implémentation
+de MapReduce pour \Ocaml\ par Yohann Padioleau~\cite{poor-man-mapreduce}.
+Elle est construite au dessus d'OCamlMPI~\cite{ocamlMPI}, alors que
+notre bibliothèque utilise son propre protocole.
+OCamlMPI permet notamment un déploiement plus flexible des ouvriers.
+D'un autre côté, notre bibliothèque offre une interface plus générique
+et surtout la tolérance aux pannes.
+Mis à part OcamlMPI, 
+il existe d'autres environnements sur lesquels une bibliothèque comme
+\functory\ pourrait être construite. Un exemple est
+\JoCaml~\cite{jocaml}. Il est cependant important de noter que
+\JoCaml\ n'offre pas de tolérance aux pannes \emph{de
+  facto}. L'utilisateur doit l'inclure dans son code, comme cela a
+déjà été démontré dans certaines expériences impliquant
+\JoCaml~\cite{mandel2008}.  Vu que l'essentiel de la complexité de
+\functory\ se situe dans le code de tolérance aux pannes, l'apport de
+\JoCaml\ serait relativement anecdotique.
 
-There are other implementations of distributed computing in the
-context of functional programming. One is the Disco
-project~\cite{disco}, which implements exactly Google's MapReduce in
-Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
-implementation of Google's MapReduce.
-There are other ways to exploit multi-core architectures. One of these
-is data parallelism, which is also relevant in the functional
-programming setting~\cite{parallel-haskell}. Our work
-does not target data parallelism at all.
+% TODO Mathias Kende
 
-\paragraph{Future Work.}
-There are still some interesting features that could be added to our
-library. 
-\begin{itemize}
-\item 
-  One is the ability to efficiently assign tasks to workers depending
-  on resource parameters, such as data locality, CPU power, memory,
-  etc. This could be achieved by providing the user with the means to control
-  task scheduling.  
-\item 
-  Another interesting feature could be the ability
-  to add or remove machines dynamically. Currently, our library assumes
-  that the list of machines to be used is given \emph{a priori}, as
-  part of the code. An alternative would be to read machine names from
-  a file, watched periodically by the master.
-\item
-  Our library provides limited support for displaying real-time
-  information about computations and communications. Processing and storing
-  information about workers and tasks locally in the master is straightforward; 
-  monitoring it in real-time could be done using
-  \textsf{Ocamlviz}~\cite{ocamlviz}. 
-\item 
-  One very nice feature of Google's MapReduce is the possibility to
-  use redundantly several idle workers on the same tasks
-  for speedup when reaching the end of computation.
-  Since we already have the fault tolerance implemented, this
-  optimization should be straightforward to add to our library.
-\end{itemize}
-We intend to enrich our library with all above features.
+Il existe également des bibliothèques de calcul distribué pour
+d'autres langages fonctionnels. L'une d'elles est le projet
+Disco~\cite{disco}, une implémentation de MapReduce dans le
+langage Erlang~\cite{erlang}. Notre bibliothèque, au contraire, n'est
+pas une implémentation de MapReduce. Une autre manière d'exploiter les
+architectures multi-\coeurs\ consiste à faire du parallélisme de
+données, et cela reste pertinent dans le contexte de la programmation
+fonctionnelle~\cite{parallel-haskell}. Le parallélisme de données
+n'est pas du tout l'objectif de \functory.
 
-%\appendix
+\paragraph{Perspectives.}
+La bibliothèque \functory\ pourrait être enrichie de nombreuses façons.
+L'une d'elles est la prise en compte, dans l'affectation des tâches
+aux ouvriers, de paramètres de ressources telles que la localité des données,
+la puissance de calcul, l'espace mémoire disponible, etc. On peut
+imaginer pour cela offrir à l'utilisateur des moyens de contrôle de
+l'affectation des tâches ou un processus plus automatique.
+% TODO tel que dans la biblio de MK
+
+Une autre amélioration consiste à faciliter le déploiement en réseau,
+en permettant l'ajout et le retrait dynamique de machines
+ouvrières. Pour l'instant notre bibliothèque utilise une liste de
+machines spécifiées \emph{a priori} mais elle pourrait par exemple lire des noms
+de machines depuis un fichier, périodiquement.
+
+\functory\ offre peu d'outils pour visualiser en temps réel l'état des
+calculs et des communications (mis à part quelques informations de
+déverminage). Calculer et stocker une telle information est immédiat,
+et en partie déjà fait, mais son observation en temps réel demande
+plus de travail. Une possibilité serait de réutiliser une partie de
+l'outil \textsf{Ocamlviz}~\cite{ocamlviz}.
+
+Enfin, une fonctionnalité appréciable du MapReduce de Google est
+l'accélération des fins de calcul en utilisant des ouvriers
+disponibles pour dupliquer les toutes dernières tâches. Vu que
+\functory\ comporte déjà tout le mécanisme de tolérance aux pannes,
+ajouter une telle fonctionnalité devrait être relativement simple.
 
@@
 % LocalWords:  sérialisation boutisme inatteignables observationnellement
-% LocalWords:  sous-modules sérialisations
+% LocalWords:  sous-modules sérialisations n'apportent d'octets d'enchaîner
+% LocalWords:  n'offre l'inclure implémente implémentation d'exploiter
+% LocalWords:  l'affectation

### 2010-10-23 doc/jfla-2011/main.tex
*** empty log message ***

@@
 \item \emph{simplicité} : on passe d'un calcul séquentiel à un
-  calcul multi-\coeurs puis à un calcul en réseau en ne modifiant
+  calcul multi-\coeurs\ puis à un calcul en réseau en ne modifiant
   que quelques lignes dans le code ;
@@
 \end{itemize}
-Bien que \functory\ a été écrite avec un soucis de généralité, elle ne
+Bien que \functory\ ait été écrite avec un soucis de généralité, elle ne
 vise pas les grandes grappes de serveurs ou les fermes de calcul. 
@@
 
-%%% derived API %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \paragraph{Fonctions dérivées.}\label{sec:derived}
@@
 parallélisme consiste à appliquer une opération sur une liste, le
-traitement de chaque élément pouvant être réaliser en parallèle.
+traitement de chaque élément pouvant être réalisé en parallèle.
 \functory\ fournit plusieurs fonctions d'ordre supérieur
@@
 
-%\medskip
 Ces cinq fonctions se dérivent facilement de la fonction \of{compute}.
@@
 \end{ocaml}
+où les variables \of{'a}, \of{'b} et \of{'c} du type de
+\of{compute} sont respectivement instanciées par 
+\of{coeff array * coeff array}, \of{coeff} et \of{int * int}.
 
@@
 Le patron est modifié de la même façon.
-The master program is modified in a similar way.
 On remplace \of{Poly} par \of{Mono} et on encode/décode les tâches et
@@
 Dessiner l'ensemble de Mandelbrot est un autre exemple de calcul
-aisément distribuable, puisque la couleur de chaque point peut être
-calculée indépendamment. Supposons donnée une région à dessiner, ainsi
+aisément distribuable.
+Il s'agit de l'ensemble des points $c$ du plan complexe telle que la
+suite définie par $z_0=0$ et $z_{n+1}=z_n^2+c$ reste de module borné.
+Pour le dessiner, on cherche le plus petit $n$ tel que $|z_n|>2$, ce
+qui assure alors la divergence, en s'arrêtant après un nombre maximal
+d'itérations fixé (par exemple 200).
+Il est clair que la couleur de chaque point peut être
+calculée indépendamment. Supposons donnée une région du plan complexe 
+à dessiner, ainsi
 que la taille $\of{w} \times \of{h}$ en pixels de l'image finale.
@@
 blocs rectangulaires.
-Chaque tâche est constituée de quatre nombre flottants définissant la
+Chaque tâche est constituée de quatre flottants définissant la
 région à dessiner, ainsi que deux entiers donnant la taille de l'image
@@
 machine, typiquement en exploitant plusieurs \coeurs. 
-Comme illustré section~\ref{sec:studies}, la
+Comme illustré section~\ref{sec:studies}, la fonction
 \of{set_number_of_cores} permet de spécifier le nombre de \coeurs\ à
@@
 passe dans l'état \of{déconnecté} ou \of{inatteignable}, la tâche $t$
-et reprogrammée, ce qui signifie qu'elle est ajoutée de nouveau à
+est reprogrammée, ce qui signifie qu'elle est ajoutée de nouveau à
 l'ensemble des tâches restant à effectuer. Lorsque le patron reçoit un
@@
 Linux Debian. L'exécution séquentielle utilise un unique \coeur. La
-version multi-\coeurs\ utilisent les 8 \coeurs\ de la machine. La
+version multi-\coeurs\ utilise les 8 \coeurs\ de la machine. La
 version réseau utilise également 8 ouvriers locaux à la machine, le
@@
 Les auteurs remercient Alain Mebsout et Johannes Kanig pour avoir testé
-la bibliothèque \functory\ et pour leurs commentaires.
+la bibliothèque \functory\ et pour leurs commentaires, ainsi que Claude
+Marché pour sa relecture attentive de cet article.
 
@@
 % LocalWords:  n'offre l'inclure implémente implémentation d'exploiter
-% LocalWords:  l'affectation
+% LocalWords:  l'affectation d'itérations

### 2010-10-24 doc/jfla-2011/main.tex
*** empty log message ***

@@
 
-\title{Functory: Une bibliothèque de calcul distribué \\ pour
+\title{Functory : Une bibliothèque de calcul distribué \\ pour
   Objective Caml\footnotetext{Ce
@@
     \hline\hline
-    2       & 10 & 15.8       (1.86) &  20.3       \\
-            & 30 & 15.7       (\underline{1.87}) &  18.7       \\
-            & 100 & 16.1      (1.83) &  19.8       \\
-            & 1000 & 19.6     (1.50) &  38.6      \\
+    2       & 10 & 15.8       (1.86$\times$) &  20.3  (1.45$\times$)      \\
+            & 30 & 15.7       (\underline{1.87}$\times$) &  18.7 (1.57$\times$)       \\
+            & 100 & 16.1      (1.83$\times$) &  19.8   (1.48$\times$)    \\
+            & 1000 & 19.6     (1.50$\times$) &  38.6    (0.76$\times$)  \\
     \hline
-    4       & 10 & 9.50       (3.09)  &  14.4       \\
-            & 30 & 8.26       (\underline{3.56})  &  11.4   \\
-            & 100 & 8.37      (3.51)  &  11.4   \\
-            & 1000 & 10.6     (2.77)  &  20.5    \\
+    4       & 10 & 9.50       (3.09$\times$)  &  14.4     (2.04$\times$)  \\
+            & 30 & 8.26       (\underline{3.56}$\times$)  &  11.4  (2.58$\times$) \\
+            & 100 & 8.37      (3.51$\times$)  &  11.4  (2.58$\times$) \\
+            & 1000 & 10.6     (2.77$\times$)  &  20.5   (1.43$\times$) \\
     \hline
-    8       & 10 & 9.40       (3.13)  &  12.6      \\
-            & 30 & 4.24       (\underline{6.93})  &   7.6      \\
-            & 100 & 4.38      (6.71)  &   7.5      \\
-            & 1000 & 6.86     (4.29)  &  11.3      \\
+    8       & 10 & 9.40       (3.13$\times$)  &  12.6    (2.33$\times$)  \\
+            & 30 & 4.24       (\underline{6.93}$\times$)  &   7.6  (3.87$\times$)    \\
+            & 100 & 4.38      (6.71$\times$)  &   7.5    (3.92$\times$)  \\
+            & 1000 & 6.86     (4.29$\times$)  &  11.3    (2.60$\times$)  \\
     \hline
@@
     \hline\hline
-\of{Sequential} & 20.3 \phantom{(0.00)} &  20.2 \phantom{(0.00)} \\
+\of{Sequential} & 20.3 \phantom{(0.00)$\times$} &  20.2 \phantom{(0.00)$\times$} \\
 \hline
  \of{Cores} 
- (2 \coeurs)     &   22.7 (0.89) &  11.3 (1.79) \\
- (4 \coeurs)     &   12.3 (1.65) &   6.1 (3.31) \\
- (6 \coeurs)     &    8.6 (2.36) &   4.3 (4.70) \\
- (8 \coeurs)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
+ (2 \coeurs)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
+ (4 \coeurs)     &   12.3 (1.65$\times$) &   6.1 (3.31$\times$) \\
+ (6 \coeurs)     &    8.6 (2.36$\times$) &   4.3 (4.70$\times$) \\
+ (8 \coeurs)     &    8.0 (2.54$\times$) &   3.5 (\underline{5.77}$\times$) \\
  \hline

### 2010-10-24 doc/jfla-2011/main.tex
*** empty log message ***

@@
 effets de bord de la fonction \of{master}.
-Nous montrerons à la fin de cette section comment en déduire des
-fonctions d'ordre supérieur usuelles telles que \of{map} ou \of{fold}.
+
+À titre d'illustration, voici comme écrire un équivalent distribué de
+la fonction \of{Array.map}, qui applique une fonction \of{f} à tous
+les éléments d'un tableau \of{a}. 
+\begin{ocaml}
+  let array_map f a =
+    let n = Array.length a in
+    let b = if n = 0 then [||] else Array.create n (f a.(0)) in
+    let tasks = ref [] in for i = 1 to n - 1 do tasks := (a.(i), i) :: !tasks done;
+    compute ~worker:f ~master:(fun (_,i) bi -> b.(i) <- bi; []) !tasks;
+    b
+\end{ocaml}
+La liste des tâches, \of{tasks}, contient 
+les couples \of{(a.(i), i)} pour tous les indices \of{i} du tableau. La
+fonction \of{worker} se réduit à la fonction \of{f}. La fonction
+\of{master} reçoit un résultat \of{bi} correspondant à un indice
+\of{i} et le stocke dans le tableau résultat \of{b}.
+La fin de cette section décrit plusieurs telles
+fonctions d'ordre supérieur fournies par \functory, toutes dérivées de
+la fonction \of{compute}.
 
@@
 \end{ocaml}
+Ces cinq fonctions se dérivent facilement de la fonction \of{compute}.
+Voici par exemple comment réaliser la fonction \of{map_fold_a}.
+On commence par introduire un type somme pour distinguer les tâches
+\of{map} des tâches \of{fold}, c'est-à-dire
+\of{type ('a, 'b) map_or_fold = Map of 'a | Fold of 'b}.
+Le code de l'ouvrier se résumé à distinguer les deux types de tâches.
+\begin{ocaml}
+  let worker = function Map x -> map x | Fold (x, y) -> fold x y
+\end{ocaml}
+Le code du patron est plus compliqué. En effet, l'opération \of{fold}
+étant seulement associative, on ne peut calculer d'expression analogue
+à (\ref{eq:map-fold}) que pour des éléments \emph{consécutifs} de la liste
+initiale $l$. 
+Le patron conserve donc, dans une table de hachage locale, tous les intervales
+$[x_i,\dots,x_j]$ de $l$ pour lesquels la valeur
+$$\of{map}~x_i \oplus \dots \oplus \of{map}~x_j$$ a déjà été calculée
+(en notant ici $\of{fold}$ par $\oplus$), ainsi que cette valeur.
+Cete table est doublement indexée, par $i$ et par $j$. Initialement
+cette table est vide et la liste des tâches ne contient que des tâches
+de type \of{Map}, une par élément de $l$.
+De manière générale, chaque tâche est associée au couple $(i,j)$
+dénotant le segment qu'elle représente.
+Lorsque le patron reçoit un résultat $r$ pour un segment
+$[x_i,\dots,x_j]$, trois cas se présentent. Soit il existe dans la
+table un résultat pour un segment de la forme $[x_k,\dots,x_{i-1}]$,
+que l'on extrait grâce à la clé $i-1$, et on produit alors une
+nouvelle tâche \of{Fold} pour calculer la valeur correspondant à
+$[x_k,\dots, x_j]$, après avoir supprimé de la table les entrées pour
+$k$ et $i-1$. Soit il existe dans la table un résultat pour un segment
+de la forme $[x_{j+1},\dots,x_k]$, que l'on traite de la même
+manière. Soit enfin il n'y a ni entrée pour $i-1$ ni entrée pour
+$j+1$ dans la table, auquel cas on se contente d'ajouter la valeur $r$
+pour les deux clés $i$ et $j$. Lorsque toutes les tâches ont été
+effectuées, la table doit contenir exactement une valeur, associée aux
+clés $1$ et $n$, sauf dans le cas $n=0$ où aucun calcul n'est nécessaire.
+Au final, la fonction \of{map_fold_a} prend la forme suivante, où
+on a seulement omis le code de fusion des résultats expliqué ci-dessus.
+\begin{ocaml}
+  let map_fold_a ~map ~fold acc l =
+    let tasks = let i = ref 0 in List.map (fun x -> incr i; Map x, (!i, !i)) l in
+    let results = Hashtbl.create 17 in 
+    let merge i j r = ...traiter le résultat r du segment [i,j]... in
+    let worker = function Map x -> map x | Fold (x, y) -> fold x y in
+    let master x r = match x with Map _,(i,_) -> merge i i r | Fold _,(i,j) -> merge i j r in
+    compute ~worker ~master tasks;
+    try let _,_,r = Hashtbl.find results 1 in r with Not_found -> acc
+\end{ocaml}
 
-Ces cinq fonctions se dérivent facilement de la fonction \of{compute}.
-On peut évidemment en imaginer d'autres, comme par exemple une variante
+On peut évidemment dériver d'autres fonctions du même genre,
+comme par exemple une variante
 où seule la fonction \of{fold} est significative.

### 2010-10-24 doc/jfla-2011/main.tex
*** empty log message ***

@@
 L'opération la plus naturelle est celle consistant à appliquer une
-fonction à tous les éléments d'une liste, c'est-à-dire
+fonction à tous les éléments d'une liste, analogue à la fonction
+\of{array_map} décrite plus haut, c'est-à-dire
 \begin{ocaml}
@@
 principale analogue, de la forme suivante :
+
+\medskip
+\begin{minipage}{1.0\linewidth}
   \begin{flushleft}
@@
   \end{flushleft}
+\end{minipage}
+
+\medskip\noindent
 La différence se situe dans la technologie utilisée pour affecter une

### 2010-12-02 doc/jfla-2011/main.tex
*** empty log message ***

@@
 Bien que \functory\ ait été écrite avec un soucis de généralité, elle ne
-vise pas les grandes grappes de serveurs ou les fermes de calcul. 
+vise pas les fermes de calcul où de grandes quantités de données sont
+manipulées, en particulier parce qu'il n'y a pas, pour l'instant, de
+moyen d'assurer la localité des données. 
 Elle s'adresse plutôt aux équipes de recherche qui souhaitent
@@
 où \of{'a} dénote le type des tâches et \of{'b} le type des résultats.
-Le patron est pour sa part une fonction à laquelle on passe la liste
-des tâches initiales, ainsi qu'une fonction pour traiter les résultats :
+Le patron est pour sa part une fonction à laquelle on passe 
+d'une part une fonction  pour traiter les résultats et d'autre part
+la liste des tâches initiales~:
 \begin{ocaml}
@@
 composante sera transmise à l'ouvrier et la seconde conservée
-par le patron. La fonction \of{worker} doit être observationnellement
-pure et sera exécutée en parallèle par tous les ouvriers. La fonction
+par le patron. La fonction \of{worker} doit être 
+pure\footnote{On entend ici \emph{observationnellement pure} mais on
+  autorise la levée d'exceptions pour signaler l'échec du calcul.}
+et sera exécutée en parallèle par tous les ouvriers. La fonction
 \of{master}, au contraire, peut être impure et sera exécutée
@@
 \end{ocaml}
-Plus subtilement, on peut combiner une fonction \of{map:'a -> 'b}
+Plus subtilement, on peut combiner une fonction \of{f:'a -> 'b}
 avec une fonction \of{fold:'c -> 'b -> 'c} pour calculer, à partir d'une liste
@@
 \begin{equation}\label{eq:map-fold}
-  \of{fold} ~ ... ~ (\of{fold} ~ (\of{fold} ~ a ~ (\of{map} ~ x_1)) ~
-  (\of{map} ~ x_2)) ~ ... ~ (\of{map} ~ x_n)
+  \of{fold} ~ ... ~ (\of{fold} ~ (\of{fold} ~ a ~ (\of{f} ~ x_1)) ~
+  (\of{f} ~ x_2)) ~ ... ~ (\of{f} ~ x_n)
 \end{equation}
@@
 On peut alors distinguer deux cas, selon que l'opération \of{fold} est
-beaucoup moins coûteuse que l'opération \of{map}, et peut être
+beaucoup moins coûteuse que l'opération \of{f}, et peut être
 effectuée localement par le patron, ou qu'au contraire elle peut être
 coûteuse et a donc intérêt à être effectuée en parallèle des
-opérations \of{map}. La bibliothèque fournit donc deux fonctions,
+opérations \of{f}. La bibliothèque fournit donc deux fonctions,
 correspondant à ces deux cas de figure :
 \begin{ocaml}
-  val map_{local,remote}_fold : map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c   
+  val map_{local,remote}_fold : f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c   
 \end{ocaml}
@@
 peut être effectuée à la fois (possiblement en parallèle d'opérations
-\of{map}), comme le montre l'équation~(\ref{eq:map-fold}).
+\of{f}), comme le montre l'équation~(\ref{eq:map-fold}).
 Il existe cependant des situations dans lesquelles plusieurs
 opérations \of{fold} peuvent être effectuées en parallèle, dès que des
-résultats intermédiaires de \of{map} sont disponibles. C'est le cas
+résultats intermédiaires de \of{f} sont disponibles. C'est le cas
 notamment lorsque l'opération \of{fold} est associative (ce qui
@@
 \begin{ocaml}
-  val map_fold_{a,ac} : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
+  val map_fold_{a,ac} : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
 \end{ocaml}
@@
 On commence par introduire un type somme pour distinguer les tâches
-\of{map} des tâches \of{fold}, c'est-à-dire
+\of{f} des tâches \of{fold}, c'est-à-dire
 \of{type ('a, 'b) map_or_fold = Map of 'a | Fold of 'b}.
@@
 \begin{ocaml}
-  let worker = function Map x -> map x | Fold (x, y) -> fold x y
+  let worker = function Map x -> f x | Fold (x, y) -> fold x y
 \end{ocaml}
@@
 $[x_i,\dots,x_j]$ de $l$ pour lesquels la valeur
-$$\of{map}~x_i \oplus \dots \oplus \of{map}~x_j$$ a déjà été calculée
+$$\of{f}~x_i \oplus \dots \oplus \of{f}~x_j$$ a déjà été calculée
 (en notant ici $\of{fold}$ par $\oplus$), ainsi que cette valeur.
@@
 de la forme $[x_{j+1},\dots,x_k]$, que l'on traite de la même
-manière. Soit enfin il n'y a ni entrée pour $i-1$ ni entrée pour
+manière. (Si les deux situations se présentent en même temps, on
+choisit arbitrairement, car il y a de toutes façons deux tâches
+\of{Fold} à effectuer.)
+Soit enfin il n'y a ni entrée pour $i-1$ ni entrée pour
 $j+1$ dans la table, auquel cas on se contente d'ajouter la valeur $r$
@@
 \begin{ocaml}
-  let map_fold_a ~map ~fold acc l =
+  let map_fold_a ~f ~fold acc l =
     let tasks = let i = ref 0 in List.map (fun x -> incr i; Map x, (!i, !i)) l in
@@
     let merge i j r = ...traiter le résultat r du segment [i,j]... in
-    let worker = function Map x -> map x | Fold (x, y) -> fold x y in
+    let worker = function Map x -> f x | Fold (x, y) -> fold x y in
     let master x r = match x with Map _,(i,_) -> merge i i r | Fold _,(i,j) -> merge i j r in
@@
       for k = 0 to p-1 do
-	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
+	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k) (* b en colonnes *)
       done
@@
 entier donnant le nombre de solutions pour cette tâche.
-On utilise la fonction \of{map_local_fold} où \of{map} effectue la
+On utilise la fonction \of{map_local_fold} où \of{f} effectue la
 recherche et \of{fold} somme les résultats intermédiaires.
@@
 est (ironiquement) le client de chaque ouvrier.
+Le patron est un programme purement séquentiel.
+En particulier, il ne peut administrer les tâches pendant l'exécution
+de la fonction \of{master}. Ce n'est pas un problème en pratique, car
+on peut supposer que la fonction \of{master} s'exécute rapidement.
+L'ouvrier, en revanche, crée un nouveau processus avec \of{fork} pour
+exécuter la tâche qui lui est confiée et peut donc continuer à
+communiquer avec le patron pendant son calcul.
 
@@
 atouts de la bibliothèque \functory.
-Les pannes concernent ici uniquement le module \of{Network} et sont de
+La tolérance aux pannes de \functory\ est limitée aux ouvriers ; la
+tolérance aux pannes du patron est laissée à l'utilisateur, par
+exemple sous la forme d'une sauvegarde régulière de l'état du patron.
+Les pannes des ouvriers sont de
 deux sortes : un ouvrier peut être stoppé et éventuellement plus tard
@@
 
+Enfin, il est important de signaler que notre bibliothèque est
+également robuste vis-à-vis des exceptions levées par la fonction
+\of{worker}. Le cas échéant, un message \of{Aborted} est envoyé au
+patron et la tâche est reprogrammée. C'est la responsabilité de
+l'utilisateur de rattraper et de traiter ces exceptions si nécessaire.
+
 \section{Résultats expérimentaux}\label{sec:experiments}
@@
 \paragraph{N-reines.}
-La table suivant montre les temps d'exécution pour différentes valeurs
+La table suivante montre les temps d'exécution pour différentes valeurs
 de $N$ et pour chacun des trois modules \of{Sequential}, \of{Cores} et
@@
 \begin{center}
-  \begin{tabular}{|r|r|r|r|r|}
+  \begin{tabular}{|r|r|r|r|r|r|}
     \hline
-    N & \#tâches  & \of{Sequential}& \of{Cores}                 & \of{Network} 
+    N & D & \#tâches  & \of{Sequential}& \of{Cores}            & \of{Network} 
     \\\hline\hline
-    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
-    \\\hline
-       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
-    \\\hline
-    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
-    \\\hline
-       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
-    \\\hline
-    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
-    \\\hline
-       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
-    \\\hline
-    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
-    \\\hline
-       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
+    16 & 1 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
+    \\\hline
+       & 2 & 210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
+    \\\hline
+    17 & 1 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
+    \\\hline
+       & 2 & 240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
+    \\\hline
+    18 & 1 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
+    \\\hline
+       & 2 & 272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
+    \\\hline
+    19 & 1 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
+    \\\hline
+       & 2 & 306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
     \\\hline
@@
 Les résultats pour les modules \of{Cores} et \of{Network} sont
-présentés dans la table ci-dessous.
+présentés figure~\ref{fig:bench:mandelbrot}.
+\begin{figure}
+  \centering
   \begin{center}
@@
   \end{center}
+
+  \caption{Résultats pour l'ensemble de Mandelbrot.}
+\label{fig:bench:mandelbrot}
+\end{figure}
 Les meilleurs résultats sont obtenus avec le module \of{Cores}, où la
@@
 milliards d'octets.
-Une réalisation moins naïve pourrait voir les ouvriers lire les deux
+Dans une réalisation moins naïve, les ouvriers pourrait lire les deux
 matrices une seule fois, par exemple depuis un fichier, et le patron
-envoyer uniquement des indices de lignes et de colonnes. Cela
+n'enverrait uniquement que des indices de lignes et de colonnes. Cela
 réduirait le volume des communications à seulement $O(\of{n}\times\of{m})$.
@@
 l'affectation des tâches ou un processus plus automatique.
+Cela permettrait d'utiliser \functory\ dans les mêmes contextes que MapReduce.
+Il est important de noter qu'à l'heure actuelle, en l'absence de toute
+information concernant les tâches, leur ordonnancement est complètement
+arbitraire. Les modules \of{Cores} et \of{Network} utilisent de
+une simple file pour les tâches restant à effectuer et les nouvelles
+tâches, produites par le patron, sont ajoutées à la fin de cette file.
 % TODO tel que dans la biblio de MK
@@
 calculs et des communications (mis à part quelques informations de
-déverminage). Calculer et stocker une telle information est immédiat,
+\emph{debugging}). Calculer et stocker une telle information est immédiat,
 et en partie déjà fait, mais son observation en temps réel demande

### 2010-12-02 doc/report/main.tex
Speed-up ratios for Mandelbrot Set experiments added [KK]
*** empty log message ***

@@
 It is worth noting that the library is not targeted at applications
-running on server farms, crunching enormous amounts of data. 
+running on server farms, crunching enormous amounts of data, mostly
+due to the lack of data locality. 
 It is rather more relevant for research labs, to make use of available
@@
 where \of{'a} denotes the type of tasks and \of{'b} the type of results.
-Then the master is given a list of initial tasks, together with a
-function to handle the results:
+Then the master is a
+function to handle the results together with
+a list of initial tasks:
 \begin{ocaml}
@@
 passed to the worker and the second component is local to the master.
-The \ocaml{worker} function should be pure and is executed in parallel
+The \ocaml{worker} function should be pure\footnote{We mean
+  \emph{observationally pure} here but we allow exceptions to be
+  raised to signal failures.} 
+and is executed in parallel
 in all worker processes. The function \ocaml{master}, on the
@@
 \begin{ocaml}
- val map_fold : map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
+ val map_fold : f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
 \end{ocaml}
@@
 \begin{equation}\label{eq:map-fold}
-  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{map} ~ x_1)) (\of{map} ~ x_2))
-  ... (\of{map} ~ x_n)
+  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{f} ~ x_1)) (\of{f} ~ x_2))
+  ... (\of{f} ~ x_n)
 \end{equation}
 for some permutation $[x_1,x_2,...,x_n]$ of the list $l$.
-We assume that the \of{map} operations are always performed in parallel.
+We assume that the \of{f} operations are always performed in parallel.
 Regarding \of{fold} operations, we distinguish two cases:
@@
 \item either \of{fold} operations are computationally less expensive
-  than \of{map} and we perform them locally;
+  than \of{f} and we perform them locally;
 \item or \of{fold} operations are computationally expensive and we
@@
 In the case of \of{map_remote_fold}, only one \of{fold} operation can
-be performed at a time (possibly in parallel with \of{map}
+be performed at a time (possibly in parallel with \of{f}
 operations), as obvious from~(\ref{eq:map-fold}). However, there are
@@
     val map_local_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+      f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
       'c -> 'a list -> 'c 
     val map_remote_fold : 
-      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
+      f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
       'c -> 'a list -> 'c 
     val map_fold_ac : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+      f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
       'b -> 'a list -> 'b 
     val map_fold_a : 
-      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+      f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
       'b -> 'a list -> 'b
@@
 In this case, combining the accumulator with an intermediate result is
-itself a task to be performed in parallel with \of{map} operations.
+itself a task to be performed in parallel with \of{f} operations.
 Thus tasks are of two different kinds and we use a sum type to make
 the distinction.
-When an intermediate \of{map} result is available, the
+When an intermediate \of{f} result is available, the
 accumulator may be already in use in a \of{fold} operation. Hence we
-need to store the intermediate \of{map} result in a table local to the
+need to store the intermediate \of{f} result in a table local to the
 master, to be
 used as soon as the accumulator becomes available.
-The \of{master} function distinguishes between \of{map} and \of{fold}
+The \of{master} function distinguishes between \of{f} and \of{fold}
 results and, accordingly, updates the table and the accumulator. For a
-\of{map} result, either we combine it with the accumulator in a new
+\of{f} result, either we combine it with the accumulator in a new
 task or we simply store it in the local table. For a \of{fold} result,
 either we simply store it into the accumulator or we immediately
-combine it with a pending \of{map} result from the table.
+combine it with a pending \of{f} result from the table.
 Whenever \of{compute} returns, the accumulator must be available and
@@
 We make use of function \of{map_local_fold} from the derived API,
-where \of{map} is performing the search and \of{fold} simply adds the
+where \of{f} is performing the search and \of{fold} simply adds the
 intermediate results.
@@
 correspond to sub-processes and idle cores are the idle cores 
-of remote workers. In addition to this, it also involves issues of message
-transfer and fault tolerance, which are subsequently described.
+of remote workers. 
+The master is purely sequential. In particular, when running the user
+\of{master} function, it is not capable of performing any task-related
+computation. This is not an issue, as we assume the \of{master}
+function not to be time-consuming.
+The worker, on the other hand, forks a new process to execute the task
+and hence can communicate with the master during its computation.
+We subsequently describe issues of message
+transfer and fault tolerance.
 
@@
 \begin{ocaml}
-  Worker.map_fold_ac : map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit -> unit
+  Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit -> unit
   Master.map_fold_ac : 'b -> 'a list -> 'b
@@
 to handle faults, which is also a distinguishing feature of our
-library.  Faults are mainly of two kinds: either a worker is stopped,
+library.  
+The fault tolerance mechanism of \functory\ is limited to workers;
+handling master failures is the responsibility of the user, for
+instance by periodically logging the master's state.
+Worker faults are mainly of two kinds: either a worker is stopped,
 and possibly later restarted; or a worker is temporarily or
@@
 \end{description}
-% TODO picture
 Whenever we receive a message from a worker, its status changes to
@@
 it was already assigned to another worker.
+
+It is worth noticing that our library is also robust w.r.t. exceptions
+raised by the user-provided \of{worker} function. In that case, an
+\of{Aborted} message is sent to the master and the task is
+rescheduled. It is the responsibility of the user to handle such
+exceptions if necessary.
 
@@
 \begin{center}
-  \begin{tabular}{|r|r|r|r|r|}
+  \begin{tabular}{|r|r|r|r|r|r|}
     \hline
-    N & \#tasks  & \of{Sequential}& \of{Cores}                 & \of{Network} 
+    N & D & \#tasks  & \of{Sequential}& \of{Cores}              & \of{Network} 
     \\\hline\hline
-    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
+    16 &1&   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
     \\\hline
-       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
+       &2&  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
     \\\hline
-    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
+    17 &1&   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
     \\\hline
-       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
+       &2&  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
     \\\hline
-    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
+    18 &1&   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
     \\\hline
-       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
+       &2&  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
     \\\hline
-    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
+    19 &1&   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
     \\\hline
-       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
+       &2&  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
     \\\hline
@@
  \of{Cores} 
- (2 cores)     &   22.7 (0.89) &  11.3 (1.79) \\
- (4 cores)     &   12.3 (1.65) &   6.1 (3.31) \\
- (6 cores)     &    8.6 (2.36) &   4.3 (4.70) \\
- (8 cores)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
+ (2 cores)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
+ (4 cores)     &   12.3 (1.65$\times$) &   6.1 (3.31$\times$) \\
+ (6 cores)     &    8.6 (2.36$\times$) &   4.3 (4.70$\times$) \\
+ (8 cores)     &    8.0 (2.54$\times$) &   3.5 (\underline{5.77}$\times$) \\
  \hline
@@
   etc. This could be achieved by providing the user with the means to control
-  task scheduling.  
+  task scheduling.  This would enable \functory\ to scale up to
+  MapReduce-like applications.
+
+  Currently, without any information about the tasks, the scheduling
+  is completely arbitrary. In both \of{Cores} and \of{Network}
+  modules, we use traditional queues for the pending tasks; in
+  particular, new tasks produced by the master are appended at the end
+  of the queue.
+
 \item 

### 2010-12-09 doc/jfla-2011/jfla_2011_final_style/main.tex
Functory paper source files in JFLA-2011 latex style - [KK]

--- 2010-12-02 doc/jfla-2011/main.tex
@@
-\documentclass[twoside]{article}
-
-\usepackage{actes,comment,amsmath,graphicx,color,url}
+\documentclass[nofirstpagebreak,cropmarks]{studia-Hermann}
+%\documentclass[english,nofirstpagebreak,cropmarks]{studia-Hermann} %%if the article is written in english language
+
+\usepackage[T1]{fontenc} % if ISO Latin 1 encoding
+%\usepackage[applemac]{inputenc} % if Mac OS Roman encoding
 \usepackage[french]{babel}
+
+%% added from the other file - KK
+% \usepackage{actes,comment,amsmath,graphicx,color,url}
 \usepackage[latin1]{inputenc}
-
+\usepackage{url}
 \newcommand{\Ocaml}{OCaml}
@@
 \newcommand{\coeurs}{c\oe urs}
-
-\title{Functory : Une bibliothèque de calcul distribué \\ pour
-  Objective Caml\footnotetext{Ce
-    travail a été réalisé en partie dans le cadre du projet U3CAT
-    (\emph{Unification of Critical C Code Analysis Techniques},
-    ANR-08-SEGI-021).}}
-
-\author{Jean-Christophe Filli\^{a}tre$^{1,2,3}$ \& K. Kalyanasundaram$^3$}
-
-\titlehead{Functory}%  a droite (page impaire)
-
-\authorhead{Filliâtre \& Kalyanasundaram}% a gauche (page paire)
-
-\affiliation{\begin{tabular}{rr} 
-\\ 1:  CNRS / LRI UMR 8623 F-91405 Orsay
-\\ 2:  Université Paris Sud F-91405 Orsay
-\\ 3:  INRIA Saclay -- Île-de-France F-91893 Orsay
-\\ {\tt filliatr@lri.fr},  
-   {\tt kalyan.krishnamani@inria.fr}
-\end{tabular}}
-
-\begin{document}
-\setcounter{page}{1}
-\maketitle
-
-\begin{abstract}
+%% added from the other file - KK
+
+\selectlanguage{french}
+%%% Commandes spécifiques à l'article
+\newcommand{\cfsect}[1]{(\textit{cf.} section~\ref{#1})}
+\newcommand{\cfsectpage}[1]{(\textit{cf.} section~\ref{#1}, page~\pageref{#1})}
+\providecommand{\figureref}[1]{\figname~\ref{#1}}
+\providecommand{\cftab}[1]{(\textit{cf.} tableau~\ref{#1})}
+\newcommand{\cmd}[1]{{\upshape\texttt{\symbol{"5C}#1}}}
+%%% Fin des commandes spécifiques à l'article
+\newtheorem{thm}{Th\'{e}or\`{e}me}[section]
+\newtheorem{theorem}{Th\'{e}or\`{e}me}[section]
+\newtheorem{defi}{D\'{e}finition}[section]
+\newtheorem{lemme}{Lemme}[section]
+\newtheorem{cor}{Corollaire}[section]
+\newtheorem{proof}{Preuve}[section]
+\newtheorem{preuve}{Preuve}[section]
+\newtheorem{prop}{Proposition}[section]
+\newtheorem{ex}{Exemple}[section]
+\newtheorem{rem}{Remarque}[section]
+\newtheorem{procedure}{Proc\'edure}[section]
+\newenvironment{Boxedminipage}
+{\begin{Sbox}\begin{minipage}}{\end{minipage}\end{Sbox}\fbox{TheSbox}}
+\def\bibfmta#1#2#3#4{{\bf #1.} {#2}. {\em #3}, #4.}
+\journal{Studia Informatica Universalis.}{-}{-}
+
+%%%%%Add your package here
+%
+%\usepackage[option]{package}
+%
+
+%%%%%%%%%Preamble, to be modified by the authors
+%
+\title[ \textit{Functory}]  {Functory : Une bibliothèque de calcul distribué pour
+  Objective Caml}
+% \subtitle{Sub-Title}
+\author{Jean-Christophe Filliâtre \fup{1}\fup{2}\fup{3}, K. Kalyanasundaram \fup{3}}
+\address{
+\fup{1}
+CNRS / LRI UMR 8623 F-91405 Orsay\\
+filliatre@lri.fr \\ [3pt]
+\fup{2} Université Paris Sud F-91405 Orsay
+\\[3pt]
+\fup{3}INRIA Saclay -- Île-de-France F-91893 Orsay\\
+kalyan.krishnamani@inria.fr\\[3pt]
+}
+\resume{
   Cet article présente \functory, une bibliothèque de calcul distribué
@@
   montre son potentiel sur de nombreux exemples.
-\end{abstract}
+}
+\motscles{mapreduce, calcul distribué, OCaml}
+%%%%%%% End Preamble
+
+\begin{document}
+\maketitlepage
+
+%%%%%Document
 
@@
     let n = Array.length a in
-    let b = if n = 0 then [||] else Array.create n (f a.(0)) in
-    let tasks = ref [] in for i = 1 to n - 1 do tasks := (a.(i), i) :: !tasks done;
-    compute ~worker:f ~master:(fun (_,i) bi -> b.(i) <- bi; []) !tasks;
+    let b = if n = 0 then [| |] 
+      else Array.create n (f a.(0)) in
+    let tasks = 
+      ref [] in for i = 1 to n - 1 do 
+         tasks := (a.(i), i) :: !tasks 
+      done;
+    compute ~worker:f 
+            ~master:(fun (_,i) bi -> 
+              b.(i) <- bi; []) !tasks;
     b
@@
 val Worker.compute : ('a -> 'b) -> unit
-val Master.compute : ('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
+val Master.compute : ('a * 'c -> 'b -> ('a * 'c) list) -> 
+                     ('a * 'c) list -> unit
 \end{ocaml}
@@
 val Worker.compute : (string -> string) -> unit
-val Master.compute : (string * 'c -> string -> (string * 'c) list) ->
+val Master.compute : (string * 'c -> string -> 
+                     (string * 'c) list) ->
                      (string * 'c) list -> unit
@@
 \begin{ocaml}
-  val map_{local,remote}_fold : f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c   
+  val map_{local,remote}_fold : f:('a -> 'b) -> 
+                               fold:('c -> 'b -> 'c) -> 
+                               'c -> 'a list -> 'c   
 \end{ocaml}
@@
 \begin{ocaml}
-  val map_fold_{a,ac} : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
+  val map_fold_{a,ac} : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
+                        'b -> 'a list -> 'b 
 \end{ocaml}
@@
 \begin{ocaml}
-  let worker = function Map x -> f x | Fold (x, y) -> fold x y
+  let worker = function Map x -> f x | 
+                                 Fold (x, y) -> fold x y
 \end{ocaml}
@@
   let map_fold_a ~f ~fold acc l =
-    let tasks = let i = ref 0 in List.map (fun x -> incr i; Map x, (!i, !i)) l in
+    let tasks = let i = ref 0 in 
+         List.map (fun x -> incr i; 
+                   Map x, (!i, !i)) l in
     let results = Hashtbl.create 17 in 
-    let merge i j r = ...traiter le résultat r du segment [i,j]... in
-    let worker = function Map x -> f x | Fold (x, y) -> fold x y in
-    let master x r = match x with Map _,(i,_) -> merge i i r | Fold _,(i,j) -> merge i j r in
+    let merge i j r = 
+            ... traiter le résultat r du segment [i,j] ... in
+    let worker = function Map x -> 
+                 f x | Fold (x, y) -> fold x y in
+    let master x r = match x with Map _,(i,_) -> 
+             merge i i r | Fold _,(i,j) -> merge i j r in
     compute ~worker ~master tasks;
-    try let _,_,r = Hashtbl.find results 1 in r with Not_found -> acc
+    try let _,_,r = Hashtbl.find results 1 in r with 
+                    Not_found -> acc
 \end{ocaml}
@@
 \of{map_fold} ci-dessus.
-
-% TODO : insister sur la simplicité de Same, pour les petits programmes
-% notamment
 
@@
 
-\paragraph{Remerciements.}
+\remerciements{
 Les auteurs remercient Alain Mebsout et Johannes Kanig pour avoir testé
@@
 Marché pour sa relecture attentive de cet article.
-
-%\vfill\pagebreak
-\bibliographystyle{plain}
+}
+
+%\subsubsection{Annexe pour le service de fabrication~: commande \cmd{publisher}}
+
+%Pour la remise finale des articles, il nous faut une version abrégée du titre courant
+%(normalement déjà placée dans l'en-tête des pages impaires) et les coordonnées
+%des auteurs pour les joindre si nécessaire. La commande \cmd{publisher} qui se
+%place au début ou à la fin de l'article fait apparaître en dernière page
+%l'annexe demandée, avec un report des informations déjà connues comme le titre
+%(version longue et abrégée), les noms d'auteurs, les références du logiciel
+%utilisé. Il ne manque que les coordonnées (numéro de téléphone, numéro de fax
+%et e-mail) pour construire cette page.
+
+%BIBLIOGRAPHIE
+\bibliographystyle{alpha}
 \bibliography{./biblio}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}
-
-% LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables OCamlMPI MapReduce généricité marshalling
-
-%%% Local Variables: 
-%%% mode: latex
-%%% ispell-local-dictionary: "francais"
-%%% TeX-master: t
-%%% End:
-
-% LocalWords:  sérialisation boutisme inatteignables observationnellement
-% LocalWords:  sous-modules sérialisations n'apportent d'octets d'enchaîner
-% LocalWords:  n'offre l'inclure implémente implémentation d'exploiter
-% LocalWords:  l'affectation d'itérations

### 2010-12-09 doc/jfla-2011/jfla_2011_final_style/main.tex
*** empty log message ***

@@
 \newcommand{\Ocaml}{OCaml}
-\newcommand{\functory}{\textsf{Functory}}
+\newcommand{\functory}{\textsc{Functory}}
 \newcommand{\JoCaml}{Jo{\&\!}Caml}
@@
 
-\paragraph{Calcul distribué.}
+\paragraph*{\bf Calcul distribué}
 Initialement inspirée par la bibliothèque
@@
 
-\paragraph{Une approche fonctionnelle.}
+\paragraph*{\bf Une approche fonctionnelle}
 Nous avons essayé de tirer partie des spécificités de la programmation
@@
 
-\paragraph{Fonctions dérivées.}\label{sec:derived}
+\paragraph*{\bf Fonctions dérivées}\label{sec:derived}
 Dans de nombreuses situations, la mise en \oeuvre\ la plus simple du
@@
 
-\paragraph{N-reines.}\label{sec:n-queens}
+\paragraph*{\bf N-reines}\label{sec:n-queens}
 Il s'agit du problème classique consistant à calculer le nombre de
@@
 
-\paragraph{L'ensemble de Mandelbrot.}
+\paragraph*{\bf L'ensemble de Mandelbrot}
 Dessiner l'ensemble de Mandelbrot est un autre exemple de calcul
@@
 
-\paragraph{Démonstrateurs automatiques.}\label{sec:SMT}
+\paragraph*{\bf Démonstrateurs automatiques}\label{sec:SMT}
 Le dernier exemple correspond à celui mentionné dans l'introduction.
@@
 
-\paragraph{Protocole.}\label{sec:protocol}
+\paragraph*{\bf Protocole}\label{sec:protocol}
 Le protocole utilisé comporte sept messages différents.
@@
 
-\subsubsection{Tolérance aux pannes}\label{sec:fault}
+\paragraph*{\bf Tolérance aux pannes}\label{sec:fault}
 
@@
 
-\paragraph{N-reines.}
+\paragraph*{\bf N-reines}
 La table suivante montre les temps d'exécution pour différentes valeurs
@@
 
-\paragraph{L'ensemble de Mandelbrot.}
+\paragraph*{\bf L'ensemble de Mandelbrot}
 Dans cette expérience, on dessine le fragment rectangulaire de l'ensemble de
@@
 
-\paragraph{Multiplication de matrices}
+\paragraph*{\bf Multiplication de matrices}
 Cette expérience a été inspirée par le concours de programmation
@@
 
-\paragraph{Démonstrateur automatiques.}
+\paragraph*{\bf Démonstrateur automatiques}
 Comme expliqué plus haut, cette expérience consiste à vérifier la
@@
 
-\paragraph{Travaux connexes.}
+\paragraph*{\bf Travaux connexes}
 Une bibliothèque relativement proche de la nôtre est l'implémentation
@@
 
-\paragraph{Perspectives.}
+\paragraph*{\bf Perspectives}
 La bibliothèque \functory\ pourrait être enrichie de nombreuses façons.
@@
 plus de travail. Une possibilité serait de réutiliser une partie de
-l'outil \textsf{Ocamlviz}~\cite{ocamlviz}.
+l'outil \textsc{Ocamlviz}~\cite{ocamlviz}.
 
@@
 Marché pour sa relecture attentive de cet article.
+Les auteurs remercient également les rapporteurs pour
+leurs commentaires, précis et pertinents.
 }

### 2010-12-09 doc/jfla-2011/jfla_2011_final_style/main-pp.tex
*** empty log message ***

--- 2010-12-09 doc/jfla-2011/jfla_2011_final_style/main.tex
@@
 
+\bigskip
 %%%%%Add your package here
@@
 
+\bigskip
 \begin{document}
@@
 
+\bigskip
 %%%%%Document
@@
 une fonction potentiellement polymorphe
-\begin{ocaml}
-  worker: 'a -> 'b
-\end{ocaml}
-où \of{'a} dénote le type des tâches et \of{'b} le type des résultats.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}worker\symbol{58}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}
+}
+
+}\noindent
+où \textsf{\ensuremath{\alpha}} dénote le type des tâches et \textsf{\ensuremath{\beta}} le type des résultats.
 Le patron est pour sa part une fonction à laquelle on passe 
@@
 la liste des tâches initiales~:
-\begin{ocaml}
-  master: ('a -> 'b -> 'a list) -> 'a list -> unit
-\end{ocaml}
-Cette fonction passée en argument à \of{master} est appliquée dès
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}master\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
+}
+
+}\noindent
+Cette fonction passée en argument à \textsf{master} est appliquée dès
 qu'un nouveau résultat est disponible. Elle peut produire de nouvelles
-tâches (d'où son type de retour \of{'a list}), qui s'ajoutent alors à
+tâches (d'où son type de retour \textsf{\ensuremath{\alpha}\hspace*{1.22ex}list}), qui s'ajoutent alors à
 la liste des tâches à effectuer.
@@
 Cette section décrit l'interface de la bibliothèque \functory.
-La fonctionnalité primitive est une fonction \of{compute} réalisant
+La fonctionnalité primitive est une fonction \textsf{compute} réalisant
 l'idée principale évoquée dans l'introduction.
-\begin{ocaml}
-  val compute : 
-    worker:('a -> 'b) -> 
-    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
-    ('a * 'c) list -> unit
-\end{ocaml}
-Les tâches sont des paires, de type \of{'a * 'c}, où la première
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{val}\hspace*{1.22ex}compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}worker\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}master\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
+}
+
+}\noindent
+Les tâches sont des paires, de type \textsf{\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}}, où la première
 composante sera transmise à l'ouvrier et la seconde conservée
-par le patron. La fonction \of{worker} doit être 
+par le patron. La fonction \textsf{worker} doit être 
 pure\footnote{On entend ici \emph{observationnellement pure} mais on
@@
 et sera exécutée en parallèle par tous les ouvriers. La fonction
-\of{master}, au contraire, peut être impure et sera exécutée
+\textsf{master}, au contraire, peut être impure et sera exécutée
 uniquement au sein d'un unique processus séquentiel.
-Cette fonction \of{master} accumule typiquement les résultats renvoyés
+Cette fonction \textsf{master} accumule typiquement les résultats renvoyés
 par les ouvriers dans une structure locale. Elle peut en outre
 produire de nouvelles tâches, 
-sous la forme d'une liste de type \of{('a * 'c) list}, 
+sous la forme d'une liste de type \textsf{(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list}, 
 qui sont alors ajoutées aux tâches restant à effectuer.
-Le troisième argument de \of{compute} est la liste des tâches
+Le troisième argument de \textsf{compute} est la liste des tâches
 initiales, qui déclenchent le calcul.
-La fonction \of{compute} rend la main lorsque toutes les tâches ont
+La fonction \textsf{compute} rend la main lorsque toutes les tâches ont
 été effectuées. Il n'y a pas de résultat renvoyé, mais uniquement des
-effets de bord de la fonction \of{master}.
+effets de bord de la fonction \textsf{master}.
 
 À titre d'illustration, voici comme écrire un équivalent distribué de
-la fonction \of{Array.map}, qui applique une fonction \of{f} à tous
-les éléments d'un tableau \of{a}. 
-\begin{ocaml}
-  let array_map f a =
-    let n = Array.length a in
-    let b = if n = 0 then [| |] 
-      else Array.create n (f a.(0)) in
-    let tasks = 
-      ref [] in for i = 1 to n - 1 do 
-         tasks := (a.(i), i) :: !tasks 
-      done;
-    compute ~worker:f 
-            ~master:(fun (_,i) bi -> 
-              b.(i) <- bi; []) !tasks;
-    b
-\end{ocaml}
-La liste des tâches, \of{tasks}, contient 
-les couples \of{(a.(i), i)} pour tous les indices \of{i} du tableau. La
-fonction \of{worker} se réduit à la fonction \of{f}. La fonction
-\of{master} reçoit un résultat \of{bi} correspondant à un indice
-\of{i} et le stocke dans le tableau résultat \of{b}.
+la fonction \textsf{Array.map}, qui applique une fonction \textsf{f} à tous
+les éléments d'un tableau \textsf{a}.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}array\_{}map\hspace*{1.22ex}f\hspace*{1.22ex}a\hspace*{1.22ex}=~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}n\hspace*{1.22ex}=\hspace*{1.22ex}Array.length\hspace*{1.22ex}a\hspace*{1.22ex}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}b\hspace*{1.22ex}=\hspace*{1.22ex}\textbf{if}\hspace*{1.22ex}n\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{then}\hspace*{1.22ex}[\ensuremath{|}\ensuremath{|}]\hspace*{1.22ex}\textbf{else}\hspace*{1.22ex}Array.create\hspace*{1.22ex}n\hspace*{1.22ex}(f\hspace*{1.22ex}a.(0))\hspace*{1.22ex}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}ref\hspace*{1.22ex}[]\hspace*{1.22ex}\textbf{in}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{for}\hspace*{1.22ex}i\hspace*{1.22ex}=\hspace*{1.22ex}1\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}n\hspace*{1.22ex}-\hspace*{1.22ex}1\hspace*{1.22ex}\textbf{do}\hspace*{1.22ex}tasks\hspace*{1.22ex}\ensuremath{:=}\hspace*{1.22ex}(a.(i),\hspace*{1.22ex}i)\hspace*{1.22ex}\symbol{58}\symbol{58}\hspace*{1.22ex}!tasks\hspace*{1.22ex}\textbf{done}\symbol{59}~\linebreak
+\noindent\hspace*{2.00em}compute\hspace*{1.22ex}\symbol{126}worker\symbol{58}f\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{6.50em}\symbol{126}master\symbol{58}(\textbf{fun}\hspace*{1.22ex}(\_{},i)\hspace*{1.22ex}bi\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}b.(i)\hspace*{1.22ex}\ensuremath{\leftarrow}\hspace*{1.22ex}bi\symbol{59}\hspace*{1.22ex}[])\hspace*{1.22ex}!tasks\symbol{59}~\linebreak
+\noindent\hspace*{2.00em}b
+}
+
+}\noindent
+La liste des tâches, \textsf{tasks}, contient 
+les couples \textsf{(a.(i),\hspace*{1.22ex}i)} pour tous les indices \textsf{i} du tableau. La
+fonction \textsf{worker} se réduit à la fonction \textsf{f}. La fonction
+\textsf{master} reçoit un résultat \textsf{bi} correspondant à un indice
+\textsf{i} et le stocke dans le tableau résultat \textsf{b}.
 La fin de cette section décrit plusieurs telles
 fonctions d'ordre supérieur fournies par \functory, toutes dérivées de
-la fonction \of{compute}.
+la fonction \textsf{compute}.
 
 En réalité, la bibliothèque \functory\ fournit \emph{cinq}
-mises en {\oe}uvre différentes de la fonction \of{compute}, correspondant à
+mises en {\oe}uvre différentes de la fonction \textsf{compute}, correspondant à
 cinq contextes d'utilisation différents.
@@
   Selon que le programme est exécuté comme le patron ou comme un
-  ouvrier, les arguments pertinents de la fonction \of{compute} sont utilisés.
+  ouvrier, les arguments pertinents de la fonction \textsf{compute} sont utilisés.
 
@@
   encore sérialiser des valeurs polymorphes.
-  En conséquence, la fonction \of{compute} est présentée sous la forme
+  En conséquence, la fonction \textsf{compute} est présentée sous la forme
   de deux fonctions, servant respectivement à réaliser le patron et
   les ouvriers :
-\vspace{-0.5em}
-\begin{ocaml}
-val Worker.compute : ('a -> 'b) -> unit
-val Master.compute : ('a * 'c -> 'b -> ('a * 'c) list) -> 
-                     ('a * 'c) list -> unit
-\end{ocaml}
+%\vspace{-0.5em}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{0.00em}\textbf{val}\hspace*{1.22ex}Worker.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit~\linebreak
+\noindent\hspace*{0.00em}\textbf{val}\hspace*{1.22ex}Master.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{10.50em}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
+}
+
+}\noindent
 
@@
     d'\Ocaml\ :} Il n'est plus possible d'utiliser la sérialisation
-  d'\Ocaml\ et la fonction \of{compute} est présentée sous la forme
+  d'\Ocaml\ et la fonction \textsf{compute} est présentée sous la forme
   de deux fonctions ne manipulant plus que des chaînes de caractères :%
-\vspace{-0.5em}
-\begin{ocaml}
-val Worker.compute : (string -> string) -> unit
-val Master.compute : (string * 'c -> string -> 
-                     (string * 'c) list) ->
-                     (string * 'c) list -> unit
-\end{ocaml}
+%\vspace{-0.5em}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{0.00em}\textbf{val}\hspace*{1.22ex}Worker.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(string\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}string)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit~\linebreak
+\noindent\hspace*{0.00em}\textbf{val}\hspace*{1.22ex}Master.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(string\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}string\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{10.50em}(string\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}~\linebreak
+\noindent\hspace*{10.50em}(string\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
+}
+
+}\noindent
 \end{enumerate}
 La bibliothèque \functory\ est donc organisée en trois modules :
-\of{Sequential} pour le calcul purement séquentiel ; \of{Cores} pour
+\textsf{Sequential} pour le calcul purement séquentiel ; \textsf{Cores} pour
 le calcul distribué sur plusieurs \coeurs\ d'une même machine ; et
-enfin \of{Network} pour le calcul en réseau. Ce dernier module
-comporte trois sous-modules, appelés respectivement \of{Same},
-\of{Poly} and \of{Mono}, correspondant aux situations 3, 4 et 5 ci-dessus.
+enfin \textsf{Network} pour le calcul en réseau. Ce dernier module
+comporte trois sous-modules, appelés respectivement \textsf{Same},
+\textsf{Poly} and \textsf{Mono}, correspondant aux situations 3, 4 et 5 ci-dessus.
 
@@
 offrant du calcul parallèle sur des listes, toutes dérivées de la fonction
-\of{compute}. En particulier, elles sont disponibles dans les cinq
+\textsf{compute}. En particulier, elles sont disponibles dans les cinq
 modules décrits ci-dessus.
@@
 fonction à tous les éléments d'une liste, analogue à la fonction
-\of{array_map} décrite plus haut, c'est-à-dire
-\begin{ocaml}
-  val map : ('a -> 'b) -> 'a list -> 'b list
-\end{ocaml}
-Plus subtilement, on peut combiner une fonction \of{f:'a -> 'b}
-avec une fonction \of{fold:'c -> 'b -> 'c} pour calculer, à partir d'une liste
+\textsf{array\_{}map} décrite plus haut, c'est-à-dire
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{val}\hspace*{1.22ex}map\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}list
+}
+
+}\noindent
+Plus subtilement, on peut combiner une fonction \textsf{f\symbol{58}\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}}
+avec une fonction \textsf{fold\symbol{58}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}} pour calculer, à partir d'une liste
 $l$  et d'un accumulateur initial $a$, la valeur finale
 \begin{equation}\label{eq:map-fold}
-  \of{fold} ~ ... ~ (\of{fold} ~ (\of{fold} ~ a ~ (\of{f} ~ x_1)) ~
-  (\of{f} ~ x_2)) ~ ... ~ (\of{f} ~ x_n)
+  \textsf{fold} ~ ... ~ (\textsf{fold} ~ (\textsf{fold} ~ a ~ (\textsf{f} ~ x_1)) ~
+  (\textsf{f} ~ x_2)) ~ ... ~ (\textsf{f} ~ x_n)
 \end{equation}
@@
 de la liste $l$.
-On peut alors distinguer deux cas, selon que l'opération \of{fold} est
-beaucoup moins coûteuse que l'opération \of{f}, et peut être
+On peut alors distinguer deux cas, selon que l'opération \textsf{fold} est
+beaucoup moins coûteuse que l'opération \textsf{f}, et peut être
 effectuée localement par le patron, ou qu'au contraire elle peut être
 coûteuse et a donc intérêt à être effectuée en parallèle des
-opérations \of{f}. La bibliothèque fournit donc deux fonctions,
+opérations \textsf{f}. La bibliothèque fournit donc deux fonctions,
 correspondant à ces deux cas de figure :
-\begin{ocaml}
-  val map_{local,remote}_fold : f:('a -> 'b) -> 
-                               fold:('c -> 'b -> 'c) -> 
-                               'c -> 'a list -> 'c   
-\end{ocaml}
-
-Dans le cas de \of{map_remote_fold}, une seule opération \of{fold}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{val}\hspace*{1.22ex}map\_{}\{local,remote\}\_{}fold\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}fold\symbol{58}(\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\hspace*{1.22ex}\hspace*{1.22ex}
+}
+
+}\noindent
+
+Dans le cas de \textsf{map\_{}remote\_{}fold}, une seule opération \textsf{fold}
 peut être effectuée à la fois (possiblement en parallèle d'opérations
-\of{f}), comme le montre l'équation~(\ref{eq:map-fold}).
+\textsf{f}), comme le montre l'équation~(\ref{eq:map-fold}).
 Il existe cependant des situations dans lesquelles plusieurs
-opérations \of{fold} peuvent être effectuées en parallèle, dès que des
-résultats intermédiaires de \of{f} sont disponibles. C'est le cas
-notamment lorsque l'opération \of{fold} est associative (ce qui
-implique que les types \of{'b} et \of{'c} sont égaux).
-Lorsque l'opération \of{fold} est de plus commutative, on peut
-effectuer encore plus d'opérations \of{fold} en parallèle.
+opérations \textsf{fold} peuvent être effectuées en parallèle, dès que des
+résultats intermédiaires de \textsf{f} sont disponibles. C'est le cas
+notamment lorsque l'opération \textsf{fold} est associative (ce qui
+implique que les types \textsf{\ensuremath{\beta}} et \textsf{\ensuremath{\gamma}} sont égaux).
+Lorsque l'opération \textsf{fold} est de plus commutative, on peut
+effectuer encore plus d'opérations \textsf{fold} en parallèle.
 Notre interface fournit donc deux autres fonctions pour ces cas
 particuliers :
-\begin{ocaml}
-  val map_fold_{a,ac} : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-                        'b -> 'a list -> 'b 
-\end{ocaml}
-Ces cinq fonctions se dérivent facilement de la fonction \of{compute}.
-Voici par exemple comment réaliser la fonction \of{map_fold_a}.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{val}\hspace*{1.22ex}map\_{}fold\_{}\{a,ac\}\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}fold\symbol{58}(\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}
+}
+
+}\noindent
+Ces cinq fonctions se dérivent facilement de la fonction \textsf{compute}.
+Voici par exemple comment réaliser la fonction \textsf{map\_{}fold\_{}a}.
 On commence par introduire un type somme pour distinguer les tâches
-\of{f} des tâches \of{fold}, c'est-à-dire
-\of{type ('a, 'b) map_or_fold = Map of 'a | Fold of 'b}.
+\textsf{f} des tâches \textsf{fold}, c'est-à-dire
+\textsf{\textbf{type}\hspace*{1.22ex}(\ensuremath{\alpha},\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}map\_{}or\_{}fold\hspace*{1.22ex}=\hspace*{1.22ex}Map\hspace*{1.22ex}\textbf{of}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{|}\hspace*{1.22ex}Fold\hspace*{1.22ex}\textbf{of}\hspace*{1.22ex}\ensuremath{\beta}}.
 Le code de l'ouvrier se résumé à distinguer les deux types de tâches.
-\begin{ocaml}
-  let worker = function Map x -> f x | 
-                                 Fold (x, y) -> fold x y
-\end{ocaml}
-Le code du patron est plus compliqué. En effet, l'opération \of{fold}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}worker\hspace*{1.22ex}=\hspace*{1.22ex}\textbf{function}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\ensuremath{|}\hspace*{1.22ex}Map\hspace*{1.22ex}x\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}f\hspace*{1.22ex}x\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\ensuremath{|}\hspace*{1.22ex}Fold\hspace*{1.22ex}(x,\hspace*{1.22ex}y)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}fold\hspace*{1.22ex}x\hspace*{1.22ex}y
+}
+
+}\noindent
+Le code du patron est plus compliqué. En effet, l'opération \textsf{fold}
 étant seulement associative, on ne peut calculer d'expression analogue
@@
 $[x_i,\dots,x_j]$ de $l$ pour lesquels la valeur
-$$\of{f}~x_i \oplus \dots \oplus \of{f}~x_j$$ a déjà été calculée
-(en notant ici $\of{fold}$ par $\oplus$), ainsi que cette valeur.
+$$\textsf{f}~x_i \oplus \dots \oplus \textsf{f}~x_j$$ a déjà été calculée
+(en notant ici $\textsf{fold}$ par $\oplus$), ainsi que cette valeur.
 Cete table est doublement indexée, par $i$ et par $j$. Initialement
 cette table est vide et la liste des tâches ne contient que des tâches
-de type \of{Map}, une par élément de $l$.
+de type \textsf{Map}, une par élément de $l$.
 De manière générale, chaque tâche est associée au couple $(i,j)$
@@
 que l'on extrait grâce à la clé $i-1$, et on produit alors une
-nouvelle tâche \of{Fold} pour calculer la valeur correspondant à
+nouvelle tâche \textsf{Fold} pour calculer la valeur correspondant à
 $[x_k,\dots, x_j]$, après avoir supprimé de la table les entrées pour
@@
 choisit arbitrairement, car il y a de toutes façons deux tâches
-\of{Fold} à effectuer.)
+\textsf{Fold} à effectuer.)
 Soit enfin il n'y a ni entrée pour $i-1$ ni entrée pour
@@
 clés $1$ et $n$, sauf dans le cas $n=0$ où aucun calcul n'est nécessaire.
-Au final, la fonction \of{map_fold_a} prend la forme suivante, où
+Au final, la fonction \textsf{map\_{}fold\_{}a} prend la forme suivante, où
 on a seulement omis le code de fusion des résultats expliqué ci-dessus.
-\begin{ocaml}
-  let map_fold_a ~f ~fold acc l =
-    let tasks = let i = ref 0 in 
-         List.map (fun x -> incr i; 
-                   Map x, (!i, !i)) l in
-    let results = Hashtbl.create 17 in 
-    let merge i j r = 
-            ... traiter le résultat r du segment [i,j] ... in
-    let worker = function Map x -> 
-                 f x | Fold (x, y) -> fold x y in
-    let master x r = match x with Map _,(i,_) -> 
-             merge i i r | Fold _,(i,j) -> merge i j r in
-    compute ~worker ~master tasks;
-    try let _,_,r = Hashtbl.find results 1 in r with 
-                    Not_found -> acc
-\end{ocaml}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}map\_{}fold\_{}a\hspace*{1.22ex}\symbol{126}f\hspace*{1.22ex}\symbol{126}fold\hspace*{1.22ex}acc\hspace*{1.22ex}l\hspace*{1.22ex}=~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{3.00em}\textbf{let}\hspace*{1.22ex}i\hspace*{1.22ex}=\hspace*{1.22ex}ref\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{in}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{3.00em}List.map\hspace*{1.22ex}(\textbf{fun}\hspace*{1.22ex}x\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}incr\hspace*{1.22ex}i\symbol{59}\hspace*{1.22ex}Map\hspace*{1.22ex}x,\hspace*{1.22ex}(!i,\hspace*{1.22ex}!i))\hspace*{1.22ex}l\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}results\hspace*{1.22ex}=\hspace*{1.22ex}Hashtbl.create\hspace*{1.22ex}17\hspace*{1.22ex}\textbf{in}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}merge\hspace*{1.22ex}i\hspace*{1.22ex}j\hspace*{1.22ex}r\hspace*{1.22ex}=\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{6.00em}...\hspace*{1.22ex}traiter\hspace*{1.22ex}le\hspace*{1.22ex}résultat\hspace*{1.22ex}r\hspace*{1.22ex}du\hspace*{1.22ex}segment\hspace*{1.22ex}[i,j]\hspace*{1.22ex}...\hspace*{1.22ex}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}worker\hspace*{1.22ex}=\hspace*{1.22ex}\textbf{function}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{3.00em}\ensuremath{|}\hspace*{1.22ex}Map\hspace*{1.22ex}x\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}f\hspace*{1.22ex}x\hspace*{1.22ex}\ensuremath{|}\hspace*{1.22ex}Fold\hspace*{1.22ex}(x,\hspace*{1.22ex}y)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}fold\hspace*{1.22ex}x\hspace*{1.22ex}y\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}master\hspace*{1.22ex}x\hspace*{1.22ex}r\hspace*{1.22ex}=\hspace*{1.22ex}\textbf{match}\hspace*{1.22ex}x\hspace*{1.22ex}\textbf{with}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{3.00em}\ensuremath{|}\hspace*{1.22ex}Map\hspace*{1.22ex}\_{},(i,\_{})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}merge\hspace*{1.22ex}i\hspace*{1.22ex}i\hspace*{1.22ex}r\hspace*{1.22ex}\ensuremath{|}\hspace*{1.22ex}Fold\hspace*{1.22ex}\_{},(i,j)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}merge\hspace*{1.22ex}i\hspace*{1.22ex}j\hspace*{1.22ex}r\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}compute\hspace*{1.22ex}\symbol{126}worker\hspace*{1.22ex}\symbol{126}master\hspace*{1.22ex}tasks\symbol{59}~\linebreak
+\noindent\hspace*{2.00em}\textbf{try}\hspace*{1.22ex}\textbf{let}\hspace*{1.22ex}\_{},\_{},r\hspace*{1.22ex}=\hspace*{1.22ex}Hashtbl.find\hspace*{1.22ex}results\hspace*{1.22ex}1\hspace*{1.22ex}\textbf{in}\hspace*{1.22ex}r\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{with}\hspace*{1.22ex}Not\_{}found\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}acc
+}
+
+}\noindent
 
@@
 comme par exemple une variante
-où seule la fonction \of{fold} est significative.
+où seule la fonction \textsf{fold} est significative.
 L'utilisateur de \functory\ ne devrait pas avoir de mal à les dériver
-lui-même de la fonction \of{compute} ou des fonctions du type
-\of{map_fold} ci-dessus.
-
+lui-même de la fonction \textsf{compute} ou des fonctions du type
+\textsf{map\_{}fold} ci-dessus.
+
+\bigskip
 \section{Études de cas}\label{sec:studies}
@@
 Comme premier exemple, on considère la multiplication de deux matrices
-\of{a} et \of{b}, de tailles respectives
-$\of{n}\times\of{p}$ et $\of{p}\times\of{m}$. 
+\textsf{a} et \textsf{b}, de tailles respectives
+$\textsf{n}\times\textsf{p}$ et $\textsf{p}\times\textsf{m}$. 
 Le résultat sera stocké dans une matrice
-\of{c} de taille $\of{n}\times\of{m}$.
-On suppose que \of{a}, \of{b} et \of{c} sont contenues dans des
+\textsf{c} de taille $\textsf{n}\times\textsf{m}$.
+On suppose que \textsf{a}, \textsf{b} et \textsf{c} sont contenues dans des
 variables globales.
-En supposant en outre \of{a} organisée en lignes et \of{b} en colonnes, une
+En supposant en outre \textsf{a} organisée en lignes et \textsf{b} en colonnes, une
 multiplication séquentielle s'écrirait ainsi
-\begin{ocaml}
-  for i = 0 to n-1 do
-    for j = 0 to m-1 do
-      for k = 0 to p-1 do
-	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k) (* b en colonnes *)
-      done
-    done
-  done  
-\end{ocaml}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{for}\hspace*{1.22ex}i\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}n-1\hspace*{1.22ex}\textbf{do}~\linebreak
+\noindent\hspace*{2.00em}\textbf{for}\hspace*{1.22ex}j\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}m-1\hspace*{1.22ex}\textbf{do}~\linebreak
+\noindent\hspace*{3.00em}\textbf{for}\hspace*{1.22ex}k\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}p-1\hspace*{1.22ex}\textbf{do}~\linebreak
+\noindent\hspace*{4.00em}c.(i).(j)\hspace*{1.22ex}\ensuremath{\leftarrow}\hspace*{1.22ex}c.(i).(j)\hspace*{1.22ex}+\hspace*{1.22ex}a.(i).(k)\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}b.(j).(k)\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{9.00em}{(*\hspace*{1.22ex}attention\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}b\hspace*{1.22ex}est\hspace*{1.22ex}en\hspace*{1.22ex}colonnes\hspace*{1.22ex}*)}~\linebreak
+\noindent\hspace*{3.00em}\textbf{done}~\linebreak
+\noindent\hspace*{2.00em}\textbf{done}~\linebreak
+\noindent\hspace*{1.00em}\textbf{done}\hspace*{1.22ex}\hspace*{1.22ex}
+}
+
+}\noindent
 l'addition et la multiplication des coefficients étant notées
 respectivement $+$ et $\times$.
-La complexité de ce code est clairement $O(\of{n}\times\of{m}\times\of{p})$,
+La complexité de ce code est clairement $O(\textsf{n}\times\textsf{m}\times\textsf{p})$,
 
 Une manière évidente de distribuer ce calcul consiste à faire de la
-boucle interne sur \of{k} une tâche élémentaire.
-On construit alors une liste de $\of{n} \times \of{m}$ tâches, de la
+boucle interne sur \textsf{k} une tâche élémentaire.
+On construit alors une liste de $\textsf{n} \times \textsf{m}$ tâches, de la
 manière suivante :
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in
-    for i = 0 to n-1 do for j = 0 to m-1 do
-      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
-    done done;
-    !l
-\end{ocaml}
-Chaque tâche est une paire formée d'une part d'une ligne \of{a.(i)} et
-d'une colonne \of{b.(j)}, et d'autre part d'une position \of{(i,j)}.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}l\hspace*{1.22ex}=\hspace*{1.22ex}ref\hspace*{1.22ex}[]\hspace*{1.22ex}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{for}\hspace*{1.22ex}i\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}n-1\hspace*{1.22ex}\textbf{do}\hspace*{1.22ex}\textbf{for}\hspace*{1.22ex}j\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}m-1\hspace*{1.22ex}\textbf{do}~\linebreak
+\noindent\hspace*{3.00em}tasks\hspace*{1.22ex}\ensuremath{:=}\hspace*{1.22ex}((a.(i),\hspace*{1.22ex}b.(j)),\hspace*{1.22ex}(i,j))\hspace*{1.22ex}\symbol{58}\symbol{58}\hspace*{1.22ex}!tasks~\linebreak
+\noindent\hspace*{2.00em}\textbf{done}\hspace*{1.22ex}\textbf{done}\symbol{59}~\linebreak
+\noindent\hspace*{2.00em}!l
+}
+
+}\noindent
+Chaque tâche est une paire formée d'une part d'une ligne \textsf{a.(i)} et
+d'une colonne \textsf{b.(j)}, et d'autre part d'une position \textsf{(i,j)}.
 L'ouvrier reçoit la première composante et calcule le produit scalaire.
-\begin{ocaml}
-  let worker (ai, bj) =
-    let c = ref 0 in
-    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
-    !c
-\end{ocaml}
-La patron est une fonction d'une ligne, qui reçoit le résultat \of{r}
-renvoyé par l'ouvrier et remplit la matrice \of{c}, en fonction de la
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}worker\hspace*{1.22ex}(ai,\hspace*{1.22ex}bj)\hspace*{1.22ex}=~\linebreak
+\noindent\hspace*{2.00em}\textbf{let}\hspace*{1.22ex}c\hspace*{1.22ex}=\hspace*{1.22ex}ref\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{in}~\linebreak
+\noindent\hspace*{2.00em}\textbf{for}\hspace*{1.22ex}k\hspace*{1.22ex}=\hspace*{1.22ex}0\hspace*{1.22ex}\textbf{to}\hspace*{1.22ex}p-1\hspace*{1.22ex}\textbf{do}\hspace*{1.22ex}c\hspace*{1.22ex}\ensuremath{:=}\hspace*{1.22ex}!c\hspace*{1.22ex}+\hspace*{1.22ex}ai.(k)\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}bj.(k)\hspace*{1.22ex}\textbf{done}\symbol{59}~\linebreak
+\noindent\hspace*{2.00em}!c
+}
+
+}\noindent
+La patron est une fonction d'une ligne, qui reçoit le résultat \textsf{r}
+renvoyé par l'ouvrier et remplit la matrice \textsf{c}, en fonction de la
 position contenue dans la seconde composante de la tâche. Aucune 
 nouvelle tâche n'est produite.
-\begin{ocaml}
-  let master (_, (i,j)) r = c.(i).(j) <- r; []
-\end{ocaml}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}master\hspace*{1.22ex}(\_{},\hspace*{1.22ex}(i,j))\hspace*{1.22ex}r\hspace*{1.22ex}=\hspace*{1.22ex}c.(i).(j)\hspace*{1.22ex}\ensuremath{\leftarrow}\hspace*{1.22ex}r\symbol{59}\hspace*{1.22ex}[]
+}
+
+}\noindent
 Au final, l'ensemble du calcul est lancé par un appel à la fonction
-\of{compute}.
-\begin{ocaml}
-  let () = compute ~worker ~master !tasks
-\end{ocaml}
-où les variables \of{'a}, \of{'b} et \of{'c} du type de
-\of{compute} sont respectivement instanciées par 
-\of{coeff array * coeff array}, \of{coeff} et \of{int * int}.
+\textsf{compute}.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}compute\hspace*{1.22ex}\symbol{126}worker\hspace*{1.22ex}\symbol{126}master\hspace*{1.22ex}!tasks
+}
+
+}\noindent
+où les variables \textsf{\ensuremath{\alpha}}, \textsf{\ensuremath{\beta}} et \textsf{\ensuremath{\gamma}} du type de
+\textsf{compute} sont respectivement instanciées par 
+\textsf{coeff\hspace*{1.22ex}array\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}coeff\hspace*{1.22ex}array}, \textsf{coeff} et \textsf{int\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}int}.
 
 Utiliser la version séquentielle de \functory\ est aussi simple que
-d'ajouter \of{open Sequential} au début du code. Le calcul est alors
+d'ajouter \textsf{\textbf{open}\hspace*{1.22ex}Sequential} au début du code. Le calcul est alors
 similaire à la multiplication usuelle donnée plus haut. Cela peut être
@@
 pour effectuer le calcul. Il suffit alors de remplacer la ligne
-\of{open Sequential} par les deux lignes suivantes.
-\begin{ocaml}
-  open Cores
-  let () = set_number_of_cores 4
-\end{ocaml}
+\textsf{\textbf{open}\hspace*{1.22ex}Sequential} par les deux lignes suivantes.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Cores~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}set\_{}number\_{}of\_{}cores\hspace*{1.22ex}4
+}
+
+}\noindent
 Le reste du code est inchangé.
@@
 Il suffit de remplacer les deux lignes de code ci-dessus par les suivantes.
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Same
-\end{ocaml}
-On utilise ici le sous-module \of{Same} de \of{Network}, qui permet
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Network~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}declare\_{}workers\hspace*{1.22ex}\symbol{126}n\symbol{58}4\hspace*{1.22ex}"orcus"~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}declare\_{}workers\hspace*{1.22ex}\symbol{126}n\symbol{58}8\hspace*{1.22ex}"belzebuth"~\linebreak
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Same
+}
+
+}\noindent
+On utilise ici le sous-module \textsf{Same} de \textsf{Network}, qui permet
 d'utiliser le même exécutable pour le patron et les ouvriers. Ce
-module fournit toujours une fonction \of{compute} de même signature
-que dans les modules \of{Sequential} et \of{Cores}, et le reste du
+module fournit toujours une fonction \textsf{compute} de même signature
+que dans les modules \textsf{Sequential} et \textsf{Cores}, et le reste du
 code est toujours inchangé. Le patron et les ouvriers sont distingués
@@
 Si les deux programmes sont compilés avec la même version d'\Ocaml, on
-utilise le module \of{Poly}.
+utilise le module \textsf{Poly}.
 Commençons par l'ouvrier. Son code prend la forme suivante.
-\begin{ocaml}
-  open Network.Poly
-  let worker (ai, bj) = ...
-  let () = Worker.compute worker ()
-\end{ocaml}
-La fonction \of{Worker.compute} entre dans une boucle qui attend les
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Network.Poly~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}worker\hspace*{1.22ex}(ai,\hspace*{1.22ex}bj)\hspace*{1.22ex}=\hspace*{1.22ex}...~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}Worker.compute\hspace*{1.22ex}worker\hspace*{1.22ex}()
+}
+
+}\noindent
+La fonction \textsf{Worker.compute} entre dans une boucle qui attend les
 tâches envoyées par le patron et renvoie les résultats calculés par
-la fonction \of{worker}. Le code du patron, quant à lui, est quasiment
-le même qu'auparavant. On commence par remplacer \of{Same} par
-\of{Poly}.
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Poly
-\end{ocaml}
-La construction des tâches et la fonction \of{master} sont inchangées.
-\begin{ocaml}
-  let tasks = ...
-  let master (_, (i,j)) r = ...
-\end{ocaml}
-Enfin on lance le calcul avec la fonction \of{Master.compute}, qui ne
-prend plus de fonction \of{worker} en argument.
-\begin{ocaml}
-  let () = Master.compute ~master tasks
-\end{ocaml}
+la fonction \textsf{worker}. Le code du patron, quant à lui, est quasiment
+le même qu'auparavant. On commence par remplacer \textsf{Same} par
+\textsf{Poly}.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Network~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}declare\_{}workers\hspace*{1.22ex}\symbol{126}n\symbol{58}4\hspace*{1.22ex}"orcus"~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}declare\_{}workers\hspace*{1.22ex}\symbol{126}n\symbol{58}8\hspace*{1.22ex}"belzebuth"~\linebreak
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Poly
+}
+
+}\noindent
+La construction des tâches et la fonction \textsf{master} sont inchangées.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}...~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}master\hspace*{1.22ex}(\_{},\hspace*{1.22ex}(i,j))\hspace*{1.22ex}r\hspace*{1.22ex}=\hspace*{1.22ex}...
+}
+
+}\noindent
+Enfin on lance le calcul avec la fonction \textsf{Master.compute}, qui ne
+prend plus de fonction \textsf{worker} en argument.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}Master.compute\hspace*{1.22ex}\symbol{126}master\hspace*{1.22ex}tasks
+}
+
+}\noindent
 
@@
 L'ouvrier modifié prend alors la forme suivante.
-\begin{ocaml}
-  open Mono
-  let worker (ai, bj) = ...
-  let worker_string s = string_of_coeff (worker (task_of_string s))
-  let () = Worker.compute worker_string ()
-\end{ocaml}
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{open}\hspace*{1.22ex}Mono~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}worker\hspace*{1.22ex}(ai,\hspace*{1.22ex}bj)\hspace*{1.22ex}=\hspace*{1.22ex}...~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}worker\_{}string\hspace*{1.22ex}s\hspace*{1.22ex}=\hspace*{1.22ex}string\_{}of\_{}coeff\hspace*{1.22ex}(worker\hspace*{1.22ex}(task\_{}of\_{}string\hspace*{1.22ex}s))~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}Worker.compute\hspace*{1.22ex}worker\_{}string\hspace*{1.22ex}()
+}
+
+}\noindent
 Le patron est modifié de la même façon.
-On remplace \of{Poly} par \of{Mono} et on encode/décode les tâches et
+On remplace \textsf{Poly} par \textsf{Mono} et on encode/décode les tâches et
 résultats.
-\begin{ocaml}
-  let tasks = ... string_of_task ...
-  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
-\end{ocaml}
-Les quatre fonctions de conversion \of{string_of_}\{\of{task,coeff}\} et
-\{\of{task,coeff}\}\of{_of_string} sont à la charge de l'utilisateur.
+
+\vspace*{-0.2em}{\small
+
+\noindent{\sffamily\parindent 0pt
+
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}...\hspace*{1.22ex}string\_{}of\_{}task\hspace*{1.22ex}...~\linebreak
+\noindent\hspace*{1.00em}\textbf{let}\hspace*{1.22ex}master\hspace*{1.22ex}(\_{},\hspace*{1.22ex}(i,j))\hspace*{1.22ex}r\hspace*{1.22ex}=\hspace*{1.22ex}c.(i).(j)\hspace*{1.22ex}\ensuremath{\leftarrow}\hspace*{1.22ex}coeff\_{}of\_{}string\hspace*{1.22ex}r\symbol{59}\hspace*{1.22ex}[]
+}
+
+}\noindent
+Les quatre fonctions de conversion \textsf{string\_{}of\_{}}\{\textsf{task,coeff}\} et
+\{\textsf{task,coeff}\}\textsf{\_{}of\_{}string} sont à la charge de l'utilisateur.
 
@@
 entier donnant le nombre de solutions pour cette tâche.
-On utilise la fonction \of{map_local_fold} où \of{f} effectue la
-recherche et \of{fold} somme les résultats intermédiaires.
+On utilise la fonction \textsf{map\_{}local\_{}fold} où \textsf{f} effectue la
+recherche et \textsf{fold} somme les résultats intermédiaires.
 
@@
 à dessiner, ainsi
-que la taille $\of{w} \times \of{h}$ en pixels de l'image finale.
-On se donne un nombre de tâches $\of{t} \ge 1$ comme paramètre. Il est
-immédiat de découper l'image en \of{t} sous-images, par exemple en
-bandes horizontales (si $\of{h} \ge \of{t}$) ou plus généralement en
+que la taille $\textsf{w} \times \textsf{h}$ en pixels de l'image finale.
+On se donne un nombre de tâches $\textsf{t} \ge 1$ comme paramètre. Il est
+immédiat de découper l'image en \textsf{t} sous-images, par exemple en
+bandes horizontales (si $\textsf{h} \ge \textsf{t}$) ou plus généralement en
 blocs rectangulaires.
@@
 correspondante. Le résultat est une matrice de pixels, de taille
-$(\of{w}\times\of{h}) / \of{t}$.
+$(\textsf{w}\times\textsf{h}) / \textsf{t}$.
 Ainsi, dessiner une image $800\times 600$ en utilisant 20 tâches
@@
 bibliothèque \functory\ introduits section~\ref{sec:API}, exception
-faire du module \of{Sequential}, dont la réalisation est immédiate.
+faire du module \textsf{Sequential}, dont la réalisation est immédiate.
 %
-Les deux modules \of{Cores} et \of{Network} contiennent une boucle
+Les deux modules \textsf{Cores} et \textsf{Network} contiennent une boucle
 principale analogue, de la forme suivante :
@@
     \quad  \quad \textbf{attendre} la fin d'une tâche \\
-    \quad  \quad \quad ajouter les nouvelles tâches renvoyées par \of{master} \\
+    \quad  \quad \quad ajouter les nouvelles tâches renvoyées par \textsf{master} \\
   \end{flushleft}
@@
 
-\subsection{\of{Cores}}
-
-Le module \of{Cores} permet de distribuer le calcul sur une unique
+\subsection{\textsf{Cores}}
+
+Le module \textsf{Cores} permet de distribuer le calcul sur une unique
 machine, typiquement en exploitant plusieurs \coeurs. 
 Comme illustré section~\ref{sec:studies}, la fonction
-\of{set_number_of_cores} permet de spécifier le nombre de \coeurs\ à
+\textsf{set\_{}number\_{}of\_{}cores} permet de spécifier le nombre de \coeurs\ à
 utiliser. Ce nombre peut être différent du nombre effectif de \coeurs\ 
@@
 
-Le module \of{Cores} est réalisé à l'aide de processus \unix, en
-utilisant la fonction \of{Unix.fork} de la bibliothèque standard d'\Ocaml.
+Le module \textsf{Cores} est réalisé à l'aide de processus \unix, en
+utilisant la fonction \textsf{Unix.fork} de la bibliothèque standard d'\Ocaml.
 Plus précisément, un compteur indique le nombre d'ouvriers disponibles 
 et l'affectation d'une tâche à un ouvrier se fait en 
-créant un nouveau sous-processus avec \of{fork}.
+créant un nouveau sous-processus avec \textsf{fork}.
 À la fin du calcul, le résultat de la tâche est transmis au processus
@@
 
-\subsection{\of{Network}}
-
-Le module \of{Network} permet de distribuer le calcul sur un réseau de
+\subsection{\textsf{Network}}
+
+Le module \textsf{Network} permet de distribuer le calcul sur un réseau de
 machines. 
 Comme illustré section~\ref{sec:studies}, la fonction 
-\of{declare_workers: n:int -> string -> unit} permet de déclarer
+\textsf{declare\_{}workers\symbol{58}\hspace*{1.22ex}n\symbol{58}int\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}string\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit} permet de déclarer
 l'ensemble des machines du réseau sur lesquels se trouvent des
@@
 ne coïncide pas nécessairement avec un nombre effectif de \coeurs).
-Le module \of{Network} est basé sur une architecture client/serveur
+Le module \textsf{Network} est basé sur une architecture client/serveur
 utilisant TCP/IP, où chaque ouvrier est un serveur et où le patron
@@
 En particulier, il ne peut administrer les tâches pendant l'exécution
-de la fonction \of{master}. Ce n'est pas un problème en pratique, car
-on peut supposer que la fonction \of{master} s'exécute rapidement.
-L'ouvrier, en revanche, crée un nouveau processus avec \of{fork} pour
+de la fonction \textsf{master}. Ce n'est pas un problème en pratique, car
+on peut supposer que la fonction \textsf{master} s'exécute rapidement.
+L'ouvrier, en revanche, crée un nouveau processus avec \textsf{fork} pour
 exécuter la tâche qui lui est confiée et peut donc continuer à
@@
 Les quatre messages possibles envoyés par le patron à un ouvrier sont : 
-\of{Assign(id:int, f:string, x:string)} pour affecter une nouvelle
-tâche \of{id} à l'ouvrier, sous la forme de deux chaînes \of{f} et \of{x} dont
+\textsf{Assign(id\symbol{58}int,\hspace*{1.22ex}f\symbol{58}string,\hspace*{1.22ex}x\symbol{58}string)} pour affecter une nouvelle
+tâche \textsf{id} à l'ouvrier, sous la forme de deux chaînes \textsf{f} et \textsf{x} dont
 le sens dépend du contexte ; 
-\of{Kill(id:int)} pour demander l'interruption de la tâche \of{id} ;
-\of{Stop} pour demander l'arrêt définitif de l'ouvrier ;
-et enfin \of{Ping} pour vérifier que l'ouvrier est toujours réactif.
+\textsf{Kill(id\symbol{58}int)} pour demander l'interruption de la tâche \textsf{id} ;
+\textsf{Stop} pour demander l'arrêt définitif de l'ouvrier ;
+et enfin \textsf{Ping} pour vérifier que l'ouvrier est toujours réactif.
 Inversement, les trois messages possibles envoyés par un ouvrier au
 patron sont :
-\of{Pong} en réponse à un message \of{Ping} ;
-\of{Completed(id:int, s:string)} pour renvoyer le résultat \of{s} de
-la tâche \of{id} ;
-et enfin \of{Aborted(id:int)} pour signifier l'échec de la tâche
-\of{id}, soit en réponse à \of{Kill}, soit à cause d'une erreur lors
-de l'exécution de la fonction \of{worker}.
+\textsf{Pong} en réponse à un message \textsf{Ping} ;
+\textsf{Completed(id\symbol{58}int,\hspace*{1.22ex}s\symbol{58}string)} pour renvoyer le résultat \textsf{s} de
+la tâche \textsf{id} ;
+et enfin \textsf{Aborted(id\symbol{58}int)} pour signifier l'échec de la tâche
+\textsf{id}, soit en réponse à \textsf{Kill}, soit à cause d'une erreur lors
+de l'exécution de la fonction \textsf{worker}.
 
@@
 l'\emph{endianness}, de la version d'\Ocaml\ et du système d'exploitation.
-Dans le sous-module \of{Same}, les arguments \of{f} et \of{x} du
-message \of{Assign} désignent respectivement
+Dans le sous-module \textsf{Same}, les arguments \textsf{f} et \textsf{x} du
+message \textsf{Assign} désignent respectivement
 la sérialisation d'une fonction et de
-son argument, et l'argument \of{s} du message \of{Completed} la
-sérialisation du résultat. Dans les sous-module \of{Poly} et
-\of{Mono}, en revanche, la
-partie \of{f} du message \of{Assign} n'est plus pertinente
-car la fonction \of{worker} est maintenant locale à l'ouvrier.
-Les chaînes \of{x} et \of{s} restent pertinentes ; pour \of{Poly},
-ce sont des sérialisations de valeurs \Ocaml\ et pour \of{Mono}  de
+son argument, et l'argument \textsf{s} du message \textsf{Completed} la
+sérialisation du résultat. Dans les sous-module \textsf{Poly} et
+\textsf{Mono}, en revanche, la
+partie \textsf{f} du message \textsf{Assign} n'est plus pertinente
+car la fonction \textsf{worker} est maintenant locale à l'ouvrier.
+Les chaînes \textsf{x} et \textsf{s} restent pertinentes ; pour \textsf{Poly},
+ce sont des sérialisations de valeurs \Ocaml\ et pour \textsf{Mono}  de
 simples chaînes de caractères.
 
+\bigskip
 \paragraph*{\bf Tolérance aux pannes}\label{sec:fault}
@@
 Il y a quatre états possibles pour un ouvrier :
-\of{déconnecté} signifie qu'il n'y a pas de connection TCP en cours
+\textsf{déconnecté} signifie qu'il n'y a pas de connection TCP en cours
 entre le patron et l'ouvrier ; 
-\of{vivant} signifie qu'un message de l'ouvrier a été reçu par le patron
+\textsf{vivant} signifie qu'un message de l'ouvrier a été reçu par le patron
 il y a moins de $T_1$ secondes ;
-\of{contacté} signifie que l'ouvrier n'a pas envoyé de message depuis
+\textsf{contacté} signifie que l'ouvrier n'a pas envoyé de message depuis
 plus de $T_1$ secondes et que le patron lui a envoyé un message
-\of{Ping} depuis moins de $T_2$ secondes ;
-enfin \of{inatteignable} signifie que l'ouvrier n'a toujours pas répondu au
-message \of{Ping} (depuis plus de $T_2$ secondes).
+\textsf{Ping} depuis moins de $T_2$ secondes ;
+enfin \textsf{inatteignable} signifie que l'ouvrier n'a toujours pas répondu au
+message \textsf{Ping} (depuis plus de $T_2$ secondes).
 Dès qu'un message est reçu en provenance d'un ouvrier, son état est
@@
 de la manière suivante. Les tâches ne sont envoyées qu'à des ouvriers
-se trouvant dans l'état \of{vivant} ou \of{contacté}.
+se trouvant dans l'état \textsf{vivant} ou \textsf{contacté}.
 D'autre part, dès qu'un ouvrier en train d'effectuer une tâche $t$
-passe dans l'état \of{déconnecté} ou \of{inatteignable}, la tâche $t$
+passe dans l'état \textsf{déconnecté} ou \textsf{inatteignable}, la tâche $t$
 est reprogrammée, ce qui signifie qu'elle est ajoutée de nouveau à
@@
 été reprogrammée et indique à tout ouvrier ayant déjà entrepris de
-recalculer $t$ de cesser son travail (avec le message \of{Kill}).
+recalculer $t$ de cesser son travail (avec le message \textsf{Kill}).
 
@@
 également robuste vis-à-vis des exceptions levées par la fonction
-\of{worker}. Le cas échéant, un message \of{Aborted} est envoyé au
+\textsf{worker}. Le cas échéant, un message \textsf{Aborted} est envoyé au
 patron et la tâche est reprogrammée. C'est la responsabilité de
@@
 La table suivante montre les temps d'exécution pour différentes valeurs
-de $N$ et pour chacun des trois modules \of{Sequential}, \of{Cores} et
-\of{Network}. L'objectif est ici de mesurer le facteur par rapport à
+de $N$ et pour chacun des trois modules \textsf{Sequential}, \textsf{Cores} et
+\textsf{Network}. L'objectif est ici de mesurer le facteur par rapport à
 l'exécution séquentielle. En conséquence, tous les calculs sont
@@
     \hline
-    N & D & \#tâches  & \of{Sequential}& \of{Cores}            & \of{Network} 
+    N & D & \#tâches  & \textsf{Sequential}& \textsf{Cores}            & \textsf{Network} 
     \\\hline\hline
@@
 \end{center}
-Il apparaît clairement que les modules \of{Cores} et \of{Network}
+Il apparaît clairement que les modules \textsf{Cores} et \textsf{Network}
 permettent d'obtenir un facteur d'accélération significatif, qui
 atteint presque 8, c'est-à-dire le nombre de \coeurs\ utilisés.
-Il apparaît également que le module \of{Network} donne de meilleurs
+Il apparaît également que le module \textsf{Network} donne de meilleurs
 résultats lorsque le temps de calcul domine largement le temps de
@@
 4.6\%\ du temps total.
-D'une manière générale, le module \of{Network} n'est intéressant que
+D'une manière générale, le module \textsf{Network} n'est intéressant que
 si le temps de calcul de chaque tâche est significatif.
@@
 Un calcul purement séquentiel prend 29,4 secondes.
-Les résultats pour les modules \of{Cores} et \of{Network} sont
+Les résultats pour les modules \textsf{Cores} et \textsf{Network} sont
 présentés figure~\ref{fig:bench:mandelbrot}.
@@
       \hline
-      \#\coeurs  &\#tâches & \of{Cores} & \of{Network} \\
+      \#\coeurs  &\#tâches & \textsf{Cores} & \textsf{Network} \\
       \hline\hline
@@
 \end{figure}
-Les meilleurs résultats sont obtenus avec le module \of{Cores}, où la
+Les meilleurs résultats sont obtenus avec le module \textsf{Cores}, où la
 communication n'intervient que localement, entre processus.
@@
 L'un des problèmes consistait à multiplier deux matrices de taille
-$100\times 100$, c'est-à-dire $\of{n} = \of{p} =
-\of{m} = 100$ avec les notations de la section~\ref{sec:matrix}.
+$100\times 100$, c'est-à-dire $\textsf{n} = \textsf{p} =
+\textsf{m} = 100$ avec les notations de la section~\ref{sec:matrix}.
 Les coefficients sont entiers mais peuvent avoir des milliers de
@@
 On compare les performances de deux programmes distribués utilisant
-\functory. Dans le premier, appelé \of{mm1}, chaque tâche consiste en
+\functory. Dans le premier, appelé \textsf{mm1}, chaque tâche consiste en
 le calcul d'un unique coefficient de la matrice résultat, exactement 
 comme décrit section~\ref{sec:matrix}. 
-Dans le second, appelé \of{mm2}, chaque tâche consiste en le calcul de
+Dans le second, appelé \textsf{mm2}, chaque tâche consiste en le calcul de
 toute une ligne de la matrice résultat.
-En conséquence, le nombre total de tâches est $\of{n}\times\of{m}=10,000$ 
-pour \of{mm1} et seulement $\of{n}=100$ pour \of{mm2}.
+En conséquence, le nombre total de tâches est $\textsf{n}\times\textsf{m}=10,000$ 
+pour \textsf{mm1} et seulement $\textsf{n}=100$ pour \textsf{mm2}.
 Les résultats expérimentaux, toujours en secondes et toujours sur la
@@
     \hline
-    & \of{mm1}       & \of{mm2}  \\
+    & \textsf{mm1}       & \textsf{mm2}  \\
     & (10,000 tâches) & (100 tâches) \\
     \hline\hline
-\of{Sequential} & 20.3 \phantom{(0.00)$\times$} &  20.2 \phantom{(0.00)$\times$} \\
+\textsf{Sequential} & 20.3 \phantom{(0.00)$\times$} &  20.2 \phantom{(0.00)$\times$} \\
 \hline
- \of{Cores} 
+ \textsf{Cores} 
  (2 \coeurs)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
@@
 coûts de calcul, donnant un temps d'exécution total supérieur à 30
-secondes. Que ce soit pour \of{mm1} ou \of{mm2}, le nombre total de
+secondes. Que ce soit pour \textsf{mm1} ou \textsf{mm2}, le nombre total de
 coefficients transmis est en effet en
-$O(\of{n}\times\of{m}\times\of{p})$, ce qui représente ici des
+$O(\textsf{n}\times\textsf{m}\times\textsf{p})$, ce qui représente ici des
 milliards d'octets.
@@
 n'enverrait uniquement que des indices de lignes et de colonnes. Cela
-réduirait le volume des communications à seulement $O(\of{n}\times\of{m})$.
+réduirait le volume des communications à seulement $O(\textsf{n}\times\textsf{m})$.
 
@@
 information concernant les tâches, leur ordonnancement est complètement
-arbitraire. Les modules \of{Cores} et \of{Network} utilisent de
+arbitraire. Les modules \textsf{Cores} et \textsf{Network} utilisent de
 une simple file pour les tâches restant à effectuer et les nouvelles
@@
 
+\bigskip
 %\subsubsection{Annexe pour le service de fabrication~: commande \cmd{publisher}}
@@
 
+\bigskip
 %BIBLIOGRAPHIE
@@
 
+\bigskip\bigskip
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-01-24 doc/jfla-2011/talk/whizzy.sh
*** empty log message ***

@@ -1,0 +1,1 @@
+export MKSLICE="latexpp -w -g color yes"

### 2011-02-25 doc/tfp-2011/main.tex
*** empty log message ***

--- 2010-12-02 doc/report/main.tex
@@
 
-\documentclass[a4paper,12pt]{article}
+\documentclass[tfpsymp,pagenumbers]{tfp07symp}
 
@@
 
-\usepackage{fullpage,comment,amsmath,graphicx,color,url}
+\usepackage{comment,amsmath,graphicx,color,url}
 
@@
 
-\title{Functory: Distributed Computing \\ for the Common Man\thanks{This
+\title{Functory: A Distributed Computing Library \\ for Objective Caml 
+\thanks{This
     research was partly supported by the French national project U3CAT
@@
     ANR-08-SEGI-021).}}
-%\subtitle{Subtitle Text, if any}
-
-\author{Jean-Christophe Filli\^{a}tre \quad K. Kalyanasundaram \\
+
+\author{Jean-Christophe Filli\^{a}tre \and K. Kalyanasundaram}
+\authorrunning{J.-C. Filli\^atre \and K. Kalyanasundaram}
+
+\institute{
   CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
   INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893 \\
-   \url{filliatr@lri.fr, kalyan.krishnamani@inria.fr}}
+   \email{filliatr@lri.fr, kalyan.krishnamani@inria.fr}}
 
@@
 \end{abstract}
+
+% PLAN
+% 1. introduction, including related work
+% 2. API
+%    2.1 low level, compute, derived APIs
+%    2.2 five different implementations
+% 3. implementation
+%    3.1 marshalling
+%    3.2 protocol
+%    3.3 fault tolerance
+% 4. experiments
 
@@
 
+\paragraph{Related Work.}
+Closest to the approach in this paper is Yohann Padioleau's MapReduce
+implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
+top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
+protocol for message passing.  Currently, we have less flexibility
+w.r.t. deployment of the user program than OCamlMPI; on the other
+hand, we provide a 
+more generic API together with fault tolerance.  There are other
+distributed computing libraries on top of which one could implement
+the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
+them. However, \JoCaml\ does not provide fault tolerance, which is
+indispensable in a distributed setting. The user has to include code
+for fault tolerance, as already demonstrated in some \JoCaml\
+experiments~\cite{mandel2008}.
+
+There are other implementations of distributed computing in the
+context of functional programming. One is the Disco
+project~\cite{disco}, which implements exactly Google's MapReduce in
+Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
+implementation of Google's MapReduce.
+There are other ways to exploit multi-core architectures. One of these
+is data parallelism, which is also relevant in the functional
+programming setting~\cite{parallel-haskell}. Our work
+does not target data parallelism at all.
+
 \medskip
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{API}\label{sec:API}
+
+\subsection{compute}
+
 In this section, we describe the
@@
 
-Actually, our library provides not just a single \ocaml{compute}
-function as above, but instead five different versions
-depending on the execution context. 
-The first two contexts are the following: 
-\begin{enumerate}
-\item \textbf{Purely sequential execution:}
-  this is mostly intended to be a reference implementation
-  for performance comparisons, as well as for debugging;
-
-\item \textbf{Several cores on the same machine:} 
-  this implementation is intended to distribute the computation over a
-  single machine and it makes use of \unix\ processes;
-\end{enumerate}
-The next three implementations are intended for distributing the
-computation over a network of machines.
-\begin{enumerate}
-\setcounter{enumi}{2}
-\item \textbf{Same executable run on master and worker machines:}
-  this implementation makes use of the ability to marshal \Ocaml\
-  closures and polymorphic values.
-  Depending on whether the program is run as a master or as a worker,
-  the relevant arguments of \ocaml{compute} are used.
-
-\item \textbf{Master and workers are different programs, compiled with
-    the same version of \Ocaml:} 
-  we can no longer marshal closures but we can still
-  marshal polymorphic values. As a consequence, 
-  the \ocaml{compute} function is split into two
-  polymorphic functions, to implement the master and workers 
-  respectively:%
-\vspace{-1em}
-\begin{ocaml}
-val Worker.compute : ('a -> 'b) -> unit -> unit
-val Master.compute : 
-  ('a * 'c -> 'b -> ('a * 'c) list) -> 
-  ('a * 'c) list -> unit
-\end{ocaml}
-
-\item \textbf{Master and workers are different programs, not even
-    compiled with the same version of \Ocaml:} we can no
-  longer use marshalling, so the
-  \ocaml{compute} function is split into two monomorphic functions
-  over strings:%
-\vspace{-1em}
-\begin{ocaml}
-val Worker.compute : (string -> string) -> unit -> unit
-val Master.compute : 
-  (string * 'c -> string -> (string * 'c) list) -> 
-  (string * 'c) list -> unit
-\end{ocaml}
-\end{enumerate}
-
-Our library is organized into three modules: \of{Sequential} for the
-pure sequential implementation, \of{Cores} for multiple cores on the
-same machine and \of{Network} for a network of machines, respectively.
-The \of{Network} module itself is organized into three sub-modules,
-called \of{Same}, \of{Poly} and \of{Mono},
-corresponding to contexts 3, 4 and 5 above. 
-The next section presents more functions provided by \functory, which
-are all derived from the generic API just presented.
-
-% The different
-% API functions are summarized in the following table (their signatures
-% being given above):
-% \begin{center}
-%   \begin{tabular}{|c|c|c|c|c|}
-%     \hline
-%     \ocaml{Sequential} & \ocaml{Cores} &
-%     \multicolumn{3}{|c|}{\ocaml{Network}} 
-%     \\\cline{3-5}
-%     &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
-%     \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
-%     \ocaml{master}  & \ocaml{master}  \\
-%     & & & \ocaml{worker}  & \ocaml{worker} \\\hline
-%   \end{tabular}
-% \end{center}
-% The following section introduces other interesting APIs that we
-% derive from this generic API.
-
 \begin{comment}
@@
 
-\section{Derived API}\label{sec:derived}
+\subsection{Low-level API}
+
+% TODO
+
+\subsection{Derived API}\label{sec:derived}
 
@@
 
+\subsection{Deployment Scenarios}
+
+Actually, our library provides not just a single \ocaml{compute}
+function as above, but instead five different versions
+depending on the execution context. 
+The first two contexts are the following: 
+\begin{enumerate}
+\item \textbf{Purely sequential execution:}
+  this is mostly intended to be a reference implementation
+  for performance comparisons, as well as for debugging;
+
+\item \textbf{Several cores on the same machine:} 
+  this implementation is intended to distribute the computation over a
+  single machine and it makes use of \unix\ processes;
+\end{enumerate}
+The next three implementations are intended for distributing the
+computation over a network of machines.
+\begin{enumerate}
+\setcounter{enumi}{2}
+\item \textbf{Same executable run on master and worker machines:}
+  this implementation makes use of the ability to marshal \Ocaml\
+  closures and polymorphic values.
+  Depending on whether the program is run as a master or as a worker,
+  the relevant arguments of \ocaml{compute} are used.
+
+\item \textbf{Master and workers are different programs, compiled with
+    the same version of \Ocaml:} 
+  we can no longer marshal closures but we can still
+  marshal polymorphic values. As a consequence, 
+  the \ocaml{compute} function is split into two
+  polymorphic functions, to implement the master and workers 
+  respectively:%
+\vspace{-1em}
+\begin{ocaml}
+val Worker.compute : ('a -> 'b) -> unit -> unit
+val Master.compute : 
+  ('a * 'c -> 'b -> ('a * 'c) list) -> 
+  ('a * 'c) list -> unit
+\end{ocaml}
+
+\item \textbf{Master and workers are different programs, not even
+    compiled with the same version of \Ocaml:} we can no
+  longer use marshalling, so the
+  \ocaml{compute} function is split into two monomorphic functions
+  over strings:%
+\vspace{-1em}
+\begin{ocaml}
+val Worker.compute : (string -> string) -> unit -> unit
+val Master.compute : 
+  (string * 'c -> string -> (string * 'c) list) -> 
+  (string * 'c) list -> unit
+\end{ocaml}
+\end{enumerate}
+
+Our library is organized into three modules: \of{Sequential} for the
+pure sequential implementation, \of{Cores} for multiple cores on the
+same machine and \of{Network} for a network of machines, respectively.
+The \of{Network} module itself is organized into three sub-modules,
+called \of{Same}, \of{Poly} and \of{Mono},
+corresponding to contexts 3, 4 and 5 above. 
+The next section presents more functions provided by \functory, which
+are all derived from the generic API just presented.
+
+% The different
+% API functions are summarized in the following table (their signatures
+% being given above):
+% \begin{center}
+%   \begin{tabular}{|c|c|c|c|c|}
+%     \hline
+%     \ocaml{Sequential} & \ocaml{Cores} &
+%     \multicolumn{3}{|c|}{\ocaml{Network}} 
+%     \\\cline{3-5}
+%     &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
+%     \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
+%     \ocaml{master}  & \ocaml{master}  \\
+%     & & & \ocaml{worker}  & \ocaml{worker} \\\hline
+%   \end{tabular}
+% \end{center}
+% The following section introduces other interesting APIs that we
+% derive from this generic API.
+
 % TODO cannot be instantiated in the monomorphic case
 
-\section{Illustrative Case Studies}\label{sec:studies}
-
-In this section, we show how to use \functory\ on several case studies.
-We focus here on the use of the API. Experimental results of these
-case studies are given in Section~\ref{sec:experiments}.
-The source code for all these case studies is contained in the
-distribution, in sub-directory \texttt{tests}.
-
-\subsection{Matrix Multiplication}\label{sec:matrix}
-
-As a first example, let us consider matrix multiplication.
-We assume two matrices \of{a} and \of{b}, respectively of size
-$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
-a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
-Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
-the standard matrix multiplication would be as follows:
-\begin{ocaml}
-  for i = 0 to n-1 do
-    for j = 0 to m-1 do
-      for k = 0 to p-1 do
-	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
-      done
-    done
-  done  
-\end{ocaml}
-and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
-assuming addition and multiplication over coefficients to be 
-denoted $+$ and $\times$ respectively.
-An easy way to distribute this computation is obviously to turn each
-computation of the inner loop over \of{k} into a task.
-To do so, we first prepare the list of tasks:
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in
-    for i = 0 to n-1 do for j = 0 to m-1 do
-      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
-    done done;
-    !l
-\end{ocaml}
-Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
-tuple, together with position \of{(i,j)} as a second tuple.
-The worker is receiving the first tuple as argument and computes
-the dot product:
-\begin{ocaml}
-  let worker (ai, bj) =
-    let c = ref 0 in
-    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
-    !c
-\end{ocaml}
-The master is a one line function which receives the result \of{r}
-from the worker and simply stores it into \of{c}, according to the
-position contained in the input task. It produces no new task.
-\begin{ocaml}
-  let master (_, (i,j)) r = c.(i).(j) <- r; []
-\end{ocaml}
-Finally, the overall computation is started by invoking \of{compute}
-as follows:
-\begin{ocaml}
-  let () = compute ~worker ~master !tasks
-\end{ocaml}
-The total number of tasks is $\of{n}\times\of{m}$.
-
-Using the sequential implementation provided by \of{Functory} is as
-simple as including the following line in the code:
-\begin{ocaml}
-  open Sequential
-\end{ocaml}
-Then the computation is roughly similar to the standard multiplication
-described above.
-
-Now, let us assume we want to make use of a 4-core machine to perform
-the same computation. 
-This is achieved by just replacing the line above by
-\begin{ocaml}
-  open Cores
-  let () = set_number_of_cores 4
-\end{ocaml}
-while the rest of the code remains unchanged.
-
-Later, we may want to use a network of machines instead, for example
-two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
-cores respectively. Similar to the above, we turn the above lines into
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Same
-\end{ocaml}
-Here, \of{Same} is a module which is to be used when master and worker
-are running the same executable. It provides a \of{compute} function
-which has the same signature as in modules \of{Sequential} and
-\of{Cores}, so that the rest of the code remains unchanged. Master and
-worker processes are distinguished at run-time using an environment
-variable \texttt{WORKER} which is set/unset.
-
-If we need to write two different programs for master and worker, for
-reason of binary incompatibility or any other reason, the library API
-is providing functions to do that. If master and worker are still
-compiled with the same version of \Ocaml, we use the \of{Poly} module
-which provides a polymorphic API. Let us start with the worker program.
-It now looks like:
-\begin{ocaml}
-  open Poly
-  let worker (ai, bj) = ...
-  let () = Worker.compute worker ()
-\end{ocaml}
-The \of{Worker.compute} function enters a loop which waits for
-tasks sent by the master and returns results computed using \of{worker}.
-The master program is almost the same as before. First, we replace
-module \of{Same} with module \of{Poly}:
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Poly
-\end{ocaml}
-Tasks and \of{master} function are unchanged:
-\begin{ocaml}
-  let tasks = ...
-  let master (_, (i,j)) r = ...
-\end{ocaml}
-Finally, we start the computation with \of{Master.compute}, which does
-not have a \of{worker} parameter anymore:
-\begin{ocaml}
-  let () = Master.compute ~master tasks
-\end{ocaml}
-
-When master and worker programs are compiled with different
-versions of the \Ocaml\ compiler, our library still provides
-a monomorphic API over strings. As a consequence, we need to convert
-tasks and results to and from strings in both master and worker.
-The modified worker program then looks as follows:
-\begin{ocaml}
-  open Mono
-  let worker (ai, bj) = ...
-  let worker_string s = string_of_coeff (worker (task_of_string s))
-  let () = Worker.compute worker_string ()
-\end{ocaml}
-The master program is modified in a similar way.
-We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
-as strings, as follows:
-\begin{ocaml}
-  let tasks = ... string_of_task ...
-  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
-\end{ocaml}
-where \of{string_of_task} and \of{task_of_string} are user-defined
-functions to convert tasks to and from strings.
-
-\subsection{N-queens}\label{sec:n-queens}
-
-The next example is the classical $N$-queens problem, where we
-compute the total number of ways to place $N$ queens on a $N\times N$
-chessboard in such a way no two queens attack each other.
-We use a standard backtracking algorithm for this problem, which
-places the queens one by one starting from the first row.
-Distributing the computation is thus quite easy: we consider all
-possible ways to place queens on the first $D$ rows and then perform
-the subsequent search in parallel. Choosing $D=1$ will result in
-exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
-so on.
-% TODO: make a picture to illustrate positioning of 2 queens on the
-% first 2 rows
-
-Each task only consists of three integers and its result is one integer,
-which is the total number of solutions for this task.
-We make use of function \of{map_local_fold} from the derived API,
-where \of{f} is performing the search and \of{fold} simply adds the
-intermediate results.
-In the network configuration, we make use of the \of{Network.Same}
-module, workers and master being the same executable.
-
-\subsection{Mandelbrot Set}
-
-Drawing the Mandelbrot set is another classical example that could be
-distributed easily, since the color of each point can be computed
-independently of the others. 
-Let us assume we are given the coordinates of the region to be drawn,
-along with the width \of{w} and height \of{h} of the resulting image
-(in pixels). 
-Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
-parameter. It is immediate to split the image into \of{t} sub-images,
-each of which is computed in parallel with and independently of the
-others. For instance, the image could be split into horizontal slices
-or, more generally, into rectangular blocks.
-
-Each task is thus four floating-point numbers denoting the region
-coordinates, together with two integers denoting the dimensions of the
-sub-image to be drawn. The result of the task is a matrix of pixels,
-of size $(\of{w}\times\of{h}) / \of{t}$. 
-For instance, drawing a $800\times 600$ image using 20 tasks will
-result in 20 sub-images of size $176,000$ bytes each,
-assuming each pixel is encoded in four bytes.
-
-In the network configuration, we deliberately choose to have two
-different programs for master and workers, using the \of{Network.Poly}
-module (actually, a single source code with a command line option).
-
-% \subsection{Polynomial Multiplication}
-% TODO?
-
-\subsection{SMT Solvers}\label{sec:SMT}
-
-Here we demonstrate the potential of our library for our application
-needs as mentioned in the introduction. In this case study,
-we consider 80 challenging verification conditions
-(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
-VC is stored in a file, which is accessible over
-NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
-which are all installed on each of the machines in use.
-
-The master program proceeds by reading the file names, turning them
-into tasks by multiplying them by the number of provers (more than
-300 tasks in total).
-Each worker in turn invokes the given prover on the given file, within
-a timeout limit.
-Each task completes with one of the four possible outcomes: \emph{valid},
-\emph{unknown} (depending on
-whether the VC is valid or undecided by the prover), 
-\emph{timeout} and \emph{failure}.
-
-The result of each computation is a pair denoting the status and the
-time spent in the prover call. The master collects these results and
-sums up the timings for each prover and each possible status.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Implementation Details}\label{sec:implem}
@@
 
+\begin{comment}
   \subsection{Multiple Cores}
 
-The \of{Cores} module implements the distributed computing library for
-multiple cores on the same machine. 
+  The \of{Cores} module implements the distributed computing library
+  for multiple cores on the same machine.
   % As illustrated in Section~\ref{sec:example},
-It provides a function
-\of{set_number_of_cores: int -> unit} to indicate the number of cores
-to be used. The number passed to this function may be different from
-the actual number of cores in the machine; it is rather the number of
-tasks to be performed simultaneously.
-% TODO comment more on the number of cores?
-% Thus a program using the \of{Cores} module typically begins with 
+  It provides a function \of{set_number_of_cores: int -> unit} to
+  indicate the number of cores to be used. The number passed to this
+  function may be different from the actual number of cores in the
+  machine; it is rather the number of tasks to be performed
+  simultaneously.
+  % TODO comment more on the number of cores?  Thus a program using
+  % the \of{Cores} module typically begins with
   % \begin{ocaml}
-% open Cores
-% let () = set_number_of_cores 8
+  %   open Cores let () = set_number_of_cores 8
   % \end{ocaml}
 
-The \of{Cores} module is implemented with \unix\ processes, using the
-\of{fork} and \of{wait} system calls provided by the \of{Unix} library
-of \Ocaml. The idea is pretty simple.  The \of{compute} function
-maintains a global table of pending tasks and keeps track of the
-number of idle cores.  Whenever there is a pending task and an idle
-core, a new sub-process is created using \of{Unix.fork}; once
-completed, the sub-process marshals the result into a 
-local file. The
-main process maintains a table mapping each sub-process ID to the
-input task and the local file name. It waits for any completed
-sub-process using \of{Unix.wait} and recovers the result from the
-local file. Then function \of{master} is applied, which may generate
-new tasks.  The main loop can be depicted in the following way:
+  The \of{Cores} module is implemented with \unix\ processes, using
+  the \of{fork} and \of{wait} system calls provided by the \of{Unix}
+  library of \Ocaml. The idea is pretty simple.  The \of{compute}
+  function maintains a global table of pending tasks and keeps track
+  of the number of idle cores.  Whenever there is a pending task and
+  an idle core, a new sub-process is created using \of{Unix.fork};
+  once completed, the sub-process marshals the result into a local
+  file. The main process maintains a table mapping each sub-process ID
+  to the input task and the local file name. It waits for any
+  completed sub-process using \of{Unix.wait} and recovers the result
+  from the local file. Then function \of{master} is applied, which may
+  generate new tasks.  The main loop can be depicted in the following
+  way:
   \begin{flushleft}
@@
   \end{flushleft}
-We also have an alternative implementation using
-\unix\ pipes instead of local files.
+  We also have an alternative implementation using \unix\ pipes
+  instead of local files.
 
   The scheduling of tasks to the different cores is left to the
-operating system, through \of{Unix.fork}. Thus it may be the case that
-two tasks are executed on the same core, even if the declared number
-of cores is less or equal than the actual number of cores on the machine.
+  operating system, through \of{Unix.fork}. Thus it may be the case
+  that two tasks are executed on the same core, even if the declared
+  number of cores is less or equal than the actual number of cores on
+  the machine.
   % TODO: comment on user-provided scheduling?
-
-\subsection{Network of Machines}
+\end{comment}
+
+\subsection{Marshaling}
+
+  As mentioned in Section~\ref{sec:API}, the \of{Network} module
+  actually provides three different implementations, according to
+  three different execution scenarios. There are provided in three
+  sub-modules, as described below.
+
+  \paragraph{\of{Same}.} This module is used when master and workers
+  are running the same executable. The master and workers have to be
+  differentiated in some manner. We use an environment variable
+  \texttt{WORKER} for this purpose. When set, it indicates that the
+  executable acts as a worker. At run-time, a worker immediately
+  enters a loop waiting for tasks from the master, without even
+  getting into the user code.  As explained in Section~\ref{sec:API},
+  the master function has the following signature.
+  \begin{ocaml}
+    val compute : worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a *
+    'c) list) -> ('a * 'c) list -> unit
+  \end{ocaml}
+  The master uses marshalling to send both a closure of type \of{'a ->
+    'b} and a task of type \of{'a} to the worker. The resulting
+  strings are passed as argument \of{f} and \of{x} in message
+  \of{Assign}. Similarly, the worker uses marshalling to send back the
+  result of the computation of type \of{'b}, which is the argument
+  \of{s} in message \of{Completed}.
+
+  Though the ability to run the same executable helps a lot in
+  deploying the program in different machines, it comes at a small
+  price. Since the worker is not getting into the user code, closures
+  which are transmitted from the master cannot refer to global
+  variables in the user code. Indeed, the initialization code for
+  these global variables is never reached on the worker side. For
+  instance, the Mandelbrot set example could be written as follows:
+  \begin{ocaml}
+    let max_iterations = 200 let worker (xmi, xma, ymi, yma, w, h) =
+    ... draw sub-image using max_iterations ...
+  \end{ocaml}
+  That is, the global function \of{worker} makes use of the global
+  variable \of{max_iterations}. The worker gets the function to
+  compute from the master, namely the closure corresponding to
+  function \of{worker} in that case, but on the worker side the
+  initialization of \of{max_iterations} is never executed.
+
+  One obvious solution is not to use global variables in the worker
+  code. This is not always possible, though.  To overcome this, the
+  \of{Same} sub-module also provides a \of{Worker.compute} function to
+  start the worker loop manually from the user code. This way, it can
+  be started at any point, in particular after the initialization of
+  the required global variables.  Master and worker are still running
+  the same executable, but are distinguished using a user-defined way
+  (command-line argument, environment variable, etc.).
+
+  \bigskip There are situations where it is not possible to run the
+  same executable for master and workers.  For instance, architectures
+  or operating systems could be different across the network.  For
+  that reason, the \of{Network} module provides two other
+  implementations.
+
+  \paragraph{\of{Poly}.} When master and workers are compiled with the
+  same version of \Ocaml, we can no longer marshal closures but we can
+  still marshal polymorphic values. Indeed, an interesting property of
+  marshalling in \Ocaml\ is to be fully architecture-independent, as
+  long as a single version of \Ocaml\ is used. It is worth pointing
+  out that absence of marshaled closures now enables the use of two
+  different programs for master and workers. This is not mandatory,
+  though, since master and workers could still be distinguished at
+  run-time as in the previous case.
+
+  On the worker side, the main loop is started manually using
+  \of{Worker.compute}. The computation to be performed on each task is
+  given as an argument to this function. It thus looks as follows:
+  \begin{ocaml}
+    Worker.compute : ('a -> 'b) -> unit -> unit
+  \end{ocaml}
+  On the master side, the \of{compute} function is simpler than in the
+  previous case, as it has one argument less, and thus has the
+  following signature.
+  \begin{ocaml}
+    Master.compute : master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a *
+    'c) list -> unit
+  \end{ocaml}
+  For realistic applications, where master and workers are completely
+  different programs, possibly written by different teams, this is the
+  module of choice in our library, since it can still pass polymorphic
+  values over the network. The issues of marshalling are automatically
+  taken care of by \Ocaml\ run-time.
+
+  The derived API presented in Section~\ref{sec:derived} is adapted to
+  deal with the absence of closures. Exactly as the \of{compute}
+  function, each API now takes two forms, one for the master and
+  another for the workers. For example, \of{map_fold_ac} takes the
+  following forms.
+  \begin{ocaml}
+    Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit
+    -> unit Master.map_fold_ac : 'b -> 'a list -> 'b
+  \end{ocaml}
+  It is the responsibility of the user to ensure type consistency
+  between master and workers.
+
+  \paragraph{\of{Mono}.} When master and workers are compiled using
+  different versions of \Ocaml, we can no longer use marshalling.  As
+  in the previous case, we split \of{compute} into two functions, one
+  for master and one for workers. In addition, values transmitted over
+  the network can only be strings. The signature thus takes the
+  following form.
+  \begin{ocaml}
+    Worker.compute : (string -> string) -> unit -> unit Master.compute
+    : master:(string * 'c -> string -> (string * 'c) list) -> (string
+    * 'c) list -> unit
+  \end{ocaml}
+  Any other datatype for tasks should be encoded to/from strings. This
+  conversion is left to the user.  Note that the second component of
+  each task is still polymorphic (of type \of{'c} here), since it is
+  local to the master.
+
+  % There is actually a single, generic implementation, which is then
+  % instantiated in three different ways, corresponding to the three
+  % scenarios.  The generic implementation assumes that only strings
+  % are being transmitted over the network, as evident from the
+  % \of{Assign} and \of{Completed} messages in
+  % Section~\ref{sec:protocol}. Depending on the implementation, these
+  % strings could be marshaled closures, marshaled values or simply
+  % user strings.
+
+\subsection{Protocol}\label{sec:protocol}
 
@@
 
-\subsubsection{Protocol}\label{sec:protocol}
+%\subsubsection{Protocol}
 
@@
 
-\subsubsection{Network Sub-modules} % TODO: change title
-
-As mentioned in Section~\ref{sec:API}, the \of{Network} module
-actually provides three different implementations, according to three
-different execution scenarios. There are provided in three
-sub-modules, as described below.
-
-\paragraph{\of{Same}.} This module is used when master and workers are
-running the same executable. The master and workers have to be
-differentiated in some manner. We use an environment variable
-\texttt{WORKER} for this purpose. When set, it indicates that the
-executable acts as a worker. At run-time, a worker immediately 
-enters a loop waiting for tasks from the master, without even getting
-into the user code. 
-As explained in Section~\ref{sec:API}, 
-the master function has the following signature.
-\begin{ocaml}
- val compute : 
-      worker:('a -> 'b) -> 
-      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit  
-\end{ocaml}
-The master uses marshalling to send both a closure of type 
-\of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting strings
-are passed as argument \of{f} and \of{x} in message
-\of{Assign}. Similarly, the worker uses marshalling to send back the
-result of the computation of type \of{'b}, which is the argument
-\of{s} in message \of{Completed}.
-
-Though the ability to run the same executable helps a lot in deploying
-the program in different machines, it comes at a small price. Since
-the worker is not getting into the user code, closures which are
-transmitted from the master cannot refer to global variables in the
-user code. Indeed, the initialization code for these global variables
-is never reached on the worker side. For instance, the Mandelbrot set
-example could be written as follows:
-\begin{ocaml}
-  let max_iterations = 200
-  let worker (xmi, xma, ymi, yma, w, h) =
-    ... draw sub-image using max_iterations ...
-\end{ocaml}
-That is, the global function \of{worker} makes use of the global
-variable \of{max_iterations}. The worker gets the function to compute
-from the master, namely the closure corresponding to function
-\of{worker} in that case, but on the worker side the initialization of
-\of{max_iterations} is never executed.
-
-One obvious solution is  not to use global variables in the
-worker code. This is not always possible, though. 
-To overcome this, the \of{Same} sub-module also provides a
-\of{Worker.compute} function to start the worker loop manually from
-the user code. This way, it can be started at any point, in particular
-after the initialization of the required global variables.
-Master and worker are still running the same executable, but are 
-distinguished using a user-defined way (command-line argument,
-environment variable, etc.).
-
-\bigskip
-There are situations where it is not possible
-to run the same executable for master and workers.
-For instance, architectures or operating systems could be different
-across the network.
-For that reason, the \of{Network} module provides two other
-implementations.
-
-\paragraph{\of{Poly}.} When master and workers are compiled with the
-same version of \Ocaml, we can no longer marshal closures but we can
-still marshal polymorphic values. Indeed, an interesting property of
-marshalling in \Ocaml\ is to be fully architecture-independent, as long
-as a single version of \Ocaml\ is used. It is worth pointing out that
-absence of marshaled closures
-now enables the use of two different programs for master and workers. This
-is not mandatory, though, since master and workers could still be
-distinguished at run-time as in the previous case.
-
-On the worker side, the main loop is started manually using
-\of{Worker.compute}. The computation to be performed on each task is
-given as an argument to this function. It thus looks as follows:
-\begin{ocaml}
-  Worker.compute : ('a -> 'b) -> unit -> unit
-\end{ocaml}
-On the master side, the \of{compute} function is simpler than in the
-previous case, as it has one argument less, and thus has the following
-signature. 
-\begin{ocaml}
-  Master.compute : master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit 
-\end{ocaml}
-For realistic applications, where master and workers are completely
-different programs, possibly written by different teams, this is the
-module of choice in our library, since it can still pass polymorphic
-values over the network. The issues of marshalling are automatically
-taken care of by \Ocaml\ run-time.
-
-The derived API presented in Section~\ref{sec:derived} is adapted to
-deal with the absence of closures. Exactly as the \of{compute}
-function, each API now takes two forms, one for the master and another
-for the workers. For example, \of{map_fold_ac} takes the following
-forms.
-\begin{ocaml}
-  Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit -> unit
-  Master.map_fold_ac : 'b -> 'a list -> 'b
-\end{ocaml}
-It is the responsibility of the user to ensure type consistency between
-master and workers.
-
-\paragraph{\of{Mono}.} When master and workers are compiled using
-different versions of \Ocaml, we can no longer use marshalling. 
-As in the previous case, we split \of{compute} into two functions, one
-for master and one for workers. In addition, values transmitted over
-the network can only be strings. The signature thus takes the
-following form.
-\begin{ocaml}
-  Worker.compute : (string -> string) -> unit -> unit
-  Master.compute : master:(string * 'c -> string -> (string * 'c) list) -> 
-                  (string * 'c) list -> unit
-\end{ocaml}
-Any other datatype for tasks should be encoded to/from strings. This
-conversion is left to the user.
-Note that the second component of each task is still polymorphic (of
-type \of{'c} here), since it is local to the master.
-
-% There is
-% actually a single, generic implementation, which is then instantiated
-% in three different ways, corresponding to the three scenarios.  The
-% generic implementation assumes that only strings are being transmitted
-% over the network, as evident from the \of{Assign} and \of{Completed}
-% messages in Section~\ref{sec:protocol}. Depending on the
-% implementation, these strings could be marshaled closures, marshaled
-% values or simply user strings.
-
-\subsubsection{Fault Tolerance}\label{sec:fault}
+\subsection{Fault Tolerance}\label{sec:fault}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Experiments}\label{sec:experiments}
@@
 case studies of Section~\ref{sec:studies}.
+
+% \section{Illustrative Case Studies}\label{sec:studies}
+
+In this section, we show how to use \functory\ on several case studies.
+We focus here on the use of the API. Experimental results of these
+case studies are given in Section~\ref{sec:experiments}.
+The source code for all these case studies is contained in the
+distribution, in sub-directory \texttt{tests}.
+
+\subsection{Matrix Multiplication}\label{sec:matrix}
+
+As a first example, let us consider matrix multiplication.
+We assume two matrices \of{a} and \of{b}, respectively of size
+$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
+a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
+Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
+the standard matrix multiplication would be as follows:
+\begin{ocaml}
+  for i = 0 to n-1 do
+    for j = 0 to m-1 do
+      for k = 0 to p-1 do
+	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
+      done
+    done
+  done  
+\end{ocaml}
+and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
+assuming addition and multiplication over coefficients to be 
+denoted $+$ and $\times$ respectively.
+An easy way to distribute this computation is obviously to turn each
+computation of the inner loop over \of{k} into a task.
+To do so, we first prepare the list of tasks:
+\begin{ocaml}
+  let tasks = 
+    let l = ref [] in
+    for i = 0 to n-1 do for j = 0 to m-1 do
+      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
+    done done;
+    !l
+\end{ocaml}
+Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
+tuple, together with position \of{(i,j)} as a second tuple.
+The worker is receiving the first tuple as argument and computes
+the dot product:
+\begin{ocaml}
+  let worker (ai, bj) =
+    let c = ref 0 in
+    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
+    !c
+\end{ocaml}
+The master is a one line function which receives the result \of{r}
+from the worker and simply stores it into \of{c}, according to the
+position contained in the input task. It produces no new task.
+\begin{ocaml}
+  let master (_, (i,j)) r = c.(i).(j) <- r; []
+\end{ocaml}
+Finally, the overall computation is started by invoking \of{compute}
+as follows:
+\begin{ocaml}
+  let () = compute ~worker ~master !tasks
+\end{ocaml}
+The total number of tasks is $\of{n}\times\of{m}$.
+
+Using the sequential implementation provided by \of{Functory} is as
+simple as including the following line in the code:
+\begin{ocaml}
+  open Sequential
+\end{ocaml}
+Then the computation is roughly similar to the standard multiplication
+described above.
+
+Now, let us assume we want to make use of a 4-core machine to perform
+the same computation. 
+This is achieved by just replacing the line above by
+\begin{ocaml}
+  open Cores
+  let () = set_number_of_cores 4
+\end{ocaml}
+while the rest of the code remains unchanged.
+
+Later, we may want to use a network of machines instead, for example
+two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
+cores respectively. Similar to the above, we turn the above lines into
+\begin{ocaml}
+  open Network
+  let () = declare_workers ~n:4 "orcus"
+  let () = declare_workers ~n:8 "belzebuth"
+  open Same
+\end{ocaml}
+Here, \of{Same} is a module which is to be used when master and worker
+are running the same executable. It provides a \of{compute} function
+which has the same signature as in modules \of{Sequential} and
+\of{Cores}, so that the rest of the code remains unchanged. Master and
+worker processes are distinguished at run-time using an environment
+variable \texttt{WORKER} which is set/unset.
+
+If we need to write two different programs for master and worker, for
+reason of binary incompatibility or any other reason, the library API
+is providing functions to do that. If master and worker are still
+compiled with the same version of \Ocaml, we use the \of{Poly} module
+which provides a polymorphic API. Let us start with the worker program.
+It now looks like:
+\begin{ocaml}
+  open Poly
+  let worker (ai, bj) = ...
+  let () = Worker.compute worker ()
+\end{ocaml}
+The \of{Worker.compute} function enters a loop which waits for
+tasks sent by the master and returns results computed using \of{worker}.
+The master program is almost the same as before. First, we replace
+module \of{Same} with module \of{Poly}:
+\begin{ocaml}
+  open Network
+  let () = declare_workers ~n:4 "orcus"
+  let () = declare_workers ~n:8 "belzebuth"
+  open Poly
+\end{ocaml}
+Tasks and \of{master} function are unchanged:
+\begin{ocaml}
+  let tasks = ...
+  let master (_, (i,j)) r = ...
+\end{ocaml}
+Finally, we start the computation with \of{Master.compute}, which does
+not have a \of{worker} parameter anymore:
+\begin{ocaml}
+  let () = Master.compute ~master tasks
+\end{ocaml}
+
+When master and worker programs are compiled with different
+versions of the \Ocaml\ compiler, our library still provides
+a monomorphic API over strings. As a consequence, we need to convert
+tasks and results to and from strings in both master and worker.
+The modified worker program then looks as follows:
+\begin{ocaml}
+  open Mono
+  let worker (ai, bj) = ...
+  let worker_string s = string_of_coeff (worker (task_of_string s))
+  let () = Worker.compute worker_string ()
+\end{ocaml}
+The master program is modified in a similar way.
+We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
+as strings, as follows:
+\begin{ocaml}
+  let tasks = ... string_of_task ...
+  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
+\end{ocaml}
+where \of{string_of_task} and \of{task_of_string} are user-defined
+functions to convert tasks to and from strings.
+
+\subsection{N-queens}\label{sec:n-queens}
+
+The next example is the classical $N$-queens problem, where we
+compute the total number of ways to place $N$ queens on a $N\times N$
+chessboard in such a way no two queens attack each other.
+We use a standard backtracking algorithm for this problem, which
+places the queens one by one starting from the first row.
+Distributing the computation is thus quite easy: we consider all
+possible ways to place queens on the first $D$ rows and then perform
+the subsequent search in parallel. Choosing $D=1$ will result in
+exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
+so on.
+% TODO: make a picture to illustrate positioning of 2 queens on the
+% first 2 rows
+
+Each task only consists of three integers and its result is one integer,
+which is the total number of solutions for this task.
+We make use of function \of{map_local_fold} from the derived API,
+where \of{f} is performing the search and \of{fold} simply adds the
+intermediate results.
+In the network configuration, we make use of the \of{Network.Same}
+module, workers and master being the same executable.
+
+\subsection{Mandelbrot Set}
+
+Drawing the Mandelbrot set is another classical example that could be
+distributed easily, since the color of each point can be computed
+independently of the others. 
+Let us assume we are given the coordinates of the region to be drawn,
+along with the width \of{w} and height \of{h} of the resulting image
+(in pixels). 
+Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
+parameter. It is immediate to split the image into \of{t} sub-images,
+each of which is computed in parallel with and independently of the
+others. For instance, the image could be split into horizontal slices
+or, more generally, into rectangular blocks.
+
+Each task is thus four floating-point numbers denoting the region
+coordinates, together with two integers denoting the dimensions of the
+sub-image to be drawn. The result of the task is a matrix of pixels,
+of size $(\of{w}\times\of{h}) / \of{t}$. 
+For instance, drawing a $800\times 600$ image using 20 tasks will
+result in 20 sub-images of size $176,000$ bytes each,
+assuming each pixel is encoded in four bytes.
+
+In the network configuration, we deliberately choose to have two
+different programs for master and workers, using the \of{Network.Poly}
+module (actually, a single source code with a command line option).
+
+% \subsection{Polynomial Multiplication}
+% TODO?
+
+\subsection{SMT Solvers}\label{sec:SMT}
+
+Here we demonstrate the potential of our library for our application
+needs as mentioned in the introduction. In this case study,
+we consider 80 challenging verification conditions
+(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
+VC is stored in a file, which is accessible over
+NFS. The purpose of the experiment is to check the validity of each VC
+using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
+which are all installed on each of the machines in use.
+
+The master program proceeds by reading the file names, turning them
+into tasks by multiplying them by the number of provers (more than
+300 tasks in total).
+Each worker in turn invokes the given prover on the given file, within
+a timeout limit.
+Each task completes with one of the four possible outcomes: \emph{valid},
+\emph{unknown} (depending on
+whether the VC is valid or undecided by the prover), 
+\emph{timeout} and \emph{failure}.
+
+The result of each computation is a pair denoting the status and the
+time spent in the prover call. The master collects these results and
+sums up the timings for each prover and each possible status.
 
@@
 
-\paragraph{Related Work.}
-Closest to the approach in this paper is Yohann Padioleau's MapReduce
-implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
-top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
-protocol for message passing.  Currently, we have less flexibility
-w.r.t. deployment of the user program than OCamlMPI; on the other
-hand, we provide a 
-more generic API together with fault tolerance.  There are other
-distributed computing libraries on top of which one could implement
-the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
-them. However, \JoCaml\ does not provide fault tolerance, which is
-indispensable in a distributed setting. The user has to include code
-for fault tolerance, as already demonstrated in some \JoCaml\
-experiments~\cite{mandel2008}.
-
-There are other implementations of distributed computing in the
-context of functional programming. One is the Disco
-project~\cite{disco}, which implements exactly Google's MapReduce in
-Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
-implementation of Google's MapReduce.
-There are other ways to exploit multi-core architectures. One of these
-is data parallelism, which is also relevant in the functional
-programming setting~\cite{parallel-haskell}. Our work
-does not target data parallelism at all.
-
 \paragraph{Future Work.}

### 2011-02-25 doc/tfp-2011/main.tex
*** empty log message ***

@@
 \section{Introduction}
+
+% TODO
+% Our library, \functory, is implemented in \Ocaml\ and is available
+% from \url{http://www.lri.fr/~filliatr/functory/}.
 
@@
 
-Our library, \functory, is implemented in \Ocaml\ and is available
-from \url{http://www.lri.fr/~filliatr/functory/}.
-In this section, we show the experimental results pertaining to the
-case studies of Section~\ref{sec:studies}.
-
-% \section{Illustrative Case Studies}\label{sec:studies}
-
-In this section, we show how to use \functory\ on several case studies.
-We focus here on the use of the API. Experimental results of these
-case studies are given in Section~\ref{sec:experiments}.
+In this section, we demonstrate the potential of using \functory\ on
+several case studies.
 The source code for all these case studies is contained in the
-distribution, in sub-directory \texttt{tests}.
-
-\subsection{Matrix Multiplication}\label{sec:matrix}
-
-As a first example, let us consider matrix multiplication.
-We assume two matrices \of{a} and \of{b}, respectively of size
-$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
-a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
-Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
-the standard matrix multiplication would be as follows:
-\begin{ocaml}
-  for i = 0 to n-1 do
-    for j = 0 to m-1 do
-      for k = 0 to p-1 do
-	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
-      done
-    done
-  done  
-\end{ocaml}
-and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
-assuming addition and multiplication over coefficients to be 
-denoted $+$ and $\times$ respectively.
-An easy way to distribute this computation is obviously to turn each
-computation of the inner loop over \of{k} into a task.
-To do so, we first prepare the list of tasks:
-\begin{ocaml}
-  let tasks = 
-    let l = ref [] in
-    for i = 0 to n-1 do for j = 0 to m-1 do
-      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
-    done done;
-    !l
-\end{ocaml}
-Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
-tuple, together with position \of{(i,j)} as a second tuple.
-The worker is receiving the first tuple as argument and computes
-the dot product:
-\begin{ocaml}
-  let worker (ai, bj) =
-    let c = ref 0 in
-    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
-    !c
-\end{ocaml}
-The master is a one line function which receives the result \of{r}
-from the worker and simply stores it into \of{c}, according to the
-position contained in the input task. It produces no new task.
-\begin{ocaml}
-  let master (_, (i,j)) r = c.(i).(j) <- r; []
-\end{ocaml}
-Finally, the overall computation is started by invoking \of{compute}
-as follows:
-\begin{ocaml}
-  let () = compute ~worker ~master !tasks
-\end{ocaml}
-The total number of tasks is $\of{n}\times\of{m}$.
-
-Using the sequential implementation provided by \of{Functory} is as
-simple as including the following line in the code:
-\begin{ocaml}
-  open Sequential
-\end{ocaml}
-Then the computation is roughly similar to the standard multiplication
-described above.
-
-Now, let us assume we want to make use of a 4-core machine to perform
-the same computation. 
-This is achieved by just replacing the line above by
-\begin{ocaml}
-  open Cores
-  let () = set_number_of_cores 4
-\end{ocaml}
-while the rest of the code remains unchanged.
-
-Later, we may want to use a network of machines instead, for example
-two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
-cores respectively. Similar to the above, we turn the above lines into
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Same
-\end{ocaml}
-Here, \of{Same} is a module which is to be used when master and worker
-are running the same executable. It provides a \of{compute} function
-which has the same signature as in modules \of{Sequential} and
-\of{Cores}, so that the rest of the code remains unchanged. Master and
-worker processes are distinguished at run-time using an environment
-variable \texttt{WORKER} which is set/unset.
-
-If we need to write two different programs for master and worker, for
-reason of binary incompatibility or any other reason, the library API
-is providing functions to do that. If master and worker are still
-compiled with the same version of \Ocaml, we use the \of{Poly} module
-which provides a polymorphic API. Let us start with the worker program.
-It now looks like:
-\begin{ocaml}
-  open Poly
-  let worker (ai, bj) = ...
-  let () = Worker.compute worker ()
-\end{ocaml}
-The \of{Worker.compute} function enters a loop which waits for
-tasks sent by the master and returns results computed using \of{worker}.
-The master program is almost the same as before. First, we replace
-module \of{Same} with module \of{Poly}:
-\begin{ocaml}
-  open Network
-  let () = declare_workers ~n:4 "orcus"
-  let () = declare_workers ~n:8 "belzebuth"
-  open Poly
-\end{ocaml}
-Tasks and \of{master} function are unchanged:
-\begin{ocaml}
-  let tasks = ...
-  let master (_, (i,j)) r = ...
-\end{ocaml}
-Finally, we start the computation with \of{Master.compute}, which does
-not have a \of{worker} parameter anymore:
-\begin{ocaml}
-  let () = Master.compute ~master tasks
-\end{ocaml}
-
-When master and worker programs are compiled with different
-versions of the \Ocaml\ compiler, our library still provides
-a monomorphic API over strings. As a consequence, we need to convert
-tasks and results to and from strings in both master and worker.
-The modified worker program then looks as follows:
-\begin{ocaml}
-  open Mono
-  let worker (ai, bj) = ...
-  let worker_string s = string_of_coeff (worker (task_of_string s))
-  let () = Worker.compute worker_string ()
-\end{ocaml}
-The master program is modified in a similar way.
-We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
-as strings, as follows:
-\begin{ocaml}
-  let tasks = ... string_of_task ...
-  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
-\end{ocaml}
-where \of{string_of_task} and \of{task_of_string} are user-defined
-functions to convert tasks to and from strings.
+distribution, in sub-directory \texttt{tests/}.
 
@@
 
-The next example is the classical $N$-queens problem, where we
+The first example is the classical $N$-queens problem, where we
 compute the total number of ways to place $N$ queens on a $N\times N$
@@
 module, workers and master being the same executable.
-
-\subsection{Mandelbrot Set}
-
-Drawing the Mandelbrot set is another classical example that could be
-distributed easily, since the color of each point can be computed
-independently of the others. 
-Let us assume we are given the coordinates of the region to be drawn,
-along with the width \of{w} and height \of{h} of the resulting image
-(in pixels). 
-Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
-parameter. It is immediate to split the image into \of{t} sub-images,
-each of which is computed in parallel with and independently of the
-others. For instance, the image could be split into horizontal slices
-or, more generally, into rectangular blocks.
-
-Each task is thus four floating-point numbers denoting the region
-coordinates, together with two integers denoting the dimensions of the
-sub-image to be drawn. The result of the task is a matrix of pixels,
-of size $(\of{w}\times\of{h}) / \of{t}$. 
-For instance, drawing a $800\times 600$ image using 20 tasks will
-result in 20 sub-images of size $176,000$ bytes each,
-assuming each pixel is encoded in four bytes.
-
-In the network configuration, we deliberately choose to have two
-different programs for master and workers, using the \of{Network.Poly}
-module (actually, a single source code with a command line option).
-
-% \subsection{Polynomial Multiplication}
-% TODO?
-
-\subsection{SMT Solvers}\label{sec:SMT}
-
-Here we demonstrate the potential of our library for our application
-needs as mentioned in the introduction. In this case study,
-we consider 80 challenging verification conditions
-(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
-VC is stored in a file, which is accessible over
-NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
-which are all installed on each of the machines in use.
-
-The master program proceeds by reading the file names, turning them
-into tasks by multiplying them by the number of provers (more than
-300 tasks in total).
-Each worker in turn invokes the given prover on the given file, within
-a timeout limit.
-Each task completes with one of the four possible outcomes: \emph{valid},
-\emph{unknown} (depending on
-whether the VC is valid or undecided by the prover), 
-\emph{timeout} and \emph{failure}.
-
-The result of each computation is a pair denoting the status and the
-time spent in the prover call. The master collects these results and
-sums up the timings for each prover and each possible status.
-
-\subsection{N-queens}
 
@@
 
+\subsection{Matrix Multiplication}\label{sec:matrix}
+
+This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
+It consists of multiplication of
+two square matrices of dimension 100 with integer coefficients.
+Coefficients have several thousands of digits, hence we use
+GMP~\cite{GMP} to handle operations over coefficients.
+
+We compare the performances of two different implementations. In the
+first one, called \of{mm1}, each task consists of the computation of a single
+coefficient of the resultant matrix. 
+In the second one, called \of{mm2}, each task consists of the
+computation of a whole row of the resultant matrix.
+As a consequence, the total number of tasks in
+$10,000$ for \of{mm1} and only $100$ for \of{mm2}.
+On the contrary, each task result for \of{mm1} is a single integer,
+while for \of{mm2} it is a row of 100 integers.
+%
+The experimental results (in seconds) are tabulated below.
+\begin{center}
+  \begin{tabular}{|r|r|r|}
+    \hline
+    & \of{mm1}       & \of{mm2}  \\
+    & (10,000 tasks) & (100 tasks) \\
+    \hline\hline
+\of{Sequential} & 20.3 &  20.2 \\
+\hline
+ \of{Cores} 
+ (2 cores)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
+ (4 cores)     &   12.3 (1.65$\times$) &   6.1 (3.31$\times$) \\
+ (6 cores)     &    8.6 (2.36$\times$) &   4.3 (4.70$\times$) \\
+ (8 cores)     &    8.0 (2.54$\times$) &   3.5 (\underline{5.77}$\times$) \\
+ \hline
+  \end{tabular}
+\end{center}
+We do not include results for the network configuration, as they do
+not achieve any benefit with respect to the sequential
+implementation. The reason is that the communication cost dominates the
+computation cost in such a way that the total execution time is
+always greater than 30 seconds. Indeed, irrespective of the
+implementation (\of{mm1} or \of{mm2}), the total size of the
+transmitted data is $10^6$ integers, which in our case amounts to
+billions of bytes.
+
+A less naive implementation would have the worker read the input matrices
+only once, \emph{e.g.} from a file, and then have the master send only
+row and column indices. This would reduce the amount of transmitted data to
+$10,000$ integers only.
+
+%             mm1     mm2
+
+% # tasks    10,000   100
+
+% sequential  20.3   20.2
+
+% cores 2     22.7   11.3
+% cores 4     12.3    6.1
+% cores 6      8.6    4.3
+% cores 8      8.0    3.5
+
+% network = workers on moloch + master on belzebuth
+
+% cores 2     93     32.3
+% cores 4     87     33.7
+% cores 6     92     33.5
+% cores 8     88     32.5
+
+\begin{comment}
+  As a first example, let us consider matrix multiplication.  We
+  assume two matrices \of{a} and \of{b}, respectively of size
+  $\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as
+  well as a matrix \of{c} of size $\of{n}\times\of{m}$ to store the
+  result.  Assuming \of{a} being a row-matrix and \of{b} a
+  column-matrix, the standard matrix multiplication would be as
+  follows:
+  \begin{ocaml}
+    for i = 0 to n-1 do for j = 0 to m-1 do for k = 0 to p-1 do
+    c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k) done done done
+  \end{ocaml}
+  and runs in $O(\of{n}\times\of{p}\times\of{m})$, assuming addition
+  and multiplication over coefficients to be denoted $+$ and $\times$
+  respectively.  An easy way to distribute this computation is
+  obviously to turn each computation of the inner loop over \of{k}
+  into a task.  To do so, we first prepare the list of tasks:
+  \begin{ocaml}
+    let tasks = let l = ref [] in for i = 0 to n-1 do for j = 0 to m-1
+    do tasks := ((a.(i), b.(j)), (i,j)) :: !tasks done done; !l
+  \end{ocaml}
+  Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
+  tuple, together with position \of{(i,j)} as a second tuple.  The
+  worker is receiving the first tuple as argument and computes the dot
+  product:
+  \begin{ocaml}
+    let worker (ai, bj) = let c = ref 0 in for k = 0 to p-1 do c := !c
+    + ai.(k) * bj.(k) done; !c
+  \end{ocaml}
+  The master is a one line function which receives the result \of{r}
+  from the worker and simply stores it into \of{c}, according to the
+  position contained in the input task. It produces no new task.
+  \begin{ocaml}
+    let master (_, (i,j)) r = c.(i).(j) <- r; []
+  \end{ocaml}
+  Finally, the overall computation is started by invoking \of{compute}
+  as follows:
+  \begin{ocaml}
+    let () = compute ~worker ~master !tasks
+  \end{ocaml}
+  The total number of tasks is $\of{n}\times\of{m}$.
+
+  Using the sequential implementation provided by \of{Functory} is as
+  simple as including the following line in the code:
+  \begin{ocaml}
+    open Sequential
+  \end{ocaml}
+  Then the computation is roughly similar to the standard
+  multiplication described above.
+
+  Now, let us assume we want to make use of a 4-core machine to
+  perform the same computation.  This is achieved by just replacing
+  the line above by
+  \begin{ocaml}
+    open Cores let () = set_number_of_cores 4
+  \end{ocaml}
+  while the rest of the code remains unchanged.
+
+  Later, we may want to use a network of machines instead, for example
+  two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and
+  8 cores respectively. Similar to the above, we turn the above lines
+  into
+  \begin{ocaml}
+    open Network let () = declare_workers ~n:4 "orcus" let () =
+    declare_workers ~n:8 "belzebuth" open Same
+  \end{ocaml}
+  Here, \of{Same} is a module which is to be used when master and
+  worker are running the same executable. It provides a \of{compute}
+  function which has the same signature as in modules \of{Sequential}
+  and \of{Cores}, so that the rest of the code remains
+  unchanged. Master and worker processes are distinguished at run-time
+  using an environment variable \texttt{WORKER} which is set/unset.
+
+  If we need to write two different programs for master and worker,
+  for reason of binary incompatibility or any other reason, the
+  library API is providing functions to do that. If master and worker
+  are still compiled with the same version of \Ocaml, we use the
+  \of{Poly} module which provides a polymorphic API. Let us start with
+  the worker program.  It now looks like:
+  \begin{ocaml}
+    open Poly let worker (ai, bj) = ...  let () = Worker.compute
+    worker ()
+  \end{ocaml}
+  The \of{Worker.compute} function enters a loop which waits for tasks
+  sent by the master and returns results computed using \of{worker}.
+  The master program is almost the same as before. First, we replace
+  module \of{Same} with module \of{Poly}:
+  \begin{ocaml}
+    open Network let () = declare_workers ~n:4 "orcus" let () =
+    declare_workers ~n:8 "belzebuth" open Poly
+  \end{ocaml}
+  Tasks and \of{master} function are unchanged:
+  \begin{ocaml}
+    let tasks = ...  let master (_, (i,j)) r = ...
+  \end{ocaml}
+  Finally, we start the computation with \of{Master.compute}, which
+  does not have a \of{worker} parameter anymore:
+  \begin{ocaml}
+    let () = Master.compute ~master tasks
+  \end{ocaml}
+
+  When master and worker programs are compiled with different versions
+  of the \Ocaml\ compiler, our library still provides a monomorphic
+  API over strings. As a consequence, we need to convert tasks and
+  results to and from strings in both master and worker.  The modified
+  worker program then looks as follows:
+  \begin{ocaml}
+    open Mono let worker (ai, bj) = ...  let worker_string s =
+    string_of_coeff (worker (task_of_string s)) let () =
+    Worker.compute worker_string ()
+  \end{ocaml}
+  The master program is modified in a similar way.  We simply replace
+  \of{Poly} with \of{Mono} and encode/decode coefficients as strings,
+  as follows:
+  \begin{ocaml}
+    let tasks = ... string_of_task ...  let master (_, (i,j)) r =
+    c.(i).(j) <- coeff_of_string r; []
+  \end{ocaml}
+  where \of{string_of_task} and \of{task_of_string} are user-defined
+  functions to convert tasks to and from strings.
+\end{comment}
+
 \subsection{Mandelbrot Set}
 
+Drawing the Mandelbrot set is another classical example that could be
+distributed easily, since the color of each point can be computed
+independently of the others. 
 This benchmark consists in drawing the fragment of the Mandelbrot set
 with lower left corner $(-1.1, 0.2)$ and upper right corner $(-0.8,
-0.4)$, as a $9,000\times6,000$ image. The sequential computation of
+0.4)$, as a $9,000\times6,000$ image. 
+%
+If the total number of tasks $\of{t} \ge 1$ is given as a
+parameter,  it is immediate to split the image into \of{t} sub-images,
+each of which is computed in parallel with and independently of the
+others. In our case, the image is split into horizontal slices.
+Each task is thus four floating-point numbers denoting the region
+coordinates, together with two integers denoting the dimensions of the
+sub-image to be drawn. The result of the task is a matrix of pixels,
+of size $54,000,000 / \of{t}$. 
+For instance, using $\of{t}=20$ tasks will
+result in 20 sub-images of size $10.3$ Mb each,
+assuming each pixel is encoded in four bytes.
+
+The sequential computation of
 this image consumes 29.4 seconds. For \of{Cores} and \of{Network}
@@
 
-\subsection{Matrix Multiplication}
-
-This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
-It consists of multiplication of
-two square matrices of dimension 100, that is $\of{n} = \of{p} =
-\of{m} = 100$, with integer coefficients.
-Coefficients have several thousands of digits, hence we use
-GMP~\cite{GMP} to handle operations over coefficients.
-
-We compare the performances of two different implementations. In the
-first one, called \of{mm1}, each task consists of the computation of a single
-coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
-In the second one, called \of{mm2}, each task consists of the
-computation of a whole row of the resultant matrix.
-As a consequence, the total number of tasks in
-$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
-The experimental results (in seconds) are tabulated below.
-\begin{center}
-  \begin{tabular}{|r|r|r|}
-    \hline
-    & \of{mm1}       & \of{mm2}  \\
-    & (10,000 tasks) & (100 tasks) \\
-    \hline\hline
-\of{Sequential} & 20.3 &  20.2 \\
-\hline
- \of{Cores} 
- (2 cores)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
- (4 cores)     &   12.3 (1.65$\times$) &   6.1 (3.31$\times$) \\
- (6 cores)     &    8.6 (2.36$\times$) &   4.3 (4.70$\times$) \\
- (8 cores)     &    8.0 (2.54$\times$) &   3.5 (\underline{5.77}$\times$) \\
- \hline
-  \end{tabular}
-\end{center}
-We do not include results for the network configuration, as they do
-not achieve any benefit with respect to the sequential
-implementation. The reason is that the communication cost dominates the
-computation cost in such a way that the total execution time is
-always greater than 30 seconds. Indeed, irrespective of the
-implementation (\of{mm1} or \of{mm2}), the total size of the
-transmitted data is 
-$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
-billions of bytes.
-
-A less naive implementation would have the worker read the input matrices
-only once, \emph{e.g.} from a file, and then have the master send only
-row and column indices. This would reduce the communication cost to
-$O(\of{n}\times\of{m})$ only.
-
-%             mm1     mm2
-
-% # tasks    10,000   100
-
-% sequential  20.3   20.2
-
-% cores 2     22.7   11.3
-% cores 4     12.3    6.1
-% cores 6      8.6    4.3
-% cores 8      8.0    3.5
-
-% network = workers on moloch + master on belzebuth
-
-% cores 2     93     32.3
-% cores 4     87     33.7
-% cores 6     92     33.5
-% cores 8     88     32.5
+\subsection{SMT Solvers}\label{sec:SMT}
+
+% ICI
+
+Here we demonstrate the potential of our library for our application
+needs as mentioned in the introduction. In this case study,
+we consider 80 challenging verification conditions
+(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
+VC is stored in a file, which is accessible over
+NFS. The purpose of the experiment is to check the validity of each VC
+using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
+which are all installed on each of the machines in use.
+
+The master program proceeds by reading the file names, turning them
+into tasks by multiplying them by the number of provers (more than
+300 tasks in total).
+Each worker in turn invokes the given prover on the given file, within
+a timeout limit.
+Each task completes with one of the four possible outcomes: \emph{valid},
+\emph{unknown} (depending on
+whether the VC is valid or undecided by the prover), 
+\emph{timeout} and \emph{failure}.
+
+The result of each computation is a pair denoting the status and the
+time spent in the prover call. The master collects these results and
+sums up the timings for each prover and each possible status.
 

### 2011-03-01 doc/tfp-2011/main.tex
experiments

@@
 
+The purpose of the following experiments is to compare the various
+deployments, namely sequential, cores and network.
+For this comparison to be fair, all
+computations are performed on the same machine, a 8 cores Intel Xeon
+3.2 GHz running Debian Linux. The sequential implementation uses a
+single core. The multi-core implementation uses up to 8 cores of
+the machine. The network implementation uses 8 workers running locally and
+a master running on a remote machine (which incurs communication cost).
+
 \subsection{N-queens}\label{sec:n-queens}
@@
 \of{Network}. The purpose of this experiment is to measure the speedup
-w.r.t. the sequential implementation. Therefore, all
-computations are performed on the same machine, a 8 cores Intel Xeon
-3.2 GHz running Debian Linux. The sequential implementation uses a
-single core. The multi-core implementation uses the 8 cores of
-the machine. The network implementation uses 8 workers running locally and
-a master running on a remote machine (which incurs communication cost).
+w.r.t. the sequential implementation.
 
@@
 
-% ICI
-
 Here we demonstrate the potential of our library for our application
-needs as mentioned in the introduction. In this case study,
-we consider 80 challenging verification conditions
+needs as mentioned in the introduction. 
+We consider 80 challenging verification conditions
 (VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
@@
 NFS. The purpose of the experiment is to check the validity of each VC
-using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
-which are all installed on each of the machines in use.
+using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3).
 
 The master program proceeds by reading the file names, turning them
-into tasks by multiplying them by the number of provers (more than
-300 tasks in total).
+into tasks by multiplying them by the number of provers, resulting in
+320 tasks in total.
 Each worker in turn invokes the given prover on the given file, within
-a timeout limit.
+a timeout limit of 1 minute.
 Each task completes with one of the four possible outcomes: \emph{valid},
@@
 \emph{timeout} and \emph{failure}.
-
 The result of each computation is a pair denoting the status and the
@@
 
-\subsection{SMT Solvers}
-
-As explained in Section~\ref{sec:SMT}, this benchmark consists of 80
-verification conditions, each being checked by 4 different SMT
-solvers. Each task is executed with a timeout limit of 1 minute.
 Our computing

### 2011-03-01 doc/tfp-2011/main.tex
high-level API

@@
 The paper is organized as follows. 
-In Section~\ref{sec:API}, we
-describe the main idea behind the generic API, and introduce
-\of{master} and \of{worker} programs. 
-In Section~\ref{sec:derived}, we
-present some additional functions provided by \functory,
-which are derived from the generic
-API. 
-Section~\ref{sec:studies} illustrates the API usage with several
-case studies.
-Section~\ref{sec:implem} delves into the implementation details
-of our library, while Section~\ref{sec:experiments} illustrates the
-potential of the presented library through experimental evaluation. We
-compare our approach with relevant related work in
-Section~\ref{sec:future} and outline our future work.
+Section~\ref{sec:API} describes the API of our library.
+Section~\ref{sec:implem} gives implementation details, focusing
+on marshaling, network protocol and fault tolerance.
+Then Section~\ref{sec:experiments} illustrates the
+potential of the presented library through experimental evaluation. 
+Finally, Section~\ref{sec:future} outlines some future work.
 
@@
 
-\subsection{compute}
-
-In this section, we describe the
-generic API which we shall extend with several specialized
-functions in the next section. The main function in our API follows the idea
+This section describes our API. We start from a generic distribution
+function, \of{compute}, as sketched in the introduction. Then we
+explain how this function is actually implemented in terms of
+low-level primitives, which are also provided in our API.
+The other way round, we also explain how the \of{compute} function can
+be used to implement high-level distribution functions for map and
+fold operations. Finally, we explain how these three APIs are
+implemented in five different ways, according to five different deployment
+scenarios.
+
+\subsection{A Generic Distribution Function}
+
+The generic distribution function in our API follows the idea
 sketched in the introduction. It has the following signature:
@@
 in all worker processes. The function \ocaml{master}, on the
-contrary, can be impure and is only executed in the master process.
+contrary, can be impure and is only executed sequentially in the master process.
 The \ocaml{master} function typically stores results 
@@
 pending tasks.
-The next section will describe the usage of this generic interface to
-perform the traditional map and fold operations.
 
@@
 
-\subsection{Low-level API}
+\subsection{Low-level Primitives}
 
@@
 
-\subsection{Derived API}\label{sec:derived}
+\subsection{High-level API}\label{sec:derived}
 
@@
 list elements are done in parallel. 
-% The previous section illustrated one such case.
 To facilitate such a processing,
-we now derive the most commonly used list operations from our generic
-API.
+our library provides most commonly used list operations, all 
+implemented using our generic \of{compute} function.
 
@@
 \begin{ocaml}
-  val map : ('a -> 'b) -> 'a list -> 'b list
-\end{ocaml}
-The next natural operation is a combination of map and fold
-operations, that is a function like
-\begin{ocaml}
- val map_fold : f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
+  val map : f:('a -> 'b) -> 'a list -> 'b list
+\end{ocaml}
+Each task consists of application of function \of{f} to a list element.
+More interesting is a combination of map and fold operations.
+For instance, we provide different flavors of function
+\begin{ocaml}
+  val map_fold :
+    f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
 \end{ocaml}
@@
 Regarding \of{fold} operations, we distinguish two cases:
-\begin{itemize}
-\item either \of{fold} operations are computationally less expensive
+either \of{fold} operations are computationally less expensive
 than \of{f} and we perform them locally;
-\item or \of{fold} operations are computationally expensive and we
+or \of{fold} operations are computationally expensive and we
 perform them in parallel.
-\end{itemize}
-We thus provide two functions \of{map_local_fold} and \of{map_remote_fold}.
+Thus we provide two functions \of{map_local_fold} and \of{map_remote_fold}.
 
@@
 operations in parallel. Thus our API provides two functions
-\of{map_fold_a} and \of{map_fold_ac} for these two particular cases.
-The five operations of the derived API are summarized
-in Figure~\ref{fig:derived}.
-\begin{figure}[t]
-  \begin{ocaml}
-    val map : 
-      f:('a -> 'b) -> 'a list -> 'b list 
-    val map_local_fold : 
-      f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_remote_fold : 
-      f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
-      'c -> 'a list -> 'c 
-    val map_fold_ac : 
-      f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b 
-    val map_fold_a : 
-      f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
-      'b -> 'a list -> 'b
-  \end{ocaml}
-  \caption{Derived API.}
-\label{fig:derived}
-\end{figure}
-
-These five functions are actually derived from the generic API.  The
-generality is achieved by implementing these derivations as a functor.
-This functor, \of{Derive}, is parameterized by the \of{compute}
-function, as follows:
-\begin{ocaml}
-module Derive
-  (X : sig
-     val compute : 
-       worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a * 'c) list) ->
-       ('a * 'c) list -> unit
-   end) = struct ... end
-\end{ocaml}
-We now explain how each function from Figure~\ref{fig:derived} is
-implemented using \of{compute} inside the functor body.
-
-\paragraph{\of{map}.} 
-Function \of{map} is easily implemented by associating an integer
-index with each element (of type \of{'a}) of the input list. Thus the
-list of tasks passed to \of{compute} is a list of pairs of type \of{'a * int}.
-As soon as an intermediate result is available, it is passed to
-\of{master}, which stores it in a local table using the index as a key
-and generates no new task. Whenever \of{compute} returns, we easily
-recover the output list from the values stored in the table.
-
-\paragraph{\of{map_local_fold}.} 
-Function \of{map_local_fold} is implemented using a local reference
-storing the current accumulator (of type \of{'c}).  As soon as an
-intermediate result is available, the \of{master} combines it with the
-accumulator using the function \of{fold}. Whenever \of{compute}
-returns, we simply return the current accumulator.
-
-\paragraph{\of{map_remote_fold}.}
-In this case, combining the accumulator with an intermediate result is
-itself a task to be performed in parallel with \of{f} operations.
-Thus tasks are of two different kinds and we use a sum type to make
-the distinction.
-When an intermediate \of{f} result is available, the
-accumulator may be already in use in a \of{fold} operation. Hence we
-need to store the intermediate \of{f} result in a table local to the
-master, to be
-used as soon as the accumulator becomes available.
-The \of{master} function distinguishes between \of{f} and \of{fold}
-results and, accordingly, updates the table and the accumulator. For a
-\of{f} result, either we combine it with the accumulator in a new
-task or we simply store it in the local table. For a \of{fold} result,
-either we simply store it into the accumulator or we immediately
-combine it with a pending \of{f} result from the table.
-Whenever \of{compute} returns, the accumulator must be available and
-we simply return it.
-
-\paragraph{\of{map_fold_ac}.}
-The function \of{map_fold_ac} is easier than the previous one, since
-we do not need to store more than one intermediate result. Indeed, as
-soon as two intermediate results are available, we immediately combine
-them using a \of{fold} operation. Thus, whenever the \of{master}
-receives some result, it either stores it or combines it with the
-accumulator, depending on the availability/non availability of the
-accumulator.  Whenever \of{compute} returns, the accumulator must be
-available and we simply return it.
-
-\paragraph{\of{map_fold_a}.}
-Function \of{map_fold_a} is definitely the trickiest. Indeed, the
-operation being only associative, we can only combine \emph{adjacent}
-intermediate results. In order to perform this, we store intermediate
-results in a local table according to indices of the input list. More
-precisely, if we have already computed
-\begin{displaymath}
-  x_i \oplus x_{i+1} \oplus \dots \oplus x_j
-\end{displaymath}
-where $\oplus$ denotes the \of{fold} operation, then we associate this
-value with indices $i,j$. As soon as a result of indices $k,i-1$ or
-$j+1,k$ is available, it can be combined with the result of $i,j$.
-Whenever \of{compute} returns, the table should only contain a result
-corresponding to the full input list and we simply return it.
-
-\medskip
-These five functions are the most natural derivations of the
-\of{compute} functions. There are obviously other possible derivations
-(e.g. a variant of \of{map_fold} where only \of{fold} is meaningful);
-in most cases, they could be derived in a straightforward way from our
-generic API.
+\of{map_fold_a} and \of{map_fold_ac} for these two particular cases,
+with types
+\begin{ocaml}
+  val map_fold_ac, map_fold_a : 
+    f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
+\end{ocaml}
+This is rather straightforward to derive these five functions from the
+generic \of{compute} function; readers interested in details could
+refer to the source code.
 
@@
 % derive from this generic API.
+
+\begin{center}
+  \begin{tabular}{|l|c|c|c|c|c|}
+    \hline
+    & & & \multicolumn{3}{|c|}{Network} \\
+    & Seq. & Cores & Same & Poly & Mono \\\hline\hline
+    low-level API & & & & & \\\hline
+    \of{compute}  & & & & & \\\hline
+    high-level API& & & & & \\\hline
+  \end{tabular}
+\end{center}
 

### 2011-03-01 doc/tfp-2011/main.tex
*** empty log message ***

@@
 
-Actually, our library provides not just a single \ocaml{compute}
-function as above, but instead five different versions
-depending on the execution context. 
-The first two contexts are the following: 
+Actually, our library provides not just one implementation for the API above,
+but instead five different implementations
+depending on the deployment scenario. 
+The first two scenarios are the following: 
 \begin{enumerate}
@@
 \end{enumerate}
-The next three implementations are intended for distributing the
+The next three scenarios are intended for distributing the
 computation over a network of machines.
@@
   closures and polymorphic values.
-  Depending on whether the program is run as a master or as a worker,
-  the relevant arguments of \ocaml{compute} are used.
+%   Depending on whether the program is run as a master or as a worker,
+%   the relevant arguments of \ocaml{compute} are used.
 
@@
   we can no longer marshal closures but we can still
-  marshal polymorphic values. As a consequence, 
-  the \ocaml{compute} function is split into two
-  polymorphic functions, to implement the master and workers 
-  respectively:%
-\vspace{-1em}
-\begin{ocaml}
-val Worker.compute : ('a -> 'b) -> unit -> unit
-val Master.compute : 
-  ('a * 'c -> 'b -> ('a * 'c) list) -> 
-  ('a * 'c) list -> unit
-\end{ocaml}
+  marshal polymorphic values. API functions are split into two sets,
+  used to implement master and workers respectively.
 
@@
     compiled with the same version of \Ocaml:} we can no
-  longer use marshalling, so the
-  \ocaml{compute} function is split into two monomorphic functions
-  over strings:%
-\vspace{-1em}
-\begin{ocaml}
-val Worker.compute : (string -> string) -> unit -> unit
-val Master.compute : 
-  (string * 'c -> string -> (string * 'c) list) -> 
-  (string * 'c) list -> unit
-\end{ocaml}
+  longer use marshalling, so API functions are restricted to work on
+  strings instead of polymorphic values.
 \end{enumerate}
-
 Our library is organized into three modules: \of{Sequential} for the
@@
 corresponding to contexts 3, 4 and 5 above.
-The next section presents more functions provided by \functory, which
-are all derived from the generic API just presented.
-
-% The different
-% API functions are summarized in the following table (their signatures
-% being given above):
-% \begin{center}
-%   \begin{tabular}{|c|c|c|c|c|}
-%     \hline
-%     \ocaml{Sequential} & \ocaml{Cores} &
-%     \multicolumn{3}{|c|}{\ocaml{Network}} 
-%     \\\cline{3-5}
-%     &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
-%     \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
-%     \ocaml{master}  & \ocaml{master}  \\
-%     & & & \ocaml{worker}  & \ocaml{worker} \\\hline
-%   \end{tabular}
-% \end{center}
-% The following section introduces other interesting APIs that we
-% derive from this generic API.
-
+
+\subsection{Functory API at a Glance} % ???
+
+From  the description above, it is clear that our library provides
+three different APIs, for five different deployment scenarios,
+resulting in fifteen possible ways of using it.
+This is summarized in the following matrix.
 \begin{center}
@@
     \hline
-    & & & \multicolumn{3}{|c|}{Network} \\
+    & & & \multicolumn{3}{c|}{Network} \\
     & Seq. & Cores & Same & Poly & Mono \\\hline\hline
@@
 \end{center}
-
-% TODO cannot be instantiated in the monomorphic case
+From the implementation point of view, there is almost no code
+duplication. We are using \Ocaml\ functors to derive the second line
+from the first, as well as the third from the second, irrespective of
+the deployment scenario.
 
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables OCamlMPI MapReduce marshalling
+% LocalWords:  genericity executables OCamlMPI MapReduce marshalling functors

### 2011-03-01 doc/tfp-2011/main.tex
*** empty log message ***

@@
 
-% TODO
+% motivation is finer control over task/workers/execution
+% example 1: remove *some* tasks when others are completed
+% example 2: interactive program (GUI, control, etc.)
+
+The function \of{compute} above can actually be implemented in terms
+of low-level primitives, such as adding a task, adding a worker,
+performing one step of communication between master and workers, etc.
+These primitives are provided in our API, such that the user can have
+finer control over the execution of the distributed computation.
+A type for distributed computations is introduced.
+\begin{ocaml}
+  type ('a, 'c) computation
+\end{ocaml}
+A computation is created with a function \of{create}, which accepts
+the same \of{worker} and \of{master} as \of{compute}:
+\begin{ocaml}
+  val create : 
+    worker:('a -> 'b) -> 
+    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a, 'c) computation
+\end{ocaml}
+Contrary to \of{compute}, it takes no list of tasks and returns
+immediately. Tasks can be added later using the following function:
+\begin{ocaml}
+  val add_task : ('a, 'c) computation -> 'a * 'c -> unit
+\end{ocaml}
+A function is provided to perform \emph{one step} of a given
+computation:
+\begin{ocaml}
+  val one_step : ('a, 'c) computation -> unit
+\end{ocaml}
+Calling this function results in one exchange of messages between
+master and workers: task assignments to workers, results returned to
+the master, etc. A few other functions are provided, such as
+\of{status} to query the status of a computation, \of{clear} to remove
+all tasks, etc.
+
+Using these low-level primitives, it is straightforward to implement
+the \of{compute} function. Basically, it is as simple as the following:
+\begin{ocaml}
+let compute ~worker ~master tasks =
+  let c = create worker master in
+  List.iter (add_task c) tasks;
+  while status c = Running do one_step c done
+\end{ocaml}
 

### 2011-03-02 doc/tfp-2011/main.tex
updated (still incomplete) related work subsection [KK]

@@
 \paragraph{Related Work.}
+
+There are several implementations of Google's Map/Reduce in functional
+programming languages. But Functory was just inspired by Google's
+Map/Reduce and is not exactly a Map/Reduce implementation. The
+simplest difference comes from the very fact that Functory does not
+operate on key/value pairs. In order to compare and distinguish
+Functory from other work with related goals and motivations, we can
+broadly classify the related work in this domain into:
+\begin{enumerate}
+\item Distributed Functional Languages (DFLs) - functional
+  languages that provide built-in primitives for
+  distribution. Examples include ML5, JoCaml, Glasgow Distributed
+  Haskell, Erlang, etc.
+\item Libraries for existing and well-known functional programming
+  languages that could be used in order to relieve the user from
+  implementing intricate details of distribution like
+  fault-tolerance, socket programming, etc.
+\end{enumerate}
+
+\JoCaml\ is one of the DFLs which provides communication primitives
+(like channels) for transmitting computations. On the other hand,
+\JoCaml\ can not in a true sense be used as a distributed computing
+library. It does not provide fault tolerance, which is indispensable
+in a distributed setting. The user has to include code for fault
+tolerance, as already demonstrated in some \JoCaml\
+experiments~\cite{mandel2008}.
+
+ML5~\cite{ML5}, a variant of ML, is a programming language for
+distributed computing, specialized for web programming. It provides
+primitives for transferring control between the client and the server,
+as well as low-level primitives for marshalling the data. In
+particular, it achieves type-safe marshalling. As in the case before,
+ML5 is a programming language that offers primitives for code
+mobility, rather than a user-friendly API for distribution of
+computation and fault-tolerance.
+
+Glasgow Distributed Haskell
+
+The above mentioned languages are programming languages themselves,
+our library, Functory, could be implemented using any of the above.
+
+We now turn to distributed computing libraries for general purpose
+functional languages and weed out the distinguishing features of
+Functory from other similar libraries.
+
+OcamlP3L
+
+-----------  BELOW THIS IS THE OLD RELATED WORK --------
+
+% -- BELOW HERE IS OLD RELATED-WORK
+
 Closest to the approach in this paper is Yohann Padioleau's MapReduce

### 2011-03-02 doc/tfp-2011/main.tex
updated (not yet complete) related work subsection [KK]

@@
 \begin{enumerate}
-\item Distributed Functional Languages (DFLs) - functional
+\item \textit{Distributed Functional Languages (DFLs)} - functional
   languages that provide built-in primitives for
@@
   Haskell, Erlang, etc.
-\item Libraries for existing and well-known functional programming
-  languages that could be used in order to relieve the user from
-  implementing intricate details of distribution like
-  fault-tolerance, socket programming, etc.
+\item \textit{Libraries for existing functional languages} - that
+  could be readily used in order to avoid implementing details like
+  task distribution, fault-tolerance, socket programming, etc.
 \end{enumerate}
@@
 tolerance, as already demonstrated in some \JoCaml\
-experiments~\cite{mandel2008}.
-
-ML5~\cite{ML5}, a variant of ML, is a programming language for
-distributed computing, specialized for web programming. It provides
-primitives for transferring control between the client and the server,
-as well as low-level primitives for marshalling the data. In
-particular, it achieves type-safe marshalling. As in the case before,
-ML5 is a programming language that offers primitives for code
-mobility, rather than a user-friendly API for distribution of
-computation and fault-tolerance.
-
-Glasgow Distributed Haskell
-
-The above mentioned languages are programming languages themselves,
-our library, Functory, could be implemented using any of the above.
+experiments~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
+programming language for distributed computing, specialized for web
+programming. It provides primitives for transferring control between
+the client and the server, as well as low-level primitives for
+marshalling the data. In particular, it achieves type-safe
+marshalling. As in the case before, ML5 is a programming language that
+offers primitives for code mobility, rather than a user-friendly API
+for distribution of computation and fault-tolerance. Glasgow
+\textbf{Distributed Haskell: TODO} Erlang~\cite{erlang} is a
+programming language which has features for distribution and fault
+tolerance. It has features for task distribution and is more
+well-known for its rich error detection primitives and the ability to
+support hot-swapping. The error detection primitives of Erlang allow
+nodes to monitor processes in other nodes and also facilitate
+automatic migration of tasks in failed nodes to recovered or active
+nodes. CamlP3l~\cite{camlP3l} mixes the features of functional
+programming with predefined patterns for parallel computation to offer
+a parallel programming environment. Again, it is a programming
+language offering primitives for distributing computation to parallel
+processes and also to merge the results from parallel executions. The
+above mentioned languages are programming languages themselves, our
+library, Functory, could be implemented using any of the above. The
+main goal of Functory is to provide the users of an existing
+functional programming language high-level user-friendly APIs that
+hides the messy details of task distribution and fault-tolerance,
+enabling a ready-to-use library for efficient utilization of computing
+resources.
 
@@
 
-OcamlP3L
-
------------  BELOW THIS IS THE OLD RELATED WORK --------
+PlasmaMR~\cite{plasma} is an implementation of Google's MapReduce on a
+distributed file system PlasmaFS. It is able to use PlasmaFS to its
+advantage - the ability of the file system to handle large files and
+query functions of the functions that implement data locality to
+optimize network traffic. % However, Plasma
+% MapReduce does not support fault tolerance.
+The Disco
+project~\cite{disco} is an implementation of Google's MapReduce in
+Erlang~\cite{erlang}. Our library is not an impementation of Google's
+MapReduce. A MapReduce implementation in \Ocaml\ is Yohann Padioleau's
+MapReduce implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is
+built on top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a
+homemade protocol for message passing.  Currently, we have less
+flexibility w.r.t. deployment of the user program than OCamlMPI; on
+the other hand, we provide a more generic API together with fault
+tolerance.
+
+\begin{comment}
 
@@
 Finally, Section~\ref{sec:future} outlines some future work.
+
+\end{comment}
 

### 2011-03-02 doc/tfp-2011/main.tex
updated (not yet complete) related work subsection [KK]

@@
 query functions of the functions that implement data locality to
-optimize network traffic. % However, Plasma
-% MapReduce does not support fault tolerance.
-The Disco
-project~\cite{disco} is an implementation of Google's MapReduce in
-Erlang~\cite{erlang}. Our library is not an impementation of Google's
-MapReduce. A MapReduce implementation in \Ocaml\ is Yohann Padioleau's
-MapReduce implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is
-built on top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a
-homemade protocol for message passing.  Currently, we have less
-flexibility w.r.t. deployment of the user program than OCamlMPI; on
-the other hand, we provide a more generic API together with fault
-tolerance.
+optimize network traffic. However, Plasma MapReduce does not support
+fault tolerance which is indespensible in any distributed computing
+application. The Disco project~\cite{disco} is an implementation of
+Google's MapReduce in Erlang~\cite{erlang}. Our library is not an
+impementation of Google's MapReduce. A MapReduce implementation in
+\Ocaml\ is Yohann Padioleau's MapReduce implementation in
+\Ocaml~\cite{poor-man-mapreduce}.  It is built on top of
+OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade protocol
+for message passing.  Currently, we have less flexibility
+w.r.t. deployment of the user program than OCamlMPI; on the other
+hand, we provide a more generic API together with fault tolerance.
 

### 2011-03-03 doc/tfp-2011/main.tex
updated (not yet complete) related work subsection [KK]

@@
 simplest difference comes from the very fact that Functory does not
-operate on key/value pairs. In order to compare and distinguish
+operate on key/value pairs. In order to compare and better distinguish
 Functory from other work with related goals and motivations, we can
@@
 \JoCaml\ is one of the DFLs which provides communication primitives
-(like channels) for transmitting computations. On the other hand,
-\JoCaml\ can not in a true sense be used as a distributed computing
-library. It does not provide fault tolerance, which is indispensable
-in a distributed setting. The user has to include code for fault
-tolerance, as already demonstrated in some \JoCaml\
+(like channels) for transmitting computations. However, it does not
+provide ready-made language features for fault tolerance, which is
+indispensable in a distributed setting. The user has to include code
+for fault tolerance, as already demonstrated in some \JoCaml\
 experiments~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
@@
 marshalling. As in the case before, ML5 is a programming language that
-offers primitives for code mobility, rather than a user-friendly API
-for distribution of computation and fault-tolerance. Glasgow
-\textbf{Distributed Haskell: TODO} Erlang~\cite{erlang} is a
-programming language which has features for distribution and fault
-tolerance. It has features for task distribution and is more
-well-known for its rich error detection primitives and the ability to
-support hot-swapping. The error detection primitives of Erlang allow
-nodes to monitor processes in other nodes and also facilitate
-automatic migration of tasks in failed nodes to recovered or active
-nodes. CamlP3l~\cite{camlP3l} mixes the features of functional
-programming with predefined patterns for parallel computation to offer
-a parallel programming environment. Again, it is a programming
-language offering primitives for distributing computation to parallel
-processes and also to merge the results from parallel executions. The
-above mentioned languages are programming languages themselves, our
-library, Functory, could be implemented using any of the above. The
-main goal of Functory is to provide the users of an existing
-functional programming language high-level user-friendly APIs that
-hides the messy details of task distribution and fault-tolerance,
-enabling a ready-to-use library for efficient utilization of computing
-resources.
+offers primitives for code mobility, and the code for distribution of
+computation and fault-tolerance has to be included by the user. It is
+worth pointing out that an existing type-safe marshalling library
+could still be used with Functory. \\ \textbf{can there exist an OCaml
+  type-safe marshalling library without changes to the
+  compiler?}. \\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
+Erlang~\cite{erlang} is a programming language which has features for
+distribution and fault tolerance. In particular, it has features for
+task distribution and is more well-known for its rich error detection
+primitives and the ability to support hot-swapping. The error
+detection primitives of Erlang allow nodes to monitor processes in
+other nodes and also facilitate automatic migration of tasks in failed
+nodes to recovered or active nodes. CamlP3l~\cite{camlP3l} mixes the
+features of functional programming with predefined patterns for
+parallel computation to offer a parallel programming
+environment. Again, it is a programming language offering primitives
+for distributing computation to parallel processes and also to merge
+the results from parallel executions. The above mentioned languages
+are programming languages themselves, our library, Functory, could be
+implemented using any of the above. The main goal of Functory is to
+provide the users of an existing functional programming language
+high-level user-friendly APIs that hides the messy details of task
+distribution and fault-tolerance, enabling a ready-to-use library for
+efficient utilization of computing resources.
 

### 2011-03-04 doc/tfp-2011/main.tex
updated (not yet complete) related work subsection [KK]

@@
 There are several implementations of Google's Map/Reduce in functional
-programming languages. But Functory was just inspired by Google's
+programming languages. But \of{Functory} was just inspired by Google's
 Map/Reduce and is not exactly a Map/Reduce implementation. The
-simplest difference comes from the very fact that Functory does not
+simplest difference comes from the very fact that \of{Functory} does not
 operate on key/value pairs. In order to compare and better distinguish
-Functory from other work with related goals and motivations, we can
+\of{Functory} from other work with related goals and motivations, we can
 broadly classify the related work in this domain into:
@@
   languages that provide built-in primitives for
-  distribution. Examples include ML5, JoCaml, Glasgow Distributed
+  distribution. Examples include ML5, \JoCaml\, Glasgow Distributed
   Haskell, Erlang, etc.
@@
 worth pointing out that an existing type-safe marshalling library
-could still be used with Functory. \\ \textbf{can there exist an OCaml
-  type-safe marshalling library without changes to the
-  compiler?}. \\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
+could still be used with \of{Functory}. Basically, the task will be
+transmitted along with the corresponding type and on unmarshalling, it
+could be matched for the correct type.
+\\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
 Erlang~\cite{erlang} is a programming language which has features for
@@
 the results from parallel executions. The above mentioned languages
-are programming languages themselves, our library, Functory, could be
-implemented using any of the above. The main goal of Functory is to
-provide the users of an existing functional programming language
-high-level user-friendly APIs that hides the messy details of task
-distribution and fault-tolerance, enabling a ready-to-use library for
-efficient utilization of computing resources.
+are programming languages themselves, our library, \of{Functory},
+could be implemented using any of the above. The main goal of
+\of{Functory} is to provide the users of an existing functional
+programming language high-level user-friendly APIs that hides the
+messy details of task distribution and fault-tolerance, enabling a
+ready-to-use library for efficient utilization of computing resources.
 
@@
 functional languages and weed out the distinguishing features of
-Functory from other similar libraries.
+\of{Functory} from other similar libraries.
 
@@
 application. The Disco project~\cite{disco} is an implementation of
-Google's MapReduce in Erlang~\cite{erlang}. Our library is not an
-impementation of Google's MapReduce. A MapReduce implementation in
-\Ocaml\ is Yohann Padioleau's MapReduce implementation in
-\Ocaml~\cite{poor-man-mapreduce}.  It is built on top of
-OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade protocol
-for message passing.  Currently, we have less flexibility
+Google's MapReduce in Erlang~\cite{erlang}. Another MapReduce
+implementation in \Ocaml\ is Yohann Padioleau's MapReduce
+implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
+top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
+protocol for message passing. Our library is not an impementation of
+Google's MapReduce. Currently, we have less flexibility
 w.r.t. deployment of the user program than OCamlMPI; on the other

### 2011-03-04 doc/tfp-2011/main.tex
related work

@@
 \paragraph{Related Work.}
-
-There are several implementations of Google's Map/Reduce in functional
-programming languages. But \of{Functory} was just inspired by Google's
-Map/Reduce and is not exactly a Map/Reduce implementation. The
-simplest difference comes from the very fact that \of{Functory} does not
-operate on key/value pairs. In order to compare and better distinguish
+In order to compare and better distinguish
 \of{Functory} from other work with related goals and motivations, we can
@@
 \begin{enumerate}
-\item \textit{Distributed Functional Languages (DFLs)} - functional
+\item \textit{Distributed Functional Languages (DFLs)} --- functional
   languages that provide built-in primitives for
@@
   Haskell, Erlang, etc.
-\item \textit{Libraries for existing functional languages} - that
+\item \textit{Libraries for existing functional languages} --- that
   could be readily used in order to avoid implementing details like
@@
 for fault tolerance, as already demonstrated in some \JoCaml\
-experiments~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
+library~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
 programming language for distributed computing, specialized for web
@@
 the client and the server, as well as low-level primitives for
-marshalling the data. In particular, it achieves type-safe
-marshalling. As in the case before, ML5 is a programming language that
+marshalling the data. As in the case before, ML5 is a programming language that
 offers primitives for code mobility, and the code for distribution of
-computation and fault-tolerance has to be included by the user. It is
-worth pointing out that an existing type-safe marshalling library
-could still be used with \of{Functory}. Basically, the task will be
+computation and fault-tolerance has to be included by the user. 
+On the contrary, ML5 implements type-safe marshalling and \functory\
+does not, though an existing type-safe marshalling library
+could be used with \functory. Basically, the task will be
 transmitted along with the corresponding type and on unmarshalling, it
@@
 for distributing computation to parallel processes and also to merge
-the results from parallel executions. The above mentioned languages
-are programming languages themselves, our library, \of{Functory},
-could be implemented using any of the above. The main goal of
-\of{Functory} is to provide the users of an existing functional
-programming language high-level user-friendly APIs that hides the
-messy details of task distribution and fault-tolerance, enabling a
-ready-to-use library for efficient utilization of computing resources.
-
+the results from parallel executions.
+
+All the above mentioned languages could be used to implement our
+library.  The main goal of \of{Functory}, though, is to provide the
+users of an \emph{existing} functional programming language high-level
+user-friendly APIs that hide the messy details of task distribution
+and fault-tolerance, enabling a ready-to-use library for efficient
+utilization of computing resources.
 We now turn to distributed computing libraries for general purpose
@@
 
-PlasmaMR~\cite{plasma} is an implementation of Google's MapReduce on a
+There are several implementations of Google's Map/Reduce in functional
+programming languages. But \of{Functory} was just inspired by Google's
+Map/Reduce and is not exactly a Map/Reduce implementation. The
+simplest difference comes from the very fact that \of{Functory} does not
+operate on key/value pairs. 
+% TODO other difference = absence of data locality
+PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's Map/Reduce on a
 distributed file system PlasmaFS. It is able to use PlasmaFS to its
-advantage - the ability of the file system to handle large files and
+advantage --- the ability of the file system to handle large files and
 query functions of the functions that implement data locality to
-optimize network traffic. However, Plasma MapReduce does not support
+optimize network traffic. However, PlasmaMR does not support
 fault tolerance which is indespensible in any distributed computing
-application. The Disco project~\cite{disco} is an implementation of
-Google's MapReduce in Erlang~\cite{erlang}. Another MapReduce
-implementation in \Ocaml\ is Yohann Padioleau's MapReduce
-implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
+application. 
+% The Disco project~\cite{disco} is an implementation of
+% Google's MapReduce in Erlang~\cite{erlang}.
+Another Map/Reduce
+implementation in \Ocaml\ is Yohann
+Padioleau's~\cite{poor-man-mapreduce}.  It is built on
 top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
-protocol for message passing. Our library is not an impementation of
-Google's MapReduce. Currently, we have less flexibility
+protocol for message passing. Currently, we have less flexibility
 w.r.t. deployment of the user program than OCamlMPI; on the other
@@
   \begin{ocaml}
-    val compute : worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a *
-    'c) list) -> ('a * 'c) list -> unit
+  val compute : 
+    worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a * 'c) list) -> 
+    ('a * 'c) list -> unit
   \end{ocaml}
@@
   \begin{ocaml}
-    let max_iterations = 200 let worker (xmi, xma, ymi, yma, w, h) =
+    let max_iterations = 200 
+    let worker (xmi, xma, ymi, yma, w, h) = 
       ... draw sub-image using max_iterations ...
@@
   \begin{ocaml}
-    Master.compute : master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a *
-    'c) list -> unit
+    Master.compute : 
+      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
   \end{ocaml}
@@
   Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit 
-    -> unit Master.map_fold_ac : 'b -> 'a list -> 'b
+  Master.map_fold_ac : 'b -> 'a list -> 'b
   \end{ocaml}
@@
   \begin{ocaml}
-    Worker.compute : (string -> string) -> unit -> unit Master.compute
-    : master:(string * 'c -> string -> (string * 'c) list) -> (string
-    * 'c) list -> unit
+  Worker.compute : (string -> string) -> unit 
+  Master.compute : 
+    master:(string * 'c -> string -> (string * 'c) list) -> 
+    (string * 'c) list -> unit
   \end{ocaml}

### 2011-03-10 doc/tfp-2011/main.tex
updated introduction section [KK]

@@
 
-This paper introduces a generic library for distributed computing,
-targeted at functional programming. 
-This work was initially motivated by the computing needs that exist in our
-own research team. Our applications include large-scale deductive
-program verification, which amounts to checking the validity of a
-large number of logical formulas using a variety of automated theorem
+This paper introduces a generic library for distributed computing, for
+a widely used functional programming language - OCaml. This work was
+initially motivated by the computing needs that exist in our own
+research team. Our applications include large-scale deductive program
+verification, which amounts to checking the validity of a large number
+of logical formulas using a variety of automated theorem
 provers~\cite{filliatre07cav}. Our computing infrastructure consists
@@
 computing infrastructure in our favorite functional programming
-language.  Hence we designed and implemented such a
-library, which is the subject of this paper.
-This library is implemented in \Ocaml, but the
-implementation should be straightforward in any functional
+language.  Hence we designed and implemented such a library, which is
+the subject of this paper.  This library is implemented in \Ocaml, but
+the implementation should be straightforward in any functional
 programming language.
@@
 facilities for reliable, distributed execution of parallelizable
-computations. In particular, it provides a set of APIs that allows the
-execution of large-scale parallelizable computations, very relevant to
-our application needs (and actually relevant to a variety of
-real-world applications), over multiple cores in the same machine or
-over a network of machines. 
-The most important features of our library are the following:
+computations. In particular, it provides a set of user-friendly APIs
+that allows distributed execution of large-scale parallelizable
+computations, very relevant to our application needs (and also
+relevant to a variety of real-world applications). Further, the
+distributed execution could be over multiple cores in the same machine
+or over a network of machines. The most important features of our
+library are the following:
 \begin{itemize}
@@
 The application domain of such a distributed computing library is manyfold. 
-It is worth noting that the library is not targeted at applications
-running on server farms, crunching enormous amounts of data, mostly
-due to the lack of data locality. 
-It is rather more relevant for research labs, to make use of available
-computing resources efficiently. It serves a variety of users and a wide
-spectrum of needs, from desktop PCs to networks of machines, and hence
-the title.
-Typical applications would involve a large number
-of computation expensive tasks.
-This is the case in our
-research endeavours, that is validating thousands of verification
-conditions using automated theorem provers.
+% It is worth noting that the library is not targeted at applications
+% running on server farms, crunching enormous amounts of data, mostly
+% due to the lack of data locality. 
+The library is very relevant for research labs, to make use of
+available computing resources efficiently. It serves a variety of
+users and a wide spectrum of needs, from desktop PCs to networks of
+machines.  Typical applications would involve executing a large number
+of computation expensive tasks in a resource-optimal and
+time-efficient manner. This is also the case in our research
+endeavours, that is validating thousands of verification conditions
+using automated theorem provers, utilizing the computing
+infrastructure to the maximum.
 
@@
 \paragraph{Distributed Computing.}
-A typical distributed computing library, as \functory,
-provides the following (we
-borrow some terminology from Google's MapReduce~\cite{mapreduce}):
+A typical distributed computing library, as \functory, provides the
+following (we borrow some terminology from Google's
+MapReduce~\cite{mapreduce}):
 \begin{itemize}

### 2011-03-11 doc/tfp-2011/main.tex
updated with documentclass{llncs} [KK]
updated [KK]

@@
 
-\documentclass[tfpsymp,pagenumbers]{tfp07symp}
+% \documentclass[tfpsymp,pagenumbers]{tfp07symp}
+\documentclass{llncs}
 
@@
 
-This paper introduces a generic library for distributed computing, for
-a widely used functional programming language - OCaml. This work was
-initially motivated by the computing needs that exist in our own
-research team. Our applications include large-scale deductive program
-verification, which amounts to checking the validity of a large number
-of logical formulas using a variety of automated theorem
-provers~\cite{filliatre07cav}. Our computing infrastructure consists
-of a few powerful multi-core machines (typically 8 to 16 cores) and
-several desktop PCs (typically dual-core). However, for our
+This paper introduces \functory, a generic library for distributed
+computing for a widely used functional programming language -
+\Ocaml. This work was initially motivated by the computing needs that
+exist in our own research team. Our applications include large-scale
+deductive program verification, which amounts to checking the validity
+of a large number of logical formulas using a variety of automated
+theorem provers~\cite{filliatre07cav}. Our computing infrastructure
+consists of a few powerful multi-core machines (typically 8 to 16
+cores) and several desktop PCs (typically dual-core). However, for our
 application needs, there is no library that helps in exploiting such a
@@
 language.  Hence we designed and implemented such a library, which is
-the subject of this paper.  This library is implemented in \Ocaml, but
-the implementation should be straightforward in any functional
+the subject of this paper.  Although this library is implemented in
+\Ocaml, the implementation should be straightforward in any functional
 programming language.
@@
   it provides automatic task distribution and
-  a robust fault-tolerance mechanism, thereby
+  a robust fault-tolerance mechanisms, thereby
   relieving the user from implementing such routines.
@@
 The application domain of such a distributed computing library is manyfold. 
-% It is worth noting that the library is not targeted at applications
-% running on server farms, crunching enormous amounts of data, mostly
-% due to the lack of data locality. 
-The library is very relevant for research labs, to make use of
-available computing resources efficiently. It serves a variety of
-users and a wide spectrum of needs, from desktop PCs to networks of
-machines.  Typical applications would involve executing a large number
-of computation expensive tasks in a resource-optimal and
-time-efficient manner. This is also the case in our research
-endeavours, that is validating thousands of verification conditions
-using automated theorem provers, utilizing the computing
-infrastructure to the maximum.
+% It is worth noting that the library is not targeted 
+\functory\ is not targeted at applications running on server farms,
+crunching enormous amounts of data and hence it does not implement
+techniques for data locality.  The library is very relevant for
+research labs, to make use of available computing resources
+efficiently. It serves a variety of users and a wide spectrum of
+needs, from desktop PCs to networks of machines.  Typical applications
+would involve executing a large number of computation expensive tasks
+in a resource-optimal and time-efficient manner. This is also the case
+in our research endeavours, that is validating thousands of
+verification conditions using automated theorem provers, utilizing the
+computing infrastructure to the maximum.
 
@@
 
+\medskip
+
+In the following, we discuss some related work in this domain and
+elucidate how \functory\ differs from the other approaches.
+
 \paragraph{Related Work.}
 In order to compare and better distinguish
-\of{Functory} from other work with related goals and motivations, we can
+\of{Functory} from others work with related goals and motivations, we can
 broadly classify the related work in this domain into:
@@
 \JoCaml\ is one of the DFLs which provides communication primitives
-(like channels) for transmitting computations. However, it does not
-provide ready-made language features for fault tolerance, which is
-indispensable in a distributed setting. The user has to include code
-for fault tolerance, as already demonstrated in some \JoCaml\
+(like channels) for facilitating transmission of
+computations. However, it does not provide ready-made language
+features for fault tolerance, which is indispensable in a distributed
+setting. The user has to include code for fault tolerance, as already
+demonstrated in some \JoCaml\
 library~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
@@
 the client and the server, as well as low-level primitives for
-marshalling the data. As in the case before, ML5 is a programming language that
-offers primitives for code mobility, and the code for distribution of
-computation and fault-tolerance has to be included by the user. 
-On the contrary, ML5 implements type-safe marshalling and \functory\
-does not, though an existing type-safe marshalling library
-could be used with \functory. Basically, the task will be
-transmitted along with the corresponding type and on unmarshalling, it
-could be matched for the correct type.
+marshalling the data. As in the case before, ML5 is a programming
+language that offers primitives for code mobility, and the code for
+distribution of computation and fault-tolerance has to be included by
+the user.  On the contrary, ML5 implements type-safe marshalling and
+\functory\ does not, though an existing type-safe marshalling library
+could be used with \functory. Basically, the task will be transmitted
+along with the corresponding type and on unmarshalling, it could be
+matched for the correct type.
 \\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
-Erlang~\cite{erlang} is a programming language which has features for
-distribution and fault tolerance. In particular, it has features for
-task distribution and is more well-known for its rich error detection
-primitives and the ability to support hot-swapping. The error
-detection primitives of Erlang allow nodes to monitor processes in
-other nodes and also facilitate automatic migration of tasks in failed
-nodes to recovered or active nodes. CamlP3l~\cite{camlP3l} mixes the
-features of functional programming with predefined patterns for
-parallel computation to offer a parallel programming
-environment. Again, it is a programming language offering primitives
-for distributing computation to parallel processes and also to merge
-the results from parallel executions.
+CamlP3l~\cite{camlP3l} mixes the features of functional programming
+with predefined patterns for parallel computation to offer a parallel
+programming environment. Again, it is a programming language offering
+primitives for distributing computation to parallel processes and also
+to merge the results from parallel executions. Erlang~\cite{erlang}
+is a programming language which has features for distribution and
+fault tolerance. In particular, it has features for task distribution
+and is more well-known for its rich error detection primitives and the
+ability to support hot-swapping. The error detection primitives of
+Erlang allow nodes to monitor processes in other nodes and also
+facilitate automatic migration of tasks in failed nodes to recovered
+or active nodes.
 
@@
 library.  The main goal of \of{Functory}, though, is to provide the
-users of an \emph{existing} functional programming language high-level
+users of an \emph{existing} functional programming language, high-level
 user-friendly APIs that hide the messy details of task distribution
@@
 % TODO other difference = absence of data locality
-PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's Map/Reduce on a
-distributed file system PlasmaFS. It is able to use PlasmaFS to its
-advantage --- the ability of the file system to handle large files and
-query functions of the functions that implement data locality to
-optimize network traffic. However, PlasmaMR does not support
-fault tolerance which is indespensible in any distributed computing
+PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's
+Map/Reduce on a distributed file system PlasmaFS. It is able to use
+PlasmaFS to its advantage --- the ability of the file system to handle
+large files and query functions that implement data locality to
+optimize network traffic. However, PlasmaMR does not support fault
+tolerance which is indespensible in any distributed computing
 application.

### 2011-03-21 doc/tfp-2011/main.tex
conclusions updated [KK]

@@
 
-We presented a distributed programming environment for functional
-programming. The main features are the genericity of the interface,
-which makes use of polymorphic higher-order functions, and the ability
-to easily switch between sequential, multi-core, and network
-implementations. In particular, \functory\ allows to use the same
-executable for master and workers, which makes the deployment of small programs
-immediate --- master and workers being only distinguished by an
-environment variable. \functory\ also allows master and workers to be
-completely different programs, which is ideal for large scale deployment.
-Another distinguishing feature of our library is a
-robust fault-tolerance mechanism which relieves
-the user of cumbersome implementation details.
-Finally, \functory\ also allows to cascade several distributed
-computations inside the same program.
+In this paper, we presented a distributed programming library for the
+functional programming language \Ocaml. The main features are the
+genericity of the interface, which makes use of polymorphic
+higher-order functions, and the ability to easily switch between
+sequential, multi-core, and network implementations. In particular,
+\functory\ allows to use the same executable for master and workers,
+which makes the deployment of small programs immediate --- master and
+workers being only distinguished by an environment
+variable. \functory\ also allows master and workers to be completely
+different programs, which is ideal for large scale deployment.
+Another distinguishing feature of our library is a robust
+fault-tolerance mechanism which relieves the user of cumbersome
+implementation details.  Yet another interesting feature of the
+library is the ability to add workers dynamically. Finally, \functory\
+also allows to cascade several distributed computations inside the
+same program.  \functory also implements two levels of API with
+different granularities which enables the user to write interactive
+programs.
 
@@
 
-\item 
-  Another interesting feature could be the ability
-  to add or remove machines dynamically. Currently, our library assumes
-  that the list of machines to be used is given \emph{a priori}, as
-  part of the code. An alternative would be to read machine names from
-  a file, watched periodically by the master.
+% \item 
+%   Another interesting feature could be the ability
+%   to add or remove machines dynamically. Currently, our library assumes
+%   that the list of machines to be used is given \emph{a priori}, as
+%   part of the code. An alternative would be to read machine names from
+%   a file, watched periodically by the master.
 \item

### 2011-03-22 doc/tfp-2011/main.tex
proofreading intro and conclusion; adding nb_tasks

@@
   We present Functory, a distributed computing library for
-  functional programming languages. The main features of this library
+  Objective Caml. The main features of this library
   include (1) a polymorphic API, (2) several implementations to
-  adapt to different computing environments such as sequential,
+  adapt to different deployment scenarios such as sequential,
   multi-core or network, and (3) a reliable fault-tolerance mechanism.
@@
 This paper introduces \functory, a generic library for distributed
-computing for a widely used functional programming language -
-\Ocaml. This work was initially motivated by the computing needs that
+computing for a widely used functional programming language,
+Objective Caml (\Ocaml\ for short). 
+This work was initially motivated by the computing needs that
 exist in our own research team. Our applications include large-scale
@@
 language.  Hence we designed and implemented such a library, which is
-the subject of this paper.  Although this library is implemented in
-\Ocaml, the implementation should be straightforward in any functional
-programming language.
+the subject of this paper.  
+% Although this library is implemented in
+% \Ocaml, the implementation should be straightforward in any functional
+% programming language.
 
@@
 \functory\ is not targeted at applications running on server farms,
-crunching enormous amounts of data and hence it does not implement
-techniques for data locality.  The library is very relevant for
+crunching enormous amounts of data, such as Google's
+MapReduce~\cite{mapreduce}.  The library is very relevant for
 research labs, to make use of available computing resources
@@
 
-The remainder of this section introduces our approach to distributed
-computing in a functional programming setting.
+In the following, we introduce our approach to distributed
+computing in a functional programming setting and distinguish it from
+related work.
 
@@
 following (we borrow some terminology from Google's
-MapReduce~\cite{mapreduce}):
+MapReduce):
 \begin{itemize}
@@
 
-\medskip
-
-In the following, we discuss some related work in this domain and
-elucidate how \functory\ differs from the other approaches.
+% \medskip
+
+% In the following, we discuss some related work in this domain and
+% elucidate how \functory\ differs from the other approaches.
 
@@
 distribution of computation and fault-tolerance has to be included by
-the user.  On the contrary, ML5 implements type-safe marshalling and
+the user.  ML5 implements type-safe marshalling and
 \functory\ does not, though an existing type-safe marshalling library
-could be used with \functory. Basically, the task will be transmitted
-along with the corresponding type and on unmarshalling, it could be
-matched for the correct type.
+could be used with \functory. 
+% Basically, the task will be transmitted
+% along with the corresponding type and on unmarshalling, it could be
+% matched for the correct type.
 \\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
@@
 
-There are several implementations of Google's Map/Reduce in functional
+There are several implementations of Google's MapReduce in functional
 programming languages. But \of{Functory} was just inspired by Google's
-Map/Reduce and is not exactly a Map/Reduce implementation. The
+MapReduce and is not exactly a MapReduce implementation. The
 simplest difference comes from the very fact that \of{Functory} does not
@@
 PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's
-Map/Reduce on a distributed file system PlasmaFS. It is able to use
+MapReduce on a distributed file system PlasmaFS. It is able to use
 PlasmaFS to its advantage --- the ability of the file system to handle
@@
 % Google's MapReduce in Erlang~\cite{erlang}.
-Another Map/Reduce
+Another MapReduce
 implementation in \Ocaml\ is Yohann
@@
 
-This section describes our API. We start from a generic distribution
-function, \of{compute}, as sketched in the introduction. Then we
-explain how this function is actually implemented in terms of
+This section describes our API. 
+We start from a simple API which is reduced to a single higher-order
+polymorphic function.
+Then we explain how this function is actually implemented in terms of
 low-level primitives, which are also provided in our API.
-The other way round, we also explain how the \of{compute} function can
+The other way round, we also explain how the same function can
 be used to implement high-level distribution functions for map and
-fold operations. Finally, we explain how these three APIs are
+fold operations. Finally, we explain how our API is
 implemented in five different ways, according to five different deployment
@@
 
+% TODO if we have space for it: small example to show the use of 'c
+% 
+%   let array_map f a =
+%     let h = Hashtbl.create 17 in
+%     let master (_, i) y = Hashtbl.add h i y in
+%     let tasks = ref [] in
+%     for i = 0 to Array.length a - 1 do tasks := (a.(i), i) :: !tasks done;
+%     compute ~worker:f ~master !tasks;
+%     Array.init n (fun i -> Hashtbl.find h i)
+
 \begin{comment}
@@
 performing one step of communication between master and workers, etc.
-These primitives are provided in our API, such that the user can have
-finer control over the execution of the distributed computation.
-A type for distributed computations is introduced.
+These primitives are provided in our API, such that the user can
+interact with the execution of the distributed computation.  For
+instance, a monitoring-like application can use these primitives to
+allow observation and modification of resources (tasks, workers)
+during the course of a computation.
+
+A type for distributed computations is introduced:
 \begin{ocaml}
@@
 
-\subsection{Functory API at a Glance} % ???
+\subsection{Several APIs in One}
 
 From  the description above, it is clear that our library provides
-three different APIs, for five different deployment scenarios,
-resulting in fifteen possible ways of using it.
-This is summarized in the following matrix.
-\begin{center}
-  \begin{tabular}{|l|c|c|c|c|c|}
-    \hline
-    & & & \multicolumn{3}{c|}{Network} \\
-    & Seq. & Cores & Same & Poly & Mono \\\hline\hline
-    low-level API & & & & & \\\hline
-    \of{compute}  & & & & & \\\hline
-    high-level API& & & & & \\\hline
-  \end{tabular}
-\end{center}
+several APIs of different granularities and different
+deployment scenarios. 
+Most combinations are meaningful, resulting in
+thirteen possible different ways of using our library.
+% This is summarized in the following matrix.
+% \begin{center}
+%   \begin{tabular}{|l|c|c|c|c|c|}
+%     \hline
+%     & & & \multicolumn{3}{c|}{Network} \\
+%     & Seq. & Cores & Same & Poly & Mono \\\hline\hline
+%     low-level API & & & & & \\\hline
+%     \of{compute}  & & & & & \\\hline
+%     high-level API& & & & & \\\hline
+%   \end{tabular}
+% \end{center}
 From the implementation point of view, there is almost no code
-duplication. We are using \Ocaml\ functors to derive the second line
-from the first, as well as the third from the second, irrespective of
-the deployment scenario.
+duplication. We are using \Ocaml\ functors to derive specific
+implementations from generic ones.
 
@@
 
-In this paper, we presented a distributed programming library for the
-functional programming language \Ocaml. The main features are the
+In this paper, we presented a distributed programming library for
+\Ocaml. The main features are the
 genericity of the interface, which makes use of polymorphic
@@
 implementation details.  Yet another interesting feature of the
-library is the ability to add workers dynamically. Finally, \functory\
+library is the ability to add workers dynamically. \functory\
 also allows to cascade several distributed computations inside the
-same program.  \functory also implements two levels of API with
-different granularities which enables the user to write interactive
-programs.
+same program.  Finally, the low-level API of \functory\ 
+can be used to write interactive programs.
 
@@
   modules, we use traditional queues for the pending tasks; in
-  particular, new tasks produced by the master are appended at the end
+  particular, new tasks produced by the master are appended to the end
   of the queue.
@@
 \item
-  Our library provides limited support for displaying real-time
+  Our library provides limited support for retrieving real-time
   information about computations and communications. Processing and storing
-  information about workers and tasks locally in the master is straightforward; 
-  monitoring it in real-time could be done using
-  \textsf{Ocamlviz}~\cite{ocamlviz}. 
+  information about workers and tasks locally in the master is
+  straightforward.
+
 \item 

### 2011-03-24 doc/tfp-2011/main.tex
updated related work [KK]

@@
 % matched for the correct type.
-\\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
+Glasgow Distributed Haskell (GdH)~\cite{GdH} is a pure distributed
+functional language that is built on top of Glasgow Haskell and
+provides features for distributed computing. It is an extension of
+both Glasgow Parallel Haskell, that supports only one process and
+multiple threads and Concurrent Haskell that supports multiple
+processes. It also offers features for fault-tolerance - error
+detection and error recovery primitives in the
+language.
+
 CamlP3l~\cite{camlP3l} mixes the features of functional programming
@@
 primitives for distributing computation to parallel processes and also
-to merge the results from parallel executions. Erlang~\cite{erlang}
-is a programming language which has features for distribution and
-fault tolerance. In particular, it has features for task distribution
-and is more well-known for its rich error detection primitives and the
+to merge the results from parallel executions. Erlang~\cite{erlang} is
+a programming language which has features for distribution and fault
+tolerance. In particular, it has features for task distribution and is
+more well-known for its rich error detection primitives and the
 ability to support hot-swapping. The error detection primitives of
@@
 
+The iTask system~\cite{iTask} is a library for the functional language
+`Clean' targeted at distributed workflow management. The library
+provides a set of combinator functions (some of which perform map/fold
+operations) that facilitate applications running in different nodes
+of a distributed system to communicate, exchange information and
+coordinate their computations.
+
 \begin{comment}
-
-% -- BELOW HERE IS OLD RELATED-WORK
-
 Closest to the approach in this paper is Yohann Padioleau's MapReduce

### 2011-03-25 doc/tfp-2011/main.tex
*** empty log message ***

@@
 \end{enumerate}
+\functory\ belongs to the second category.
+For reasons of completeness, though, we first describe some existing DFLs
+related to functional programming.
 
@@
 
-All the above mentioned languages could be used to implement our
-library.  The main goal of \of{Functory}, though, is to provide the
-users of an \emph{existing} functional programming language, high-level
+Any DFL above could have been used to implement our library. 
+Our motivation, though, was neither to implement our system using any
+existing DFL nor to come up with a new DFL.
+The goal of \of{Functory} is rather to provide the
+users of an \emph{existing} general-purpose functional programming
+language, namely \Ocaml, high-level 
 user-friendly APIs that hide the messy details of task distribution
-and fault-tolerance, enabling a ready-to-use library for efficient
-utilization of computing resources.
+and fault-tolerance.
 We now turn to distributed computing libraries for general purpose
 functional languages and weed out the distinguishing features of
-\of{Functory} from other similar libraries.
+\of{Functory}.
 
@@
 \begin{center}
-  \includegraphics{state.mps}
+  \includegraphics[scale=0.7]{state.mps}
 \end{center}

### 2011-03-30 doc/tfp-2011/main.tex
*** empty log message ***

@@
 
-\usepackage{comment,amsmath,graphicx,color,url}
+\usepackage{comment,amsmath,graphicx,color,url,hyperref}
 
@@
 cores) and several desktop PCs (typically dual-core). However, for our
-application needs, there is no library that helps in exploiting such a
-computing infrastructure in our favorite functional programming
-language.  Hence we designed and implemented such a library, which is
+application needs, no existing library provides a polymorphic API with
+usual map/fold higher-order operations, built-in fault-tolerance, and
+the ability to easily switch between multi-core and network
+infrastructures.
+Hence we designed and implemented such a library, which is
 the subject of this paper.  
-% Although this library is implemented in
-% \Ocaml, the implementation should be straightforward in any functional
-% programming language.
+The library is available at 
+\href{http://functory.lri.fr/}{\url{http://functory.lri.fr/}}.
 
@@
 The application domain of such a distributed computing library is manyfold. 
-% It is worth noting that the library is not targeted 
-\functory\ is not targeted at applications running on server farms,
-crunching enormous amounts of data, such as Google's
-MapReduce~\cite{mapreduce}.  The library is very relevant for
-research labs, to make use of available computing resources
-efficiently. It serves a variety of users and a wide spectrum of
+It serves a variety of users and a wide spectrum of
 needs, from desktop PCs to networks of machines.  Typical applications
@@
 computing infrastructure to the maximum.
+It is worth noting that 
+\functory\ is not targeted at applications running on server farms,
+crunching enormous amounts of data, such as Google's
+MapReduce~\cite{mapreduce}.
 
@@
 In addition to the above, distributed computing environments also
-implement mechanisms for fault tolerance, efficient storage, and
+implement mechanisms for fault-tolerance, efficient storage, and
 distribution of tasks. This is required to handle network failures
@@
 The function passed to the master is applied whenever a result is
-available and may in turn generate new tasks (hence the return type
-\of{'a list}).  The master is executed as long as there are pending
-tasks.
+available. The first argument is the task (of type \of{'a}) and the
+second one its result (of type \of{'b}). 
+It may in turn generate new tasks, hence the return type
+\of{'a list}.  
+The master is executed as long as there are pending tasks.
 
@@
 computations. However, it does not provide ready-made language
-features for fault tolerance, which is indispensable in a distributed
-setting. The user has to include code for fault tolerance, as already
+features for fault-tolerance, which is indispensable in a distributed
+setting. The user has to include code for fault-tolerance, as already
 demonstrated in some \JoCaml\
@@
 to merge the results from parallel executions. Erlang~\cite{erlang} is
-a programming language which has features for distribution and fault
-tolerance. In particular, it has features for task distribution and is
+a programming language which has features for distribution and 
+fault-tolerance. In particular, it has features for task distribution and is
 more well-known for its rich error detection primitives and the
@@
 large files and query functions that implement data locality to
-optimize network traffic. However, PlasmaMR does not support fault
-tolerance which is indespensible in any distributed computing
+optimize network traffic. However, PlasmaMR does not support 
+fault-tolerance which is indespensible in any distributed computing
 application.
@@
 w.r.t. deployment of the user program than OCamlMPI; on the other
-hand, we provide a more generic API together with fault tolerance.
+hand, we provide a more generic API together with fault-tolerance.
 
@@
 hand, we provide a 
-more generic API together with fault tolerance.  There are other
+more generic API together with fault-tolerance.  There are other
 distributed computing libraries on top of which one could implement
 the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
-them. However, \JoCaml\ does not provide fault tolerance, which is
+them. However, \JoCaml\ does not provide fault-tolerance, which is
 indispensable in a distributed setting. The user has to include code
-for fault tolerance, as already demonstrated in some \JoCaml\
+for fault-tolerance, as already demonstrated in some \JoCaml\
 experiments~\cite{mandel2008}.
@@
 Section~\ref{sec:implem} gives implementation details, focusing
-on marshaling, network protocol and fault tolerance.
+on marshaling, network protocol and fault-tolerance.
 Then Section~\ref{sec:experiments} illustrates the
@@
 low-level primitives, which are also provided in our API.
-The other way round, we also explain how the same function can
+Conversely, we also explain how the same function can
 be used to implement high-level distribution functions for map and
@@
 \begin{ocaml}
-  val compute : 
+  val compute: 
     worker:('a -> 'b) -> 
-    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
-    ('a * 'c) list -> unit
+    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
 \end{ocaml}
@@
 \of{Cores}, so that the rest of the code remains unchanged. Master and
-worker processes are distinguished at run-time using an environment
+worker processes are distinguished at runtime using an environment
 variable \texttt{WORKER} which is set/unset.
@@
 of low-level primitives, such as adding a task, adding a worker,
-performing one step of communication between master and workers, etc.
+performing some communication between master and workers, etc.
 These primitives are provided in our API, such that the user can
@@
 during the course of a computation.
-
 A type for distributed computations is introduced:
@@
 \begin{ocaml}
-  val create : 
-    worker:('a -> 'b) -> 
+  val create: worker:('a -> 'b) -> 
     master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a, 'c) computation
@@
 \begin{ocaml}
-  val add_task : ('a, 'c) computation -> 'a * 'c -> unit
+  val add_task: ('a, 'c) computation -> 'a * 'c -> unit
 \end{ocaml}
@@
 \begin{ocaml}
-  val one_step : ('a, 'c) computation -> unit
+  val one_step: ('a, 'c) computation -> unit
 \end{ocaml}
@@
 The most obvious operation is the traditional map operation over
-lists, that is:
-\begin{ocaml}
-  val map : f:('a -> 'b) -> 'a list -> 'b list
-\end{ocaml}
-Each task consists of application of function \of{f} to a list element.
+lists, that is \of{val map: f:('a -> 'b) -> 'a list -> 'b list}.
+Each task consists of the application of function \of{f} to a list element.
 More interesting is a combination of map and fold operations.
@@
 \begin{ocaml}
-  val map_fold :
-    f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
+  val map_fold: f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
 \end{ocaml}
@@
 \begin{ocaml}
-  val map_fold_ac, map_fold_a : 
+  val map_fold_ac, map_fold_a: 
     f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
 \end{ocaml}
-This is rather straightforward to derive these five functions from the
+It is rather straightforward to derive these five functions from the
 generic \of{compute} function; readers interested in details could
@@
 
-\subsection{Several APIs in One}
+\subsection{Several Libraries in One}
 
 From  the description above, it is clear that our library provides
-several APIs of different granularities and different
-deployment scenarios. 
+several APIs of different granularities, as well as several
+implementations for various deployment scenarios. 
 Most combinations are meaningful, resulting in
 thirteen possible different ways of using our library.
-% This is summarized in the following matrix.
-% \begin{center}
-%   \begin{tabular}{|l|c|c|c|c|c|}
-%     \hline
-%     & & & \multicolumn{3}{c|}{Network} \\
-%     & Seq. & Cores & Same & Poly & Mono \\\hline\hline
-%     low-level API & & & & & \\\hline
-%     \of{compute}  & & & & & \\\hline
-%     high-level API& & & & & \\\hline
-%   \end{tabular}
-% \end{center}
+For instance, one may use the low-level API on a single multi-core
+machine, or use the high-level API on a network of machines all
+running the same executable, etc.
 From the implementation point of view, there is almost no code
@@
 
-We now describe the implementations of the various modules introduced
-in Section~\ref{sec:API}. The implementation of the \of{Sequential}
+The implementation of the \of{Sequential}
 module is straightforward and does not require any explanation.
+The \of{Cores} module is implemented with \unix\ processes, using
+the \of{fork} and \of{wait} system calls provided by the \of{Unix}
+library of \Ocaml. We do not describe this implementation but
+rather focus on the more interesting module \of{Network}.
 
@@
   \texttt{WORKER} for this purpose. When set, it indicates that the
-  executable acts as a worker. At run-time, a worker immediately
+  executable acts as a worker. At runtime, a worker immediately
   enters a loop waiting for tasks from the master, without even
@@
   \begin{ocaml}
-  val compute : 
-    worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a * 'c) list) -> 
-    ('a * 'c) list -> unit
+  val compute: worker:('a -> 'b) -> 
+    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
   \end{ocaml}
@@
   these global variables is never reached on the worker side. For
-  instance, the Mandelbrot set example could be written as follows:
+  instance, some code for drawing Mandelbrot's set
+  could be written as follows:
   \begin{ocaml}
     let max_iterations = 200 
-    let worker (xmi, xma, ymi, yma, w, h) = 
-      ... draw sub-image using max_iterations ...
+    let worker si = ... draw sub-image si using max_iterations ...
   \end{ocaml}
@@
 
-  \bigskip There are situations where it is not possible to run the
+  There are situations where it is not possible to run the
   same executable for master and workers.  For instance, architectures
@@
   though, since master and workers could still be distinguished at
-  run-time as in the previous case.
+  runtime as in the previous case.
 
@@
   \begin{ocaml}
-    Worker.compute : ('a -> 'b) -> unit -> unit
+    Worker.compute: ('a -> 'b) -> unit -> unit
   \end{ocaml}
@@
   \begin{ocaml}
-    Master.compute : 
+    Master.compute: 
       master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
@@
   values over the network. The issues of marshalling are automatically
-  taken care of by \Ocaml\ run-time.
+  taken care of by the \Ocaml\ runtime.
 
@@
   \begin{ocaml}
-  Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit 
-  Master.map_fold_ac : 'b -> 'a list -> 'b
-  \end{ocaml}
-  It is the responsibility of the user to ensure type consistency
+  Worker.map_fold_ac: f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit 
+  Master.map_fold_ac: 'b -> 'a list -> 'b
+  \end{ocaml}
+  It is the responsibility of the user to ensure consistency
   between master and workers.
@@
   \begin{ocaml}
-  Worker.compute : (string -> string) -> unit 
-  Master.compute : 
-    master:(string * 'c -> string -> (string * 'c) list) -> 
+  Worker.compute: (string -> string) -> unit 
+  Master.compute: master:(string * 'c -> string -> (string * 'c) list) -> 
     (string * 'c) list -> unit
@@
 We subsequently describe issues of message
-transfer and fault tolerance.
+transfer and fault-tolerance.
 
@@
 
-\subsection{Fault Tolerance}\label{sec:fault}
+\subsection{Fault-Tolerance}\label{sec:fault}
 
@@
 library.  
-The fault tolerance mechanism of \functory\ is limited to workers;
+The fault-tolerance mechanism of \functory\ is limited to workers;
 handling master failures is the responsibility of the user, for
@@
 and possibly later restarted; or a worker is temporarily or
-permanently unreachable on the network. To provide fault tolerance,
+permanently unreachable on the network. To provide fault-tolerance,
 our master implementation is keeping track of the status of each
@@
 For this comparison to be fair, all
-computations are performed on the same machine, a 8 cores Intel Xeon
+computations are performed on the same machine, a 8 core Intel Xeon
 3.2 GHz running Debian Linux. The sequential implementation uses a
@@
 the machine. The network implementation uses 8 workers running locally and
-a master running on a remote machine (which incurs communication cost).
+a master running on a remote machine over a LAN 
+(which incurs communication cost).
 
@@
 the subsequent search in parallel. Choosing $D=1$ will result in
-exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
-so on.
+exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; 
+greater values for $D$ would result in two many tasks.
+
 % TODO: make a picture to illustrate positioning of 2 queens on the
@@
 \end{center}
+The difference in the number of tasks explains the differences in the
+speedup ratios above.
 We do not include results for the network configuration, as they do
@@
   and \of{Cores}, so that the rest of the code remains
-  unchanged. Master and worker processes are distinguished at run-time
+  unchanged. Master and worker processes are distinguished at runtime
   using an environment variable \texttt{WORKER} which is set/unset.
@@
 infrastructure for this experiment consists of 3 machines with 4, 8 and 8 cores
-respectively. 
+respectively, the master being run on a fourth machine. 
 The figure below shows the total time in minutes spent by each prover
@@
   for speedup when reaching the end of computation.
-  Since we already have the fault tolerance implemented, this
+  Since we already have the fault-tolerance implemented, this
   optimization should be straightforward to add to our library.

### 2011-03-31 doc/tfp-2011/main.tex
spell-checked [KK]

@@
 optimize network traffic. However, PlasmaMR does not support 
-fault-tolerance which is indespensible in any distributed computing
+fault-tolerance which is indispensable in any distributed computing
 application.
@@
 
-\section{Conclusion and Future Work}\label{sec:future}
+\section{Conclusions and Future Work}\label{sec:future}
 

### 2011-05-10 doc/tfp-2011/slides/slides-pp.tex
TFP-2011 slides [KK]

% \documentclass[xcolor=dvipsnames]{beamer}
% % \documentclass[handout]{beamer}
% \usecolortheme[named=Maroon]{structure} 
% \usetheme{Boadilla}
% \setbeamertemplate{beaver}
\documentclass[xcolor=dvipsnames,8pt]{beamer} 
\usecolortheme[named=MidnightBlue]{structure} 
\usetheme[height=7mm]{Rochester} 
\usepackage{fontenc}
\renewcommand*\familydefault{\sfdefault}
\usepackage{times}
% \usepackage[latin1]{inputenc}
\usepackage{alltt}
\renewcommand{\iff}{\leftrightarrow}
\newcommand{\tr}{\textcolor{red}}
\newcommand{\tb}{\textcolor{MidnightBlue}}
\newcommand{\intt}{\textcolor{blue}{\textbf{int }}}
\newcommand{\while}{\textcolor{blue}{\textbf{while }}}
\newcommand{\ret}{\textcolor{blue}{\textbf{return }}}
\newcommand{\asrt}{\textcolor{blue}{\textbf{//@assert }}}
\newcommand{\ts}[1]{\scriptsize{#1}}
\newcommand{\hl}[1]{\colorbox{yellow}{#1}}
%\usepackage[french]{babel}

\beamertemplatenavigationsymbolsempty

\bigskip
\definecolor{MyGray}{rgb}{0.86,0.87,0.98}
\makeatletter\newenvironment{bbox}{%
   \begin{lrbox}
     {\@tempboxa}\begin{minipage}
       {\columnwidth}}
     {\end{minipage}
   \end{lrbox}%
   \colorbox{MyGray}{\usebox{\@tempboxa}}
}\makeatother
% \definecolor{MyGray}{rgb}{0.96,0.97,0.98}

\definecolor{kwblue}{rgb}{0.67,0.12,0.92}
\definecolor{ceruleanblue}{rgb}{0, 0.48, 0.65}
\definecolor{lightpink}{rgb}{1., 0.71, 0.75}
\definecolor{lightblue}{rgb}{0.8,0.8,1}
\definecolor{lightred}{rgb}{1,0.8,0.8}

\let\emph\alert
\newcommand{\coeurs}{c\oe urs}

\begin{document}

%% -------------------------------------------------------------------- %%

\title{FUNCTORY \\ A Distributed Computing Library for Objective Caml}

\author[Kalyan]{Jean-Christophe Filli\^{a}tre \& Kalyan Krishnamani\\
  \- \\ 
  INRIA-Saclay \^{I}le-de-France}
\date {}

\bigskip
\begin{frame}
  \titlepage
  \begin{center}
    \includegraphics[scale=0.11]{cnrs-logo2.jpg}
    \hspace{4mm}
    \includegraphics[scale=0.09]{lrilogo.jpg}
    \hspace{4mm}
    \includegraphics[scale=0.09]{upsudlogo.jpg}
    \hspace{4mm}
    \includegraphics[scale=0.4]{inria-saclay.jpeg}
    % \pgfimage[height=6mm]{saclay}
  \end{center}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Outline}
  \begin{itemize}
  \item Motivation
  \item The Basics
  \item API
  \item Implementation Details
  \item Experiments
  \item Related Work
  \item Conclusions
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Motivation}
  \begin{itemize}
  \item Why Platform - A set of tools for deductive program verification
  \item Generates numerous verification conditions (VCs)
  \item VCs discharged by various automated provers
  \item Typically takes hours to complete
  \end{itemize}
  \vspace{1cm}
  \begin{itemize}
  \item Some multi-core machines at our disposal
  \item How to make the best possible use of them?
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {A Distributed Computing Library}
  Requirements
  \begin{itemize}
  \item Ideally in a functional language (OCaml)
  \item Robust fault-tolerance
  \item User-friendly API
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Basic Design}

Inspired by Google's Map/Reduce (OSDI 2004)

\begin{itemize}
\item Workers in parallel
\item Master
\end{itemize}

\begin{center}
  \includegraphics{master_workers_1.mps}  
\end{center}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {API}
A generic distribution function \texttt{compute}:\\

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{2.00em}{\color{blue}val}\hspace*{1.22ex}compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{2.00em}worker\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{2.00em}master\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{2.00em}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{2.00em}unit
\end{minipage}}}

\medskip\noindent
  \begin{itemize}
  \item A task is of type \textsf{\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}}, its result of type
    \textsf{\ensuremath{\beta}}
  \item A completed task may in turn generate new tasks
  \item \texttt{compute} returns when there is no more task
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Low-level Primitives}
How to implement \texttt{compute} using low-level primitives?
\begin{itemize}
\item \textsf{{\color{blue}type}\hspace*{1.22ex}(\ensuremath{\alpha},\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}computation}
\item Computation is created using \texttt{create}:
\end{itemize}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}create\symbol{58}\hspace*{1.22ex}worker\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{7.00em}master\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})list)\hspace*{1.22ex}\ensuremath{\rightarrow}~\linebreak
\noindent\hspace*{7.00em}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}computation
\end{minipage}}}

\medskip\noindent
  \begin{itemize}
  \item \texttt{create} does not add tasks. However, tasks could be added using
  \end{itemize}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}add\_{}task\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}computation\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
\end{minipage}}}

\medskip\noindent
  \begin{itemize}
  \item One step of the computation can be performed using
  \end{itemize}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}one\_{}step\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}computation\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
\end{minipage}}}

\medskip\noindent
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Low-level Primitives}
  \begin{itemize}
  \item \texttt{compute} could be implemented using low-level primitives:
  \end{itemize}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}let}\hspace*{1.22ex}\symbol{126}compute\hspace*{1.22ex}\symbol{126}worker\hspace*{1.22ex}\symbol{126}master\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}~\linebreak
\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}c\hspace*{1.22ex}=\hspace*{1.22ex}create\hspace*{1.22ex}worker\hspace*{1.22ex}master\hspace*{1.22ex}{\color{blue}in}~\linebreak
\noindent\hspace*{1.00em}List.iter\hspace*{1.22ex}(add\_{}task\hspace*{1.22ex}c)\hspace*{1.22ex}tasks~\linebreak
\noindent\hspace*{1.00em}{\color{blue}while}\hspace*{1.22ex}status\hspace*{1.22ex}c\hspace*{1.22ex}=\hspace*{1.22ex}Running\hspace*{1.22ex}{\color{blue}do}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{2.50em}one\_{}step\hspace*{1.22ex}c\hspace*{1.22ex}~\linebreak
\noindent\hspace*{1.00em}{\color{blue}done}\hspace*{1.22ex}\hspace*{1.22ex}\hspace*{1.22ex}\hspace*{1.22ex}
\end{minipage}}}

\medskip\noindent
  \begin{itemize}
  \item User can interact with the execution of the distributed
    computation
  \item Monitoring applications: observation of consumption of
  resources, etc
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {High-level API}
  \begin{itemize}
  \item Implementation of the most common map/fold operations over lists
  \end{itemize}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\symbol{58}\hspace*{1.22ex}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}list~\linebreak
\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\_{}fold\symbol{58}\hspace*{1.22ex}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{8.00em}fold\symbol{58}(\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}
\end{minipage}}}

\medskip\noindent
  \begin{itemize}
  \item \texttt{f} operations are always in parallel.
  \item Two flavours: \texttt{map\_local\_fold} and
    \texttt{map\_remote\_fold}
  \item More parallelism in \texttt{map\_remote\_fold} when fold is
    associative and commutative
  \end{itemize}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\_{}fold\_{}ac,\hspace*{1.22ex}map\_{}fold\_{}a\symbol{58}~\linebreak
\noindent\hspace*{0.00em}f\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{0.00em}fold\symbol{58}\hspace*{1.22ex}(\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}
\end{minipage}}}

\medskip\noindent
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Deployment Scenarios}
  \begin{itemize}
  \item Purely sequential execution: for debugging and comparison
    purposes
  \item Several cores on the same machine
  \item Computation distributed over a network of machines
  \end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Cores Implementation: \texttt{Functory.Cores}}
  Uses \texttt{Unix.fork} (no control over scheduling)

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{1.00em}{\color{blue}open}\hspace*{1.22ex}Cores~\linebreak
\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}set\_{}number\_{}of\_{}cores\hspace*{1.22ex}4
\end{minipage}}}

\medskip\noindent

  \begin{center}
    \includegraphics{master_workers_cores.mps}
  \end{center}

  master maintains a queue of pending tasks

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Network Implementation: \texttt{Functory.Network}}
  based on
  \begin{itemize}
  \item TCP/IP client/server architecture
  \item Ocaml's marshaling capabilities
  \end{itemize}

  \vfill
  marshaling considerations
  \begin{enumerate}
  \item \emph{same binary}: we can marshal closures
  \item \emph{same version of Ocaml}: we can only marshal values
  \item \emph{otherwise}: we can only marshal strings
  \end{enumerate}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {\texttt{Functory.Network}: Three Scenarios}

\emph{Same binary}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}{\color{red}(*\hspace*{1.22ex}same\hspace*{1.22ex}as\hspace*{1.22ex}before\hspace*{1.22ex}*)}\hspace*{1.22ex}...
\end{minipage}}}

\medskip\noindent

\bigskip
\emph{Same version of Caml}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}Worker.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit~\linebreak
\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}Master.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{1.00em}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}~\linebreak
\noindent\hspace*{1.00em}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
\end{minipage}}}

\medskip\noindent

\bigskip
\emph{Otherwise}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}Worker.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}(string\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}string)\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit~\linebreak
\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}Master.compute\hspace*{1.22ex}\symbol{58}\hspace*{1.22ex}~\linebreak
\noindent\hspace*{1.00em}(string\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}string\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}(string\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list)\hspace*{1.22ex}\ensuremath{\rightarrow}~\linebreak
\noindent\hspace*{1.00em}(string\hspace*{1.22ex}\ensuremath{\times}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}unit
\end{minipage}}}

\medskip\noindent

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Details: \texttt{Functory.Network}}

\medskip\noindent
{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt

\noindent\hspace*{1.00em}{\color{blue}open}\hspace*{1.22ex}Network~\linebreak
\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}declare\_{}workers\hspace*{1.22ex}\symbol{126}n\symbol{58}3\hspace*{1.22ex}"moloch"~\linebreak
\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}declare\_{}workers\hspace*{1.22ex}\symbol{126}n\symbol{58}2\hspace*{1.22ex}"orcus"
\end{minipage}}}

\medskip\noindent

  \begin{center}
    \includegraphics{master_workers_network.mps}
  \end{center}
  Each worker behaves as a server, the master being the client

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Underneath: Protocol}
  \only<1>{Master sends a task to a worker}
  \only<2>{Worker computes and sends back a result}
  \only<3-4>{Master and workers exchange \emph{ping}/\emph{pong} messages}
  \only<5>{Master sends another task to a worker}
  \only<6>{In case of a disconnection...}
  \only<7>{The task is \emph{rescheduled} to another worker}
  \only<8>{Whenever one completes...}
  \only<9>{The other one is stopped}
  \only<10>{The master is notified when a computation fails}
  \only<11>{At the very end, the master may ask the workers to stop}
  \vfill
  \begin{center}
    \only<1,5>{\includegraphics{master_workers_assign.mps}}
    \only<2>{\includegraphics{master_workers_completed.mps}}
    \only<3>{\includegraphics{master_workers_ping.mps}}
    \only<4>{\includegraphics{master_workers_pong.mps}}
    \only<6>{\includegraphics{master_workers_disconnection.mps}}
    \only<7>{\includegraphics{master_workers_assign_2.mps}}
    \only<8>{\includegraphics{master_workers_completed_2.mps}}
    \only<9>{\includegraphics{master_workers_kill.mps}}
    \only<10>{\includegraphics{master_workers_aborted.mps}}
    \only<11>{\includegraphics{master_workers_stop.mps}}
  \end{center}
  \begin{center}
    \only<1>{\textsf{Assign(42,\hspace*{1.22ex}f,\hspace*{1.22ex}a)}}
    \only<2>{\textsf{Completed\hspace*{1.22ex}(42,\hspace*{1.22ex}b)}}
    \only<3>{\textsf{Ping}}
    \only<4>{\textsf{Pong}}
    \only<5>{\textsf{Assign(43,\hspace*{1.22ex}f,\hspace*{1.22ex}a)}}
    \only<6>{\textsf{}} % disconnection
    \only<7>{\textsf{Assign(43,\hspace*{1.22ex}f,\hspace*{1.22ex}a)}}
    \only<8>{\textsf{Completed\hspace*{1.22ex}(43,\hspace*{1.22ex}b)}}
    \only<9>{\textsf{Kill\hspace*{1.22ex}43}}
    \only<10>{\textsf{Aborted\hspace*{1.22ex}44}}
    \only<11>{\textsf{Stop}}
  \end{center}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Feature: Fault Tolerance}

Master knows the state of each worker through ping/pong messages

  \begin{center}
    \includegraphics[scale=0.7]{state.mps}
  \end{center}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Potential: Experimental Results}
  Motivating example
  \begin{itemize}
  \item 80 verification conditions
  \item 4 provers
  \item one minute timeout per VC
  \item network of 3 machines (4, 8 and 8 cores)
  \end{itemize}

\bigskip
  \begin{itemize}
  \item sequential computation: \emph{$>$ 6 hours}
  \item with Functory: \emph{22.5 minutes}
  \item speedup ratio = \emph{16} (optimal is 20)
  \end{itemize}
  \begin{center}
  More results in the paper !!    
  \end{center}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Related Work}
  Distributed Functional Languages (DFL)
  \begin{itemize}
  \item Jo\&Caml - rich communication primitives, no easy features
    for fault-tolerance
  \item ML5 - code mobility, marshalling, etc, but no
    easy features for fault-tolerance
  \item Glasgow Distributed Haskell - features for fault-tolerance,
    error detection/recovery
  \end{itemize}
  Libraries for existing functional languages
  \begin{itemize}
  \item Plasma MR - OCaml implementation of Google's Map/Reduce on
    Plasma FS
  \item iTask - a library for 'Clean' language for distributed workflow management
  \end{itemize}

Our library could be implemented using any of the DFLs, but our goals
are:
\begin{itemize}
\item Provide users of an existing language, user-friendly APIs
\item Provide robust fault-tolerance mechanism 
\end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {What's next}
  \begin{itemize}
  \item Better Scheduling
  \item Real-time visualization - OcamlViz?
  \item Any gains from Plasma-FS - data locality?
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Thanks}

\bigskip
Check out:
\begin{center}
 \textcolor{blue}{\url{http://functory.lri.fr/}}  
\end{center}

Feedback, comments welcome!

\end{frame}

%% -------------------------------------------------------------------- %%
\end{document}

\bigskip\bigskip\bigskip
\begin{frame}\frametitle{The Talk}

\bigskip
\begin{bbox}
\begin{alltt}\intt main ()\\
\{\\
  \- \- \intt x = 0;\\
  \- \- \intt y = 1;\\
  \- \- \while (x < 100) \{\\
  \- \- \- \- x = x + y;\\
  \- \- \- \- y = x;\\
  \- \- \}\\
  \- \- \asrt (x < 200);\\
  \- \- \ret 0;\\
\}
\end{alltt}%
\end{bbox}
\pause

Can we automatically: \\ 
\begin{itemize}
  \itemsep=0.3cm
\item Generate a loop invariant ?
\item Generate a loop invariant that helps in proving the assertion ?

\end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Abstraction - Intuition}
  \centering
  \includegraphics[scale=0.3]{Diagram1_new1.png}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Abstraction - Intuition}
  \centering
  \includegraphics[scale=0.3]{Diagram2_new1.png}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Predicate Abstraction}
The Problem: \\ \- \\ Given a concrete program and a set of predicates, track
the behaviour of the predicates corresponding to the behaviour of the
concrete program \\ \- \\
\pause
Ingredients:
  \begin{itemize}
    \itemsep=0.3cm
  \item Concrete program: $R(\vec{x}, \vec{x}')$
  \item A set of predicates: \{$\vec{P} \iff \alpha (\vec{x})$, $\vec{P}' \iff
    \alpha (\vec{x'})$ \} 
  \item Behaviour of the predicates w.r.t the concrete program:
    $$
    \exists. \vec{x}, \vec{x}'.\ R(\vec{x}, \vec{x}') \land \vec{P} =
    \alpha (\vec{x}) \land \vec{P}' = \alpha (\vec{x'})
    $$
  \item Existential quantification - solved using a SMT solver
  \item Abstract interpretation with Abstract domain =
      vector of Booleans
  \end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Predicate Abstraction - Example}
  \begin{columns}
    \begin{column}{0.65\textwidth}
      \begin{itemize}
        \itemsep=0.3cm
      \item Concrete initial states : \\ $I_c(x, y) = (x=0 \land y=1)$
      \item Concrete transitions : \\ $R_c(x, x', y, y') = \newline (x<100) \implies
        (x'=x+y \land y'=x)$
      \item Predicate : $\vec{P} = \{(b_0: (x<200))\}$
      \item<2-> Abstract initial states: \\
        $I_a(b_0) = \exists x,y. (x=0 \land y=1) \land (b_0 \iff x<200) $
      \item<2-> Abstract transitions: \\
        $R_a(b_0, b_0') = \exists x,y,x',y'. \newline ((x<100) \implies (x'=x+y
        \land y'=x) \land \newline
        (b_0 \iff x<200) \land (b_0' \iff x'<200)$
      \end{itemize}
    \end{column}
    \begin{column}{0.35\textwidth}
      \uncover<1->{
      \begin{bbox}
        \begin{alltt}          \intt x = 0;\\
          \intt y = 1;\\
          \while (x < 100) \{\\
          \- \- x = x + y;\\
          \- \- y = x;\\
          \}\\
        \end{alltt}%
      \end{bbox}}
      \newline
      \uncover<2->{
      \ts{abstract initial state:} \\
      \includegraphics[scale=0.25]{init.png} \\
      \ts{abstract transitions:} \\
      \includegraphics[scale=0.25]{trans.png}
    }
    \end{column}
  \end{columns}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Abstraction at Program Points}
  \begin{itemize}
    \itemsep=0.4cm
  \item Compute CFG of the concrete program
  \item Choose a set of predicates
  \item Compute abstractions at each program point in the CFG
    \pause
  \item Intuitively,
    \pause
    \begin{itemize}
      \itemsep=0.3cm
    \item The predicate values at each program point is an invariant
      at the program point 
      \pause
    \item Concretization of the predicates correspond to
      the invariant of the concrete program
    \end{itemize}
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle {Let's begin}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{bbox}
        \begin{alltt}          \intt main ()\\
          \{\\
          \- \- \intt x = 0;\\
          \- \- \intt y = 1;\\
          \- \- \while (x < 100) \{\\
          \- \- \- \- x = x + y;\\
          \- \- \- \- y = x;\\
          \- \- \}\\
          \- \- \ret 0;\\
          \}
        \end{alltt}%
      \end{bbox}
      \begin{itemize}
        \itemsep=0.3cm
      \item compute the CFG
      \item abstract it at program points
      \end{itemize}
    \end{column}

    \begin{column}{0.5\textwidth}
      \includegraphics[scale=0.3]{c.pdf}  
    \end{column}
  \end{columns}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Abstract Transitions}
\begin{minipage}{6.5cm}
  \begin{itemize}
\itemsep=0.4cm
  \item Consider the first concrete transition
  \item Assume predicates: \\ 
    \begin{bbox}
      {\scriptsize{
          $b_0: (x < 200), b_1: (x < 100), b_2: (x + y < 200)$      
        }}
    \end{bbox}
  \item<2-> Compute abstraction w.r.t the predicates:

    \begin{bbox}
      {\scriptsize{
          $
          \exists x_0, x_1, y_0, y_1.\ \newline  
          (x_1=0 \land y_1=1)\ \land \newline
          (b_0 \iff (x_0 < 200) \ \land \ (b_0' \iff (x_1 < 200) \ \land \newline 
          (b_1 \iff (x_0 < 100) \ \land \ (b_1' \iff (x_1 < 100) \ \land  \newline
          (b_2 \iff (x_0 + y_0 < 200) \ \land \ (b_2' \iff (x_1 + y_1 < 200)
          $
        }}
    \end{bbox}

  \item<3-> solve the above using alt-ergo
  \item<3-> solutions: \\ 
    \begin{bbox}
      {\scriptsize{
        $(b_0' = true), (b_1' = true), (b_2' = true)$      
      }}
    \end{bbox}
  \end{itemize}

\end{minipage}
\hspace{1cm}
\begin{minipage}{3cm}
\alt<3->
  { \includegraphics[scale=0.3]{abs-ex-1.pdf} }
  { \includegraphics[scale=0.3]{conc-1.pdf} }
\end{minipage}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Abstract Program - Loop Invariant?}
\begin{columns}
  \begin{column}{0.65\textwidth}
    \begin{itemize}
      \itemsep=0.3cm
    \item<1-> Final abstract system
      \item<2-> Collect the predicates at the loop head (here, 2)
      \item<3-> $(b_0 \land b_1 \land b_2)\  \lor\ (\lnot b_0 \land b_1
        \land \lnot b_2)$ \\ is a loop invariant
      \item<4-> which is: \\
        $ 
        ((x<100) \land (x<200) \land (x+y<200)) \ \lor \newline
        (\lnot (x<100) \land (x<200) \land \lnot (x+y<200))
        $
      \end{itemize}
  \end{column}
\begin{column}{0.35\textwidth}

  \alt<2->{\includegraphics[scale=0.3]{abs-final2.png}}{\includegraphics[scale=0.3]{abs-final.png}}
\end{column}
\end{columns}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Demo}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Part - II}
  \begin{itemize}
    \itemsep=0.4cm
  \item What predicates are interesting?
  \item How to discover new predicates?
  \item How did we collect predicates?
  \end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Predicate Discovery}
  When proving properties about programs, we need \\ \- \\
  \begin{itemize}
    \itemsep=0.4cm
  \item Loop invariants that will help in proving the property
  \item Predicates (interesting) to generate such an invariant
  \item Interesting predicates = useful information about the program
    w.r.t property
  \item To be able to construct abstractions incrementally
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Abstraction - Recap}
  \centering
  \includegraphics[scale=0.3]{Diagram2_new1.png}
\end{frame}

\begin{frame}{Refinement - Intuition}
  \centering
  \includegraphics[scale=0.3]{Diagram3.png}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Predicate Discovery: WP Computation}
  Weakest Precondition Computation \\
  \begin{itemize}
    \itemsep=0.3cm
  \item Given a property P and a loop body $B$
  \item Compute $P_{new} = WP(P, B)$ 
  \item Add the atomic predicates in $P_{new}$ as new predicates
  \item Intuitively, $P_{new}$ is some information relevant to the property
    and driven by the statements in $B$
  \end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Predicate Discovery: Other Techniques}
  Craig Interpolants\\ 
  \begin{itemize}
    \itemsep=0.3cm
  \item Given a pair of mutually inconsistent formulas $\phi_1$ and
    $\phi_2$, there is an interpolant $\psi$, with the property
    \begin{itemize}
      \itemsep=0.2cm
    \item $\phi_1 \rightarrow \psi$
    \item $\psi \land \phi_2$ is unsatisfiable
    \item $\psi$ refers only to the variables common to both $\phi_1$
      and $\phi_2$
  \end{itemize}
\item Intuitively, the interpolant $\psi$ extracts the relevant parts
  of $\phi_1$.
\item If a set of first-order formulas is inconsistent, an interpolant
  always exists and can be computed from the proof of unsatisfiability
  of the candidate formulas.

\end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Predicate Discovery}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      In our example, \\ \- \\
      Suppose we want to prove the assertion $(x < 200)$ \\ \-
      \newline
      \uncover<2->{
      Predicates:}
      \begin{itemize}
      \item<2-> Guards
      \item<2-> The property to be proved
      \item<3-> WP of the property w.r.t loop body $B$
        $$
        WP (x<200, B) =\hl{$x+y < 200$} 
        $$
      \end{itemize}
    \end{column}
    \begin{column}{0.4\textwidth}
      \alt<2->{
      \begin{bbox}
        \begin{alltt}          \intt main ()\\
          \{\\
          \- \- \intt x = 0;\\
          \- \- \intt y = 1;\\
          \- \- \while \hl{(x < 100)} \{\\
          \- \- \- \- x = x + y;\\
          \- \- \- \- y = x;\\
          \- \- \}\\
          \- \- \asrt \hl{(x < 200)};\\
          \- \- \ret 0;\\
          \}
        \end{alltt}%
      \end{bbox}}
    {
      \begin{bbox}
        \begin{alltt}          \intt main ()\\
          \{\\
          \- \- \intt x = 0;\\
          \- \- \intt y = 1;\\
          \- \- \while (x < 100) \{\\
          \- \- \- \- x = x + y;\\
          \- \- \- \- y = x;\\
          \- \- \}\\
          \- \- \asrt (x < 200);\\
          \- \- \ret 0;\\
          \}
        \end{alltt}%
      \end{bbox}
    }
    \end{column}
  \end{columns}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Other Tools/Techniques}
  \begin{itemize}
    \itemsep=0.3cm
  \item Most abstraction techniques abstract the whole program
  \item SLAM abstracts at each program statement but does not
    infer invariants
  \item Most abstraction techniques also implement refinement
    framework \\ (BLAST, CBMC, NuSMT)
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{BLAST - Lazy Abstraction}
  BLAST - Lazy Abstraction
  \begin{itemize}
  \item Most generic predicate is used to start with
  \item Iteratively refined 
  \item Abstraction is coarse - so more iterations
  \end{itemize}

\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{NuSMT: SMT + BDD}
  NuSMT - SMT + BDD-based Abstraction
  \begin{itemize}
  \item Tight integration of BDD enumeration and SMT reasoning
  \item CEGAR loop with predicate discovery techniques
  \item Computes precise abstraction - less iterations
  \item No invariant generation
  \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{SAT-based Abstraction}
CBMC - SAT-based Abstraction
  \begin{itemize}
  \item SAT based abstraction
    \item Encoding in SAT an issue 
    \item Imprecise abstractions - more iterations of CEGAR loop
    \end{itemize}
\end{frame}

%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{Directions for Future Work}
  \begin{itemize}
    \itemsep=0.4cm
  \item Use WP computations to discover predicates
  \item Generating quantified invariants automatically
  \item Iterative abstraction techniques - model-checker?
  \item Use information from abstract interpretation techniques
    in Frama-C \\(APRON library?)
  \end{itemize}

\end{frame}

\bigskip\bigskip
%% -------------------------------------------------------------------- %%

\begin{frame}\frametitle{\begin{center} Merci! \end{center}}

\end{frame}

%% -------------------------------------------------------------------- %%

\bigskip
\end{document}

\bigskip
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2011-05-10 doc/tfp-2011/slides/slides-pp.tex
*** empty log message ***

@@
 % \setbeamertemplate{beaver}
-\documentclass[xcolor=dvipsnames,8pt]{beamer} 
+\documentclass[xcolor=dvipsnames]{beamer} 
 \usecolortheme[named=MidnightBlue]{structure} 
@@
 
-\title{FUNCTORY \\ A Distributed Computing Library for Objective Caml}
+\title{FUNCTORY \\ A Distributed Computing Library \\ for Objective Caml}
 
-\author[Kalyan]{Jean-Christophe Filli\^{a}tre \& Kalyan Krishnamani\\
+\author[Kalyan]{Jean-Christophe Filli\^{a}tre \& K. Kalyanasundaram\\
   \- \\ 
-  INRIA-Saclay \^{I}le-de-France}
+  CNRS / INRIA Saclay -- \^{I}le-de-France}
 \date {}
@@
 
-\begin{frame}\frametitle{Outline}
-  \begin{itemize}
-  \item Motivation
-  \item The Basics
-  \item API
-  \item Implementation Details
-  \item Experiments
-  \item Related Work
-  \item Conclusions
-  \end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
 \begin{frame}\frametitle{Motivation}
   \begin{itemize}
-  \item Why Platform - A set of tools for deductive program verification
-  \item Generates numerous verification conditions (VCs)
-  \item VCs discharged by various automated provers
-  \item Typically takes hours to complete
+  \item In our team, we do deductive program verification
+  \item Generates numerous verification conditions
+  \item Discharged by various automated provers
+  \item Typically takes \emph{hours} to complete
   \end{itemize}
@@
   \begin{itemize}
-  \item Ideally in a functional language (OCaml)
-  \item Robust fault-tolerance
+  \item Fault-tolerance
   \item User-friendly API
+  \item In our favorite programming language (OCaml)
   \end{itemize}
@@
 
-\begin{frame}\frametitle {API}
-A generic distribution function \texttt{compute}:\\
+\begin{frame}\frametitle{Outline}
+  \begin{columns}
+    \column{0.8\textwidth}
+    \begin{itemize}
+    \item API
+      \begin{itemize}
+      \item general-purpose \texttt{compute} function
+      \item high-level: map/fold operations
+      \item low-level: micro-step computations
+      \end{itemize}
+
+\bigskip
+    \item Deployment Scenarios
+      \begin{itemize}
+      \item Sequential
+      \item Cores
+      \item Network
+      \end{itemize}
+    \end{itemize}
+
+    \column{0.2\textwidth}
+  \begin{tabular}{|p{0.2em}|p{0.2em}|p{0.2em}|}
+    \hline
+    && \\\hline
+    && \\\hline
+    && \\\hline
+  \end{tabular}
+\end{columns}
+\end{frame}
+
+%% -------------------------------------------------------------------- %%
+
+\begin{frame}\frametitle{A General-Purpose \texttt{compute} Function}
 
 \medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
+{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\ttfamily\parindent 0pt
 
@@
 
-\begin{frame}\frametitle {Low-level Primitives}
-How to implement \texttt{compute} using low-level primitives?
+\begin{frame}\frametitle {High-Level API}
   \begin{itemize}
-\item \textsf{{\color{blue}type}\hspace*{1.22ex}(\ensuremath{\alpha},\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}computation}
-\item Computation is created using \texttt{create}:
+  \item most common map/fold operations over lists
   \end{itemize}
@@
 \medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
+{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\ttfamily\parindent 0pt
+
+\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\symbol{58}\hspace*{1.22ex}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}list~\linebreak
+\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\_{}fold\symbol{58}\hspace*{1.22ex}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{7.00em}fold\symbol{58}(\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}
+\end{minipage}}}
+
+\medskip\noindent
+  \begin{itemize}
+  \item \texttt{f} operations always in parallel
+  \item Two flavours: \texttt{map\_local\_fold} and
+    \texttt{map\_remote\_fold}
+  \item More parallelism when fold is associative and commutative
+  \end{itemize}
+
+\medskip\noindent
+{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\ttfamily\parindent 0pt
+
+\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\_{}fold\_{}ac,\hspace*{1.22ex}map\_{}fold\_{}a\symbol{58}~\linebreak
+\noindent\hspace*{1.00em}f\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
+\noindent\hspace*{1.00em}fold\symbol{58}\hspace*{1.22ex}(\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}
+\end{minipage}}}
+
+\medskip\noindent
+\end{frame}
+
+%% -------------------------------------------------------------------- %%
+
+\begin{frame}\frametitle {Low-Level API}
+  \begin{itemize}
+  \item User can interact with the execution of the distributed
+    computation
+
+\bigskip
+  \item Examples:
+    \begin{itemize}
+    \item Monitoring applications: observation of consumption of
+      resources, etc
+    \item Interactive programs
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+%% -------------------------------------------------------------------- %%
+
+\begin{frame}\frametitle {Low-Level API}
+\begin{itemize}
+\item \textsf{{\color{blue}type}\hspace*{1.22ex}(\ensuremath{\alpha},\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}computation}
+\item creation
+\end{itemize}
+
+\medskip\noindent
+{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\ttfamily\parindent 0pt
 
@@
   \begin{itemize}
-  \item \texttt{create} does not add tasks. However, tasks could be added using
+  \item adding new tasks
   \end{itemize}
@@
 \medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
+{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\ttfamily\parindent 0pt
 
@@
   \begin{itemize}
-  \item One step of the computation can be performed using
+  \item performing one step of the computation
   \end{itemize}
@@
 \medskip\noindent
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Low-level Primitives}
 \begin{itemize}
-  \item \texttt{compute} could be implemented using low-level primitives:
-  \end{itemize}
-
-\medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
-
-\noindent\hspace*{0.00em}{\color{blue}let}\hspace*{1.22ex}\symbol{126}compute\hspace*{1.22ex}\symbol{126}worker\hspace*{1.22ex}\symbol{126}master\hspace*{1.22ex}tasks\hspace*{1.22ex}=\hspace*{1.22ex}~\linebreak
-\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}c\hspace*{1.22ex}=\hspace*{1.22ex}create\hspace*{1.22ex}worker\hspace*{1.22ex}master\hspace*{1.22ex}{\color{blue}in}~\linebreak
-\noindent\hspace*{1.00em}List.iter\hspace*{1.22ex}(add\_{}task\hspace*{1.22ex}c)\hspace*{1.22ex}tasks~\linebreak
-\noindent\hspace*{1.00em}{\color{blue}while}\hspace*{1.22ex}status\hspace*{1.22ex}c\hspace*{1.22ex}=\hspace*{1.22ex}Running\hspace*{1.22ex}{\color{blue}do}\hspace*{1.22ex}~\linebreak
-\noindent\hspace*{2.50em}one\_{}step\hspace*{1.22ex}c\hspace*{1.22ex}~\linebreak
-\noindent\hspace*{1.00em}{\color{blue}done}\hspace*{1.22ex}\hspace*{1.22ex}\hspace*{1.22ex}\hspace*{1.22ex}
-\end{minipage}}}
-
-\medskip\noindent
-  \begin{itemize}
-  \item User can interact with the execution of the distributed
-    computation
-  \item Monitoring applications: observation of consumption of
-  resources, etc
+\item etc.
 \end{itemize}
@@
 
-\begin{frame}\frametitle {High-level API}
+\begin{frame}\frametitle{Outline}
     \begin{itemize}
-  \item Implementation of the most common map/fold operations over lists
+    \item API
+      \begin{itemize}
+      \item general-purpose \texttt{compute} function
+      \item high-level: map/fold operations
+      \item low-level: micro-step computations
       \end{itemize}
 
-\medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
-
-\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\symbol{58}\hspace*{1.22ex}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}list~\linebreak
-\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\_{}fold\symbol{58}\hspace*{1.22ex}f\symbol{58}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
-\noindent\hspace*{8.00em}fold\symbol{58}(\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\gamma}
-\end{minipage}}}
-
-\medskip\noindent
+\bigskip
+    \item \emph{Deployment Scenarios}
       \begin{itemize}
-  \item \texttt{f} operations are always in parallel.
-  \item Two flavours: \texttt{map\_local\_fold} and
-    \texttt{map\_remote\_fold}
-  \item More parallelism in \texttt{map\_remote\_fold} when fold is
-    associative and commutative
+      \item Sequential
+      \item Cores
+      \item Network
+      \end{itemize}
     \end{itemize}
 
-\medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
+\end{frame}
+% \begin{frame}\frametitle {Deployment Scenarios}
+%   \begin{itemize}
+%   \item Purely sequential execution: for debugging and comparison
+%     purposes
+%   \item Several cores on the same machine
+%   \item Computation distributed over a network of machines
+%   \end{itemize}
 
-\noindent\hspace*{0.00em}{\color{blue}val}\hspace*{1.22ex}map\_{}fold\_{}ac,\hspace*{1.22ex}map\_{}fold\_{}a\symbol{58}~\linebreak
-\noindent\hspace*{0.00em}f\symbol{58}\hspace*{1.22ex}(\ensuremath{\alpha}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}~\linebreak
-\noindent\hspace*{0.00em}fold\symbol{58}\hspace*{1.22ex}(\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta})\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\alpha}\hspace*{1.22ex}list\hspace*{1.22ex}\ensuremath{\rightarrow}\hspace*{1.22ex}\ensuremath{\beta}
-\end{minipage}}}
-
-\medskip\noindent
-\end{frame}
+% \end{frame}
 
@@
 
-\begin{frame}\frametitle {Deployment Scenarios}
-  \begin{itemize}
-  \item Purely sequential execution: for debugging and comparison
-    purposes
-  \item Several cores on the same machine
-  \item Computation distributed over a network of machines
-  \end{itemize}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Cores Implementation: \texttt{Functory.Cores}}
+\begin{frame}\frametitle {Cores Implementation}
   Uses \texttt{Unix.fork} (no control over scheduling)
@@
 
-\begin{frame}\frametitle {Network Implementation: \texttt{Functory.Network}}
+\begin{frame}\frametitle {Network Implementation}
   based on
@@
 
-\begin{frame}\frametitle {\texttt{Functory.Network}: Three Scenarios}
+\begin{frame}\frametitle {Three Implementations of Network}
 
@@
 
-\begin{frame}\frametitle {Details: \texttt{Functory.Network}}
+\begin{frame}\frametitle {Network Implementation Details}
 
 \medskip\noindent
-{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\tt\parindent 0pt
+{\definecolor{tmpcolor}{rgb}{0.80,0.80,1.00}\colorbox{tmpcolor}{\begin{minipage}{\textwidth}\ttfamily\parindent 0pt
 
@@
 
-\begin{frame}\frametitle {Feature: Fault Tolerance}
+\begin{frame}\frametitle {Fault Tolerance}
 
@@
   \begin{center}
-    \includegraphics[scale=0.7]{state.mps}
+    \includegraphics{state.mps}
   \end{center}
@@
 
-\begin{frame}\frametitle {Potential: Experimental Results}
+\begin{frame}\frametitle {Experimental Results}
   Motivating example
   \begin{itemize}
-  \item 80 verification conditions
-  \item 4 provers
-  \item one minute timeout per VC
+  \item 80 verification conditions / 4 provers = 320 tasks
   \item network of 3 machines (4, 8 and 8 cores)
@@
 
-\bigskip
   \begin{itemize}
@@
   \end{itemize}
-  \begin{center}
-  More results in the paper !!    
-  \end{center}
 
+\bigskip
+  More experimental results in the paper: \par N-queens, Mandelbrot set,
+  matrix multiplication
 \end{frame}
@@
 \item Provide robust fault-tolerance mechanism 
+\item Relieve the users from coding messy details
 \end{itemize}

### 2011-05-10 doc/tfp-2011/slides/slides-pp.tex
*** empty log message ***

@@
   \end{itemize}
-  Libraries for existing functional languages
+
+\bigskip
+  Libraries for existing functional languages --- like Functory
   \begin{itemize}
-  \item Plasma MR - OCaml implementation of Google's Map/Reduce on
-    Plasma FS
-  \item iTask - a library for 'Clean' language for distributed workflow management
+  \item Plasma MR - OCaml implementation of Map/Reduce
+  \item iTask - Clean library for distributed workflow management
   \end{itemize}
 
-Our library could be implemented using any of the DFLs, but our goals
-are:
-\begin{itemize}
-\item Provide users of an existing language, user-friendly APIs
-\item Provide robust fault-tolerance mechanism 
-\item Relieve the users from coding messy details
-\end{itemize}
+% Our library could be implemented using any of the DFLs, but our goals
+% are:
+% \begin{itemize}
+% \item Provide users of an existing language, user-friendly APIs
+% \item Provide robust fault-tolerance mechanism 
+% \item Relieve the users from coding messy details
+% \end{itemize}
 
@@
   \begin{itemize}
-  \item Better Scheduling
-  \item Real-time visualization - OcamlViz?
-  \item Any gains from Plasma-FS - data locality?
+  \item User-controlled scheduling
+  \item Real-time visualization --- OcamlViz?
+  \item Any gains from Plasma-FS --- data locality?
   \end{itemize}
@@
 
-\bigskip\bigskip\bigskip
-\begin{frame}\frametitle{The Talk}
-
-\bigskip
-\begin{bbox}
-\begin{alltt}\intt main ()\\
-\{\\
-  \- \- \intt x = 0;\\
-  \- \- \intt y = 1;\\
-  \- \- \while (x < 100) \{\\
-  \- \- \- \- x = x + y;\\
-  \- \- \- \- y = x;\\
-  \- \- \}\\
-  \- \- \asrt (x < 200);\\
-  \- \- \ret 0;\\
-\}
-\end{alltt}%
-\end{bbox}
-\pause
-
-Can we automatically: \\ 
-\begin{itemize}
-  \itemsep=0.3cm
-\item Generate a loop invariant ?
-\item Generate a loop invariant that helps in proving the assertion ?
-
-\end{itemize}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Abstraction - Intuition}
-  \centering
-  \includegraphics[scale=0.3]{Diagram1_new1.png}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Abstraction - Intuition}
-  \centering
-  \includegraphics[scale=0.3]{Diagram2_new1.png}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Predicate Abstraction}
-The Problem: \\ \- \\ Given a concrete program and a set of predicates, track
-the behaviour of the predicates corresponding to the behaviour of the
-concrete program \\ \- \\
-\pause
-Ingredients:
-  \begin{itemize}
-    \itemsep=0.3cm
-  \item Concrete program: $R(\vec{x}, \vec{x}')$
-  \item A set of predicates: \{$\vec{P} \iff \alpha (\vec{x})$, $\vec{P}' \iff
-    \alpha (\vec{x'})$ \} 
-  \item Behaviour of the predicates w.r.t the concrete program:
-    $$
-    \exists. \vec{x}, \vec{x}'.\ R(\vec{x}, \vec{x}') \land \vec{P} =
-    \alpha (\vec{x}) \land \vec{P}' = \alpha (\vec{x'})
-    $$
-  \item Existential quantification - solved using a SMT solver
-  \item Abstract interpretation with Abstract domain =
-      vector of Booleans
-  \end{itemize}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Predicate Abstraction - Example}
-  \begin{columns}
-    \begin{column}{0.65\textwidth}
-      \begin{itemize}
-        \itemsep=0.3cm
-      \item Concrete initial states : \\ $I_c(x, y) = (x=0 \land y=1)$
-      \item Concrete transitions : \\ $R_c(x, x', y, y') = \newline (x<100) \implies
-        (x'=x+y \land y'=x)$
-      \item Predicate : $\vec{P} = \{(b_0: (x<200))\}$
-      \item<2-> Abstract initial states: \\
-        $I_a(b_0) = \exists x,y. (x=0 \land y=1) \land (b_0 \iff x<200) $
-      \item<2-> Abstract transitions: \\
-        $R_a(b_0, b_0') = \exists x,y,x',y'. \newline ((x<100) \implies (x'=x+y
-        \land y'=x) \land \newline
-        (b_0 \iff x<200) \land (b_0' \iff x'<200)$
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.35\textwidth}
-      \uncover<1->{
-      \begin{bbox}
-        \begin{alltt}          \intt x = 0;\\
-          \intt y = 1;\\
-          \while (x < 100) \{\\
-          \- \- x = x + y;\\
-          \- \- y = x;\\
-          \}\\
-        \end{alltt}%
-      \end{bbox}}
-      \newline
-      \uncover<2->{
-      \ts{abstract initial state:} \\
-      \includegraphics[scale=0.25]{init.png} \\
-      \ts{abstract transitions:} \\
-      \includegraphics[scale=0.25]{trans.png}
-    }
-    \end{column}
-  \end{columns}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Abstraction at Program Points}
-  \begin{itemize}
-    \itemsep=0.4cm
-  \item Compute CFG of the concrete program
-  \item Choose a set of predicates
-  \item Compute abstractions at each program point in the CFG
-    \pause
-  \item Intuitively,
-    \pause
-    \begin{itemize}
-      \itemsep=0.3cm
-    \item The predicate values at each program point is an invariant
-      at the program point 
-      \pause
-    \item Concretization of the predicates correspond to
-      the invariant of the concrete program
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle {Let's begin}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-      \begin{bbox}
-        \begin{alltt}          \intt main ()\\
-          \{\\
-          \- \- \intt x = 0;\\
-          \- \- \intt y = 1;\\
-          \- \- \while (x < 100) \{\\
-          \- \- \- \- x = x + y;\\
-          \- \- \- \- y = x;\\
-          \- \- \}\\
-          \- \- \ret 0;\\
-          \}
-        \end{alltt}%
-      \end{bbox}
-      \begin{itemize}
-        \itemsep=0.3cm
-      \item compute the CFG
-      \item abstract it at program points
-      \end{itemize}
-    \end{column}
-
-    \begin{column}{0.5\textwidth}
-      \includegraphics[scale=0.3]{c.pdf}  
-    \end{column}
-  \end{columns}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Abstract Transitions}
-\begin{minipage}{6.5cm}
-  \begin{itemize}
-\itemsep=0.4cm
-  \item Consider the first concrete transition
-  \item Assume predicates: \\ 
-    \begin{bbox}
-      {\scriptsize{
-          $b_0: (x < 200), b_1: (x < 100), b_2: (x + y < 200)$      
-        }}
-    \end{bbox}
-  \item<2-> Compute abstraction w.r.t the predicates:
-
-    \begin{bbox}
-      {\scriptsize{
-          $
-          \exists x_0, x_1, y_0, y_1.\ \newline  
-          (x_1=0 \land y_1=1)\ \land \newline
-          (b_0 \iff (x_0 < 200) \ \land \ (b_0' \iff (x_1 < 200) \ \land \newline 
-          (b_1 \iff (x_0 < 100) \ \land \ (b_1' \iff (x_1 < 100) \ \land  \newline
-          (b_2 \iff (x_0 + y_0 < 200) \ \land \ (b_2' \iff (x_1 + y_1 < 200)
-          $
-        }}
-    \end{bbox}
-
-  \item<3-> solve the above using alt-ergo
-  \item<3-> solutions: \\ 
-    \begin{bbox}
-      {\scriptsize{
-        $(b_0' = true), (b_1' = true), (b_2' = true)$      
-      }}
-    \end{bbox}
-  \end{itemize}
-
-\end{minipage}
-\hspace{1cm}
-\begin{minipage}{3cm}
-\alt<3->
-  { \includegraphics[scale=0.3]{abs-ex-1.pdf} }
-  { \includegraphics[scale=0.3]{conc-1.pdf} }
-\end{minipage}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Abstract Program - Loop Invariant?}
-\begin{columns}
-  \begin{column}{0.65\textwidth}
-    \begin{itemize}
-      \itemsep=0.3cm
-    \item<1-> Final abstract system
-      \item<2-> Collect the predicates at the loop head (here, 2)
-      \item<3-> $(b_0 \land b_1 \land b_2)\  \lor\ (\lnot b_0 \land b_1
-        \land \lnot b_2)$ \\ is a loop invariant
-      \item<4-> which is: \\
-        $ 
-        ((x<100) \land (x<200) \land (x+y<200)) \ \lor \newline
-        (\lnot (x<100) \land (x<200) \land \lnot (x+y<200))
-        $
-      \end{itemize}
-  \end{column}
-\begin{column}{0.35\textwidth}
-
-  \alt<2->{\includegraphics[scale=0.3]{abs-final2.png}}{\includegraphics[scale=0.3]{abs-final.png}}
-\end{column}
-\end{columns}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Demo}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Part - II}
-  \begin{itemize}
-    \itemsep=0.4cm
-  \item What predicates are interesting?
-  \item How to discover new predicates?
-  \item How did we collect predicates?
-  \end{itemize}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Predicate Discovery}
-  When proving properties about programs, we need \\ \- \\
-  \begin{itemize}
-    \itemsep=0.4cm
-  \item Loop invariants that will help in proving the property
-  \item Predicates (interesting) to generate such an invariant
-  \item Interesting predicates = useful information about the program
-    w.r.t property
-  \item To be able to construct abstractions incrementally
-  \end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Abstraction - Recap}
-  \centering
-  \includegraphics[scale=0.3]{Diagram2_new1.png}
-\end{frame}
-
-\begin{frame}{Refinement - Intuition}
-  \centering
-  \includegraphics[scale=0.3]{Diagram3.png}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Predicate Discovery: WP Computation}
-  Weakest Precondition Computation \\
-  \begin{itemize}
-    \itemsep=0.3cm
-  \item Given a property P and a loop body $B$
-  \item Compute $P_{new} = WP(P, B)$ 
-  \item Add the atomic predicates in $P_{new}$ as new predicates
-  \item Intuitively, $P_{new}$ is some information relevant to the property
-    and driven by the statements in $B$
-  \end{itemize}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Predicate Discovery: Other Techniques}
-  Craig Interpolants\\ 
-  \begin{itemize}
-    \itemsep=0.3cm
-  \item Given a pair of mutually inconsistent formulas $\phi_1$ and
-    $\phi_2$, there is an interpolant $\psi$, with the property
-    \begin{itemize}
-      \itemsep=0.2cm
-    \item $\phi_1 \rightarrow \psi$
-    \item $\psi \land \phi_2$ is unsatisfiable
-    \item $\psi$ refers only to the variables common to both $\phi_1$
-      and $\phi_2$
-  \end{itemize}
-\item Intuitively, the interpolant $\psi$ extracts the relevant parts
-  of $\phi_1$.
-\item If a set of first-order formulas is inconsistent, an interpolant
-  always exists and can be computed from the proof of unsatisfiability
-  of the candidate formulas.
-
-\end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Predicate Discovery}
-  \begin{columns}
-    \begin{column}{0.6\textwidth}
-      In our example, \\ \- \\
-      Suppose we want to prove the assertion $(x < 200)$ \\ \-
-      \newline
-      \uncover<2->{
-      Predicates:}
-      \begin{itemize}
-      \item<2-> Guards
-      \item<2-> The property to be proved
-      \item<3-> WP of the property w.r.t loop body $B$
-        $$
-        WP (x<200, B) =\hl{$x+y < 200$} 
-        $$
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.4\textwidth}
-      \alt<2->{
-      \begin{bbox}
-        \begin{alltt}          \intt main ()\\
-          \{\\
-          \- \- \intt x = 0;\\
-          \- \- \intt y = 1;\\
-          \- \- \while \hl{(x < 100)} \{\\
-          \- \- \- \- x = x + y;\\
-          \- \- \- \- y = x;\\
-          \- \- \}\\
-          \- \- \asrt \hl{(x < 200)};\\
-          \- \- \ret 0;\\
-          \}
-        \end{alltt}%
-      \end{bbox}}
-    {
-      \begin{bbox}
-        \begin{alltt}          \intt main ()\\
-          \{\\
-          \- \- \intt x = 0;\\
-          \- \- \intt y = 1;\\
-          \- \- \while (x < 100) \{\\
-          \- \- \- \- x = x + y;\\
-          \- \- \- \- y = x;\\
-          \- \- \}\\
-          \- \- \asrt (x < 200);\\
-          \- \- \ret 0;\\
-          \}
-        \end{alltt}%
-      \end{bbox}
-    }
-    \end{column}
-  \end{columns}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Other Tools/Techniques}
-  \begin{itemize}
-    \itemsep=0.3cm
-  \item Most abstraction techniques abstract the whole program
-  \item SLAM abstracts at each program statement but does not
-    infer invariants
-  \item Most abstraction techniques also implement refinement
-    framework \\ (BLAST, CBMC, NuSMT)
-  \end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{BLAST - Lazy Abstraction}
-  BLAST - Lazy Abstraction
-  \begin{itemize}
-  \item Most generic predicate is used to start with
-  \item Iteratively refined 
-  \item Abstraction is coarse - so more iterations
-  \end{itemize}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{NuSMT: SMT + BDD}
-  NuSMT - SMT + BDD-based Abstraction
-  \begin{itemize}
-  \item Tight integration of BDD enumeration and SMT reasoning
-  \item CEGAR loop with predicate discovery techniques
-  \item Computes precise abstraction - less iterations
-  \item No invariant generation
-  \end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{SAT-based Abstraction}
-CBMC - SAT-based Abstraction
-  \begin{itemize}
-  \item SAT based abstraction
-    \item Encoding in SAT an issue 
-    \item Imprecise abstractions - more iterations of CEGAR loop
-    \end{itemize}
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{Directions for Future Work}
-  \begin{itemize}
-    \itemsep=0.4cm
-  \item Use WP computations to discover predicates
-  \item Generating quantified invariants automatically
-  \item Iterative abstraction techniques - model-checker?
-  \item Use information from abstract interpretation techniques
-    in Frama-C \\(APRON library?)
-  \end{itemize}
-
-\end{frame}
-
 \bigskip\bigskip
-%% -------------------------------------------------------------------- %%
-
-\begin{frame}\frametitle{\begin{center} Merci! \end{center}}
-
-\end{frame}
-
-%% -------------------------------------------------------------------- %%
-
-\bigskip
-\end{document}
-
-\bigskip
 %%% Local Variables: 

### 2011-09-14 doc/tfp_2011_final/slides/slides-pp.tex
for the final version KK

--- 2011-05-10 doc/tfp-2011/slides/slides-pp.tex
@@
   \- \\ 
-  CNRS / INRIA Saclay -- \^{I}le-de-France}
+  CNRS / INRIA Saclay -- \^{I}le-de-France
+\\
+\vspace{0.5cm}
+TFP 2011, Madrid, Spain
+}
+
+\bigskip
 \date {}
@@
   \end{itemize}
+
+\bigskip
+Also
+  \begin{itemize}
+  \item General purpose library
+  \item Portability
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}\frametitle{Outline}
-  \begin{columns}
-    \column{0.8\textwidth}
     \begin{itemize}
@@
       \item Cores
-      \item Network
-      \end{itemize}
+      \item Network (3 flavours)
       \end{itemize}
 
-    \column{0.2\textwidth}
-  \begin{tabular}{|p{0.2em}|p{0.2em}|p{0.2em}|}
-    \hline
-    && \\\hline
-    && \\\hline
-    && \\\hline
-  \end{tabular}
-\end{columns}
+\bigskip
+    \item Many libraries in one
+
+    \end{itemize}
+
 \end{frame}
@@
 \noindent\hspace*{1.00em}{\color{blue}open}\hspace*{1.22ex}Cores~\linebreak
-\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}set\_{}number\_{}of\_{}cores\hspace*{1.22ex}4
+\noindent\hspace*{1.00em}{\color{blue}let}\hspace*{1.22ex}()\hspace*{1.22ex}=\hspace*{1.22ex}set\_{}number\_{}of\_{}cores\hspace*{1.22ex}3
 \end{minipage}}}
@@
 
-\begin{frame}\frametitle {Underneath: Protocol}
+\begin{frame}\frametitle {Protocol}
   \only<1>{Master sends a task to a worker}
@@
     \only<8>{\textsf{Completed\hspace*{1.22ex}(43,\hspace*{1.22ex}b)}}
-    \only<9>{\textsf{Kill\hspace*{1.22ex}43}}
-    \only<10>{\textsf{Aborted\hspace*{1.22ex}44}}
+    \only<9>{\textsf{Kill\hspace*{1.22ex}42}}
+    \only<10>{\textsf{Aborted\hspace*{1.22ex}42}}
     \only<11>{\textsf{Stop}}
@@
   \end{center}
+
+  \begin{itemize}
+  \item Time-out values for deciding the status of the worker
+  \end{itemize}
 
@@
   \begin{itemize}
-  \item Jo\&Caml - rich communication primitives, no easy features
+  \item Jo\&Caml - rich communication primitives, no primitive features
     for fault-tolerance
-  \item ML5 - code mobility, marshalling, etc, but no
-    easy features for fault-tolerance
+  \item ML5 - code mobility, type-safe marshalling, etc, but no
+    primitive features for fault-tolerance
   \item Glasgow Distributed Haskell - features for fault-tolerance,
@@
 
-\begin{frame}\frametitle {What's next}
+\begin{frame}\frametitle {Future Work}
   \begin{itemize}
-  \item User-controlled scheduling
-  \item Real-time visualization --- OcamlViz?
-  \item Any gains from Plasma-FS --- data locality?
+  \item More user control
+    \begin{itemize}
+    \item Scheduling of tasks, etc
+    \end{itemize}
+
+\bigskip
+  \item Real-time visualization 
+    \begin{itemize}
+    \item Resource consumption, task distribution patterns, etc.
+    \end{itemize}
+
+\bigskip
+  \item Speeding up using idle workers
   \end{itemize}

### 2011-09-14 doc/tfp_2011_final/main.tex
for the final version

--- 2011-03-31 doc/tfp-2011/main.tex
@@
 `Clean' targeted at distributed workflow management. The library
-provides a set of combinator functions (some of which perform map/fold
+provides a set of combinators (some of which perform map/fold
 operations) that facilitate applications running in different nodes
@@
 
-\paragraph{Acknowledgements.}
+\paragraph{Acknowledgments.}
 The authors are grateful to the ProVal team for support and comments
@@
 % LocalWords:  genericity executables OCamlMPI MapReduce marshalling functors
+% LocalWords:  workflow combinators granularities

### 2011-09-16 doc/tfp_2011_final/main.tex
started revision KK

@@
 % \documentclass[tfpsymp,pagenumbers]{tfp07symp}
+
 \documentclass{llncs}
@@
 \newcommand{\functory}{\textsf{Functory}}
-\newcommand{\JoCaml}{Jo{\&\!}Caml}
+%\newcommand{\JoCaml}{Jo{\&\!}Caml}
 \newcommand{\unix}{\textsc{Unix}}
@@
   languages that provide built-in primitives for
-  distribution. Examples include ML5, \JoCaml\, Glasgow Distributed
+  distribution. Examples include ML5, JoCaml, Glasgow Distributed
   Haskell, Erlang, etc.
@@
 
-\JoCaml\ is one of the DFLs which provides communication primitives
+JoCaml is one of the DFLs which provides communication primitives
 (like channels) for facilitating transmission of
@@
 setting. The user has to include code for fault-tolerance, as already
-demonstrated in some \JoCaml\
+demonstrated in some JoCaml
 library~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
@@
 of a distributed system to communicate, exchange information and
-coordinate their computations.
+coordinate their computations in a type-safe manner.
 
@@
 distributed computing libraries on top of which one could implement
-the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
-them. However, \JoCaml\ does not provide fault-tolerance, which is
+the library discussed in this paper. JoCaml~\cite{jocaml} is one of
+them. However, JoCaml does not provide fault-tolerance, which is
 indispensable in a distributed setting. The user has to include code
-for fault-tolerance, as already demonstrated in some \JoCaml\
+for fault-tolerance, as already demonstrated in some JoCaml
 experiments~\cite{mandel2008}.
@@
 
-In most cases, the easiest way to parallelize an execution it to make
+In most cases, the easiest way to parallelize an execution is to make
 use of operations over lists, where processing of the
@@
 It is rather straightforward to derive these five functions from the
-generic \of{compute} function; readers interested in details could
+generic \of{compute} function; we invite readers interested in details to
 refer to the source code.
@@
   As mentioned in Section~\ref{sec:API}, the \of{Network} module
-  actually provides three different implementations, according to
-  three different execution scenarios. There are provided in three
-  sub-modules, as described below.
+  actually provides three different implementations as sub-modules, according to
+  three different execution scenarios, the details of which are
+  presented below:
 
@@
   \end{ocaml}
-  The master uses marshalling to send both a closure of type \of{'a ->
-    'b} and a task of type \of{'a} to the worker. The resulting
+  The master uses marshalling to send both a closure of type \of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting
   strings are passed as argument \of{f} and \of{x} in message
@@
 For this comparison to be fair, all
-computations are performed on the same machine, a 8 core Intel Xeon
+computations are performed on the same machine, an 8 core Intel Xeon
 3.2 GHz running Debian Linux. The sequential implementation uses a
@@
 exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; 
-greater values for $D$ would result in two many tasks.
+greater values for $D$ would result in too many tasks.
 
@@
 computation of a whole row of the resultant matrix.
-As a consequence, the total number of tasks in
+As a consequence, the total number of tasks is
 $10,000$ for \of{mm1} and only $100$ for \of{mm2}.
@@
 If the total number of tasks $\of{t} \ge 1$ is given as a
-parameter,  it is immediate to split the image into \of{t} sub-images,
+parameter,  it is straight forward to split the image into \of{t} sub-images,
 each of which is computed in parallel with and independently of the
@@
 easily than in the case of n-queens. We experimentally figured out the
-optimal number of tasks to be 30. One the other hand, each computation
+optimal number of tasks to be 30. On the other hand, each computation
 result is an image, rather than just an integer as in the case of

### 2011-09-16 doc/tfp_2011_final/main.tex
started revision KK

@@
   it provides automatic task distribution and
-  a robust fault-tolerance mechanisms, thereby
+  a robust fault-tolerance mechanism, thereby
   relieving the user from implementing such routines.
@@
 needs, from desktop PCs to networks of machines.  Typical applications
-would involve executing a large number of computation expensive tasks
+would involve executing a large number of computationally expensive tasks
 in a resource-optimal and time-efficient manner. This is also the case

### 2011-09-18 doc/tfp_2011_final/main.tex
updated main KK

@@
 % 3. implementation
-%    3.1 marshalling
+%    3.1 marshaling
 %    3.2 protocol
@@
 
-Our library makes use of \Ocaml's marshalling capabilities as much as
+Our library makes use of \Ocaml's marshaling capabilities as much as
 possible. Whenever master and worker executables are exactly the same,
@@
 the client and the server, as well as low-level primitives for
-marshalling the data. As in the case before, ML5 is a programming
+marshaling the data. As in the case before, ML5 is a programming
 language that offers primitives for code mobility, and the code for
 distribution of computation and fault-tolerance has to be included by
-the user.  ML5 implements type-safe marshalling and
-\functory\ does not, though an existing type-safe marshalling library
+the user.  ML5 implements type-safe marshaling and
+\functory\ does not, though an existing type-safe marshaling library
 could be used with \functory. 
 % Basically, the task will be transmitted
-% along with the corresponding type and on unmarshalling, it could be
+% along with the corresponding type and on unmarshaling, it could be
 % matched for the correct type.
@@
     compiled with the same version of \Ocaml:} we can no
-  longer use marshalling, so API functions are restricted to work on
+  longer use marshaling, so API functions are restricted to work on
   strings instead of polymorphic values.
@@
   \end{ocaml}
-  The master uses marshalling to send both a closure of type \of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting
+  The master uses marshaling to send both a closure of type \of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting
   strings are passed as argument \of{f} and \of{x} in message
-  \of{Assign}. Similarly, the worker uses marshalling to send back the
+  \of{Assign}. Similarly, the worker uses marshaling to send back the
   result of the computation of type \of{'b}, which is the argument
-  \of{s} in message \of{Completed}.
+  \of{s} in message \of{Completed}. These messages are described in
+  detail in Section~\ref{sec:protocol}.
 
@@
   still marshal polymorphic values. Indeed, an interesting property of
-  marshalling in \Ocaml\ is to be fully architecture-independent, as
+  marshaling in \Ocaml\ is to be fully architecture-independent, as
   long as a single version of \Ocaml\ is used. It is worth pointing
@@
   module of choice in our library, since it can still pass polymorphic
-  values over the network. The issues of marshalling are automatically
+  values over the network. The issues of marshaling are automatically
   taken care of by the \Ocaml\ runtime.
@@
   \paragraph{\of{Mono}.} When master and workers are compiled using
-  different versions of \Ocaml, we can no longer use marshalling.  As
+  different versions of \Ocaml, we can no longer use marshaling.  As
   in the previous case, we split \of{compute} into two functions, one
@@
 16$\times$ is already a significant improvement for our day-to-day
-experiments.
+experiments. Further a large parallelizable computation could be
+distributed by just adding 3-4 lines of code (to just specify the module
+to be used and the tasks) which is an important user-friendly feature
+of the library.
 
@@
 same program.  Finally, the low-level API of \functory\ 
-can be used to write interactive programs.
+can be used to write interactive programs where one can adjust certain
+parameters in a GUI, like increasing or decreasing the number of
+workers, to observe the progress in computation, resource consumption, etc.
 
@@
 % LocalWords:  parallelize functor parameterized indices endianness monomorphic
-% LocalWords:  genericity executables OCamlMPI MapReduce marshalling functors
+% LocalWords:  genericity executables OCamlMPI MapReduce marshaling functors
 % LocalWords:  workflow combinators granularities

### 2011-09-29 doc/tfp_2011_final/main.tex
updated KK

@@
 w.r.t. deployment of the user program than OCamlMPI; on the other
-hand, we provide a more generic API together with fault-tolerance.
+hand, we provide a more generic API together with fault-tolerance. We
+feel that an indispensible need for any distributed computing library
+is fault-tolerance, and using a homemade protocol enables us to
+tune our implementation to our needs of fault-tolerance.
 
@@
 \end{center}
+
 These figures sum up to more than 6 hours if provers were executed
@@
 experiments. Further a large parallelizable computation could be
-distributed by just adding 3-4 lines of code (to just specify the module
-to be used and the tasks) which is an important user-friendly feature
-of the library.
+distributed by just adding 3-4 lines of code (to just specify the
+module to be used and the tasks) which is an important user-friendly
+feature of the library. Further we assume files available over
+NFS. Intelligent distribution of data over a network is in itself an
+area of research which is beyond the scope of our work.
 

### 2011-09-29 doc/tfp_2011_final/main.tex
*** empty log message ***

@@
 
-\author{Jean-Christophe Filli\^{a}tre \and K. Kalyanasundaram}
+\author{Jean-Christophe Filli\^{a}tre and K. Kalyanasundaram}
 \authorrunning{J.-C. Filli\^atre \and K. Kalyanasundaram}
 
-\institute{
-  CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
+\institute{CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
   INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893 \\
@@
 %\vfill\pagebreak
-\bibliographystyle{plain}
-\bibliography{./biblio}
+%\bibliographystyle{plain}
+%\bibliography{./biblio}
+\begin{thebibliography}{10}
+
+\bibitem{camlP3l}
+{CamlP3l}.
+\newblock \url{http://camlp3l.inria.fr/}.
+
+\bibitem{plasma}
+{Plasma}.
+\newblock \url{http://plasma.camlcity.org/plasma}.
+
+\bibitem{erlang}
+{The Erlang Programming Language}.
+\newblock \url{http://www.erlang.org/}.
+
+\bibitem{GMP}
+{The GNU Multiple Precision Arithmetic Library}.
+\newblock \url{http://gmplib.org/}.
+
+\bibitem{PASCO}
+{Parallel Symbolic Computation 2010 (PASCO)}, 2010.
+\newblock \url{http://pasco2010.imag.fr/}.
+
+\bibitem{mapreduce}
+Jeffrey Dean and Sanjay Ghemawat.
+\newblock {MapReduce: Simplified Data Processing on Large Clusters}.
+\newblock In {\em OSDI}, pages 137--150, 2004.
+
+\bibitem{filliatre07cav}
+Jean-Christophe Filli\^atre and Claude March\'e.
+\newblock {The {Why/Krakatoa/Caduceus} Platform for Deductive Program
+  Verification}.
+\newblock In Werner Damm and Holger Hermanns, editors, {\em 19th International
+  Conference on Computer Aided Verification}, volume 4590 of {\em Lecture Notes
+  in Computer Science}, pages 173--177, Berlin, Germany, July 2007. Springer.
+
+\bibitem{iTask}
+Jan~Martin Jansen, Rinus Plasmeijer, Pieter Koopman, and Peter Achten.
+\newblock {Embedding a Web-based Workflow Management System in a Functional
+  Language}.
+\newblock In {\em Proceedings of the Tenth Workshop on Language Descriptions,
+  Tools and Applications}, LDTA '10, pages 7:1--7:8, New York, NY, USA, 2010.
+  ACM.
+
+\bibitem{ocamlMPI}
+Xavier Leroy.
+\newblock {OCamlMPI: Interface with the MPI Message-passing Interface}.
+\newblock \url{http://pauillac.inria.fr/~xleroy/software.html}.
+
+\bibitem{mandel2008}
+Louis Mandel and Luc Maranget.
+\newblock Programming in {JoCaml} (tool demonstration).
+\newblock In {\em 17th European Symposium on Programming ({ESOP 2008})}, pages
+  108--111, Budapest, Hungary, April 2008.
+
+\bibitem{ML5}
+Tom Murphy, VII., Karl Crary, and Robert Harper.
+\newblock {Type-safe Distributed Programming with ML5}.
+\newblock In {\em Proceedings of the 3rd conference on Trustworthy global
+  computing}, TGC'07, pages 108--123, Berlin, Heidelberg, 2008.
+  Springer-Verlag.
+
+\bibitem{poor-man-mapreduce}
+Yoann Padioleau.
+\newblock {A Poor Man's MapReduce for OCaml}, 2009.
+\newblock \url{http://www.padator.org/ocaml/mapreduce.pdf}.
+
+\bibitem{GdH}
+Robert~F. Pointon, Philip~W. Trinder, and Hans-Wolfgang Loidl.
+\newblock {The Design and Implementation of Glasgow Distributed Haskell}.
+\newblock In {\em Selected Papers from the 12th International Workshop on
+  Implementation of Functional Languages}, IFL '00, pages 53--70, London, UK,
+  2001.
+
+\end{thebibliography}
 

### 2011-09-29 doc/tfp_2011_final/main.tex
*** empty log message ***

@@
 w.r.t. the sequential implementation.
-
 The first column shows the value of $N$.  The number of tasks is shown in
@@
   \end{center}
+\begin{figure}[t]
+  \vspace*{-4em}
+  \begin{center}
+    \includegraphics[width=\textwidth]{chart1.pdf}
+  \end{center}
+  \vspace*{-5em}
+  \caption{Speedup ratios for the $N$-queens experiment.}
+\label{fig:queens}
+\end{figure}
 From the table above, it is clear that the \of{Cores} and \of{Network}
@@
 \paragraph{Acknowledgments.}
-The authors are grateful to the ProVal team for support and comments
+We are grateful to the ProVal team for support and comments
 on early versions of the library and of this paper.
+We thank the anonymous reviewers for their helpful comments and
+suggestions.
 

### 2011-09-29 doc/tfp_2011_final/main.tex
*** empty log message ***

@@
 seconds for the three implementations. The figures within brackets
-show the speedup w.r.t. sequential implementation.
+show the speedup w.r.t. sequential implementation. Speedup ratios are
+also displayed in Fig.~\ref{fig:queens} (note the logarithmic scale).
   \begin{center}
@@
 \end{figure}
-From the table above, it is clear that the \of{Cores} and \of{Network}
+From the table above and Fig.~\ref{fig:queens},
+it is clear that the \of{Cores} and \of{Network}
 implementations provide a significant speedup. As evident from the
